### 2017-01-24 paper/paper.tex
Paper

\documentclass[letterpaper]{article}
\usepackage[pass,showframe]{geometry}

\usepackage{ijcai17}
\usepackage{times}

\title{A Partitioning Algorithm for Maximum Common (Connected) Subgraphs\thanks{This work
was supported by the Engineering and Physical Sciences Research Council [grant
numbers EP/K503058/1 and EP/M508056/1]}}
\author{Ciaran McCreesh \and Patrick Prosser \and James Trimble \\
University of Glasgow, Glasgow, Scotland \\
j.trimble.1@research.gla.ac.uk}

\begin{document}

\maketitle

\begin{abstract}
    James came up with an awesome new algorithm for the maximum common (connected)
    subgraph problem.
\end{abstract}

\section{Introduction}

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{paper}

\end{document}

### 2017-01-25 code/james-c/mcsp.c
Fixed a segfault bug where too few lists were preallocated

@@ -569,7 +569,7 @@ int main(int argc, char** argv) {
     }
 
     start = clock();
-    preallocated_lists = malloc(g0->n * sizeof(struct BidomainList));
+    preallocated_lists = malloc((g0->n+1) * sizeof(struct BidomainList));
     calculate_all_degrees(g0);
     calculate_all_degrees(g1);
 

### 2017-01-25 code/james-c/mcsp.c
Added big-first option

@@ -58,6 +58,7 @@ static struct argp_option options[] = {
     {"lad", 'l', 0, 0, "Read LAD format"},
     {"timeout", 't', "TIMEOUT", 0, "Set timeout of TIMEOUT seconds"},
     {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
     { 0 }
 };
 
@@ -65,6 +66,7 @@ static struct {
     bool quiet;
     bool verbose;
     bool connected;
+    bool big_first;
     bool dimacs;
     bool lad;
     int timeout;
@@ -77,6 +79,7 @@ void set_default_arguments() {
     arguments.quiet = false;
     arguments.verbose = false;
     arguments.connected = false;
+    arguments.big_first = false;
     arguments.dimacs = false;
     arguments.lad = false;
     arguments.timeout = 0;
@@ -99,6 +102,9 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             break;
         case 'q':
             arguments.quiet = true;
+            break;
+        case 'b':
+            arguments.big_first = true;
             break;
         case 'v':
             arguments.verbose = true;
@@ -189,6 +195,7 @@ struct D {
     struct VtxPairList *incumbent;
     struct VtxPairList *current;
     struct BidomainList *domains;
+    int matching_size_goal;
 };
 
 struct BidomainList *preallocated_lists;
@@ -408,7 +415,11 @@ void solve(struct D d, int level) {
         if (!arguments.quiet) printf("Incumbent size: %d\n", d.incumbent->len);
     }
 
-    if (d.current->len + calc_bound(d.domains) <= d.incumbent->len)
+    int bound = d.current->len + calc_bound(d.domains);
+    if (bound <= d.incumbent->len || bound < d.matching_size_goal)
+        return;
+
+    if (arguments.big_first && d.incumbent->len==d.matching_size_goal)
         return;
 
     struct Bidomain *bd = select_bidomain(d.domains, d.current->len);
@@ -425,6 +436,7 @@ void solve(struct D d, int level) {
 
     struct D new_d = d;
     new_d.domains = &preallocated_lists[level];
+    new_d.domains->len = 0;
 
[…]

### 2017-01-25 code/james-cpp/mcsp.c
Added a first attempt at a C++ version of James's code

@@ -1,0 +1,381 @@
+#include "graph.h"
+
+#include <argp.h>
+#include <limits.h>
+#include <locale.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <time.h>
+
+#include <iostream>
+#include <string>
+#include <utility>
+#include <vector>
+
+using std::vector;
+
+static void fail(std::string msg) {
+    std::cerr << msg << std::endl;
+    exit(1);
+}
+
+/*******************************************************************************
+                             Command-line arguments
+*******************************************************************************/
+
+static char doc[] = "Find a maximum clique in a graph in DIMACS format";
+static char args_doc[] = "FILENAME1 FILENAME2";
+static struct argp_option options[] = {
+    {"quiet", 'q', 0, 0, "Quiet output"},
+    {"verbose", 'v', 0, 0, "Verbose output"},
+    {"dimacs", 'd', 0, 0, "Read DIMACS format"},
+    {"lad", 'l', 0, 0, "Read LAD format"},
+    {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    { 0 }
+};
+
+static struct {
+    bool quiet;
+    bool verbose;
+    bool connected;
+    bool dimacs;
+    bool lad;
+    char *filename1;
+    char *filename2;
+    int arg_num;
+} arguments;
+
+void set_default_arguments() {
+    arguments.quiet = false;
+    arguments.verbose = false;
+    arguments.connected = false;
+    arguments.dimacs = false;
+    arguments.lad = false;
+    arguments.filename1 = NULL;
+    arguments.filename2 = NULL;
+    arguments.arg_num = 0;
+}
+
+static error_t parse_opt (int key, char *arg, struct argp_state *state) {
+    switch (key) {
+        case 'd':
+            if (arguments.lad)
+                fail("The -d and -l options cannot be used together.\n");
+            arguments.dimacs = true;
+            break;
+        case 'l':
+            if (arguments.dimacs)
+                fail("The -d and -l options cannot be used together.\n");
+            arguments.lad = true;
+            break;
+        case 'q':
+            arguments.quiet = true;
+            break;
+        case 'v':
+            arguments.verbose = true;
+            break;
+        case 'c':
+            arguments.connected = true;
+            break;
+        case ARGP_KEY_ARG:
+            if (arguments.arg_num == 0) {
+                arguments.filename1 = arg;
[…]

### 2017-01-25 code/james-c/mcsp.c
Show upper bound's progress whnen -b option is used

@@ -524,6 +524,7 @@ struct VtxPairList mcs(struct Graph *g0, struct Graph *g1) {
             int goal = g0->n - k;
             build_domains_and_solve(g0, g1, &incumbent, goal, all_labels, all_labels_len);
             if (incumbent.len == goal) break;
+            printf("Upper bound: %d\n", goal - 1);
         }
     } else {
         build_domains_and_solve(g0, g1, &incumbent, 1, all_labels, all_labels_len);

### 2017-01-25 code/james-c/mcsp.c
Tidied up partitioning slightly

@@ -244,9 +244,6 @@ struct Bidomain *select_bidomain(struct BidomainList *domains, int current_match
 int partition(int *vv, int vv_len, unsigned char *adjrow) {
     int i=0;
     for (int j=0; j<vv_len; j++) {
-        // The following lines of code are intended to do the same as the following,
-        // but without branching:
-        //// if (adjrow[vv[j]]) { swap(&vv[i], &vv[j]); i++; }
 #ifdef LABELLED
         int adj = adjrow[vv[j]] != 0;
 #else
@@ -305,12 +302,10 @@ void filter_domains(struct BidomainList *d, struct BidomainList *new_d,
 #ifdef LABELLED
         int left_len = left_len_edge;
         int right_len = right_len_edge;
-        for (unsigned char i=1; i<max_label && left_len && right_len; i++) {
-            // TODO: try moving vertices to left instead of right
-            int left_len_no =
-                labelled_partition(old_bd->left_vv, left_len, g0->adjmat[v], i);
-            int right_len_no =
-                labelled_partition(old_bd->right_vv, right_len, g1->adjmat[w], i);
+        while (left_len && right_len) {
+            unsigned char label = g0->adjmat[v][old_bd->left_vv[0]];
+            int left_len_no = labelled_partition(old_bd->left_vv, left_len, g0->adjmat[v], label);
+            int right_len_no = labelled_partition(old_bd->right_vv, right_len, g1->adjmat[w], label);
             int left_len_yes = left_len - left_len_no;
             int right_len_yes = right_len - right_len_no;
             if (left_len_yes && right_len_yes) {
@@ -321,10 +316,6 @@ void filter_domains(struct BidomainList *d, struct BidomainList *new_d,
             //printf("%d %d %d\n", left_len_no, left_len_yes, left_len);
             left_len = left_len_no;
             right_len = right_len_no;
-        }
-        if (left_len && right_len) {
-            add_bidomain(new_d, old_bd->left_vv, old_bd->right_vv,
-                    left_len, right_len, true);
         }
 #else
         if (left_len_edge && right_len_edge) {

### 2017-01-25 code/james-c/graph.c
Tidied up slightly, and added command-line flag for directed (not yet implemented)

@@ -120,8 +120,7 @@ int read_word(FILE *fp) {
 }
 
 // Precondition: *g is already zeroed out
-// returns max edge label
-int readBinaryGraph(char* filename, struct Graph* g, bool labelled) {
+void readBinaryGraph(char* filename, struct Graph* g, bool labelled) {
     FILE* f;
     
     if ((f=fopen(filename, "rb"))==NULL)
@@ -149,16 +148,13 @@ int readBinaryGraph(char* filename, struct Graph* g, bool labelled) {
             g->label[i] |= label;
     }
 
-    int max_label = 0;
     for (int i=0; i<nvertices; i++) {
         int len = read_word(f);
         for (int j=0; j<len; j++) {
             int target = read_word(f);
             int label = (read_word(f) >> (16-k1)) + 1;
-            max_label = label > max_label ? label : max_label;
             add_edge(g, i, target, labelled ? label : 1);
         }
     }
     fclose(f);
-    return max_label;
 }

### 2017-01-25 code/james-c/graph.h

@@ -21,7 +21,7 @@ struct Graph {
 void readGraph(char* filename, struct Graph* g, bool labelled);
 
 // Precondition: *g is already zeroed out
-int readBinaryGraph(char* filename, struct Graph* g, bool labelled);
+void readBinaryGraph(char* filename, struct Graph* g, bool labelled);
 
 // Precondition: *g is already zeroed out
 void readLadGraph(char* filename, struct Graph* g);

### 2017-01-25 code/james-c/mcsp.c

@@ -58,6 +58,7 @@ static struct argp_option options[] = {
     {"lad", 'l', 0, 0, "Read LAD format"},
     {"timeout", 't', "TIMEOUT", 0, "Set timeout of TIMEOUT seconds"},
     {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    {"directed", 'r', 0, 0, "Solve max common CONNECTED subgraph problem"},
     {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
     { 0 }
 };
@@ -66,6 +67,7 @@ static struct {
     bool quiet;
     bool verbose;
     bool connected;
+    bool directed;
     bool big_first;
     bool dimacs;
     bool lad;
@@ -79,6 +81,7 @@ void set_default_arguments() {
     arguments.quiet = false;
     arguments.verbose = false;
     arguments.connected = false;
+    arguments.directed = false;
     arguments.big_first = false;
     arguments.dimacs = false;
     arguments.lad = false;
@@ -110,7 +113,14 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             arguments.verbose = true;
             break;
         case 'c':
+            if (arguments.directed)
+                fail("The connected and directed options cannot be used together.\n");
             arguments.connected = true;
+            break;
+        case 'r':
+            if (arguments.connected)
+                fail("The connected and directed options cannot be used together.\n");
+            arguments.directed = true;
             break;
         case 't':
             arguments.timeout = atoi(arg);
@@ -163,10 +173,6 @@ void calculate_all_degrees(struct Graph *g) {
 /*******************************************************************************
                                  MCS functions
 *******************************************************************************/
-
-#ifdef LABELLED
-static int max_label;
-#endif
 
 struct VtxPair {
     int v;
@@ -560,32 +566,23 @@ int main(int argc, char** argv) {
     struct Graph* g0 = calloc(1, sizeof(*g0));
     struct Graph* g1 = calloc(1, sizeof(*g1));
 
-    if (arguments.dimacs) {
 #ifdef LABELLED
-        readGraph(arguments.filename1, g0, true);
-        readGraph(arguments.filename2, g1, true);
-        max_label = 1;
+    bool labelled = true;
 #else
-        readGraph(arguments.filename1, g0, false);
[…]

### 2017-01-25 code/james-c/graph.c
Added labelled version

@@ -11,10 +11,11 @@ static void fail(char* msg) {
     exit(1);
 }
 
-void add_edge(struct Graph *g, int v, int w, unsigned char edge_val) {
+void add_edge(struct Graph *g, int v, int w, unsigned char edge_val, bool directed) {
     if (v != w) {
         g->adjmat[v][w] = edge_val;
-        g->adjmat[w][v] = edge_val;
+        if (!directed)
+            g->adjmat[w][v] = edge_val;
     } else if (edge_val != 1) {
         fail("Unexpected edge val != 1 on a loop\n");
     } else {
@@ -35,7 +36,7 @@ void induced_subgraph(struct Graph *g, struct Graph *subg, int *vv, int vv_len) 
 }
 
 // Precondition: *g is already zeroed out
-void readGraph(char* filename, struct Graph* g, bool labelled) {
+void readGraph(char* filename, struct Graph* g, bool directed, bool labelled) {
     FILE* f;
     
     if ((f=fopen(filename, "r"))==NULL)
@@ -64,7 +65,7 @@ void readGraph(char* filename, struct Graph* g, bool labelled) {
             case 'e':
                 if (sscanf(line, "e %d %d", &v, &w)!=2)
                     fail("Error reading a line beginning with e.\n");
-                add_edge(g, v-1, w-1, 1);
+                add_edge(g, v-1, w-1, 1, directed);
                 edges_read++;
                 break;
             case 'n':
@@ -83,7 +84,7 @@ void readGraph(char* filename, struct Graph* g, bool labelled) {
 }
 
 // Precondition: *g is already zeroed out
-void readLadGraph(char* filename, struct Graph* g) {
+void readLadGraph(char* filename, struct Graph* g, bool directed) {
     FILE* f;
     
     if ((f=fopen(filename, "r"))==NULL)
@@ -105,7 +106,7 @@ void readLadGraph(char* filename, struct Graph* g) {
         for (int j=0; j<edge_count; j++) {
             if (fscanf(f, "%d", &w) != 1)
                 fail("An edge was not read correctly.\n");
-            add_edge(g, i, w, 1);
+            add_edge(g, i, w, 1, directed);
         }
     }
 
@@ -120,7 +121,7 @@ int read_word(FILE *fp) {
 }
 
 // Precondition: *g is already zeroed out
-void readBinaryGraph(char* filename, struct Graph* g, bool labelled) {
[…]

### 2017-01-25 code/james-c/graph.h

@@ -18,12 +18,12 @@ struct Graph {
 };
 
 // Precondition: *g is already zeroed out
-void readGraph(char* filename, struct Graph* g, bool labelled);
+void readGraph(char* filename, struct Graph* g, bool directed, bool labelled);
 
 // Precondition: *g is already zeroed out
-void readBinaryGraph(char* filename, struct Graph* g, bool labelled);
+void readBinaryGraph(char* filename, struct Graph* g, bool directed, bool labelled);
 
 // Precondition: *g is already zeroed out
-void readLadGraph(char* filename, struct Graph* g);
+void readLadGraph(char* filename, struct Graph* g, bool directed);
 
 void induced_subgraph(struct Graph *g, struct Graph *subg, int *vv, int vv_len);

### 2017-01-25 code/james-c/mcsp.c

@@ -289,7 +289,7 @@ void add_bidomain(struct BidomainList *bd_list, int *left_vv, int *right_vv,
     };
 }
 
-void filter_domains(struct BidomainList *d, struct BidomainList *new_d,
+void filter_domains_undirected(struct BidomainList *d, struct BidomainList *new_d,
         struct Graph *g0, struct Graph *g1, int v, int w)
 {
     new_d->len=0;
@@ -330,6 +330,54 @@ void filter_domains(struct BidomainList *d, struct BidomainList *new_d,
         }
 #endif
     }
+}
+
+int directed_labelled_partition(int *vv, int vv_len, int v, unsigned char (*adjmat)[MAX_N], unsigned int label) {
+    int i=0;
+    for (int j=0; j<vv_len; j++) {
+        int nonadj = (adjmat[v][vv[j]]<<CHAR_BIT | adjmat[vv[j]][v]) != label;
+        if (nonadj) {
+            swap(&vv[i], &vv[j]);
+            i++;
+        }
+    }
+    return i;
+}
+
+void filter_domains_directed(struct BidomainList *d, struct BidomainList *new_d,
+        struct Graph *g0, struct Graph *g1, int v, int w)
+{
+    new_d->len=0;
+    for (int j=0; j<d->len; j++) {
+        struct Bidomain *old_bd = &d->vals[j];
+        int left_len = old_bd->left_len;
+        int right_len = old_bd->right_len;
+        while (left_len && right_len) {
+            unsigned int label = (g0->adjmat[v][old_bd->left_vv[0]]<<CHAR_BIT) |
+                                  g0->adjmat[old_bd->left_vv[0]][v];
+            int left_len_no = directed_labelled_partition(old_bd->left_vv, left_len, v, g0->adjmat, label);
+            int right_len_no = directed_labelled_partition(old_bd->right_vv, right_len, w, g1->adjmat, label);
+            int left_len_yes = left_len - left_len_no;
+            int right_len_yes = right_len - right_len_no;
+            if (left_len_yes && right_len_yes) {
+                int *left_vv = old_bd->left_vv + left_len_no;
+                int *right_vv = old_bd->right_vv + right_len_no;
+                add_bidomain(new_d, left_vv, right_vv, left_len_yes, right_len_yes, true);
+            }
+            //printf("%d %d %d\n", left_len_no, left_len_yes, left_len);
+            left_len = left_len_no;
+            right_len = right_len_no;
+        }
+    }
+}
+
[…]

### 2017-01-26 code/james-c/mcsp.c
Show edge count in solution

@@ -578,6 +578,28 @@ struct VtxPairList mcs(struct Graph *g0, struct Graph *g1) {
     return incumbent;
 }
 
+// Count edges in the subgraph
+int count_edges(struct Graph *g0, struct Graph *g1, struct VtxPairList *solution) {
+    int count = 0;
+    for (int i=0; i<solution->len; i++) {
+        struct VtxPair p0 = solution->vals[i];
+        for (int j=i+1; j<solution->len; j++) {
+            struct VtxPair p1 = solution->vals[j];
+            if (g0->adjmat[p0.v][p1.v]) count++;
+            if (arguments.directed && g0->adjmat[p1.v][p0.v]) count++;
+        }
+    }
+//    for (int i=0; i<solution->len; i++) {
+//        struct VtxPair p0 = solution->vals[i];
+//        for (int j=0; j<solution->len; j++) {
+//            struct VtxPair p1 = solution->vals[j];
+//            printf("%s ", g0->adjmat[p0.v][p1.v]!=0 ? "1" : ".");
+//        }
+//        printf("\n");
+//    }
+    return count;
+}
+
 bool check_sol(struct Graph *g0, struct Graph *g1, struct VtxPairList *solution) {
     bool used_left[MAX_N];
     bool used_right[MAX_N];
@@ -678,6 +700,7 @@ int main(int argc, char** argv) {
         printf("TIMEOUT\n");
 
     printf("Solution size %d\n", solution.len);
+    printf("Number of edges in subgraph %d\n", count_edges(g0, g1, &solution));
     INSERTION_SORT(struct VtxPair, solution.vals, solution.len,
             (solution.vals[j-1].v > solution.vals[j].v))
     for (int i=0; i<solution.len; i++)

### 2017-01-26 code/james-c/mcsp.c
Better validity checking of ssolution for directed graphs

@@ -613,8 +613,8 @@ bool check_sol(struct Graph *g0, struct Graph *g1, struct VtxPairList *solution)
         used_right[p0.w] = true;
         for (int j=i+1; j<solution->len; j++) {
             struct VtxPair p1 = solution->vals[j];
-            if (g0->adjmat[p0.v][p1.v] != g1->adjmat[p0.w][p1.w])
-                return false;
+            if (g0->adjmat[p0.v][p1.v] != g1->adjmat[p0.w][p1.w]) return false;
+            if (g0->adjmat[p1.v][p0.v] != g1->adjmat[p1.w][p0.w]) return false;
         }
     }
     return true;

### 2017-01-26 code/james-c/mcsp.c
When calculating degree for directed graphs, calculate total degree rather than out-degree

@@ -165,8 +165,11 @@ void initialise_stats() {
 void calculate_all_degrees(struct Graph *g) {
     for (int v=0; v<g->n; v++) {
         g->degree[v] = 0;
-        for (int w=0; w<g->n; w++)
+        for (int w=0; w<g->n; w++) {
             g->degree[v] += (g->adjmat[v][w] != 0);
+            if (arguments.directed)
+                g->degree[v] += (g->adjmat[w][v] != 0);
+        }
     }
 }
 

### 2017-01-26 code/graph-utils/remove_labels.c
Added a program for removing labels

@@ -1,0 +1,132 @@
+#define _GNU_SOURCE
+#define _POSIX_SOURCE
+
+#include <argp.h>
+#include <limits.h>
+#include <locale.h>
+#include <stdbool.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <time.h>
+
+static void fail(char* msg) {
+    printf("%s\n", msg);
+    exit(1);
+}
+
+/*******************************************************************************
+                             Command-line arguments
+*******************************************************************************/
+
+static char doc[] = "Remove vertex or edge labels from a VF-format graph";
+static char args_doc[] = "INPUT-FILENAME OUTPUT-FILENAME";
+static struct argp_option options[] = {
+    {"no-edge-labels", 'e', 0, 0, "Remove edge labels"},
+    {"no-vtx-labels", 'v', 0, 0, "Remove vertex labels"},
+    { 0 }
+};
+
+static struct {
+    bool use_edge_labels;
+    bool use_vtx_labels;
+    char *filename1;
+    char *filename2;
+    int arg_num;
+} arguments;
+
+void set_default_arguments() {
+    arguments.use_edge_labels = true;
+    arguments.use_vtx_labels = true;
+    arguments.filename1 = NULL;
+    arguments.filename2 = NULL;
+    arguments.arg_num = 0;
+}
+
+static error_t parse_opt (int key, char *arg, struct argp_state *state) {
+    switch (key) {
+        case 'e':
+            arguments.use_edge_labels = false;
+            break;
+        case 'v':
+            arguments.use_vtx_labels = false;
+            break;
+        case ARGP_KEY_ARG:
+            if (arguments.arg_num == 0) {
+                arguments.filename1 = arg;
+            } else if (arguments.arg_num == 1) {
+                arguments.filename2 = arg;
+            } else {
+                argp_usage(state);
+            }
+            arguments.arg_num++;
+            break;
+        case ARGP_KEY_END:
+            if (arguments.arg_num == 0)
+                argp_usage(state);
+            break;
+        default: return ARGP_ERR_UNKNOWN;
+    }
+    return 0;
+}
+
+static struct argp argp = { options, parse_opt, args_doc, doc };
+
+/*******************************************************************************/
+
+int read_word(FILE *fp) {
+    unsigned char a[2];
[…]

### 2017-01-26 code/bigraphs/convert_bigraph.py
Added script to convert bigraph files

@@ -1,0 +1,80 @@
+import sys
+
+class Graph(object):
+    def __init__(self, adjmat, labels):
+        self.adjmat = adjmat
+        self.labels = labels
+
+    def n(self):
+        return len(self.adjmat)
+
+    def __repr__(self):
+        return ("\n".join("".join("1" if x else "0" for x in row) for row in self.adjmat)               + "\n" + "\n".join(self.labels))
+
+def parse_vertex(vertex):
+    """Takes a string like '0, Area(2):1' and returns a pair like
+       (0, 'Area(2):1')
+    """
+    tokens = vertex.split(",")
+    return int(tokens[0]), tokens[1].strip()
+    
+
+def read_graph(filename):
+    with open(filename, "r") as f:
+        lines = [line for line in f.readlines()]
+
+    l0 = lines[0]
+
+    vertices = []
+
+    vertex_str = ""
+    paren_depth = 0
+    for char in l0:
+        if char == ")":
+            paren_depth -= 1
+            if paren_depth == 0:
+                vertices.append(parse_vertex(vertex_str))
+                vertex_str = ""
+        if paren_depth >= 1:
+            vertex_str += char
+        if char == "(":
+            paren_depth += 1
+
+    vertex_idx_to_label = dict(vertices)
+
+    nroots, nvertices, nsites = [int(x) for x in lines[1].strip().split()]
+
+    adjmatstr = [line.strip() for line in lines[nroots+2:nroots+nvertices+2]]
+    adjmat = [[ch=="1" for ch in line[:nvertices]] for line in adjmatstr]
+
+    return Graph(adjmat, [vertex_idx_to_label[i] for i in range(nvertices)])
+
+def write_word(f, word):
+    f.write(bytearray([word&0xFF, (word&0xFF00)>>8]))
+
+def write_graph(g, label_name_to_number, filename):
+    with open(filename, "wb") as f:
+        write_word(f, g.n())
+        for label in g.labels:
+            write_word(f, label_name_to_number[label])
+        for row in g.adjmat:
+            write_word(f, sum(row))
+            for i in range(g.n()):
+                if row[i]:
+                    write_word(f, i)
+                    write_word(f, 1)   # edge label
+
+if __name__=="__main__":
+    if len(sys.argv) != 5:
+        print "Usage:", sys.argv[0], "IN-FILE-1 IN-FILE-2 OUT-FILE-1 OUT-FILE-2"
[…]

### 2017-01-26 code/bigraphs/convert_bigraph.py
Fixed a line break

@@ -9,7 +9,8 @@ class Graph(object):
         return len(self.adjmat)
 
     def __repr__(self):
-        return ("\n".join("".join("1" if x else "0" for x in row) for row in self.adjmat)               + "\n" + "\n".join(self.labels))
+        return ("\n".join("".join("1" if x else "0" for x in row) for row in self.adjmat)
+                + "\n" + "\n".join(self.labels))
 
 def parse_vertex(vertex):
     """Takes a string like '0, Area(2):1' and returns a pair like

### 2017-01-26 code/bigraphs/convert_bigraph.py
Exit if not enough arguments are given

@@ -68,6 +68,7 @@ def write_graph(g, label_name_to_number, filename):
 if __name__=="__main__":
     if len(sys.argv) != 5:
         print "Usage:", sys.argv[0], "IN-FILE-1 IN-FILE-2 OUT-FILE-1 OUT-FILE-2"
+        sys.exit(1)
     g0 = read_graph(sys.argv[1])
     g1 = read_graph(sys.argv[2])
     print g0

### 2017-01-27 code/james-c/graph.h
Removed some unused definitions

@@ -3,11 +3,6 @@
 #include <limits.h>
 #include <stdbool.h>
 
-typedef unsigned long long ULL;
-
-#define BYTES_PER_WORD sizeof(ULL)
-#define BITS_PER_WORD (CHAR_BIT * BYTES_PER_WORD)
-
 #define BITS_PER_UNSIGNED_INT (CHAR_BIT * sizeof(unsigned int))
 
 struct Graph {

### 2017-01-27 code/james-c/graph.c
For directed graphs, combine inward and outward edge labels in the same value

@@ -11,11 +11,34 @@ static void fail(char* msg) {
     exit(1);
 }
 
-void add_edge(struct Graph *g, int v, int w, unsigned char edge_val, bool directed) {
+// Unlabelled, undirected: if there is an edge v-w, then adjmat[v][w]==adjmat[w][v]==1
+// Labelled, undirected:   if there is an edge v-w, then adjmat[v][w]==adjmat[w][v]==label
+// Unlabelled, directed:   if there is an edge v->w, then adjmat[v][w]&1 == (adjmat[w][v]>>1)&1 == 1
+// Labelled, directed:     if there is an edge v->w, then the least significant two bytes of adjmat[v][w]
+//                          and the most significant two bytes of adjmat[w][v] equal the label
+void add_edge(struct Graph *g, int v, int w, edge_label_t edge_val, bool directed) {
     if (v != w) {
-        g->adjmat[v][w] = edge_val;
-        if (!directed)
+#ifdef LABELLED
+        if (edge_val > 0xFFFFu)
+            fail("An edge label is too large.");
+        if (directed) {
+            if (g->adjmat[v][w] & 0xFFFFu)
+                fail("Duplicate edge.");
+            g->adjmat[v][w] |= edge_val;
+            g->adjmat[w][v] |= (edge_val << 16);
+        } else {
+            g->adjmat[v][w] = edge_val;
             g->adjmat[w][v] = edge_val;
+        }
+#else
+        if (directed) {
+            g->adjmat[v][w] |= 1;
+            g->adjmat[w][v] |= 2;
+        } else {
+            g->adjmat[v][w] = 1;
+            g->adjmat[w][v] = 1;
+        }
+#endif
     } else if (edge_val != 1) {
         fail("Unexpected edge val != 1 on a loop\n");
     } else {

### 2017-01-27 code/james-c/graph.h

@@ -5,10 +5,16 @@
 
 #define BITS_PER_UNSIGNED_INT (CHAR_BIT * sizeof(unsigned int))
 
+#ifdef LABELLED
+typedef unsigned int edge_label_t;
+#else
+typedef unsigned char edge_label_t;
+#endif
+
 struct Graph {
     int n;
     int degree[MAX_N];
-    unsigned char adjmat[MAX_N][MAX_N];
+    edge_label_t adjmat[MAX_N][MAX_N];
     unsigned int label[MAX_N];
 };
 

### 2017-01-27 code/james-c/mcsp.c

@@ -17,8 +17,6 @@ typedef unsigned long long ULL;
 #define MIN(X, Y) (((X) < (Y)) ? (X) : (Y))
 #define MAX(X, Y) (((X) > (Y)) ? (X) : (Y))
 
-#define DEG_SEQ_CHECK_SIZE_LIMIT 15
-
 #define INSERTION_SORT(type, arr, arr_len, swap_condition) do { \
     for (int i=1; i<arr_len; i++) {                             \
         for (int j=i; j>=1; j--) {                              \
@@ -166,10 +164,15 @@ void calculate_all_degrees(struct Graph *g) {
     for (int v=0; v<g->n; v++) {
         g->degree[v] = 0;
         for (int w=0; w<g->n; w++) {
-            g->degree[v] += (g->adjmat[v][w] != 0);
-            if (arguments.directed)
-                g->degree[v] += (g->adjmat[w][v] != 0);
+#ifdef LABELLED
+            if  (g->adjmat[v][w]&0xFFFFu) g->degree[v]++;
+            if  (g->adjmat[v][w]&0xFFFF0000u) g->degree[v]++; // inward edges, for directed case
+#else
+            if  (g->adjmat[v][w]&1) g->degree[v]++;
+            if  (g->adjmat[v][w]&2) g->degree[v]++;   // inward edges, for directed case
+#endif
         }
+//        printf("%d\n", g->degree[v]);
     }
 }
 
@@ -250,15 +253,10 @@ struct Bidomain *select_bidomain(struct BidomainList *domains, int current_match
 }
 
 // Returns length of left half of array
-int partition(int *vv, int vv_len, unsigned char *adjrow) {
+int partition(int *vv, int vv_len, edge_label_t *adjrow) {
     int i=0;
     for (int j=0; j<vv_len; j++) {
-#ifdef LABELLED
-        int adj = adjrow[vv[j]] != 0;
-#else
-        int adj = adjrow[vv[j]];
-#endif
-        if (adj) {
+        if (adjrow[vv[j]]) {
             swap(&vv[i], &vv[j]);
             i++;
         }
@@ -268,7 +266,7 @@ int partition(int *vv, int vv_len, unsigned char *adjrow) {
 
 // Swaps everything with an edge of the correct label to the end of vv
 // Returns length of left half of array
-int labelled_partition(int *vv, int vv_len, unsigned char *adjrow, unsigned char label) {
[…]

### 2017-01-27 code/james-cpp-cilk/graph.h
Start on cilk version

@@ -1,0 +1,15 @@
+#include <limits.h>
+#include <stdbool.h>
+
+#include <vector>
+
+struct Graph {
+    int n;
+    std::vector<std::vector<unsigned char>> adjmat;
+    std::vector<std::vector<unsigned char>> antiadjmat;
+    std::vector<unsigned int> label;
+    Graph(unsigned int n);
+};
+
+Graph readGraph(char* filename, char format);
+

### 2017-01-27 code/james-cpp-cilk/mcsp.c

@@ -1,0 +1,390 @@
+#include "graph.h"
+
+#include <argp.h>
+#include <limits.h>
+#include <locale.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <time.h>
+
+#include <iostream>
+#include <string>
+#include <utility>
+#include <vector>
+#include <chrono>
+
+#include <cilk/cilk.h>
+
+using std::vector;
+using std::cout;
+using std::endl;
+
+static void fail(std::string msg) {
+    std::cerr << msg << std::endl;
+    exit(1);
+}
+
+/*******************************************************************************
+                             Command-line arguments
+*******************************************************************************/
+
+static char doc[] = "Find a maximum clique in a graph in DIMACS format";
+static char args_doc[] = "FILENAME1 FILENAME2";
+static struct argp_option options[] = {
+    {"quiet", 'q', 0, 0, "Quiet output"},
+    {"verbose", 'v', 0, 0, "Verbose output"},
+    {"dimacs", 'd', 0, 0, "Read DIMACS format"},
+    {"lad", 'l', 0, 0, "Read LAD format"},
+    {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    { 0 }
+};
+
+static struct {
+    bool quiet;
+    bool verbose;
+    bool connected;
+    bool dimacs;
+    bool lad;
+    char *filename1;
+    char *filename2;
+    int arg_num;
+} arguments;
+
+void set_default_arguments() {
+    arguments.quiet = false;
+    arguments.verbose = false;
+    arguments.connected = false;
+    arguments.dimacs = false;
+    arguments.lad = false;
+    arguments.filename1 = NULL;
+    arguments.filename2 = NULL;
+    arguments.arg_num = 0;
+}
+
+static error_t parse_opt (int key, char *arg, struct argp_state *state) {
+    switch (key) {
+        case 'd':
+            if (arguments.lad)
+                fail("The -d and -l options cannot be used together.\n");
+            arguments.dimacs = true;
+            break;
+        case 'l':
+            if (arguments.dimacs)
+                fail("The -d and -l options cannot be used together.\n");
+            arguments.lad = true;
+            break;
+        case 'q':
+            arguments.quiet = true;
+            break;
+        case 'v':
+            arguments.verbose = true;
+            break;
+        case 'c':
+            arguments.connected = true;
+            break;
+        case ARGP_KEY_ARG:
[…]

### 2017-01-27 code/james-c/mcsp.c
Tidied up partitioning function

@@ -297,36 +297,37 @@ void filter_domains(struct BidomainList *d, struct BidomainList *new_d,
     new_d->len=0;
     for (int j=0; j<d->len; j++) {
         struct Bidomain *old_bd = &d->vals[j];
-        int left_len_edge = partition(old_bd->left_vv, old_bd->left_len, g0->adjmat[v]);
-        int right_len_edge = partition(old_bd->right_vv, old_bd->right_len, g1->adjmat[w]);
-        int left_len_noedge = old_bd->left_len - left_len_edge;
-        int right_len_noedge = old_bd->right_len - right_len_edge;
+        // After these two partitions, left_len and right_len are the lengths of the
+        // arrays of vertices with edges from v or w (int the directed case, edges
+        // either from or to v or w)
+        int left_len = partition(old_bd->left_vv, old_bd->left_len, g0->adjmat[v]);
+        int right_len = partition(old_bd->right_vv, old_bd->right_len, g1->adjmat[w]);
+        int left_len_noedge = old_bd->left_len - left_len;
+        int right_len_noedge = old_bd->right_len - right_len;
         if (left_len_noedge && right_len_noedge) {
-            int *left_vv = old_bd->left_vv + left_len_edge;
-            int *right_vv = old_bd->right_vv + right_len_edge;
-            add_bidomain(new_d, left_vv, right_vv, left_len_noedge, right_len_noedge,
-                    old_bd->is_adjacent);
+            add_bidomain(new_d, old_bd->left_vv+left_len, old_bd->right_vv+right_len,
+                    left_len_noedge, right_len_noedge, old_bd->is_adjacent);
         }
         if (multiway) {
-            int left_len = left_len_edge;
-            int right_len = right_len_edge;
             while (left_len && right_len) {
                 edge_label_t label = g0->adjmat[v][old_bd->left_vv[0]];
-                int left_len_no = labelled_partition(old_bd->left_vv, left_len, g0->adjmat[v], label);
-                int right_len_no = labelled_partition(old_bd->right_vv, right_len, g1->adjmat[w], label);
-                int left_len_yes = left_len - left_len_no;
-                int right_len_yes = right_len - right_len_no;
+                // These partitions move vertices whose edges have label 'label' to the end of
+                // the old_bd->left_vv and old_bd->right_vv arrays. left_len becomes the length
+                // of the remaining array; that is, the left part of old_bd->left_vv that
+                // contains vertices with other edge labels
[…]

### 2017-01-27 code/james-cpp-cilk/mcsp.c
Less awful

@@ -13,6 +13,7 @@
 #include <utility>
 #include <vector>
 #include <chrono>
+#include <atomic>
 
 #include <cilk/cilk.h>
 
@@ -108,7 +109,7 @@ static struct argp argp = { options, parse_opt, args_doc, doc };
                                      Stats
 *******************************************************************************/
 
-static long nodes = 0;
+std::atomic<unsigned long long> nodes{ 0 };
 
 /*******************************************************************************
                                  MCS functions
@@ -150,30 +151,6 @@ bool check_sol(struct Graph *g0, struct Graph *g1, vector<VtxPair>& solution) {
         }
     }
     return true;
-}
-
-void show(const vector<VtxPair>& current, const vector<Bidomain> &domains,
-        const vector<int>& left, const vector<int>& right)
-{
-    printf("Nodes: %ld\n", nodes);
-    printf("Length of current assignment: %ld\n", current.size());
-    printf("Current assignment:");
-    for (unsigned int i=0; i<current.size(); i++) {
-        printf("  %d->%d", current[i].v, current[i].w);
-    }
-    printf("\n");
-    for (unsigned int i=0; i<domains.size(); i++) {
-        struct Bidomain bd = domains[i];
-        printf("Left  ");
-        for (int j=0; j<bd.left_len; j++)
-            printf("%d ", left[bd.l + j]);
-        printf("\n");
-        printf("Right  ");
-        for (int j=0; j<bd.right_len; j++)
-            printf("%d ", right[bd.r + j]);
-        printf("\n");
-    }
-    printf("\n\n");
 }
 
 int calc_bound(const vector<Bidomain>& domains) {
@@ -285,7 +262,6 @@ void solve(struct Graph *g0, struct Graph *g1, vector<VtxPair>& incumbent,
         const vector<VtxPair>& current, const vector<Bidomain> &domains,
         const vector<int>& left, const vector<int>& right, int level)
 {
-    if (arguments.verbose) show(current, domains, left, right);
     nodes++;
 
     if (current.size() > incumbent.size()) {
@@ -385,6 +361,6 @@ int main(int argc, char** argv) {
     printf("\n");
 
     setlocale(LC_NUMERIC, "");
-    printf("Nodes:                      %'15ld\n", nodes);
-    printf("CPU time (ms):              %15ld\n", time_elapsed);
+    cout << "Nodes:                      " << nodes << endl;
+    cout << "CPU time (ms):              " << time_elapsed << endl;
 }

### 2017-01-27 code/james-c/mcsp.c
Made the node count an unsigned long long

@@ -148,13 +148,7 @@ static struct argp argp = { options, parse_opt, args_doc, doc };
                                      Stats
 *******************************************************************************/
 
-static struct {
-    long nodes;
-} stats;
-
-void initialise_stats() {
-    stats.nodes = 0;
-}
+static unsigned long long nodes = 0;
 
 /*******************************************************************************
                                 Graph functions
@@ -339,7 +333,7 @@ void remove_bidomain(struct BidomainList *list, struct Bidomain *b) {
 }
 
 void show(struct D d) {
-    printf("Nodes: %ld\n", stats.nodes);
+    printf("Nodes: %llu\n", nodes);
     printf("Length of current assignment: %d\n", d.current->len);
     printf("Current assignment:");
     for (int i=0; i<d.current->len; i++) {
@@ -400,8 +394,8 @@ void solve(struct D d, int level) {
 //    printf(" mcsp --- Nodes: %ld\n", stats.nodes);
     if (arguments.verbose) show(d);
 
-    stats.nodes++;
-    if (arguments.timeout && stats.nodes%100000==0 && 
+    nodes++;
+    if (arguments.timeout && nodes%100000==0 && 
             (clock()-start)*1000/CLOCKS_PER_SEC > arguments.timeout*1000) {
         arguments.timeout = -1;
     }
@@ -587,8 +581,6 @@ int main(int argc, char** argv) {
     set_default_arguments();
     argp_parse(&argp, argc, argv, 0, 0, 0);
 
-    initialise_stats();
-
     struct Graph* g0 = calloc(1, sizeof(*g0));
     struct Graph* g1 = calloc(1, sizeof(*g1));
 
@@ -664,7 +656,7 @@ int main(int argc, char** argv) {
     printf("\n");
 
     setlocale(LC_NUMERIC, "");
-    printf("Nodes:                      %'15ld\n", stats.nodes);
+    printf("Nodes:                      %'15llu\n", nodes);
     printf("CPU time (ms):              %15ld\n", time_elapsed * 1000 / CLOCKS_PER_SEC);
 
     free(preallocated_lists);

### 2017-01-27 code/james-c/mcsp.c
Simplified variable selection code

@@ -226,21 +226,29 @@ int find_min(int *vv, int len) {
 struct Bidomain *select_bidomain(struct BidomainList *domains, int current_matching_size) {
     // Select the bidomain with the smallest max(leftsize, rightsize), breaking
     // ties on the smallest vertex index in the left set
+    struct Bidomain *bds[MAX_N];  // a list of bidomains of minimum size
+    int bds_len = 0;
     int min_size = 999999999;
-    int min_idx = 999999999;
-    struct Bidomain *best = NULL;
     for (int i=0; i<domains->len; i++) {
         struct Bidomain *bd = &domains->vals[i];
         if (arguments.connected && current_matching_size>0 && !bd->is_adjacent) continue;
-        int len = MAX(bd->left_len, bd->right_len);
-        int idx;
-        if (len < min_size) {
-            min_size = len;
-            min_idx = find_min(bd->left_vv, bd->left_len);
-            best = &domains->vals[i];
-        } else if (len==min_size && (idx=find_min(bd->left_vv, bd->left_len))<min_idx) {
-            min_idx = idx;
-            best = &domains->vals[i];
+        int size = MAX(bd->left_len, bd->right_len);
+        if (size == min_size) {
+            bds[bds_len++] = bd;
+        } else if (size < min_size) {
+            min_size = size;
+            bds[0] = bd;
+            bds_len = 1;
+        }
+    }
+    int min_vtx_idx = 999999999;
+    struct Bidomain *best = NULL;
+    for (int i=0; i<bds_len; i++) {
+        struct Bidomain *bd = bds[i];
+        int vtx_idx = find_min(bd->left_vv, bd->left_len);
+        if (vtx_idx < min_vtx_idx) {
+            min_vtx_idx = vtx_idx;
+            best = bd;
         }
     }
     return best;

### 2017-01-27 code/james-c/mcsp.c
Tidied use of preallocated lists

@@ -398,6 +398,12 @@ int index_of_next_smallest(int *arr, int len, int w) {
     return idx;
 }
 
+struct BidomainList *get_preallocated_list(int level) {
+    struct BidomainList *list = &preallocated_lists[level];
+    list->len = 0;
+    return list;
+}
+
 void solve(struct D d, int level) {
 //    printf(" mcsp --- Nodes: %ld\n", stats.nodes);
     if (arguments.verbose) show(d);
@@ -434,8 +440,7 @@ void solve(struct D d, int level) {
     }
 
     struct D new_d = d;
-    new_d.domains = &preallocated_lists[level];
-    new_d.domains->len = 0;
+    new_d.domains = get_preallocated_list(level);
 
     // Try assigning v to each vertex w in bd->right_vv, in turn
     bd->right_len--;
@@ -473,8 +478,7 @@ void sort_and_remove_dups(unsigned int *arr, int *len) {
 void build_domains_and_solve(struct Graph *g0, struct Graph *g1, struct VtxPairList *incumbent,
         int matching_size_goal, unsigned int *all_labels, int all_labels_len)
 {
-    struct BidomainList *domains = &preallocated_lists[0];
-    domains->len = 0;
+    struct BidomainList *domains = get_preallocated_list(0);
 
     int left[MAX_N];  // the buffer of vertex indices for the left partitions
     int right[MAX_N];  // the buffer of vertex indices for the right partitions

### 2017-01-27 code/james-c/mcsp.c
O(n log n) partitioning for labelled and directed cases

@@ -266,20 +266,6 @@ int partition(int *vv, int vv_len, edge_label_t *adjrow) {
     return i;
 }
 
-// Swaps everything with an edge of the correct label to the end of vv
-// Returns length of left half of array
-int labelled_partition(int *vv, int vv_len, edge_label_t *adjrow, edge_label_t label) {
-    int i=0;
-    for (int j=0; j<vv_len; j++) {
-        int nonadj = adjrow[vv[j]] != label;
-        if (nonadj) {
-            swap(&vv[i], &vv[j]);
-            i++;
-        }
-    }
-    return i;
-}
-
 void add_bidomain(struct BidomainList *bd_list, int *left_vv, int *right_vv,
         int left_len, int right_len, bool is_adjacent)
 {
@@ -292,6 +278,14 @@ void add_bidomain(struct BidomainList *bd_list, int *left_vv, int *right_vv,
     };
 }
 
+int compare_by_edge_labels(const void *a, const void *b, void *ar)
+{
+    int v = *(int*) a;
+    int w = *(int*) b;
+    edge_label_t *adjrow = (edge_label_t *) ar;
+    return (adjrow[v]>adjrow[w]) - (adjrow[v]<adjrow[w]);
+}
+
 // multiway must be true iff we are solving a labelled or directed problem
 void filter_domains(struct BidomainList *d, struct BidomainList *new_d,
         struct Graph *g0, struct Graph *g1, int v, int w, bool multiway)
@@ -299,37 +293,43 @@ void filter_domains(struct BidomainList *d, struct BidomainList *new_d,
     new_d->len=0;
     for (int j=0; j<d->len; j++) {
         struct Bidomain *old_bd = &d->vals[j];
+        int *l = old_bd->left_vv;
+        int *r = old_bd->right_vv;
         // After these two partitions, left_len and right_len are the lengths of the
         // arrays of vertices with edges from v or w (int the directed case, edges
         // either from or to v or w)
-        int left_len = partition(old_bd->left_vv, old_bd->left_len, g0->adjmat[v]);
-        int right_len = partition(old_bd->right_vv, old_bd->right_len, g1->adjmat[w]);
+        int left_len = partition(l, old_bd->left_len, g0->adjmat[v]);
+        int right_len = partition(r, old_bd->right_len, g1->adjmat[w]);
         int left_len_noedge = old_bd->left_len - left_len;
         int right_len_noedge = old_bd->right_len - right_len;
[…]

### 2017-01-31 code/tests/generate.cpp
Added some hacky test stuff

@@ -1,0 +1,88 @@
+#include <fstream>
+#include <iostream>
+#include <numeric>
+#include <random>
+#include <sstream>
+
+int read_int(std::string s) {
+    std::istringstream ss(s);
+    int retval;
+    if (!(ss >> retval))
+        return 0;
+    return retval;
+}
+
+void write_word(int word, std::ofstream &fout) {
+    unsigned char bytes[2];
+    bytes[0] = word & 0xFF;
+    bytes[1] = (word>>8) & 0xFF;
+    fout.write((char *) bytes, 2);
+}
+
+void generate(int n,
+              int p,
+              int min_vtx_label,
+              int max_vtx_label,
+              int min_edge_label,
+              int max_edge_label,
+              int loops_policy,
+              std::string filename)
+{
+    std::ofstream fout;
+    fout.open(filename, std::ios::binary | std::ios::out);
+
+    write_word(n, fout);
+
+    std::random_device seeder;
+    std::mt19937 rng(seeder());
+    std::uniform_int_distribution<int> vtx_label_gen(min_vtx_label, max_vtx_label);
+    std::uniform_int_distribution<int> edge_label_gen(min_edge_label, max_edge_label);
+    std::uniform_int_distribution<int> edge_exists_dist(1, 100);
+
+    for (int i=0; i<n; i++)
+        write_word(vtx_label_gen(rng), fout);
+
+    for (int i=0; i<n; i++) {
+        std::vector<std::pair<int, int>> edges;    // (target, label) pairs
+        for (int j=0; j<n; j++) {
+            if (i != j) {
+                if (edge_exists_dist(rng) <= p) {
+                    edges.push_back({j, edge_label_gen(rng)});
+                }
+            }
+        }
+        if (loops_policy==2 || (loops_policy==1 && i%2)) {
+            edges.push_back({i, 1});
+        }
+        write_word(edges.size(), fout);
+        for (auto edge : edges) {
+            write_word(edge.first, fout);   // target
+            write_word(edge.second, fout);  // label
+        }
+    }
+
+    fout.close();
+}
+
+int main(int argc, char **argv) {
+    if (argc != 9) {
+        std::cerr << "Usage: " << argv[0];
+        std::cerr << " n p min_vtx_label max_vtx_label min_edge_label max_edge_label loops_policy filename" << std::endl;
+        std::cerr << "Where p is an integer from 0 to 100" << std::endl;
[…]

### 2017-02-01 paper/paper.tex
Plotting!
References
Better plottery

@@
 \usepackage{times}
+\usepackage{graphicx}
+
+\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
+\newcommand{\citep}[1]{\cite{#1}}
 
@@
 
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
+\citet{UpcomingAAAIPaper}
+
+\citet{DBLP:conf/cp/McCreeshNPS16}
+
+\citet{DBLP:conf/cp/NdiayeS11}
+
+\section{Computational Experiments}
+
+Dual Xeon E5-2640 v2 CPUs, 64GBytes RAM. Parallel experiments 16 threads (one per hardware core), and 32 with hyper-threading.
+
+\paragraph{No labels, undirected}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-plain-cumulative.pdf}
+    \caption{MCS Plain instances, cumulative runtimes}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-33ved-cumulative.pdf}
+    \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}
+\end{figure}
+
+\paragraph{Vertex and edge labels, directed}
+
+\paragraph{No labels, undirected, connected}
+
+\paragraph{Vertex and edge labels, connected}
+
+\paragraph{Large subgraph isomorphism instances}
+
+\paragraph{Compared to CP FC}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-plain-james-versus-cp-fc-scatter.pdf}
+    \caption{MCS Plain instances, James vs CP-FC, Runtimes}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-scatter.pdf}
+    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Runtimes}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
+    \caption{MCS Plain instances, James vs CP-FC, Nodes}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
+    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Nodes}
+\end{figure}
+
 \bibliographystyle{named}

### 2017-02-01 paper/paper.tex
Plot tweaks

@@
 \usepackage{graphicx}
+\usepackage{cleveref}
 
@@
 
-\paragraph{No labels, undirected}
+\paragraph{No labels, undirected} are in \cref{figure:plain-cumulative}.
+
+\paragraph{Vertex and edge labels, directed} are in \cref{figure:33ved-cumulative}.
+
+\paragraph{No labels, undirected, connected} have yet to be run.
+
+\paragraph{Vertex and edge labels, connected} have yet to be run.
+
+\paragraph{Large subgraph isomorphism instances} have yet to be run.
+
+\paragraph{Compared to CP FC} Runtimes are in \cref{figure:plain-james-versus-cp-fc-scatter} and \cref{figure:33ved-james-versus-cp-fc-scatter}, and nodes are in \cref{figure:plain-james-versus-cp-fc-nodes-scatter} and \cref{figure:33ved-james-versus-cp-fc-nodes-scatter}.
 
@@
     \includegraphics*{gen-graph-plain-cumulative.pdf}
-    \caption{MCS Plain instances, cumulative runtimes}
+    \caption{MCS Plain instances, cumulative runtimes}\label{figure:plain-cumulative}
 \end{figure}
@@
     \includegraphics*{gen-graph-33ved-cumulative.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}
+    \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}\label{figure:33ved-cumulative}
 \end{figure}
-
-\paragraph{Vertex and edge labels, directed}
-
-\paragraph{No labels, undirected, connected}
-
-\paragraph{Vertex and edge labels, connected}
-
-\paragraph{Large subgraph isomorphism instances}
-
-\paragraph{Compared to CP FC}
 
@@
     \includegraphics*{gen-graph-plain-james-versus-cp-fc-scatter.pdf}
-    \caption{MCS Plain instances, James vs CP-FC, Runtimes}
+    \caption{MCS Plain instances, James vs CP-FC, Runtimes (TODO?? heatmapify this)}\label{figure:plain-james-versus-cp-fc-scatter}
 \end{figure}
@@
     \includegraphics*{gen-graph-33ved-james-versus-cp-fc-scatter.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Runtimes}
+    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Runtimes (TODO?? heatmapify this)}\label{figure:33ved-james-versus-cp-fc-scatter}
 \end{figure}
@@
     \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{MCS Plain instances, James vs CP-FC, Nodes}
+    \caption{MCS Plain instances, James vs CP-FC, Nodes (TODO?? heatmapify this)}\label{figure:plain-james-versus-cp-fc-nodes-scatter}
 \end{figure}
@@
     \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Nodes}
+    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Nodes (TODO?? heatmapify this)}\label{figure:33ved-james-versus-cp-fc-nodes-scatter}
 \end{figure}

### 2017-02-02 code/james-cpp-cilk/graph.c
Bring up to date with james-cpp

@@ -6,6 +6,8 @@
 #include <iostream>
 #include <string>
 
+constexpr int BITS_PER_UNSIGNED_INT (CHAR_BIT * sizeof(unsigned int));
+
 static void fail(std::string msg) {
     std::cerr << msg << std::endl;
     exit(1);
@@ -14,25 +16,37 @@ static void fail(std::string msg) {
 Graph::Graph(unsigned int n) {
     this->n = n;
     label = std::vector<unsigned int>(n, 0u);
-    adjmat = {n, std::vector<unsigned char>(n, false)};
-    antiadjmat = {n, std::vector<unsigned char>(n, true)};
-    for (unsigned int i=0; i<n; i++)
-        antiadjmat[i][i] = false;
+    adjmat = {n, std::vector<unsigned int>(n, false)};
 }
 
-void add_edge(struct Graph *g, int v, int w) {
+Graph induced_subgraph(struct Graph& g, std::vector<int> vv) {
+    Graph subg(vv.size());
+    for (int i=0; i<subg.n; i++)
+        for (int j=0; j<subg.n; j++)
+            subg.adjmat[i][j] = g.adjmat[vv[i]][vv[j]];
+
+    for (int i=0; i<subg.n; i++)
+        subg.label[i] = g.label[vv[i]];
+    return subg;
+}
+
+void add_edge(Graph& g, int v, int w, bool directed=false, unsigned int val=1) {
     if (v != w) {
-        g->adjmat[v][w] = 1;
-        g->adjmat[w][v] = 1;
-        g->antiadjmat[v][w] = 0;
-        g->antiadjmat[w][v] = 0;
+        if (directed) {
+            g.adjmat[v][w] |= val;
+            g.adjmat[w][v] |= (val<<16);
+        } else {
+            g.adjmat[v][w] = val;
+            g.adjmat[w][v] = val;
+        }
     } else {
-        // To indicate that a vertex has a loop, we set its label to 1
-        g->label[v] = 1;
+        // To indicate that a vertex has a loop, we set the most
+        // significant bit of its label to 1
+        g.label[v] |= (1u << (BITS_PER_UNSIGNED_INT-1));
     }
 }
 
-struct Graph readDimacsGraph(char* filename) {
+struct Graph readDimacsGraph(char* filename, bool directed, bool labelled) {
     struct Graph g(0);
 
     FILE* f;
@@ -47,6 +61,7 @@ struct Graph readDimacsGraph(char* filename) {
[…]

### 2017-02-02 code/james-cpp-cilk/graph.h

@@ -5,11 +5,12 @@
 
 struct Graph {
     int n;
-    std::vector<std::vector<unsigned char>> adjmat;
-    std::vector<std::vector<unsigned char>> antiadjmat;
+    std::vector<std::vector<unsigned int>> adjmat;
     std::vector<unsigned int> label;
     Graph(unsigned int n);
 };
 
-Graph readGraph(char* filename, char format);
+Graph induced_subgraph(struct Graph& g, std::vector<int> vv);
+
+Graph readGraph(char* filename, char format, bool directed, bool labelled);
 

### 2017-02-02 code/james-cpp-cilk/mcsp.c

@@ -1,21 +1,23 @@
 #include "graph.h"
 
-#include <argp.h>
-#include <limits.h>
-#include <locale.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <time.h>
-
+#include <algorithm>
+#include <numeric>
+#include <chrono>
 #include <iostream>
+#include <set>
 #include <string>
 #include <utility>
 #include <vector>
-#include <chrono>
+#include <mutex>
+#include <thread>
+#include <condition_variable>
 #include <atomic>
 
-#include <cilk/cilk.h>
+#include <argp.h>
+#include <limits.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
 
 using std::vector;
 using std::cout;
@@ -26,40 +28,57 @@ static void fail(std::string msg) {
     exit(1);
 }
 
+enum Heuristic { min_max, min_product };
+
 /*******************************************************************************
                              Command-line arguments
 *******************************************************************************/
 
-static char doc[] = "Find a maximum clique in a graph in DIMACS format";
-static char args_doc[] = "FILENAME1 FILENAME2";
+static char doc[] = "Find a maximum clique in a graph in DIMACS format\vHEURISTIC can be min_max or min_product";
+static char args_doc[] = "HEURISTIC FILENAME1 FILENAME2";
 static struct argp_option options[] = {
     {"quiet", 'q', 0, 0, "Quiet output"},
     {"verbose", 'v', 0, 0, "Verbose output"},
     {"dimacs", 'd', 0, 0, "Read DIMACS format"},
     {"lad", 'l', 0, 0, "Read LAD format"},
     {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    {"directed", 'i', 0, 0, "Use directed graphs"},
+    {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
+    {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
+    {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
     { 0 }
 };
 
 static struct {
     bool quiet;
     bool verbose;
-    bool connected;
     bool dimacs;
     bool lad;
+    bool connected;
+    bool directed;
+    bool labelled;
+    bool big_first;
+    Heuristic heuristic;
     char *filename1;
     char *filename2;
+    int timeout;
     int arg_num;
[…]

### 2017-02-03 more-analysis/kup-vs-kdown.py
Added comparison of k-ascending with k-descending

@@ -1,0 +1,60 @@
+import struct
+import sys
+import re
+
+def read_n(filename):
+    with open(filename, "rb") as f:
+        return struct.unpack("<H", f.read(2))[0]
+
+def read_sol_size(filename):
+    with open(filename, "r") as f:
+        for line in f:
+            if line.startswith("Solution size"):
+                return int(line.strip().split()[2])
+
+def analyse_instance(instance, runtime_kup, runtime_kdown):
+    n0 = read_n(instance[1])
+    n1 = read_n(instance[2])
+    solsize_kup = read_sol_size(
+            "../experiments/gpgnode-results/mcsplain/james-cpp-max/" + instance[0] + ".out")
+    solsize_kdown = read_sol_size(
+            "../experiments/gpgnode-results/mcsplain/james-cpp-max/" + instance[0] + ".out")
+    if solsize_kup != solsize_kdown:
+        sys.stderr.write("Solution sizes differ between runs!\n")
+        sys.exit(1)
+    runtime_ratio = float(runtime_kup) / runtime_kdown
+    print instance[0], n0, n1, runtime_kup, runtime_kdown, runtime_ratio, solsize_kup, \
+            float(solsize_kup) / n0
+    
+def print_header():
+    print "instance", \
+            "n_pattern", \
+            "n_target", \
+            "runtime_kup", \
+            "runtime_kdown", \
+            "runtime_ratio", \
+            "solsize", \
+            "solsize_over_n_pattern"
+def go():
+    print_header()
+    runtimes = {}
+    with open("../experiments/gpgnode-results/mcsplain/runtimes.data", "r") as f:
+        lines = [line.strip().split() for line in f]
+    for tokens in lines[1:]:
+        kup_time = int(tokens[7])
+        kdown_time = int(tokens[8])
+        if kup_time == 0:
+            kup_time = 1
+        if kdown_time == 0:
+            kdown_time = 1
+        runtimes[tokens[0]] = {"kup":kup_time, "kdown":kdown_time}
+
+    with open("../experiments/mcsplaininstances.txt", "r") as f:
+        instances = [line.strip().split() for line in f]
+    for instance in instances:
+        runtime = runtimes[instance[0]]
+        if runtime["kup"]<1000000 and runtime["kdown"]<1000000:
+            analyse_instance(instance, runtime["kup"], runtime["kdown"])
+
+if __name__=="__main__":
+    go()

### 2017-02-03 more-analysis/kup-vs-kdown.py
Added colour to k up vs k down scatterplot, and removed confusing plot

@@ -22,8 +22,7 @@ def analyse_instance(instance, runtime_kup, runtime_kdown):
     if solsize_kup != solsize_kdown:
         sys.stderr.write("Solution sizes differ between runs!\n")
         sys.exit(1)
-    runtime_ratio = float(runtime_kup) / runtime_kdown
-    print instance[0], n0, n1, runtime_kup, runtime_kdown, runtime_ratio, solsize_kup, \
+    print instance[0], n0, n1, runtime_kup, runtime_kdown, solsize_kup, \
             float(solsize_kup) / n0
     
 def print_header():
@@ -32,7 +31,6 @@ def print_header():
             "n_target", \
             "runtime_kup", \
             "runtime_kdown", \
-            "runtime_ratio", \
             "solsize", \
             "solsize_over_n_pattern"
 def go():
@@ -43,10 +41,6 @@ def go():
     for tokens in lines[1:]:
         kup_time = int(tokens[7])
         kdown_time = int(tokens[8])
-        if kup_time == 0:
-            kup_time = 1
-        if kdown_time == 0:
-            kdown_time = 1
         runtimes[tokens[0]] = {"kup":kup_time, "kdown":kdown_time}
 
     with open("../experiments/mcsplaininstances.txt", "r") as f:

### 2017-02-03 more-analysis/kup-vs-kdown.py
Added k up vs k down for labelled graphs

@@ -12,13 +12,15 @@ def read_sol_size(filename):
             if line.startswith("Solution size"):
                 return int(line.strip().split()[2])
 
-def analyse_instance(instance, runtime_kup, runtime_kdown):
+def analyse_instance(instance, runtime_kup, runtime_kdown, dataset_name):
     n0 = read_n(instance[1])
     n1 = read_n(instance[2])
     solsize_kup = read_sol_size(
-            "../experiments/gpgnode-results/mcsplain/james-cpp-max/" + instance[0] + ".out")
+            "../experiments/gpgnode-results/{}/james-cpp-max/{}.out".format(
+                dataset_name, instance[0]))
     solsize_kdown = read_sol_size(
-            "../experiments/gpgnode-results/mcsplain/james-cpp-max/" + instance[0] + ".out")
+            "../experiments/gpgnode-results/{}/james-cpp-max-down/{}.out".format(
+                dataset_name, instance[0]))
     if solsize_kup != solsize_kdown:
         sys.stderr.write("Solution sizes differ between runs!\n")
         sys.exit(1)
@@ -33,22 +35,23 @@ def print_header():
             "runtime_kdown", \
             "solsize", \
             "solsize_over_n_pattern"
-def go():
+
+def go(dataset_name, kup_time_column, kdown_time_column):
     print_header()
     runtimes = {}
-    with open("../experiments/gpgnode-results/mcsplain/runtimes.data", "r") as f:
+    with open("../experiments/gpgnode-results/{}/runtimes.data".format(dataset_name), "r") as f:
         lines = [line.strip().split() for line in f]
     for tokens in lines[1:]:
-        kup_time = int(tokens[7])
-        kdown_time = int(tokens[8])
+        kup_time = int(tokens[kup_time_column])
+        kdown_time = int(tokens[kdown_time_column])
         runtimes[tokens[0]] = {"kup":kup_time, "kdown":kdown_time}
 
-    with open("../experiments/mcsplaininstances.txt", "r") as f:
+    with open("../experiments/{}instances.txt".format(dataset_name), "r") as f:
         instances = [line.strip().split() for line in f]
     for instance in instances:
         runtime = runtimes[instance[0]]
         if runtime["kup"]<1000000 and runtime["kdown"]<1000000:
-            analyse_instance(instance, runtime["kup"], runtime["kdown"])
[…]

### 2017-02-04 paper/paper.tex
More plotting

@@
 
-Dual Xeon E5-2640 v2 CPUs, 64GBytes RAM. Parallel experiments 16 threads (one per hardware core), and 32 with hyper-threading.
+Dual Xeon E5-2640 v2 CPUs, 64GBytes RAM. Parallel experiments 32 threads, using
+all 16 physical cores with hyper-threading enabled.
 
@@
 
-\paragraph{Large subgraph isomorphism instances} have yet to be run.
+\paragraph{Large subgraph isomorphism instances} are in \cref{figure:sip-cumulative}.
 
-\paragraph{Compared to CP FC} Runtimes are in \cref{figure:plain-james-versus-cp-fc-scatter} and \cref{figure:33ved-james-versus-cp-fc-scatter}, and nodes are in \cref{figure:plain-james-versus-cp-fc-nodes-scatter} and \cref{figure:33ved-james-versus-cp-fc-nodes-scatter}.
+\paragraph{Compared to CP FC} Nodes are in \cref{figure:plain-james-versus-cp-fc-nodes-scatter} and \cref{figure:33ved-james-versus-cp-fc-nodes-scatter}.
+
+\paragraph{Nodes compared to kdown on SIP} have yet to be run.
 
@@
     \centering
-    \includegraphics*{gen-graph-plain-james-versus-cp-fc-scatter.pdf}
-    \caption{MCS Plain instances, James vs CP-FC, Runtimes (TODO?? heatmapify this)}\label{figure:plain-james-versus-cp-fc-scatter}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-scatter.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Runtimes (TODO?? heatmapify this)}\label{figure:33ved-james-versus-cp-fc-scatter}
+    \includegraphics*{gen-graph-sip-cumulative.pdf}
+    \caption{SIP instances, cumulative runtimes}\label{figure:sip-cumulative}
 \end{figure}
@@
     \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{MCS Plain instances, James vs CP-FC, Nodes (TODO?? heatmapify this)}\label{figure:plain-james-versus-cp-fc-nodes-scatter}
+    \caption{MCS Plain instances, James vs CP-FC, Nodes (TODO?? heatmapify this
+    and exclude timeouts)}\label{figure:plain-james-versus-cp-fc-nodes-scatter}
 \end{figure}
@@
     \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, James vs CP-FC, Nodes (TODO?? heatmapify this)}\label{figure:33ved-james-versus-cp-fc-nodes-scatter}
+    \caption{MCS 33\% vertex and edge labelled directed instances, James vs
+    CP-FC, Nodes (TODO?? heatmapify this and exclude
+    timeouts)}\label{figure:33ved-james-versus-cp-fc-nodes-scatter}
 \end{figure}

### 2017-02-04 paper/paper.tex
Plot plot plot
More plotting

@@
 
-\paragraph{Vertex and edge labels, connected} have yet to be run.
+\paragraph{Vertex and edge labels, undirected, connected} are in \cref{figure:33ve-connected-cumulative}.
 
@@
     \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}\label{figure:33ved-cumulative}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-33ve-connected-cumulative.pdf}
+    \caption{MCS 33\% vertex and edge labelled undirected instances, connected, cumulative runtimes}\label{figure:33ve-connected-cumulative}
 \end{figure}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
+    \caption{SIP instances, James vs kdown, Nodes (TODO?? heatmapify this and exclude
+    timeouts)}\label{figure:sip-james-versus-kdown-nodes-scatter}
+\end{figure}
+
 \bibliographystyle{named}

### 2017-02-05 code/james-cpp-cilk/mcsp.c
Hacky atomic incumbent

@@ -19,9 +19,44 @@
 #include <stdlib.h>
 #include <string.h>
 
+#include <cilk/reducer_opadd.h>
+
 using std::vector;
 using std::cout;
 using std::endl;
+
+struct AtomicIncumbent
+{
+    std::atomic<unsigned> value;
+
+    AtomicIncumbent()
+    {
+        value.store(0, std::memory_order_seq_cst);
+    }
+
+    bool update(unsigned v)
+    {
+        while (true) {
+            unsigned cur_v = value.load(std::memory_order_seq_cst);
+            if (v > cur_v) {
+                if (value.compare_exchange_strong(cur_v, v, std::memory_order_seq_cst))
+                    return true;
+            }
+            else
+                return false;
+        }
+    }
+
+    bool beaten_by(unsigned v) const
+    {
+        return v > value.load(std::memory_order_seq_cst);
+    }
+
+    unsigned get() const
+    {
+        return value.load(std::memory_order_relaxed);
+    }
+};
 
 static void fail(std::string msg) {
     std::cerr << msg << std::endl;
@@ -151,7 +186,7 @@ static struct argp argp = { options, parse_opt, args_doc, doc };
                                      Stats
 *******************************************************************************/
 
-unsigned long long nodes{ 0 };
+cilk::reducer<cilk::op_add<unsigned long long> > nodes{ 0 };
 
 /*******************************************************************************
                                  MCS functions
@@ -178,7 +213,7 @@ struct Bidomain {
 void show(const vector<VtxPair>& current, const vector<Bidomain> &domains,
         const vector<int>& left, const vector<int>& right)
 {
-    cout << "Nodes: " << nodes << std::endl;
+    cout << "Nodes: " << nodes.get_value() << std::endl;
     cout << "Length of current assignment: " << current.size() << std::endl;
     cout << "Current assignment:";
     for (unsigned int i=0; i<current.size(); i++) {
@@ -358,7 +393,7 @@ void remove_bidomain(vector<Bidomain>& domains, int idx) {
     domains.pop_back();
 }
 
-void solve(const Graph & g0, const Graph & g1, vector<VtxPair> & incumbent,
+void solve(const Graph & g0, const Graph & g1, AtomicIncumbent & incumbent,
         vector<VtxPair> & current, vector<Bidomain> & domains,
[…]

### 2017-02-05 code/james-cpp-cilk/graph.cc
Fix James's annoying file extensions

@@ -1,4 +1,4 @@
-#include "graph.h"
+#include "graph.hh"
 
 #include <stdio.h>
 #include <stdlib.h>

### 2017-02-05 code/james-cpp-cilk/mcsp.cc

@@ -1,4 +1,4 @@
-#include "graph.h"
+#include "graph.hh"
 
 #include <algorithm>
 #include <numeric>

### 2017-02-05 code/james-cpp-cilk/mcsp.cc
constant current

@@ -394,7 +394,7 @@ void remove_bidomain(vector<Bidomain>& domains, int idx) {
 }
 
 void solve(const Graph & g0, const Graph & g1, AtomicIncumbent & incumbent,
-        vector<VtxPair> & current, vector<Bidomain> & domains,
+        const vector<VtxPair> & current, vector<Bidomain> & domains,
         vector<int> & left, vector<int> & right, unsigned int matching_size_goal)
 {
     if (abort_due_to_timeout)
@@ -433,9 +433,10 @@ void solve(const Graph & g0, const Graph & g1, AtomicIncumbent & incumbent,
 
         auto new_domains = filter_domains(domains, left, right, g0, g1, v, w,
                 arguments.directed || arguments.labelled);
-        current.push_back(VtxPair(v, w));
-        solve(g0, g1, incumbent, current, new_domains, left, right, matching_size_goal);
-        current.pop_back();
+
+        auto new_current = current;
+        new_current.push_back(VtxPair(v, w));
+        solve(g0, g1, incumbent, new_current, new_domains, left, right, matching_size_goal);
     }
     bd.right_len++;
     if (bd.left_len == 0)

### 2017-02-06 paper/paper.tex
We have those
Added Patrick's pseudocode

@@
 \usepackage{cleveref}
+
+\usepackage[ruled,vlined]{algorithm2e}
+\usepackage{MnSymbol,wasysym}
+\usepackage{latexsym}
 
@@
 \begin{abstract}
-    James came up with an awesome new algorithm for the maximum common (connected)
-    subgraph problem.
+    We came up with an important new algorithm for the maximum common (connected)
+    subgraph problem. This is an important problem, with at least two important
+    applications.
 \end{abstract}
@@
 \citet{DBLP:conf/cp/NdiayeS11}
+
+\section{The Lilybank Algorithm}
+
+\begin{algorithm}
+\DontPrintSemicolon
+\nl $expand(\mathit{future},matching)$ \;
+\nl \Begin{
+\nl \lIf {$future = \emptyset$ \bf{and} $|matching|  > |incumbent|$}{$save(matching)$}
+\nl \lIf {$future = \emptyset$}{return}
+\nl $bestFuture \gets 0$ \;
+\nl \lFor {$\langle P,T \rangle \in future$}{$bestFuture \gets bestFuture + min(|P|,|T|)$}
+\nl \lIf {$|matching|  + bestFuture \leq |incumbent|$}{return}
+\bigskip
+\nl $\langle P,T \rangle \gets select(future)$ \;
+\nl $v \gets select(P)$ \;
+\nl \For {$w \in T$}{
+\nl    $matching \gets matching + (v,w)$\;
+\nl    \bf{Let} $future^{'} \gets \emptyset$ \;
+\nl    \For {$\langle P^{'},T'^{'}\rangle \in future$}{
+\nl        \bf{Let} $P^{''} \gets P^{'} \cap N(v,G) \setminus \{v\}$ \;
+\nl        \bf{Let} $T^{''} \gets T^{'} \cap N(w,H) \setminus \{w\}$ \;
+\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
+\nl        \bf{Let} $P^{''} \gets P^{'} \cap N(v,\overline{G}) \setminus \{v\}$ \;
+\nl        \bf{Let} $T^{''} \gets T^{'} \cap N(w,\overline{H}) \setminus \{w\}$ \;
+\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
+       }
+\nl   $expand(future^{'},matching)$ \;
+\nl   $matching \gets matching \setminus \{(v,w)\}$ \;
+  }
+\nl $P \gets P \setminus \{v\}$\;
+\nl \lIf {$P = \emptyset$} {$future \gets future \setminus \{\langle P,T \rangle \}$}
+\nl $expand(future,matching)$ \;
+}
+\;
+\nl $Split(G,H)$ \;
+\nl \Begin{
+\nl $incumbent \gets \emptyset$ \;
+\nl $expand(\{\langle V(G),V(H) \rangle \},\emptyset)$ \;
+\nl return $|incubent|$ \;
+}
+\caption{Lilybank splitting algorithm}
+\label{jtAlg}
+\end{algorithm}
+
+\noindent
+Now, let's try and explain this. First, assume we have two graph G and H and there are no more vertices in G than in H (think of G as the pattern graph and H as the target).
+future is a list of pairs, where the first element of a pair is a set of pattern vertices P and the second is a set of target vertices T. It is assumed that each pattern vertex in P 
+might be matched with any (all) of the target vertices in T, or be unmatched.
+Therefore we start our search via the function LilybankSplit (line 26).
+This returns an integer, the size of the largest matching.
+In line 29 a call is made to expand, where we have as arguments a list of pairs (called future, line 1) and a set of pairs of the form (v,w) where pattern vertex v is matched to target vertex w.
+
+\bigskip
+\noindent
+We now focus on expand (line (1)). This is initially called (line 29) with the future as a list containing the single pair $\langle V(G),V(H) \rangle$. Lines 4 to 7 define the conditions 
+where we return, i.e. when the incumbent cannot be improved. In line 3, a new incumbent might be saved. Lines 8 to 24 perform the search.
+
+\bigskip
+\noindent
+A pair is selected from the future (line 8) using some heuristic.
+Note that this pair is not removed from the future, and that P and T can be considered as pointers to sets.
+A pattern vertex is then selected, possibly using a heuristic (line 9).
+We now iterate over all possible target vertices in T (lines 10 to 21).
+A target vertex w is selected and a matching produced (line 11).
+A new future is created (using {\bf Let} to signify a new variable, line 12).
+Every pair in the future can now be split (lines 13 to 19) into pattern vertices adjacent to the selected pattern vertex v and target vertices adjacent to the selected target vertex w (lines 14 to 15) and this is added to the new future if both sets are not null (line 16) i.e.
+there is a potential for more matchings.
+This is then repeated symmetrically for non-adjacency (lines 17 to 19) using the compliments of the graphs.
+A recursive call is then made using the new future and the increased matching (line 20).
+On return from that call we remove the matching (v,w) (line 21).
+Having explored all matchings with v in its target set we now consider what happens if the selected pattern vertex v is not matched (line 22 to 24), where that set of matchings does not contain a pair (v,*).
+
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\noindent
+Phew!
 
@@
 
-\paragraph{Nodes compared to kdown on SIP} have yet to be run.
+\paragraph{Nodes compared to kdown on SIP} are in \cref{figure:sip-james-versus-kdown-nodes-scatter}.
 

### 2017-02-06 paper/paper.tex
Added some stuff from James's short note

@@
 
+\usepackage{tikz}
+
+\usetikzlibrary{graphs}
+
 \newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
 \newcommand{\citep}[1]{\cite{#1}}
+
+\newcommand{\examplexG}[5] {
+    \begin{minipage}{.3\textwidth}
+    \tikz {
+        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1cm,
+                clockwise=5] {
+                    1[label=90:#1],2[label=0:#2],3[label=0:#3],4[label=180:#4],5[label=180:#5];
+            1--4; 1--5; 2--3; 2--5; 3--5;
+        };
+    }
+    \end{minipage}
+}
+\newcommand{\examplexH}[6] {
+    \begin{minipage}{.3\textwidth}
+    \tikz {
+        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1.1cm,
+                clockwise=6, phase=60] {
+                    a[label=0:#1],b[label=0:#2],c[label=0:#3],d[label=180:#4],e[label=180:#5],f[label=180:#6];
+            a--b; a--c; a--e; b--d; b--f; c--d; c--e; c--f; d--f; e--f;
+        };
+    }
+    \end{minipage}
+}
 
@@
 \citet{DBLP:conf/cp/NdiayeS11}
+
+Given graphs $G=(V^G, E^G)$ and $H=(V^H, E^H)$, the \emph{Maximum Common
+Subgraph (MCS)} problem is to find a graph with as many vertices as possible
+that is isomorphic to induced subgraphs of both $G$ and $H$. Equivalently, MCS
+is the problem of finding a maximum-cardinality matching $M = \{(v_1, w_1),
+\dots, (v_{|M|}, w_{|M|})\}$, such that for each pair $i,j$ ($1 \leq i \leq j
+\leq |M|$), $v_i$ and $v_j$ are adjacent in $G$ if and only if $w_i$ and $w_j$
+are adjacent in $H$.
+
+The most successful existing approaches to optimally solving MCS are
+reformulation to the maximum clique problem and constraint programming
+\cite{DBLP:conf/cp/McCreeshNPS16}. In this paper, we introduce a new branch and bound
+algorithm for MCS which does not require a domain to be explicity maintained in
+memory for each vertex in $V^G$. The algorithm maintains labellings of the
+unmatched vertices in $V^G$ and $V^H$ such that a pair $(v, w)$ with $v \in V^G$
+and $w \in V^H$ can be added to the matching if and only if $v$ and $w$ have the
+same label.
+
+\section{Preliminaries}
+
+Let $N(v, G)$ denote the neighbourhood of a vertex $v$ in a graph $G$.  We
+extend this notation to denote the set of vertices that are adjacent to all
+vertices in a set $U$: $N(U, G) = \bigcap_{v \in U} N_G(v)$.
+
+\section{Walkthrough}
+
+Suppose we are given the graphs $G$ and $H$ in Figure~\ref{fig:alg1}.  Graphs $G$ and $H$
+have no common subgraph with five vertices. However, they do have common
+subgraphs with four vertices; the mapping $\{1a, 2f, 3d, 5b\}$ is one example.
+
+\begin{figure}[ht]
+\centering
+    \examplexG{}{}{}{}{}
+    \examplexH{}{}{}{}{}{}
+\caption{Step 1}
+\label{fig:alg1}
+\end{figure}
+
+The algorithm proceeds by building up a mapping $M$, starting with the empty
+mapping $\{\}$. Begin by (arbitrarily) mapping vertex $1$ in $G$ to vertex $a$
+in $H$, giving $M=\{1a\}$.  Label each unmatched vertex according to whether it
+is adjacent to $1$ (in $G$) or $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.
+Adjacent vertices have label 1; non-adjacent vertices have label 0.  It is
+straightforward to see that we can extend $M$ with a vertex mapping $vw$, with
+$v \in V^G$ and $w \in V^H$, if and only if $v$ and $w$ have the same label.  This
+property is the algorithm's key invariant.
+
+\begin{figure}[ht]
+\centering
+    \examplexG{$(1a)$}{0}{0}{1}{1}
+    \examplexH{$(1a)$}{1}{1}{0}{1}{0}
+\caption{Step 2}
+\label{fig:alg2}
+\end{figure}
+
+Next, extend the mapping by pairing a vertex in $G$ with a vertex in $H$ of the
+same label; we will choose to map vertex $2$ to vertex $d$, giving $M=\{1a,
+2d\}$ (Figure~\ref{fig:alg3}).  Each unmatched vertex $v \in V^G$ is labelled
+with a two-character bit string, indicating whether $v$ is adjacent to each of
+the two matched vertices in $V^G$ (vertices $1$ and $2$).  For example, vertex
+$3$ is labelled $01$, indicating that it is not adjacent to $1$ but is adjacent
+to $2$.  Labels are given to vertices in $V^H$ in a similar fashion, showing
+whether each vertex is adjacent to $a$ and $d$.
+
+Observe that our invariant is maintained: we can extend $M$ by a vertex pairing if
+and only if the two vertices have the same colour.
+
+\begin{figure}[ht]
+\centering
+    \examplexG{$(1a)$}{$(2d)$}{01}{10}{11}
+    \examplexH{$(1a)$}{11}{11}{$(2d)$}{10}{01}
+\caption{Step 3}
+\label{fig:alg3}
+\end{figure}
+
+The algorithm proceeds by backtracking search. When we have explored the full
+search space of matchings containing $\{1a, 2d\}$, we try reassigning $2$ to
+$f$.  Since $d$ and $f$ are the only vertices to which $2$ can be matched given
+the decision to match $1$ to $a$, we lastly explore the possibility that $2$ is
+left unmatched, by giving $2$ the label $\bot$ and selecting another vertex in
+$V^G$ to assign.
 

### 2017-02-07 paper/paper.tex
Put in partial 33ve results for now
Plain connected plots

@@
 
-\paragraph{No labels, undirected, connected} have yet to be run.
+\paragraph{Vertex and edge labels, not directed} are in \cref{figure:33ve-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
+
+\paragraph{No labels, undirected, connected} are in \cref{figure:plain-connected-cumulative}.
 
@@
     \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}\label{figure:33ved-cumulative}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-33ve-cumulative.pdf}
+    \caption{MCS 33\% vertex and edge labelled undirected instances, cumulative runtimes}\label{figure:33ve-cumulative}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-plain-connected-cumulative.pdf}
+    \caption{Unlabelled undirected instances, connected, cumulative runtimes}\label{figure:plain-connected-cumulative}
 \end{figure}

### 2017-02-07 code/james-cpp/graph.c
Added option to use vertex labels but not edge labels

@@ -46,7 +46,7 @@ void add_edge(Graph& g, int v, int w, bool directed=false, unsigned int val=1) {
     }
 }
 
-struct Graph readDimacsGraph(char* filename, bool directed, bool labelled) {
+struct Graph readDimacsGraph(char* filename, bool directed, bool vertex_labelled) {
     struct Graph g(0);
 
     FILE* f;
@@ -80,7 +80,7 @@ struct Graph readDimacsGraph(char* filename, bool directed, bool labelled) {
             case 'n':
                 if (sscanf(line, "n %d %d", &v, &label)!=2)
                     fail("Error reading a line beginning with n.\n");
-                if (labelled)
+                if (vertex_labelled)
                     g.label[v-1] |= label;
                 break;
             }
@@ -129,7 +129,9 @@ int read_word(FILE *fp) {
     return (int)a[0] | (((int)a[1]) << 8);
 }
 
-struct Graph readBinaryGraph(char* filename, bool directed, bool labelled) {
+struct Graph readBinaryGraph(char* filename, bool directed, bool edge_labelled,
+        bool vertex_labelled)
+{
     struct Graph g(0);
     FILE* f;
     
@@ -153,7 +155,7 @@ struct Graph readBinaryGraph(char* filename, bool directed, bool labelled) {
     
     for (int i=0; i<nvertices; i++) {
         int label = (read_word(f) >> (16-k1));
-        if (labelled)
+        if (vertex_labelled)
             g.label[i] |= label;
     }
 
@@ -162,18 +164,18 @@ struct Graph readBinaryGraph(char* filename, bool directed, bool labelled) {
         for (int j=0; j<len; j++) {
             int target = read_word(f);
             int label = (read_word(f) >> (16-k1)) + 1;
-            add_edge(g, i, target, directed, labelled ? label : 1);
+            add_edge(g, i, target, directed, edge_labelled ? label : 1);
         }
     }
     fclose(f);
     return g;
 }
 
-struct Graph readGraph(char* filename, char format, bool directed, bool labelled) {
+struct Graph readGraph(char* filename, char format, bool directed, bool edge_labelled, bool vertex_labelled) {
     struct Graph g(0);
-    if (format=='D') g = readDimacsGraph(filename, directed, labelled);
+    if (format=='D') g = readDimacsGraph(filename, directed, vertex_labelled);
     else if (format=='L') g = readLadGraph(filename, directed);
[…]

### 2017-02-07 code/james-cpp/graph.h

@@ -12,5 +12,5 @@ struct Graph {
 
 Graph induced_subgraph(struct Graph& g, std::vector<int> vv);
 
-Graph readGraph(char* filename, char format, bool directed, bool labelled);
+Graph readGraph(char* filename, char format, bool directed, bool edge_labelled, bool vertex_labelled);
 

### 2017-02-07 code/james-cpp/mcsp.c

@@ -44,6 +44,7 @@ static struct argp_option options[] = {
     {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
     {"directed", 'i', 0, 0, "Use directed graphs"},
     {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
+    {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
     {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
     {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
     { 0 }
@@ -57,6 +58,7 @@ static struct {
     bool connected;
     bool directed;
     bool labelled;
+    bool vertex_labelled_only;
     bool big_first;
     Heuristic heuristic;
     char *filename1;
@@ -75,6 +77,7 @@ void set_default_arguments() {
     arguments.connected = false;
     arguments.directed = false;
     arguments.labelled = false;
+    arguments.vertex_labelled_only = false;
     arguments.big_first = false;
     arguments.filename1 = NULL;
     arguments.filename2 = NULL;
@@ -111,7 +114,14 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             arguments.directed = true;
             break;
         case 'a':
+            if (arguments.vertex_labelled_only)
+                fail("The -a and -x options can't be used together.");
             arguments.labelled = true;
+            break;
+        case 'x':
+            if (arguments.labelled)
+                fail("The -a and -x options can't be used together.");
+            arguments.vertex_labelled_only = true;
             break;
         case 'b':
             arguments.big_first = true;
@@ -488,8 +498,10 @@ int main(int argc, char** argv) {
     argp_parse(&argp, argc, argv, 0, 0, 0);
 
     char format = arguments.dimacs ? 'D' : arguments.lad ? 'L' : 'B';
-    struct Graph g0 = readGraph(arguments.filename1, format, arguments.directed, arguments.labelled);
-    struct Graph g1 = readGraph(arguments.filename2, format, arguments.directed, arguments.labelled);
+    struct Graph g0 = readGraph(arguments.filename1, format, arguments.directed,
+            arguments.labelled, arguments.labelled || arguments.vertex_labelled_only);
+    struct Graph g1 = readGraph(arguments.filename2, format, arguments.directed,
[…]

### 2017-02-07 code/james-cpp/mcsp.c
Changed variable names for vertex and edge labelled options

@@ -57,8 +57,8 @@ static struct {
     bool lad;
     bool connected;
     bool directed;
-    bool labelled;
-    bool vertex_labelled_only;
+    bool edge_labelled;
+    bool vertex_labelled;
     bool big_first;
     Heuristic heuristic;
     char *filename1;
@@ -76,8 +76,8 @@ void set_default_arguments() {
     arguments.lad = false;
     arguments.connected = false;
     arguments.directed = false;
-    arguments.labelled = false;
-    arguments.vertex_labelled_only = false;
+    arguments.edge_labelled = false;
+    arguments.vertex_labelled = false;
     arguments.big_first = false;
     arguments.filename1 = NULL;
     arguments.filename2 = NULL;
@@ -114,14 +114,15 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             arguments.directed = true;
             break;
         case 'a':
-            if (arguments.vertex_labelled_only)
+            if (arguments.vertex_labelled)
                 fail("The -a and -x options can't be used together.");
-            arguments.labelled = true;
+            arguments.edge_labelled = true;
+            arguments.vertex_labelled = true;
             break;
         case 'x':
-            if (arguments.labelled)
+            if (arguments.edge_labelled)
                 fail("The -a and -x options can't be used together.");
-            arguments.vertex_labelled_only = true;
+            arguments.vertex_labelled = true;
             break;
         case 'b':
             arguments.big_first = true;
@@ -410,7 +411,7 @@ void solve(const Graph & g0, const Graph & g1, vector<VtxPair> & incumbent,
         right[bd.r + bd.right_len] = w;
 
         auto new_domains = filter_domains(domains, left, right, g0, g1, v, w,
-                arguments.directed || arguments.labelled);
+                arguments.directed || arguments.edge_labelled);
         current.push_back(VtxPair(v, w));
         solve(g0, g1, incumbent, current, new_domains, left, right, matching_size_goal);
         current.pop_back();
@@ -499,9 +500,9 @@ int main(int argc, char** argv) {
 
     char format = arguments.dimacs ? 'D' : arguments.lad ? 'L' : 'B';
     struct Graph g0 = readGraph(arguments.filename1, format, arguments.directed,
-            arguments.labelled, arguments.labelled || arguments.vertex_labelled_only);
+            arguments.edge_labelled, arguments.vertex_labelled);
[…]

### 2017-02-07 paper/paper.tex
Added titles of some other sections in paper

@@
 
+\DeclareMathOperator{\V}{V}
+\DeclareMathOperator{\E}{E}
+
 \newcommand{\examplexG}[5] {
-    \begin{minipage}{.3\textwidth}
+    \begin{minipage}{.2\textwidth}
     \tikz {
@@
 \newcommand{\examplexH}[6] {
-    \begin{minipage}{.3\textwidth}
+    \begin{minipage}{.2\textwidth}
     \tikz {
@@
 
-Given graphs $G=(V^G, E^G)$ and $H=(V^H, E^H)$, the \emph{Maximum Common
+Given graphs $G=(\V(G), \E(G))$ and $H=(\V(H), \E(H))$, the \emph{Maximum Common
 Subgraph (MCS)} problem is to find a graph with as many vertices as possible
 that is isomorphic to induced subgraphs of both $G$ and $H$. Equivalently, MCS
-is the problem of finding a maximum-cardinality matching $M = \{(v_1, w_1),
-\dots, (v_{|M|}, w_{|M|})\}$, such that for each pair $i,j$ ($1 \leq i \leq j
+is the problem of finding a maximum-cardinality set of pairs $M = \{(v_1, w_1),
+\dots, (v_{|M|}, w_{|M|})\}$, with $v_i \in V(G)$ and $v_i \in V(H)$ for all $i$,
+such that for each pair $i,j$ ($1 \leq i < j
 \leq |M|$), $v_i$ and $v_j$ are adjacent in $G$ if and only if $w_i$ and $w_j$
@@
 algorithm for MCS which does not require a domain to be explicity maintained in
-memory for each vertex in $V^G$. The algorithm maintains labellings of the
-unmatched vertices in $V^G$ and $V^H$ such that a pair $(v, w)$ with $v \in V^G$
-and $w \in V^H$ can be added to the matching if and only if $v$ and $w$ have the
+memory for each vertex in $\V(G)$. The algorithm maintains labellings of the
+unmatched vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in \V(G)$
+and $w \in \V(H)$ can be added to the matching if and only if $v$ and $w$ have the
 same label.
@@
     \examplexH{}{}{}{}{}{}
-\caption{Step 1}
+\caption{Graphs $G$ and $H$}
 \label{fig:alg1}
@@
 straightforward to see that we can extend $M$ with a vertex mapping $vw$, with
-$v \in V^G$ and $w \in V^H$, if and only if $v$ and $w$ have the same label.  This
+$v \in \V(G)$ and $w \in \V(H)$, if and only if $v$ and $w$ have the same label.  This
 property is the algorithm's key invariant.
@@
 same label; we will choose to map vertex $2$ to vertex $d$, giving $M=\{1a,
-2d\}$ (Figure~\ref{fig:alg3}).  Each unmatched vertex $v \in V^G$ is labelled
+2d\}$ (Figure~\ref{fig:alg3}).  Each unmatched vertex $v \in \V(G)$ is labelled
 with a two-character bit string, indicating whether $v$ is adjacent to each of
-the two matched vertices in $V^G$ (vertices $1$ and $2$).  For example, vertex
+the two matched vertices in $\V(G)$ (vertices $1$ and $2$).  For example, vertex
 $3$ is labelled $01$, indicating that it is not adjacent to $1$ but is adjacent
-to $2$.  Labels are given to vertices in $V^H$ in a similar fashion, showing
+to $2$.  Labels are given to vertices in $\V(H)$ in a similar fashion, showing
 whether each vertex is adjacent to $a$ and $d$.
@@
 left unmatched, by giving $2$ the label $\bot$ and selecting another vertex in
-$V^G$ to assign.
+$\V(G)$ to assign.
 
@@
 Phew!
+
+\section{Heuristics}
+
+\section{Extensions}
+
+So far, we have considered only unlabelled, undirected graphs. In this section, we
+describe some straightforward changes to the algorithm that relax these restrictions.
+
+For the case of unlabelled, directed graphs, each adjacency-matrix entry for a vertex
+pair $(u,v)$ in graph $G$ or $H$ takes the value 0 if the vertices are not adjacent,
+1 if there is an edge $u \rightarrow v$ but no edge $v \rightarrow u$, 2 if there is a
+edge in the opposite direction, and 3 if there are edges in both directions. In lines
+14-19 of the algorithm .......
+
+Labelled, undirected .......
+
+Labelled, directed ......
+
+Connected ........
+
+\section{Future Work}
+
+Instead of branching on vertices in $\V(G)$, it would be equally valid to branch on
+a vertex in $\V(H)$ in lines 9-23 of the algorithm, since our data structure treats the
+two graphs symmetrically. More interestingly, we could choose which graph to branch on
+at each search node using some heuristic (perhaps choosing based on whether the $P$ or
+$T$ set is smaller).
+
+We could create a hybrid of Lilybank and clique (for example, using Lilybank near the
+top of the search tree, but switching to a clique encoding when fewer than some threshold
+number of vertices remain to be selected). This might give us most of the benefits of
+the clique encoding for labelled graphs, while avoiding the high memory cost and colouring
+time of encoding the full instance.
 
@@
 
+\section{Conclusion}
+
 \bibliographystyle{named}

### 2017-02-07 code/james-cpp-for-bigraphs/graph.c
Added a C++ version that doesn't bit-shift labels, for use on bigraphs

@@ -1,0 +1,169 @@
+#include "graph.h"
+
+#include <stdio.h>
+#include <stdlib.h>
+
+#include <iostream>
+#include <string>
+
+constexpr int BITS_PER_UNSIGNED_INT (CHAR_BIT * sizeof(unsigned int));
+
+static void fail(std::string msg) {
+    std::cerr << msg << std::endl;
+    exit(1);
+}
+
+Graph::Graph(unsigned int n) {
+    this->n = n;
+    label = std::vector<unsigned int>(n, 0u);
+    adjmat = {n, std::vector<unsigned int>(n, false)};
+}
+
+Graph induced_subgraph(struct Graph& g, std::vector<int> vv) {
+    Graph subg(vv.size());
+    for (int i=0; i<subg.n; i++)
+        for (int j=0; j<subg.n; j++)
+            subg.adjmat[i][j] = g.adjmat[vv[i]][vv[j]];
+
+    for (int i=0; i<subg.n; i++)
+        subg.label[i] = g.label[vv[i]];
+    return subg;
+}
+
+void add_edge(Graph& g, int v, int w, bool directed=false, unsigned int val=1) {
+    if (v != w) {
+        if (directed) {
+            g.adjmat[v][w] |= val;
+            g.adjmat[w][v] |= (val<<16);
+        } else {
+            g.adjmat[v][w] = val;
+            g.adjmat[w][v] = val;
+        }
+    } else {
+        // To indicate that a vertex has a loop, we set the most
+        // significant bit of its label to 1
+        g.label[v] |= (1u << (BITS_PER_UNSIGNED_INT-1));
+    }
+}
+
+struct Graph readDimacsGraph(char* filename, bool directed, bool vertex_labelled) {
+    struct Graph g(0);
+
+    FILE* f;
+    
+    if ((f=fopen(filename, "r"))==NULL)
+        fail("Cannot open file");
+
+    char* line = NULL;
+    size_t nchar = 0;
+
+    int nvertices = 0;
+    int medges = 0;
+    int v, w;
+    int edges_read = 0;
+    int label;
+
+    while (getline(&line, &nchar, f) != -1) {
+        if (nchar > 0) {
+            switch (line[0]) {
+            case 'p':
+                if (sscanf(line, "p edge %d %d", &nvertices, &medges)!=2)
+                    fail("Error reading a line beginning with p.\n");
+                g = Graph(nvertices);
+                break;
+            case 'e':
+                if (sscanf(line, "e %d %d", &v, &w)!=2)
[…]

### 2017-02-07 code/james-cpp-for-bigraphs/mcsp.c

@@ -1,0 +1,589 @@
+#include "graph.h"
+
+#include <algorithm>
+#include <numeric>
+#include <chrono>
+#include <iostream>
+#include <set>
+#include <string>
+#include <utility>
+#include <vector>
+#include <mutex>
+#include <thread>
+#include <condition_variable>
+#include <atomic>
+
+#include <argp.h>
+#include <limits.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+using std::vector;
+using std::cout;
+using std::endl;
+
+static void fail(std::string msg) {
+    std::cerr << msg << std::endl;
+    exit(1);
+}
+
+enum Heuristic { min_max, min_product };
+
+/*******************************************************************************
+                             Command-line arguments
+*******************************************************************************/
+
+static char doc[] = "Find a maximum clique in a graph in DIMACS format\vHEURISTIC can be min_max or min_product";
+static char args_doc[] = "HEURISTIC FILENAME1 FILENAME2";
+static struct argp_option options[] = {
+    {"quiet", 'q', 0, 0, "Quiet output"},
+    {"verbose", 'v', 0, 0, "Verbose output"},
+    {"dimacs", 'd', 0, 0, "Read DIMACS format"},
+    {"lad", 'l', 0, 0, "Read LAD format"},
+    {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    {"directed", 'i', 0, 0, "Use directed graphs"},
+    {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
+    {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
+    {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
+    {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
+    { 0 }
+};
+
+static struct {
+    bool quiet;
+    bool verbose;
+    bool dimacs;
+    bool lad;
+    bool connected;
+    bool directed;
+    bool edge_labelled;
+    bool vertex_labelled;
+    bool big_first;
+    Heuristic heuristic;
+    char *filename1;
+    char *filename2;
+    int timeout;
+    int arg_num;
+} arguments;
+
+static std::atomic<bool> abort_due_to_timeout;
+
+void set_default_arguments() {
+    arguments.quiet = false;
+    arguments.verbose = false;
+    arguments.dimacs = false;
[…]

### 2017-02-07 code/james-cpp-for-bigraphs/mcsp.c
Added MCS enumeration for bigraph version

@@ -45,7 +45,6 @@ static struct argp_option options[] = {
     {"directed", 'i', 0, 0, "Use directed graphs"},
     {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
     {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
-    {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
     {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
     { 0 }
 };
@@ -59,7 +58,6 @@ static struct {
     bool directed;
     bool edge_labelled;
     bool vertex_labelled;
-    bool big_first;
     Heuristic heuristic;
     char *filename1;
     char *filename2;
@@ -78,7 +76,6 @@ void set_default_arguments() {
     arguments.directed = false;
     arguments.edge_labelled = false;
     arguments.vertex_labelled = false;
-    arguments.big_first = false;
     arguments.filename1 = NULL;
     arguments.filename2 = NULL;
     arguments.timeout = 0;
@@ -123,9 +120,6 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             if (arguments.edge_labelled)
                 fail("The -a and -x options can't be used together.");
             arguments.vertex_labelled = true;
-            break;
-        case 'b':
-            arguments.big_first = true;
             break;
         case 't':
             arguments.timeout = std::stoi(arg);
@@ -370,8 +364,8 @@ void remove_bidomain(vector<Bidomain>& domains, int idx) {
 }
 
 void solve(const Graph & g0, const Graph & g1, vector<VtxPair> & incumbent,
-        vector<VtxPair> & current, vector<Bidomain> & domains,
-        vector<int> & left, vector<int> & right, unsigned int matching_size_goal)
+        vector<VtxPair> & current, vector<vector<VtxPair>>& solutions,
+        vector<Bidomain> & domains, vector<int> & left, vector<int> & right)
 {
     if (abort_due_to_timeout)
         return;
@@ -379,16 +373,15 @@ void solve(const Graph & g0, const Graph & g1, vector<VtxPair> & incumbent,
     if (arguments.verbose) show(current, domains, left, right);
     nodes++;
 
-    if (current.size() > incumbent.size()) {
+    if (current.size() >= incumbent.size()) {
+        if (current.size() > incumbent.size())
[…]

### 2017-02-07 paper/paper.tex
Even more gratuitous plotting

@@
 
-\paragraph{Vertex and edge labels, directed} are in \cref{figure:33ved-cumulative}.
+\paragraph{Vertex labels but no edge labels, undirected} are in \cref{figure:33v-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
 
 \paragraph{Vertex and edge labels, not directed} are in \cref{figure:33ve-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
+
+\paragraph{Vertex and edge labels, directed} are in \cref{figure:33ved-cumulative}.
 
@@
     \centering
-    \includegraphics*{gen-graph-33ved-cumulative.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}\label{figure:33ved-cumulative}
+    \includegraphics*{gen-graph-33v-cumulative.pdf}
+    \caption{MCS 33\% vertex labelled, no edge labels, undirected instances, cumulative runtimes}\label{figure:33v-cumulative}
 \end{figure}
@@
     \caption{MCS 33\% vertex and edge labelled undirected instances, cumulative runtimes}\label{figure:33ve-cumulative}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics*{gen-graph-33ved-cumulative.pdf}
+    \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}\label{figure:33ved-cumulative}
 \end{figure}

### 2017-02-08 paper/paper.tex
Added more drafty stuff to paper

@@
 
+% TODO: use operatorname for these
 \DeclareMathOperator{\V}{V}
 \DeclareMathOperator{\E}{E}
+
+\newcommand{\BigO}[1]{\ensuremath{\operatorname{O}\left(#1\right)}}
 
@@
 
-\noindent
 Now, let's try and explain this. First, assume we have two graph G and H and there are no more vertices in G than in H (think of G as the pattern graph and H as the target).
@@
 
-\bigskip
-\noindent
 We now focus on expand (line (1)). This is initially called (line 29) with the future as a list containing the single pair $\langle V(G),V(H) \rangle$. Lines 4 to 7 define the conditions 
@@
 
-\bigskip
-\noindent
 A pair is selected from the future (line 8) using some heuristic.
@@
 \bigskip
-\bigskip
-\bigskip
-\bigskip
 \noindent
@@
 \section{Heuristics}
+
+We ran a small experiment comparing lots of heuristics such as \dots on random
+instances. The one where you choose a colour class with as small a max(left size,
+right size) as possible seemed to work best. ? Why?
+
+\section{Comparison with Existing Algorithms}
+
+? Describe CP-FC and clique encodings
+
+The Lilybank algorithm explores exactly the same search tree as CP-FC, because \dots
+
+The new algorithm takes $\BigO{n}$ time at each search node. CP-FC takes $\BigO{?}$ time,
+and clique takes $\BigO{n^4}$ time.
+
+Our data structure lets us compute good heuristics cheaply, which is why we take
+slightly fewer search nodes than CP-FC.
+
+Clique gives us very strong propagation.
+
+Our algorithm needs less space than CP-FC ($\BigO{n^2}$ rather than $\BigO{n^3}$?) and
+much less than clique ($\BigO{n^4}$?).
 
@@
 
-For the case of unlabelled, directed graphs, each adjacency-matrix entry for a vertex
-pair $(u,v)$ in graph $G$ or $H$ takes the value 0 if the vertices are not adjacent,
-1 if there is an edge $u \rightarrow v$ but no edge $v \rightarrow u$, 2 if there is a
-edge in the opposite direction, and 3 if there are edges in both directions. In lines
-14-19 of the algorithm .......
+If we have vertex labels, we replace ${\langle V(G),V(H) \rangle}$ in line 28 of
+the algorithm with a set containing a colour class for each label that appears in
+both $\V(G)$ and $\V(H)$. In the remainder of this section, \textit{labels} refers to
+edge labels.
 
-Labelled, undirected .......
+For the unlabelled, directed case, let $A_G$ and $A_H$ be the adjacency matrices
+of $G$ and $H$ respectively, as stored in memory. For each
+vertex pair $(t,u)$ in graph $G$ or $H$, the adjacency matrix entry 
+takes the value 0 if the vertices are
+not adjacent, 1 if the two vertices share a single edge in the direction $t
+\rightarrow u$, 2 if they share a single edge in the direction $u \rightarrow
+t$, and 3 if there are edges in both directions. Where lines 14-19 of the
+basic algorithm split the pair $\langle P',T' \rangle$ two ways, according to whether
+vertices are ajacent to $v$ (resp. $w$), we now perform a four-way split where
+each vertex is classified according to the label on its adjacency matrix from $v$
+(resp. $w$). This is shown in Algorithm~\ref{labDirAlg}, where $L=\{0,1,2,3\}$.
 
-Labelled, directed ......
+\begin{algorithm}
+\DontPrintSemicolon
+\nl    \For {$l \in L$}{
+\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} \mid u \neq v \wedge A_G[v][u] = l \}$ \;
+\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} \mid u \neq w \wedge A_H[w][u] = l \}$ \;
+\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
+       }
+\caption{Replacement for lines 14--19 in directed in labelled cases}
+\label{labDirAlg}
+\end{algorithm}
 
-Connected ........
+In the labelled, undirected case, each adjacency-matrix entry contains the edge
+label, or a null entry 0 if no edge exists between the pair of vertices.  we
+can also use Algorithm~\ref{labDirAlg}, by letting $L$ be the union of $\{0\}$
+with the set of all labels that appear in the input graphs. Since there may be
+up to $n_G + n_H$ distinct labels, the loop in Algorithm~\ref{labDirAlg} may
+execute up to $n_G + n_H$ times, and this variation of the algorithm therefore
+requires $\BigO{n^2}$ time per search node.  It is possible to modify the
+algorithms to run in $\BigO{n \log n}$ per search node; we will give a brief
+outline of the method. First, run lines 17-19 of Algorithm~\ref{jtAlg} to
+create a new label-class of vertices that are not adjacent to $v$ or $w$, and
+remove these vertices from $\langle P',T' \rangle$. Next, sort $P'$ and $T'$ in
+ascending order of the label on the edge from $v$ or $w$ to each vertex. We can
+then create the colour classes corresponding to each edge label by
+simulataneously traversing $P'$ and $T'$ from left to right.
+
+The labelled, directed case is similar to its undirected counterpart, except that
+each element $A[u][v]$ in an adjacency matrix is a pair $(l_1, l_2)$, where
+$l_1$ is the label on the edge $u \rightarrow v$ (or 0 if no edge exists) and $l_2$
+is the label on the reverse edge.
+
+For the variant in which the common subgraph must be connected, we consider
+only undirected graphs. We modify the algorithm by permitting branching only on
+a vertex $v$ that has at least one non-zero element in its bitstring label.
+(Cite the first paper that did this.) ? Should we say that smarter propagation
+might be possible?
 

### 2017-02-09 code/james-cpp-cilk/graph.cc
Remove, so we can start again with the new code

@@ -1,179 +1,0 @@
-#include "graph.hh"
-
-#include <stdio.h>
-#include <stdlib.h>
-
-#include <iostream>
-#include <string>
-
-constexpr int BITS_PER_UNSIGNED_INT (CHAR_BIT * sizeof(unsigned int));
-
-static void fail(std::string msg) {
-    std::cerr << msg << std::endl;
-    exit(1);
-}
-
-Graph::Graph(unsigned int n) {
-    this->n = n;
-    label = std::vector<unsigned int>(n, 0u);
-    adjmat = {n, std::vector<unsigned int>(n, false)};
-}
-
-Graph induced_subgraph(struct Graph& g, std::vector<int> vv) {
-    Graph subg(vv.size());
-    for (int i=0; i<subg.n; i++)
-        for (int j=0; j<subg.n; j++)
-            subg.adjmat[i][j] = g.adjmat[vv[i]][vv[j]];
-
-    for (int i=0; i<subg.n; i++)
-        subg.label[i] = g.label[vv[i]];
-    return subg;
-}
-
-void add_edge(Graph& g, int v, int w, bool directed=false, unsigned int val=1) {
-    if (v != w) {
-        if (directed) {
-            g.adjmat[v][w] |= val;
-            g.adjmat[w][v] |= (val<<16);
-        } else {
-            g.adjmat[v][w] = val;
-            g.adjmat[w][v] = val;
-        }
-    } else {
-        // To indicate that a vertex has a loop, we set the most
-        // significant bit of its label to 1
-        g.label[v] |= (1u << (BITS_PER_UNSIGNED_INT-1));
-    }
-}
-
-struct Graph readDimacsGraph(char* filename, bool directed, bool labelled) {
-    struct Graph g(0);
-
-    FILE* f;
-    
-    if ((f=fopen(filename, "r"))==NULL)
-        fail("Cannot open file");
-
-    char* line = NULL;
-    size_t nchar = 0;
-
-    int nvertices = 0;
-    int medges = 0;
-    int v, w;
-    int edges_read = 0;
-    int label;
-
-    while (getline(&line, &nchar, f) != -1) {
-        if (nchar > 0) {
-            switch (line[0]) {
-            case 'p':
-                if (sscanf(line, "p edge %d %d", &nvertices, &medges)!=2)
-                    fail("Error reading a line beginning with p.\n");
-                g = Graph(nvertices);
-                break;
-            case 'e':
-                if (sscanf(line, "e %d %d", &v, &w)!=2)
[…]

### 2017-02-09 code/james-cpp-cilk/mcsp.cc

@@ -1,609 +1,0 @@
-#include "graph.hh"
-
-#include <algorithm>
-#include <numeric>
-#include <chrono>
-#include <iostream>
-#include <set>
-#include <string>
-#include <utility>
-#include <vector>
-#include <mutex>
-#include <thread>
-#include <condition_variable>
-#include <atomic>
-
-#include <argp.h>
-#include <limits.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-
-#include <cilk/reducer_opadd.h>
-
-using std::vector;
-using std::cout;
-using std::endl;
-
-struct AtomicIncumbent
-{
-    std::atomic<unsigned> value;
-
-    AtomicIncumbent()
-    {
-        value.store(0, std::memory_order_seq_cst);
-    }
-
-    bool update(unsigned v)
-    {
-        while (true) {
-            unsigned cur_v = value.load(std::memory_order_seq_cst);
-            if (v > cur_v) {
-                if (value.compare_exchange_strong(cur_v, v, std::memory_order_seq_cst))
-                    return true;
-            }
-            else
-                return false;
-        }
-    }
-
-    bool beaten_by(unsigned v) const
-    {
-        return v > value.load(std::memory_order_seq_cst);
-    }
-
-    unsigned get() const
-    {
-        return value.load(std::memory_order_relaxed);
-    }
-};
-
-static void fail(std::string msg) {
-    std::cerr << msg << std::endl;
-    exit(1);
-}
-
-enum Heuristic { min_max, min_product };
-
-/*******************************************************************************
-                             Command-line arguments
-*******************************************************************************/
-
-static char doc[] = "Find a maximum clique in a graph in DIMACS format\vHEURISTIC can be min_max or min_product";
-static char args_doc[] = "HEURISTIC FILENAME1 FILENAME2";
-static struct argp_option options[] = {
-    {"quiet", 'q', 0, 0, "Quiet output"},
-    {"verbose", 'v', 0, 0, "Verbose output"},
-    {"dimacs", 'd', 0, 0, "Read DIMACS format"},
-    {"lad", 'l', 0, 0, "Read LAD format"},
-    {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
-    {"directed", 'i', 0, 0, "Use directed graphs"},
-    {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
[…]

### 2017-02-09 code/james-cpp-parallel/graph.cc
Here we go again

@@ -1,0 +1,181 @@
+#include "graph.hh"
+
+#include <stdio.h>
+#include <stdlib.h>
+
+#include <iostream>
+#include <string>
+
+constexpr int BITS_PER_UNSIGNED_INT (CHAR_BIT * sizeof(unsigned int));
+
+static void fail(std::string msg) {
+    std::cerr << msg << std::endl;
+    exit(1);
+}
+
+Graph::Graph(unsigned int n) {
+    this->n = n;
+    label = std::vector<unsigned int>(n, 0u);
+    adjmat = {n, std::vector<unsigned int>(n, false)};
+}
+
+Graph induced_subgraph(struct Graph& g, std::vector<int> vv) {
+    Graph subg(vv.size());
+    for (int i=0; i<subg.n; i++)
+        for (int j=0; j<subg.n; j++)
+            subg.adjmat[i][j] = g.adjmat[vv[i]][vv[j]];
+
+    for (int i=0; i<subg.n; i++)
+        subg.label[i] = g.label[vv[i]];
+    return subg;
+}
+
+void add_edge(Graph& g, int v, int w, bool directed=false, unsigned int val=1) {
+    if (v != w) {
+        if (directed) {
+            g.adjmat[v][w] |= val;
+            g.adjmat[w][v] |= (val<<16);
+        } else {
+            g.adjmat[v][w] = val;
+            g.adjmat[w][v] = val;
+        }
+    } else {
+        // To indicate that a vertex has a loop, we set the most
+        // significant bit of its label to 1
+        g.label[v] |= (1u << (BITS_PER_UNSIGNED_INT-1));
+    }
+}
+
+struct Graph readDimacsGraph(char* filename, bool directed, bool vertex_labelled) {
+    struct Graph g(0);
+
+    FILE* f;
+    
+    if ((f=fopen(filename, "r"))==NULL)
+        fail("Cannot open file");
+
+    char* line = NULL;
+    size_t nchar = 0;
+
+    int nvertices = 0;
+    int medges = 0;
+    int v, w;
+    int edges_read = 0;
+    int label;
+
+    while (getline(&line, &nchar, f) != -1) {
+        if (nchar > 0) {
+            switch (line[0]) {
+            case 'p':
+                if (sscanf(line, "p edge %d %d", &nvertices, &medges)!=2)
+                    fail("Error reading a line beginning with p.\n");
+                g = Graph(nvertices);
+                break;
+            case 'e':
+                if (sscanf(line, "e %d %d", &v, &w)!=2)
[…]

### 2017-02-09 code/james-cpp-parallel/mcsp.cc

@@ -1,0 +1,589 @@
+#include "graph.hh"
+
+#include <algorithm>
+#include <numeric>
+#include <chrono>
+#include <iostream>
+#include <set>
+#include <string>
+#include <utility>
+#include <vector>
+#include <mutex>
+#include <thread>
+#include <condition_variable>
+#include <atomic>
+
+#include <argp.h>
+#include <limits.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+using std::vector;
+using std::cout;
+using std::endl;
+
+static void fail(std::string msg) {
+    std::cerr << msg << std::endl;
+    exit(1);
+}
+
+enum Heuristic { min_max, min_product };
+
+/*******************************************************************************
+                             Command-line arguments
+*******************************************************************************/
+
+static char doc[] = "Find a maximum clique in a graph in DIMACS format\vHEURISTIC can be min_max or min_product";
+static char args_doc[] = "HEURISTIC FILENAME1 FILENAME2";
+static struct argp_option options[] = {
+    {"quiet", 'q', 0, 0, "Quiet output"},
+    {"verbose", 'v', 0, 0, "Verbose output"},
+    {"dimacs", 'd', 0, 0, "Read DIMACS format"},
+    {"lad", 'l', 0, 0, "Read LAD format"},
+    {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    {"directed", 'i', 0, 0, "Use directed graphs"},
+    {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
+    {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
+    {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
+    {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
+    { 0 }
+};
+
+static struct {
+    bool quiet;
+    bool verbose;
+    bool dimacs;
+    bool lad;
+    bool connected;
+    bool directed;
+    bool edge_labelled;
+    bool vertex_labelled;
+    bool big_first;
+    Heuristic heuristic;
+    char *filename1;
+    char *filename2;
+    int timeout;
+    int arg_num;
+} arguments;
+
+static std::atomic<bool> abort_due_to_timeout;
+
+void set_default_arguments() {
+    arguments.quiet = false;
+    arguments.verbose = false;
[…]

### 2017-02-09 paper/paper.tex
More writing

@@
 \documentclass[letterpaper]{article}
-\usepackage[pass,showframe]{geometry}
+%\usepackage[pass,showframe]{geometry}
 
@@
 \usepackage{latexsym}
+\usepackage{algpseudocode}
 
@@
 \newcommand{\citep}[1]{\cite{#1}}
+
+\newcommand{\McSplit}{\textproc{McSplit}}
+
+\newcommand{\nmax}{n_{\max}}
 
@@
 \begin{abstract}
-    We came up with an important new algorithm for the maximum common (connected)
-    subgraph problem. This is an important problem, with at least two important
-    applications.
+    Maximium Common Subgraph (MCS) is the problem of finding a
+    graph of maximium size that is an induced subgraph of each of the two input
+    graphs.  We introduce \McSplit, a branch and bound algorithm that
+    uses a vertex-labelling scheme to compactly represent the subsets of
+    vertices that may be matched together to create a larger common subgraph.
+    The new algorithm explores the same search space as the existing CP-FC
+    algorithm, but requires only \BigO{n^2} space in total and \BigO{n} time
+    for each recursive call, greatly improving on all existing algorithms for
+    the problem.
+
+    Experimental results using two large graph databases show that \McSplit\ is
+    more than an order of magnitude faster than the leading existing algorithms
+    on instances involving undirected, unlabelled graphs, including the problem
+    variant in which the subgraph must be connected.
 \end{abstract}
@@
 Subgraph (MCS)} problem is to find a graph with as many vertices as possible
-that is isomorphic to induced subgraphs of both $G$ and $H$. Equivalently, MCS
-is the problem of finding a maximum-cardinality set of pairs $M = \{(v_1, w_1),
-\dots, (v_{|M|}, w_{|M|})\}$, with $v_i \in V(G)$ and $v_i \in V(H)$ for all $i$,
-such that for each pair $i,j$ ($1 \leq i < j
-\leq |M|$), $v_i$ and $v_j$ are adjacent in $G$ if and only if $w_i$ and $w_j$
-are adjacent in $H$.
+that is isomorphic to induced subgraphs of both $G$ and $H$.
 
-The most successful existing approaches to optimally solving MCS are
-reformulation to the maximum clique problem and constraint programming
-\cite{DBLP:conf/cp/McCreeshNPS16}. In this paper, we introduce a new branch and bound
-algorithm for MCS which does not require a domain to be explicity maintained in
-memory for each vertex in $\V(G)$. The algorithm maintains labellings of the
-unmatched vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in \V(G)$
-and $w \in \V(H)$ can be added to the matching if and only if $v$ and $w$ have the
-same label.
+MCS has numerous applications, including finding the similarity between organic
+molecules, \dots
 
 \section{Preliminaries}
+
+MCS is the problem of finding a maximum-cardinality set of pairs $M = \{(v_1,
+w_1), \dots, (v_{|M|}, w_{|M|})\}$, with $v_i \in V(G)$ and $v_i \in V(H)$ for
+all $i$, such that for each pair $i,j$ ($1 \leq i < j \leq |M|$), $v_i$ and
+$v_j$ are adjacent in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$.
 
@@
 
+\section{Prior Work}
+
+The most successful existing approaches to solving MCS are reformulation to the
+maximum clique problem on one hand, and constraint programming
+\cite{DBLP:conf/cp/McCreeshNPS16} on the other. In the clique
+\textit{association graph} encoding, a graph is created with one node for each
+pair in the set $\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and edges representing
+pairs of mappings that may be chosen together. A maximum clique in this graph
+represents a maximum common subgraph of $G$ and $H$.
+
+In the constraint programming approach, we have a variable corresponding to each
+vertex in $G$, and each variable initially has a value in its domain for each vertex
+in $H$ to which it may be mapped, plus an additional value $\bot$ signifying that
+the vertex remains unmapped. A backtracking search is performed in which values
+are assigned to variables, and the domains of unassigned variables are pruned.
+
+In this paper, we introduce a
+new branch and bound algorithm for MCS which does not require a domain to be
+explicity maintained in memory for each vertex in $\V(G)$. The algorithm
+maintains labellings of the unmatched vertices in $\V(G)$ and $\V(H)$ such that
+a pair $(v, w)$ with $v \in \V(G)$ and $w \in \V(H)$ can be added to the
+mapping if and only if $v$ and $w$ have the same label.
+
 \section{Walkthrough}
 
-Suppose we are given the graphs $G$ and $H$ in Figure~\ref{fig:alg1}.  Graphs $G$ and $H$
-have no common subgraph with five vertices. However, they do have common
-subgraphs with four vertices; the mapping $\{1a, 2f, 3d, 5b\}$ is one example.
+We will use the graphs $G$ and $H$ in Figure~\ref{fig:alg1} to demonstrate the
+algorithm.  These graphs have a maximum common subgraph with four vertices; the
+mapping $\{1a, 2f, 3d, 5b\}$ is one example.
 
@@
 
-\bigskip
-\noindent
-Phew!
-
-\section{Heuristics}
-
-We ran a small experiment comparing lots of heuristics such as \dots on random
-instances. The one where you choose a colour class with as small a max(left size,
-right size) as possible seemed to work best. ? Why?
+\paragraph{Heuristics} We ran a small experiment comparing lots of heuristics
+such as \dots on random instances. The one where you choose a colour class with
+as small a max(left size, right size) as possible seemed to work best. ? Why?
 
@@
 might be possible?
-
-\section{Future Work}
-
-Instead of branching on vertices in $\V(G)$, it would be equally valid to branch on
-a vertex in $\V(H)$ in lines 9-23 of the algorithm, since our data structure treats the
-two graphs symmetrically. More interestingly, we could choose which graph to branch on
-at each search node using some heuristic (perhaps choosing based on whether the $P$ or
-$T$ set is smaller).
-
-We could create a hybrid of Lilybank and clique (for example, using Lilybank near the
-top of the search tree, but switching to a clique encoding when fewer than some threshold
-number of vertices remain to be selected). This might give us most of the benefits of
-the clique encoding for labelled graphs, while avoiding the high memory cost and colouring
-time of encoding the full instance.
 
@@
 
+We have introduced the \McSplit\ algorithm for Max Common Subgraph, and shown
+that it explores the same search tree as CP-FC with greatly reduced effort at
+each search node.  The new algorithm is more than an order of
+magnitude faster than the previous state of the art for unlabelled, undirected
+instances.
+
+\paragraph{Future work} We could create a hybrid of Lilybank and clique (for
+example, using Lilybank near the top of the search tree, but switching to a
+clique encoding when fewer than some threshold number of vertices remain to be
+selected). This might give us most of the benefits of the clique encoding for
+labelled graphs, while avoiding the high memory cost and colouring time of
+encoding the full instance.
+
+Instead of branching on vertices in $\V(G)$, it would be equally valid to branch on
+a vertex in $\V(H)$ in lines 9-23 of the algorithm, since our data structure treats the
+two graphs symmetrically. More interestingly, we could choose which graph to branch on
+at each search node using some heuristic (perhaps choosing based on whether the $P$ or
+$T$ set is smaller).
+
 \bibliographystyle{named}

### 2017-02-09 paper/paper.tex
Small edits

@@
     more than an order of magnitude faster than the leading existing algorithms
-    on instances involving undirected, unlabelled graphs, including the problem
+    on undirected, unlabelled graphs, including the problem
     variant in which the subgraph must be connected.
@@
 \cite{DBLP:conf/cp/McCreeshNPS16} on the other. In the clique
-\textit{association graph} encoding, a graph is created with one node for each
+encoding, an \textit{association graph} is created with one node for each
 pair in the set $\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and edges representing

### 2017-02-09 code/james-cpp-parallel/mcsp.cc
Threadify incumbent and nodes

@@ -23,6 +23,32 @@ using std::vector;
 using std::cout;
 using std::endl;
 
+namespace
+{
+    struct AtomicIncumbent
+    {
+        std::atomic<unsigned> value;
+
+        AtomicIncumbent()
+        {
+            value.store(0, std::memory_order_seq_cst);
+        }
+
+        bool update(unsigned v)
+        {
+            while (true) {
+                unsigned cur_v = value.load(std::memory_order_seq_cst);
+                if (v > cur_v) {
+                    if (value.compare_exchange_strong(cur_v, v, std::memory_order_seq_cst))
+                        return true;
+                }
+                else
+                    return false;
+            }
+        }
+    };
+}
+
 static void fail(std::string msg) {
     std::cerr << msg << std::endl;
     exit(1);
@@ -162,7 +188,7 @@ static struct argp argp = { options, parse_opt, args_doc, doc };
                                      Stats
 *******************************************************************************/
 
-unsigned long long nodes{ 0 };
+std::atomic<unsigned long long> global_nodes{ 0 };
 
 /*******************************************************************************
                                  MCS functions
@@ -185,30 +211,6 @@ struct Bidomain {
             right_len (right_len),
             is_adjacent (is_adjacent) { };
 };
-
-void show(const vector<VtxPair>& current, const vector<Bidomain> &domains,
-        const vector<int>& left, const vector<int>& right)
-{
-    cout << "Nodes: " << nodes << std::endl;
-    cout << "Length of current assignment: " << current.size() << std::endl;
-    cout << "Current assignment:";
-    for (unsigned int i=0; i<current.size(); i++) {
-        cout << "  (" << current[i].v << " -> " << current[i].w << ")";
-    }
-    cout << std::endl;
-    for (unsigned int i=0; i<domains.size(); i++) {
-        struct Bidomain bd = domains[i];
-        cout << "Left  ";
-        for (int j=0; j<bd.left_len; j++)
-            cout << left[bd.l + j] << " ";
-        cout << std::endl;
-        cout << "Right  ";
-        for (int j=0; j<bd.right_len; j++)
-            cout << right[bd.r + j] << " ";
-        cout << std::endl;
-    }
-    cout << "\n" << std::endl;
-}
 
 bool check_sol(const Graph & g0, const Graph & g1 , const vector<VtxPair> & solution) {
     return true;
[…]

### 2017-02-09 code/james-cpp-parallel/mcsp.cc
Copy

@@ -416,7 +416,13 @@ void solve(const Graph & g0, const Graph & g1, AtomicIncumbent & global_incumben
         auto new_domains = filter_domains(domains, left, right, g0, g1, v, w,
                 arguments.directed || arguments.edge_labelled);
         current.push_back(VtxPair(v, w));
-        solve(g0, g1, global_incumbent, this_thread_incumbent, current, new_domains, left, right, matching_size_goal, this_thread_nodes);
+
+        auto left_copy = left;
+        auto right_copy = right;
+        auto current_copy = current;
+
+        solve(g0, g1, global_incumbent, this_thread_incumbent, current_copy, new_domains, left_copy, right_copy, matching_size_goal, this_thread_nodes);
+
         current.pop_back();
     }
     bd.right_len++;

### 2017-02-09 code/james-cpp-parallel/mcsp.cc
Cilk version for now

@@ -12,6 +12,8 @@
 #include <thread>
 #include <condition_variable>
 #include <atomic>
+#include <list>
+#include <functional>
 
 #include <argp.h>
 #include <limits.h>
@@ -19,35 +21,12 @@
 #include <stdlib.h>
 #include <string.h>
 
+#include <cilk/cilk.h>
+#include <cilk/reducer_max.h>
+
 using std::vector;
 using std::cout;
 using std::endl;
-
-namespace
-{
-    struct AtomicIncumbent
-    {
-        std::atomic<unsigned> value;
-
-        AtomicIncumbent()
-        {
-            value.store(0, std::memory_order_seq_cst);
-        }
-
-        bool update(unsigned v)
-        {
-            while (true) {
-                unsigned cur_v = value.load(std::memory_order_seq_cst);
-                if (v > cur_v) {
-                    if (value.compare_exchange_strong(cur_v, v, std::memory_order_seq_cst))
-                        return true;
-                }
-                else
-                    return false;
-            }
-        }
-    };
-}
 
 static void fail(std::string msg) {
     std::cerr << msg << std::endl;
@@ -212,6 +191,67 @@ struct Bidomain {
             is_adjacent (is_adjacent) { };
 };
 
+namespace
+{
+    struct AtomicIncumbent
+    {
+        std::atomic<unsigned> value;
+
+        AtomicIncumbent()
+        {
+            value.store(0, std::memory_order_seq_cst);
+        }
+
+        bool update(unsigned v)
+        {
+            while (true) {
+                unsigned cur_v = value.load(std::memory_order_seq_cst);
+                if (v > cur_v) {
+                    if (value.compare_exchange_strong(cur_v, v, std::memory_order_seq_cst))
+                        return true;
+                }
+                else
+                    return false;
+            }
+        }
+    };
+
+    struct ThisThreadIncumbent
+    {
+        struct View
+        {
+            vector<VtxPair> v;
+        };
+
+        struct Monoid : cilk::monoid_base<View>
+        {
+            void identity(View * v) const
+            {
+                new (v) View();
+            };
+
+            void reduce(View * left, View * right)
+            {
+                if (right->v.size() > left->v.size())
+                    *left = *right;
+            }
+        };
+
+        cilk::reducer<Monoid> impl;
+
+        void update(const vector<VtxPair> & v)
+        {
[…]

### 2017-02-09 code/james-cpp-cilk/mcsp.cc
Switch to reducer

@@ -23,6 +23,7 @@
 
 #include <cilk/cilk.h>
 #include <cilk/reducer_max.h>
+#include <cilk/reducer_opadd.h>
 
 using std::vector;
 using std::cout;
@@ -167,7 +168,7 @@ static struct argp argp = { options, parse_opt, args_doc, doc };
                                      Stats
 *******************************************************************************/
 
-std::atomic<unsigned long long> global_nodes{ 0 };
+unsigned long long global_nodes{ 0 };
 
 /*******************************************************************************
                                  MCS functions
@@ -415,12 +416,12 @@ void solve(const Graph & g0, const Graph & g1, AtomicIncumbent & global_incumben
         ThisThreadIncumbent & this_thread_incumbent,
         vector<VtxPair> & current, vector<Bidomain> & domains,
         vector<int> & left, vector<int> & right, unsigned int matching_size_goal,
-        unsigned long long & this_thread_nodes)
+        cilk::reducer<cilk::op_add<unsigned long long> > & this_thread_nodes)
 {
     if (abort_due_to_timeout)
         return;
 
-    this_thread_nodes++;
+    *this_thread_nodes += 1;
 
     this_thread_incumbent.update(current);
     global_incumbent.update(current.size());
@@ -517,21 +518,21 @@ vector<VtxPair> mcs(const Graph & g0, const Graph & g1) {
             auto domains_copy = domains;
             vector<VtxPair> current;
             ThisThreadIncumbent this_thread_incumbent;
-            unsigned long long this_thread_nodes = 0;
+            cilk::reducer<cilk::op_add<unsigned long long> > this_thread_nodes{ 0 };
             solve(g0, g1, global_incumbent, this_thread_incumbent, current, domains_copy, left_copy, right_copy, goal, this_thread_nodes);
             incumbent = this_thread_incumbent.get_value();
-            global_nodes += this_thread_nodes;
+            global_nodes += this_thread_nodes.get_value();
             if (global_incumbent.value == goal) break;
             if (!arguments.quiet) cout << "Upper bound: " << goal-1 << std::endl;
         }
 
     } else {
         vector<VtxPair> current;
-        unsigned long long this_thread_nodes = 0;
+        cilk::reducer<cilk::op_add<unsigned long long> > this_thread_nodes{ 0 };
         ThisThreadIncumbent this_thread_incumbent;
[…]

### 2017-02-09 code/james-cpp-parallel/mcsp.cc
Super hacky parallel version

@@ -1,0 +1,765 @@
+#include "graph.hh"
+
+#include <algorithm>
+#include <numeric>
+#include <chrono>
+#include <iostream>
+#include <set>
+#include <string>
+#include <utility>
+#include <vector>
+#include <mutex>
+#include <thread>
+#include <condition_variable>
+#include <atomic>
+#include <list>
+#include <functional>
+#include <map>
+#include <cassert>
+
+#include <argp.h>
+#include <limits.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+using std::vector;
+using std::cout;
+using std::endl;
+
+static void fail(std::string msg) {
+    std::cerr << msg << std::endl;
+    exit(1);
+}
+
+enum Heuristic { min_max, min_product };
+
+/*******************************************************************************
+                             Command-line arguments
+*******************************************************************************/
+
+static char doc[] = "Find a maximum clique in a graph in DIMACS format\vHEURISTIC can be min_max or min_product";
+static char args_doc[] = "HEURISTIC FILENAME1 FILENAME2";
+static struct argp_option options[] = {
+    {"quiet", 'q', 0, 0, "Quiet output"},
+    {"verbose", 'v', 0, 0, "Verbose output"},
+    {"dimacs", 'd', 0, 0, "Read DIMACS format"},
+    {"lad", 'l', 0, 0, "Read LAD format"},
+    {"connected", 'c', 0, 0, "Solve max common CONNECTED subgraph problem"},
+    {"directed", 'i', 0, 0, "Use directed graphs"},
+    {"labelled", 'a', 0, 0, "Use edge and vertex labels"},
+    {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
+    {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
+    {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
+    {"threads", 'T', "threads", 0, "Specify how many threads to use"},
+    { 0 }
+};
+
+static struct {
+    bool quiet;
+    bool verbose;
+    bool dimacs;
+    bool lad;
+    bool connected;
+    bool directed;
+    bool edge_labelled;
+    bool vertex_labelled;
+    bool big_first;
+    Heuristic heuristic;
+    char *filename1;
+    char *filename2;
+    int timeout;
+    int threads;
+    int arg_num;
+} arguments;
+
[…]

### 2017-02-09 code/james-cpp-parallel/mcsp.cc
Maybe betterer parallel code

@@ -221,9 +221,9 @@ namespace
         }
     };
 
-    using ThisThreadIncumbent = vector<VtxPair>;
+    using PerThreadIncumbents = std::map<std::thread::id, vector<VtxPair> >;
 
-    const constexpr int split_levels = 8;
+    const constexpr int split_levels = 3;
 
     struct Position
     {
@@ -267,11 +267,14 @@ namespace
 
         std::vector<std::thread> threads;
 
+        std::list<std::chrono::milliseconds> times;
+
         HelpMe(int n_threads) :
             finish(false)
         {
             for (int t = 0 ; t < n_threads ; ++t)
                 threads.emplace_back([this, n_threads, t] {
+                        std::chrono::milliseconds total_work_time = std::chrono::milliseconds::zero();
                         while (! finish.load()) {
                             std::unique_lock<std::mutex> guard(mutex);
                             bool did_something = false;
@@ -281,7 +284,12 @@ namespace
                                     ++task->second.pending;
                                     guard.unlock();
 
+                                    auto start_work_time = std::chrono::steady_clock::now(); // local start time
+
                                     (*f)();
+
+                                    auto work_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::steady_clock::now() - start_work_time);
+                                    total_work_time += work_time;
 
                                     guard.lock();
                                     task->second.func = nullptr;
@@ -296,6 +304,9 @@ namespace
                             if (! did_something)
                                 cv.wait(guard);
                         }
+
+                        std::unique_lock<std::mutex> guard(mutex);
+                        times.push_back(total_work_time);
                         });
         }
 
@@ -311,6 +322,14 @@ namespace
                 t.join();
 
             threads.clear();
+
+            if (! times.empty()) {
+                cout << "Worker thread times (ms):  ";
+                for (auto & t : times)
+                    cout << " " << t.count();
+                cout << endl;
+                times.clear();
+            }
         }
 
         ~HelpMe()
@@ -502,8 +521,64 @@ void remove_bidomain(vector<Bidomain>& domains, int idx) {
     domains.pop_back();
 }
 
[…]

### 2017-02-09 code/james-cpp-parallel/mcsp.cc
Deadlock

@@ -301,7 +301,7 @@ namespace
                                 }
                             }
 
-                            if (! did_something)
+                            if ((! did_something) && (! finish.load()))
                                 cv.wait(guard);
                         }
 

### 2017-02-09 paper/paper.tex
Added another future work item
Changed some figures of algorithm walkthrough to tables

@@
 \DeclareMathOperator{\E}{E}
+\DeclareMathOperator{\N}{N}
 
@@
 
-Given graphs $G=(\V(G), \E(G))$ and $H=(\V(H), \E(H))$, the \emph{Maximum Common
-Subgraph (MCS)} problem is to find a graph with as many vertices as possible
-that is isomorphic to induced subgraphs of both $G$ and $H$.
+Given graphs $G$ and $H$, the \emph{Maximum Common Subgraph (MCS)} problem is
+to find a graph with as many vertices as possible that is isomorphic to induced
+subgraphs of both $G$ and $H$.
 
@@
 \centering
-    \examplexG{}{}{}{}{}
-    \examplexH{}{}{}{}{}{}
+    \includegraphics{graph_G}
+    \includegraphics{graph_H}
+%    \examplexG{}{}{}{}{}
+%    \examplexH{}{}{}{}{}{}
 \caption{Graphs $G$ and $H$}
@@
 \centering
-    \examplexG{$(1a)$}{0}{0}{1}{1}
-    \examplexH{$(1a)$}{1}{1}{0}{1}{0}
-\caption{Step 2}
+\begin{minipage}[t]{0.15\linewidth}
+    Mapping
+
+    \bigskip
+
+    $\{1a\}$
+\end{minipage}
+\quad
+\begin{minipage}[t]{0.3\linewidth}
+    Labelling of $G$
+    \begin{tabular}[t]{cc}
+    \hline
+        Vertex & Label\\
+    \hline
+        $2$ & 0 \\
+        $3$ & 0 \\
+        $4$ & 1 \\
+        $5$ & 1 \\
+    \hline
+    \end{tabular}
+\end{minipage}
+\quad
+\begin{minipage}[t]{0.3\linewidth}
+    Labelling of $H$
+    \begin{tabular}[t]{cc}
+    \hline
+        Vertex & Label\\
+    \hline
+        $b$ & 1 \\
+        $c$ & 1 \\
+        $d$ & 0 \\
+        $e$ & 1 \\
+        $f$ & 0 \\
+    \hline
+    \end{tabular}
+\end{minipage}
+    \caption{Labels on vertices after mapping $1$ to $a$}
 \label{fig:alg2}
@@
 \centering
-    \examplexG{$(1a)$}{$(2d)$}{01}{10}{11}
-    \examplexH{$(1a)$}{11}{11}{$(2d)$}{10}{01}
+\begin{minipage}[t]{0.15\linewidth}
+    Mapping
+
+    \bigskip
+
+    $\{1a, 2d\}$
+\end{minipage}
+\quad
+\begin{minipage}[t]{0.3\linewidth}
+    Labelling of $G$
+    \begin{tabular}[t]{cc}
+    \hline
+        Vertex & Label\\
+    \hline
+        $3$ & 01 \\
+        $4$ & 10 \\
+        $5$ & 11 \\
+    \hline
+    \end{tabular}
+\end{minipage}
+\quad
+\begin{minipage}[t]{0.3\linewidth}
+    Labelling of $H$
+    \begin{tabular}[t]{cc}
+    \hline
+        Vertex & Label\\
+    \hline
+        $b$ & 11 \\
+        $c$ & 11 \\
+        $e$ & 10 \\
+        $f$ & 01 \\
+    \hline
+    \end{tabular}
+\end{minipage}
 \caption{Step 3}
@@
 
+Parallel versions: bitset and multithreaded
+
 \bibliographystyle{named}

### 2017-02-10 paper/paper.tex
Worked on algorithm description

@@
 \DeclareMathOperator{\N}{N}
+\DeclareMathOperator{\vtxlabel}{label}
 
@@
 
-\newcommand{\examplexG}[5] {
-    \begin{minipage}{.2\textwidth}
-    \tikz {
-        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1cm,
-                clockwise=5] {
-                    1[label=90:#1],2[label=0:#2],3[label=0:#3],4[label=180:#4],5[label=180:#5];
-            1--4; 1--5; 2--3; 2--5; 3--5;
-        };
-    }
-    \end{minipage}
-}
-\newcommand{\examplexH}[6] {
-    \begin{minipage}{.2\textwidth}
-    \tikz {
-        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1.1cm,
-                clockwise=6, phase=60] {
-                    a[label=0:#1],b[label=0:#2],c[label=0:#3],d[label=180:#4],e[label=180:#5],f[label=180:#6];
-            a--b; a--c; a--e; b--d; b--f; c--d; c--e; c--f; d--f; e--f;
-        };
-    }
-    \end{minipage}
-}
+%\newcommand{\examplexG}[5] {
+%    \begin{minipage}{.2\textwidth}
+%    \tikz {
+%        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1cm,
+%                clockwise=5] {
+%                    1[label=90:#1],2[label=0:#2],3[label=0:#3],4[label=180:#4],5[label=180:#5];
+%            1--4; 1--5; 2--3; 2--5; 3--5;
+%        };
+%    }
+%    \end{minipage}
+%}
+%\newcommand{\examplexH}[6] {
+%    \begin{minipage}{.2\textwidth}
+%    \tikz {
+%        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1.1cm,
+%                clockwise=6, phase=60] {
+%                    a[label=0:#1],b[label=0:#2],c[label=0:#3],d[label=180:#4],e[label=180:#5],f[label=180:#6];
+%            a--b; a--c; a--e; b--d; b--f; c--d; c--e; c--f; d--f; e--f;
+%        };
+%    }
+%    \end{minipage}
+%}
 
@@
 
-MCS is the problem of finding a maximum-cardinality set of pairs $M = \{(v_1,
-w_1), \dots, (v_{|M|}, w_{|M|})\}$, with $v_i \in V(G)$ and $v_i \in V(H)$ for
-all $i$, such that for each pair $i,j$ ($1 \leq i < j \leq |M|$), $v_i$ and
-$v_j$ are adjacent in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$.
+We assume that $G$ and $H$ are unlabelled, undirected graphs without loops
+(Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
 
-Let $N(v, G)$ denote the neighbourhood of a vertex $v$ in a graph $G$.  We
-extend this notation to denote the set of vertices that are adjacent to all
-vertices in a set $U$: $N(U, G) = \bigcap_{v \in U} N_G(v)$.
+The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.
+The set of vertices adjacent to vertex $v$ in graph $G$ is called the
+\emph{neighbourhood} of $v$, and denoted $\N(v, G)$.
+
+We use $n_G$ and $n_H$ to denote $|\V(G)|$ and $|\V(H)|$. We will occasionally use
+and $n$ in complexity results to denote $n_G + n_H$.
 
@@
 maximum clique problem on one hand, and constraint programming
-\cite{DBLP:conf/cp/McCreeshNPS16} on the other. In the clique
-encoding, an \textit{association graph} is created with one node for each
-pair in the set $\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and edges representing
-pairs of mappings that may be chosen together. A maximum clique in this graph
-represents a maximum common subgraph of $G$ and $H$.
+\cite{DBLP:conf/cp/McCreeshNPS16} on the other. In the clique encoding, an
+\textit{association graph} is created with one node for each pair in the set
+$\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and and edges for each pair of
+mappings that may be chosen together. A maximum clique in this graph represents
+a maximum common subgraph of $G$ and $H$.
 
-In the constraint programming approach, we have a variable corresponding to each
-vertex in $G$, and each variable initially has a value in its domain for each vertex
-in $H$ to which it may be mapped, plus an additional value $\bot$ signifying that
-the vertex remains unmapped. A backtracking search is performed in which values
-are assigned to variables, and the domains of unassigned variables are pruned.
+In the constraint programming approach, we have a variable corresponding to
+each vertex in $G$, and each variable initially has one value in its domain for
+each vertex in $H$ to which it may be mapped, plus an additional value $\bot$
+signifying that the vertex remains unmapped. A backtracking search is performed
+in which values are assigned to variables, and propagation rules are used to
+remove values from the domains of unassigned variables.
 
-In this paper, we introduce a
-new branch and bound algorithm for MCS which does not require a domain to be
-explicity maintained in memory for each vertex in $\V(G)$. The algorithm
-maintains labellings of the unmatched vertices in $\V(G)$ and $\V(H)$ such that
-a pair $(v, w)$ with $v \in \V(G)$ and $w \in \V(H)$ can be added to the
-mapping if and only if $v$ and $w$ have the same label.
+\section{The \McSplit\ algorithm}
 
-\section{Walkthrough}
+We introduce a branch and bound algorithm, \McSplit, which uses a new
+vertex-labelling scheme to avoid explicitly maintaining the set of feasible
+mappings for each vertex in $\V(G)$. As we will explore in
+Section~\ref{sec:comparison}, the algorithm may be viewed as a version of the
+constraint programming forward-checking (CP-FC) algorithm using a compact data
+structure that exploits the nature of the MCS problem.  It is sufficiently
+different, however, that it is helpful to describe the algorithm without
+reference to previous approaches.
+
+%At each node of the search tree, the algorithm
+%labels vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in
+%\V(G)$ and $w \in \V(H)$ can be added to the mapping if and only if $v$ and $w$
+%have the same label.
+
+The \McSplit\ algorithm finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
+(v_{|M^*|}, w_{|M^*|})\}$, where the $v_i$ are distinct members of $\V(G)$ and the
+$w_i$ are distinct members of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
+in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. It is clear that,
+given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{|M^*|}\}$
+and the subgraph of $H$ induced by $\{w_i, \dots, w_{|M^*|}\}$ are a maximum
+common subgraph.
 
@@
 algorithm.  These graphs have a maximum common subgraph with four vertices; the
-mapping $\{1a, 2f, 3d, 5b\}$ is one example.
+mapping $\{1a, 2f, 3d, 5b\}$ (where vertex $1$ is assigned to vertex $a$ and so
+on) is one example.
 
@@
 The algorithm proceeds by building up a mapping $M$, starting with the empty
-mapping $\{\}$. Begin by (arbitrarily) mapping vertex $1$ in $G$ to vertex $a$
-in $H$, giving $M=\{1a\}$.  Label each unmatched vertex according to whether it
-is adjacent to $1$ (in $G$) or $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.
-Adjacent vertices have label 1; non-adjacent vertices have label 0.  It is
-straightforward to see that we can extend $M$ with a vertex mapping $vw$, with
-$v \in \V(G)$ and $w \in \V(H)$, if and only if $v$ and $w$ have the same label.  This
-property is the algorithm's key invariant.
+mapping $\{\}$. Select a vertex in $\V(G)$ as the first vertex to
+be mapped; we will arbitrarily choose vertex $1$. Each of the vertices in $\V(H)$
+to which $1$ may be mapped will be tried in turn, and finally the possibility where
+$1$ remains unmatched will be tried.
+
+We will begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Label
+each unmatched vertex according to whether it is adjacent to $1$ (in $G$) or
+$a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent vertices have label
+1; non-adjacent vertices have label 0.  It is straightforward to see that we
+can extend $M$ with a vertex mapping $vw$, with $v \in \V(G)$ and $w \in
+\V(H)$, if and only if $v$ and $w$ have the same label.  This property that
+two vertices may be mapped together if and only if they share a label is the
+algorithm's main invariant.
 
@@
 Observe that our invariant is maintained: we can extend $M$ by a vertex pairing if
-and only if the two vertices have the same colour.
+and only if the two vertices have the same label. Note also that each label is prefixed
+by the label at the previous level of the search tree (Figure~\ref{fig:alg2}).
 
@@
 \end{minipage}
-\caption{Step 3}
+\caption{After mapping $2$ to $d$}
 \label{fig:alg3}
@@
 
-The algorithm proceeds by backtracking search. When we have explored the full
-search space of matchings containing $\{1a, 2d\}$, we try reassigning $2$ to
-$f$.  Since $d$ and $f$ are the only vertices to which $2$ can be matched given
-the decision to match $1$ to $a$, we lastly explore the possibility that $2$ is
-left unmatched, by giving $2$ the label $\bot$ and selecting another vertex in
-$\V(G)$ to assign.
+The algorithm backtracks when the incumbent best mapping is at least as large
+as a calculated bound given $M$ and the current labelling. To demonstrate how
+this bound is calculated, we consider the situation one level deeper in the search
+tree shown in Figure~\ref{fig:alg4}. Three vertex labels are used: 100, 101, and 111.
+The first two of these only appear in one graph, and therefore vertices with these
+labels cannot be added to the mapping. The final label, 111, appears once in $G$ and
+twice in $H$, and therefore at most one pair with this label can be added to $M$.
+Thus, the upper bound on matching size is $4$. In general, for each label $l$
+we add to the bound $\min(|\{ v \in \V(G) : \vtxlabel(v)=l\}|,
+|\{ v \in \V(H) : \vtxlabel(v)=l \}|)$.
 
-\section{The Lilybank Algorithm}
+\begin{figure}[ht]
+\centering
+\begin{minipage}[t]{0.2\linewidth}
+    Mapping
+
+    \bigskip
+
+    $\{1a, 2d, 3f\}$
+\end{minipage}
+\quad
+\begin{minipage}[t]{0.3\linewidth}
+    Labelling of $G$
+    \begin{tabular}[t]{cc}
+    \hline
+        Vertex & Label\\
+    \hline
+        $4$ & 100 \\
+        $5$ & 111 \\
+    \hline
+    \end{tabular}
+\end{minipage}
+\quad
+\begin{minipage}[t]{0.3\linewidth}
+    Labelling of $H$
+    \begin{tabular}[t]{cc}
+    \hline
+        Vertex & Label\\
+    \hline
+        $b$ & 111 \\
+        $c$ & 111 \\
+        $e$ & 101 \\
+    \hline
+    \end{tabular}
+\end{minipage}
+\caption{After mapping $3$ to $f$}
+\label{fig:alg4}
+\end{figure}
+
+%When we have explored the full
+%search space of matchings containing $\{1a, 2d\}$, we try reassigning $2$ to
+%$f$.  Since $d$ and $f$ are the only vertices to which $2$ can be matched given
+%the decision to match $1$ to $a$, we lastly explore the possibility that $2$ is
+%left unmatched, by giving $2$ the label $\bot$ and selecting another vertex in
+%$\V(G)$ to assign.
+
+\paragraph{Efficient representation of label classes} The algorithm as
+described so far requires $\BigO{n^2}$ space at each node of the search tree,
+since labels may be up to $n_G$ in length. We can reduce this to $\BigO{n}$
+by storing a pair $\langle P,T \rangle$ for each label $l$ that is used, where $P$
+is the set of vertices in $\V(G)$ labelled $l$, and $T$ is the set of vertices
+in $\V(H)$ labelled $l$. Since there are $n_G + n_H$ vertices, at most this number
+of label-classes can exist at once, and the total length of the $P$ and $T$ sets
+is at most $n_G + n_H$. We can gain further efficiency by discarding during search
+any label-class that has an empty $P$ or $T$ set.
 
@@
 \DontPrintSemicolon
-\nl $expand(\mathit{future},matching)$ \;
+\nl $\FuncSty{expand}(\mathit{future},matching)$ \;
 \nl \Begin{
@@
        }
-\nl   $expand(future^{'},matching)$ \;
+\nl   $\FuncSty{expand}(future^{'},matching)$ \;
 \nl   $matching \gets matching \setminus \{(v,w)\}$ \;
@@
 \nl \lIf {$P = \emptyset$} {$future \gets future \setminus \{\langle P,T \rangle \}$}
-\nl $expand(future,matching)$ \;
+\nl $\FuncSty{expand}(future,matching)$ \;
 }
 \;
-\nl $Split(G,H)$ \;
+\nl $\FuncSty{McSplit}(G,H)$ \;
 \nl \Begin{
 \nl $incumbent \gets \emptyset$ \;
-\nl $expand(\{\langle V(G),V(H) \rangle \},\emptyset)$ \;
+\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \;
 \nl return $|incubent|$ \;
 }
-\caption{Lilybank splitting algorithm}
+\caption{\McSplit\ algorithm for Maximum Common Subgraph}
 \label{jtAlg}
@@
 
-Now, let's try and explain this. First, assume we have two graph G and H and there are no more vertices in G than in H (think of G as the pattern graph and H as the target).
-future is a list of pairs, where the first element of a pair is a set of pattern vertices P and the second is a set of target vertices T. It is assumed that each pattern vertex in P 
-might be matched with any (all) of the target vertices in T, or be unmatched.
-Therefore we start our search via the function LilybankSplit (line 26).
-This returns an integer, the size of the largest matching.
-In line 29 a call is made to expand, where we have as arguments a list of pairs (called future, line 1) and a set of pairs of the form (v,w) where pattern vertex v is matched to target vertex w.
+The full algorithm is shown as Algorithm~\ref{jtAlg}, with input graphs $G$ and $H$.
+The main recursive procedure, future is a list of pairs, where the first element of a pair
+is a set of pattern vertices P and the second is a set of target vertices T. It
+is assumed that each pattern vertex in P might be matched with any (all) of the
+target vertices in T, or be unmatched. Therefore we start our search via the
+function LilybankSplit (line 26). This returns an integer, the size of the
+largest matching.  In line 29 a call is made to expand, where we have as
+arguments a list of pairs (called future, line 1) and a set of pairs of the
+form (v,w) where pattern vertex v is matched to target vertex w.
 
-We now focus on expand (line (1)). This is initially called (line 29) with the future as a list containing the single pair $\langle V(G),V(H) \rangle$. Lines 4 to 7 define the conditions 
-where we return, i.e. when the incumbent cannot be improved. In line 3, a new incumbent might be saved. Lines 8 to 24 perform the search.
+We now focus on expand (line (1)). This is initially called (line 29) with the
+future as a list containing the single pair $\langle V(G),V(H) \rangle$. Lines
+4 to 7 define the conditions where we return, i.e. when the incumbent cannot be
+improved. In line 3, a new incumbent might be saved. Lines 8 to 24 perform the
+search.
 
-A pair is selected from the future (line 8) using some heuristic.
-Note that this pair is not removed from the future, and that P and T can be considered as pointers to sets.
-A pattern vertex is then selected, possibly using a heuristic (line 9).
-We now iterate over all possible target vertices in T (lines 10 to 21).
-A target vertex w is selected and a matching produced (line 11).
-A new future is created (using {\bf Let} to signify a new variable, line 12).
-Every pair in the future can now be split (lines 13 to 19) into pattern vertices adjacent to the selected pattern vertex v and target vertices adjacent to the selected target vertex w (lines 14 to 15) and this is added to the new future if both sets are not null (line 16) i.e.
-there is a potential for more matchings.
-This is then repeated symmetrically for non-adjacency (lines 17 to 19) using the compliments of the graphs.
-A recursive call is then made using the new future and the increased matching (line 20).
-On return from that call we remove the matching (v,w) (line 21).
-Having explored all matchings with v in its target set we now consider what happens if the selected pattern vertex v is not matched (line 22 to 24), where that set of matchings does not contain a pair (v,*).
+A pair is selected from the future (line 8) using some heuristic. Note that
+this pair is not removed from the future, and that P and T can be considered as
+pointers to sets. A pattern vertex is then selected, possibly using a heuristic
+(line 9). We now iterate over all possible target vertices in T (lines 10 to
+21). A target vertex w is selected and a matching produced (line 11). A new
+future is created (using {\bf Let} to signify a new variable, line 12). Every
+pair in the future can now be split (lines 13 to 19) into pattern vertices
+adjacent to the selected pattern vertex v and target vertices adjacent to the
+selected target vertex w (lines 14 to 15) and this is added to the new future
+if both sets are not null (line 16) i.e. there is a potential for more
+matchings. This is then repeated symmetrically for non-adjacency (lines 17 to
+19) using the compliments of the graphs.  A recursive call is then made using
+the new future and the increased matching (line 20). On return from that call
+we remove the matching (v,w) (line 21). Having explored all matchings with v in
+its target set we now consider what happens if the selected pattern vertex v is
+not matched (line 22 to 24), where that set of matchings does not contain a
+pair (v,*).
 
 \paragraph{Heuristics} We ran a small experiment comparing lots of heuristics
-such as \dots on random instances. The one where you choose a colour class with
+such as \dots on random instances. The one where you choose a label class with
 as small a max(left size, right size) as possible seemed to work best. ? Why?
 
-\section{Comparison with Existing Algorithms}
+\section{Comparison with Existing Algorithms}\label{sec:comparison}
 
@@
 
-\section{Extensions}
+? Mention partition backtracking
+
+\section{Extensions}\label{sec:extensions}
 
@@
 If we have vertex labels, we replace ${\langle V(G),V(H) \rangle}$ in line 28 of
-the algorithm with a set containing a colour class for each label that appears in
+the algorithm with a set containing a label class for each label that appears in
 both $\V(G)$ and $\V(H)$. In the remainder of this section, \textit{labels} refers to
@@
 ascending order of the label on the edge from $v$ or $w$ to each vertex. We can
-then create the colour classes corresponding to each edge label by
+then create the label classes corresponding to each edge label by
 simulataneously traversing $P'$ and $T'$ from left to right.
@@
 
+Application to induced subgraph isomorphism (with almost no changes). Applications
+of a similar technique to other problems???
+
 \bibliographystyle{named}

### 2017-02-10 paper/paper.tex
More algorithm-writing

@@
 
+\newcommand{\AlgVar}[1]{\mathit{#1}}
+
 \newcommand{\McSplit}{\textproc{McSplit}}
@@
 \newcommand{\nmax}{n_{\max}}
+
+\newcommand{\lineref}[1]{line~\ref{#1}}
+\newcommand{\twolinesref}[2]{lines~\ref{#1} and~\ref{#2}}
+\newcommand{\Lineref}[1]{Line~\ref{#1}}
+\newcommand{\Twolinesref}[2]{Lines~\ref{#1} and~\ref{#2}}
 
@@
 \DontPrintSemicolon
-\nl $\FuncSty{expand}(\mathit{future},matching)$ \;
+\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
 \nl \Begin{
-\nl \lIf {$future = \emptyset$ \bf{and} $|matching|  > |incumbent|$}{$save(matching)$}
-\nl \lIf {$future = \emptyset$}{return}
-\nl $bestFuture \gets 0$ \;
-\nl \lFor {$\langle P,T \rangle \in future$}{$bestFuture \gets bestFuture + min(|P|,|T|)$}
-\nl \lIf {$|matching|  + bestFuture \leq |incumbent|$}{return}
+%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
+\nl \lIf {$|M| > |\AlgVar{incumbent}|$}
+    {$save(matching)$} \label{StoreIncumbent}
+%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
+\nl $bestFuture \gets 0$ \label{SetBestFutureToZero} \;
+\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}
+    {$\AlgVar{bestFuture} \gets \AlgVar{bestFuture} + \min(|P|,|T|)$}
+\nl \lIf {$|M|  + \AlgVar{bestFuture} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
 \bigskip
-\nl $\langle P,T \rangle \gets select(future)$ \;
-\nl $v \gets select(P)$ \;
-\nl \For {$w \in T$}{
-\nl    $matching \gets matching + (v,w)$\;
-\nl    \bf{Let} $future^{'} \gets \emptyset$ \;
-\nl    \For {$\langle P^{'},T'^{'}\rangle \in future$}{
-\nl        \bf{Let} $P^{''} \gets P^{'} \cap N(v,G) \setminus \{v\}$ \;
-\nl        \bf{Let} $T^{''} \gets T^{'} \cap N(w,H) \setminus \{w\}$ \;
-\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
-\nl        \bf{Let} $P^{''} \gets P^{'} \cap N(v,\overline{G}) \setminus \{v\}$ \;
-\nl        \bf{Let} $T^{''} \gets T^{'} \cap N(w,\overline{H}) \setminus \{w\}$ \;
-\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
+\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
+\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
+\nl \For {$w \in T$ \label{WLoop}} {
+\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
+\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
+\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
+\nl        \bf{Let} $P'' \gets P' \cap N(v,G) \setminus \{v\}$ \label{NewPWithEdge} \;
+\nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
+\nl        \lIf {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
+        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
+\nl        \bf{Let} $P'' \gets P' \cap N(v,\overline{G}) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
+\nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
+\nl        \lIf {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
+        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
        }
-\nl   $\FuncSty{expand}(future^{'},matching)$ \;
-\nl   $matching \gets matching \setminus \{(v,w)\}$ \;
+\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
+\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
   }
-\nl $P \gets P \setminus \{v\}$\;
-\nl \lIf {$P = \emptyset$} {$future \gets future \setminus \{\langle P,T \rangle \}$}
-\nl $\FuncSty{expand}(future,matching)$ \;
+\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
+\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\} \cup \{\langle P', T \rangle\}$\;
+\nl \lIf {$P' = \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P',T \rangle \}$}
+\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
 }
 \;
-\nl $\FuncSty{McSplit}(G,H)$ \;
+\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
 \nl \Begin{
-\nl $incumbent \gets \emptyset$ \;
-\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \;
-\nl return $|incubent|$ \;
+\nl $\AlgVar{incumbent} \gets \emptyset$ \;
+\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
+\nl return $|\AlgVar{incubent}|$ \;
 }
@@
 
-The full algorithm is shown as Algorithm~\ref{jtAlg}, with input graphs $G$ and $H$.
-The main recursive procedure, future is a list of pairs, where the first element of a pair
-is a set of pattern vertices P and the second is a set of target vertices T. It
-is assumed that each pattern vertex in P might be matched with any (all) of the
-target vertices in T, or be unmatched. Therefore we start our search via the
-function LilybankSplit (line 26). This returns an integer, the size of the
-largest matching.  In line 29 a call is made to expand, where we have as
-arguments a list of pairs (called future, line 1) and a set of pairs of the
-form (v,w) where pattern vertex v is matched to target vertex w.
+\paragraph{Full algorithm} The full algorithm is shown as
+Algorithm~\ref{jtAlg}.  The main recursive procedure, $\FuncSty{expand}$, has
+two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
+each represented as a $\langle P, T \rangle$ pair as described above.  The
+parameter $M$ is the current mapping of vertices.  On each call to
+$\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
+if and only if $v$ and $w$ belong to the same label class in $M$.
 
-We now focus on expand (line (1)). This is initially called (line 29) with the
-future as a list containing the single pair $\langle V(G),V(H) \rangle$. Lines
-4 to 7 define the conditions where we return, i.e. when the incumbent cannot be
-improved. In line 3, a new incumbent might be saved. Lines 8 to 24 perform the
-search.
+\Lineref{StoreIncumbent} stores the current mapping $M$ if it is large enough
+to unseat the incumbent.  \Twolinesref{SetBestFutureToZero}{PruneSearch} prune
+the search when a calculated upper bound ($\AlgVar{bestFuture}$) is not larger
+than the incumbent.
 
-A pair is selected from the future (line 8) using some heuristic. Note that
-this pair is not removed from the future, and that P and T can be considered as
-pointers to sets. A pattern vertex is then selected, possibly using a heuristic
-(line 9). We now iterate over all possible target vertices in T (lines 10 to
-21). A target vertex w is selected and a matching produced (line 11). A new
-future is created (using {\bf Let} to signify a new variable, line 12). Every
-pair in the future can now be split (lines 13 to 19) into pattern vertices
-adjacent to the selected pattern vertex v and target vertices adjacent to the
-selected target vertex w (lines 14 to 15) and this is added to the new future
-if both sets are not null (line 16) i.e. there is a potential for more
-matchings. This is then repeated symmetrically for non-adjacency (lines 17 to
-19) using the compliments of the graphs.  A recursive call is then made using
-the new future and the increased matching (line 20). On return from that call
-we remove the matching (v,w) (line 21). Having explored all matchings with v in
-its target set we now consider what happens if the selected pattern vertex v is
-not matched (line 22 to 24), where that set of matchings does not contain a
-pair (v,*).
+The remainder of $\FuncSty{expand}$ performs the search.  A label-class
+$\langle P, T \rangle$ is selected from $\AlgVar{future}$
+(\lineref{SelectClass}) using some heuristic.  From this class, a pattern
+vertex $v$ is selected using a heuristic (\lineref{SelectVertex}). We now
+iterate over all possible vertices in $T$ (\twolinesref{WLoop}{ShrinkM}). A
+vertex $w \in \V(H)$ is selected mapping $M$ is extended (\lineref{GrowM}). A
+new set of label-classes, $\AlgVar{future'}$, is created (\lineref{NewFuture}).
+Every label-class in $\AlgVar{future}$ can now be split
+(\twolinesref{InnerLoop}{InnerLoopEnd}) into two new classes. The first of
+these classes (\twolinesref{NewPWithEdge}{AddToFutureWithEdge}) contains
+pattern vertices adjacent to $v$ and target vertices adjacent to $w$.  This is
+added to $\AlgVar{future'}$ if both sets contain at least one vertex.  This is
+then repeated symmetrically for non-adjacency
+(\twolinesref{NewPWithoutEdge}{InnerLoopEnd}) using the compliments of the
+graphs.  A recursive call is then made (\lineref{ExpandWithV}), on return from
+which we remove the mapping $(v,w)$.  Having explored all possible mappings of
+$v$ its target set we now consider what happens if $v$ is not matched
+(\twolinesref{RemoveV}{ExpandWithoutV}). Note that removing $v$ from its label
+class is equivalent in effect to assigning it the label $\bot$, which is never
+taken by any vertex in $\V(H)$.
+
+We start our search at the function $\FuncSty{McSplit}$ (\lineref{McSplitFun}),
+with graphs $G$ and $H$ as inputs.  This function returns the size of the
+largest mapping.  In \lineref{FirstExpandCall} a call is made to
+$\FuncSty{expand}$; at this point we have a single label-class containing all
+vertices, and the mapping $M$ is empty.
 

### 2017-02-10 paper/paper.tex
Removed a line
Corrected line-range grammar

@@
 \newcommand{\lineref}[1]{line~\ref{#1}}
-\newcommand{\twolinesref}[2]{lines~\ref{#1} and~\ref{#2}}
+\newcommand{\linerangeref}[2]{lines~\ref{#1} to~\ref{#2}}
 \newcommand{\Lineref}[1]{Line~\ref{#1}}
-\newcommand{\Twolinesref}[2]{Lines~\ref{#1} and~\ref{#2}}
+\newcommand{\Linerangeref}[2]{Lines~\ref{#1} to~\ref{#2}}
 
@@
 \Lineref{StoreIncumbent} stores the current mapping $M$ if it is large enough
-to unseat the incumbent.  \Twolinesref{SetBestFutureToZero}{PruneSearch} prune
+to unseat the incumbent.  \Linerangeref{SetBestFutureToZero}{PruneSearch} prune
 the search when a calculated upper bound ($\AlgVar{bestFuture}$) is not larger
@@
 vertex $v$ is selected using a heuristic (\lineref{SelectVertex}). We now
-iterate over all possible vertices in $T$ (\twolinesref{WLoop}{ShrinkM}). A
+iterate over all possible vertices in $T$ (\linerangeref{WLoop}{ShrinkM}). A
 vertex $w \in \V(H)$ is selected mapping $M$ is extended (\lineref{GrowM}). A
@@
 Every label-class in $\AlgVar{future}$ can now be split
-(\twolinesref{InnerLoop}{InnerLoopEnd}) into two new classes. The first of
-these classes (\twolinesref{NewPWithEdge}{AddToFutureWithEdge}) contains
+(\linerangeref{InnerLoop}{InnerLoopEnd}) into two new classes. The first of
+these classes (\linerangeref{NewPWithEdge}{AddToFutureWithEdge}) contains
 pattern vertices adjacent to $v$ and target vertices adjacent to $w$.  This is
@@
 then repeated symmetrically for non-adjacency
-(\twolinesref{NewPWithoutEdge}{InnerLoopEnd}) using the compliments of the
+(\linerangeref{NewPWithoutEdge}{InnerLoopEnd}) using the compliments of the
 graphs.  A recursive call is then made (\lineref{ExpandWithV}), on return from
@@
 $v$ its target set we now consider what happens if $v$ is not matched
-(\twolinesref{RemoveV}{ExpandWithoutV}). Note that removing $v$ from its label
+(\linerangeref{RemoveV}{ExpandWithoutV}). Note that removing $v$ from its label
 class is equivalent in effect to assigning it the label $\bot$, which is never
@@
 \section{Comparison with Existing Algorithms}\label{sec:comparison}
-
-? Describe CP-FC and clique encodings
 

### 2017-02-10 paper/paper.tex
Editing

@@
 \usepackage{algpseudocode}
+\usepackage{amsmath}
 
@@
 
+This paper introduces a new branch and bound algorithm for MCS that improves
+state-of-the-art run times on unlabelled, undirected instances by well over an
+order of magnitude, and uses only quadratic memory in the order of the input
+graphs, allowing us to solve larger instances than previously possible.
+
 \section{Preliminaries}
@@
 \cite{DBLP:conf/cp/McCreeshNPS16} on the other. In the clique encoding, an
-\textit{association graph} is created with one node for each pair in the set
-$\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and and edges for each pair of
-mappings that may be chosen together. A maximum clique in this graph represents
+association graph is created with one node for each pair in the set $\{(u,v)
+\mid u \in \V(G), v \in \V(H) \}$ and an edge for each pair of mappings that
+may be chosen together. A maximum clique in this graph corresponds directly to
 a maximum common subgraph of $G$ and $H$.
@@
 in which values are assigned to variables, and propagation rules are used to
-remove values from the domains of unassigned variables.
+remove infeasible values from domains.
 
@@
 
-We introduce a branch and bound algorithm, \McSplit, which uses a new
-vertex-labelling scheme to avoid explicitly maintaining the set of feasible
-mappings for each vertex in $\V(G)$. As we will explore in
+This paper's main contribution is a branch and bound algorithm, \McSplit,
+which uses a new vertex-labelling scheme to avoid explicitly maintaining the
+set of feasible mappings for each vertex in $\V(G)$. As we will explore in
 Section~\ref{sec:comparison}, the algorithm may be viewed as a version of the
@@
 We will use the graphs $G$ and $H$ in Figure~\ref{fig:alg1} to demonstrate the
-algorithm.  These graphs have a maximum common subgraph with four vertices; the
-mapping $\{1a, 2f, 3d, 5b\}$ (where vertex $1$ is assigned to vertex $a$ and so
-on) is one example.
+algorithm.  These graphs have a maximum common subgraph with four vertices; one
+example is the mapping $\{1a, 2f, 3d, 5b\}$ (where vertex $1$ is assigned to
+vertex $a$, $2$ is assigned to $f$, and so on).
 
@@
 
-The algorithm proceeds by building up a mapping $M$, starting with the empty
-mapping $\{\}$. Select a vertex in $\V(G)$ as the first vertex to
-be mapped; we will arbitrarily choose vertex $1$. Each of the vertices in $\V(H)$
-to which $1$ may be mapped will be tried in turn, and finally the possibility where
-$1$ remains unmatched will be tried.
+The algorithm builds up a mapping $M$, starting with the empty mapping
+$\emptyset$. Select a vertex in $\V(G)$ as the first vertex to be mapped; in
+our example we will arbitrarily choose vertex $1$. Each of the vertices in
+$\V(H)$ to which $1$ may be mapped will be tried in turn, and finally the
+possibility where $1$ remains unmatched will be tried.
 
@@
 same label; we will choose to map vertex $2$ to vertex $d$, giving $M=\{1a,
-2d\}$ (Figure~\ref{fig:alg3}).  Each unmatched vertex $v \in \V(G)$ is labelled
+2d\}$ (Figure~\ref{fig:alg3}).  Each unmapped vertex $v \in \V(G)$ is labelled
 with a two-character bit string, indicating whether $v$ is adjacent to each of
-the two matched vertices in $\V(G)$ (vertices $1$ and $2$).  For example, vertex
+the two mapped vertices in $\V(G)$ (vertices $1$ and $2$).  For example, vertex
 $3$ is labelled $01$, indicating that it is not adjacent to $1$ but is adjacent
-to $2$.  Labels are given to vertices in $\V(H)$ in a similar fashion, showing
-whether each vertex is adjacent to $a$ and $d$.
+to $2$.  Labels are given to unmapped vertices in $\V(H)$ in a similar fashion,
+showing whether each vertex is adjacent to $a$ and $d$.
 
-Observe that our invariant is maintained: we can extend $M$ by a vertex pairing if
+Our invariant is maintained: we can extend $M$ by a vertex pairing if
 and only if the two vertices have the same label. Note also that each label is prefixed
@@
 as a calculated bound given $M$ and the current labelling. To demonstrate how
-this bound is calculated, we consider the situation one level deeper in the search
-tree shown in Figure~\ref{fig:alg4}. Three vertex labels are used: 100, 101, and 111.
-The first two of these only appear in one graph, and therefore vertices with these
-labels cannot be added to the mapping. The final label, 111, appears once in $G$ and
-twice in $H$, and therefore at most one pair with this label can be added to $M$.
-Thus, the upper bound on matching size is $4$. In general, for each label $l$
-we add to the bound $\min(|\{ v \in \V(G) : \vtxlabel(v)=l\}|,
-|\{ v \in \V(H) : \vtxlabel(v)=l \}|)$.
+this bound is calculated, we consider the situation one level deeper in the
+search tree shown in Figure~\ref{fig:alg4}.
 
@@
 
+Three vertex labels are used: 100,
+101, and 111.  The first two of these only appear in one graph, and therefore
+there is no way to add a pair of vertices with label 100 or 101 to the mapping.
+The final label, 111, appears once in $G$ and twice in $H$, and therefore at
+most one pair with this label can be added to $M$.  Thus, the upper bound on
+matching size is $|M| + 1 = 4$. The general formula for the upper bound is
+as follows, where $L$ is the set of labels used in both graphs.
+
+\begin{multline*}
+    \mathit{bound} = |M| + \sum_{l \in L} \min\big(|\{ v \in \V(G) : \vtxlabel(v)=l\}|, \\
+        |\{ v \in \V(H) : \vtxlabel(v)=l \}|\big)
+\end{multline*}
+
 %When we have explored the full
@@
 \paragraph{Efficient representation of label classes} The algorithm as
-described so far requires $\BigO{n^2}$ space at each node of the search tree,
-since labels may be up to $n_G$ in length. We can reduce this to $\BigO{n}$
-by storing a pair $\langle P,T \rangle$ for each label $l$ that is used, where $P$
-is the set of vertices in $\V(G)$ labelled $l$, and $T$ is the set of vertices
-in $\V(H)$ labelled $l$. Since there are $n_G + n_H$ vertices, at most this number
-of label-classes can exist at once, and the total length of the $P$ and $T$ sets
-is at most $n_G + n_H$. We can gain further efficiency by discarding during search
-any label-class that has an empty $P$ or $T$ set.
+described so far requires $\BigO{n^2}$ space at each level of the search tree,
+since labels may be up to $n_G$ bits in length. We can reduce the memory
+requirement to $\BigO{n}$ per level by storing a pair $\langle P,T \rangle$ for
+each label $l$ that is used, where $P$ is the set of vertices in $\V(G)$
+labelled $l$, and $T$ is the set of vertices in $\V(H)$ labelled $l$. Since
+there are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$
+label-classes can exist at once, and the total length of the $P$ and $T$ sets
+is at most $n_G + n_H$.  We can gain further efficiency by discarding during
+search any label-class that has an empty $P$ or $T$ set.
 
@@
 \paragraph{Full algorithm} The full algorithm is shown as
-Algorithm~\ref{jtAlg}.  The main recursive procedure, $\FuncSty{expand}$, has
+Algorithm~\ref{jtAlg}.  The recursive procedure, $\FuncSty{expand}$, has
 two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
@@
 with graphs $G$ and $H$ as inputs.  This function returns the size of the
-largest mapping.  In \lineref{FirstExpandCall} a call is made to
+largest mapping.  In \lineref{FirstExpandCall} the initial call is made to
 $\FuncSty{expand}$; at this point we have a single label-class containing all
@@
 
-If we have vertex labels, we replace ${\langle V(G),V(H) \rangle}$ in line 28 of
-the algorithm with a set containing a label class for each label that appears in
-both $\V(G)$ and $\V(H)$. In the remainder of this section, \textit{labels} refers to
-edge labels.
+\paragraph{Vertex labels} If we have vertex labels, we replace ${\langle
+V(G),V(H) \rangle}$ in \lineref{FirstExpandCall} of the algorithm with a set
+containing a label class for each label that appears in both $\V(G)$ and
+$\V(H)$. Loops on vertices can be handled without difficulty; assuming without
+loss of generality that all vertex labels are positive integers, we can negate
+the labels of all vertices that have loops.
 
-For the unlabelled, directed case, let $A_G$ and $A_H$ be the adjacency matrices
+The remainder of this section considers edges that are directed or labelled.
+
+\paragraph{Directed, without edge labels} Let $A_G$ and $A_H$ be the adjacency matrices
 of $G$ and $H$ respectively, as stored in memory. For each
@@
 
-In the labelled, undirected case, each adjacency-matrix entry contains the edge
-label, or a null entry 0 if no edge exists between the pair of vertices.  we
-can also use Algorithm~\ref{labDirAlg}, by letting $L$ be the union of $\{0\}$
-with the set of all labels that appear in the input graphs. Since there may be
-up to $n_G + n_H$ distinct labels, the loop in Algorithm~\ref{labDirAlg} may
-execute up to $n_G + n_H$ times, and this variation of the algorithm therefore
-requires $\BigO{n^2}$ time per search node.  It is possible to modify the
-algorithms to run in $\BigO{n \log n}$ per search node; we will give a brief
-outline of the method. First, run lines 17-19 of Algorithm~\ref{jtAlg} to
-create a new label-class of vertices that are not adjacent to $v$ or $w$, and
-remove these vertices from $\langle P',T' \rangle$. Next, sort $P'$ and $T'$ in
-ascending order of the label on the edge from $v$ or $w$ to each vertex. We can
-then create the label classes corresponding to each edge label by
-simulataneously traversing $P'$ and $T'$ from left to right.
+\paragraph{Undirected, with edge labels} In this case, each adjacency-matrix
+entry contains the edge label, or a null entry 0 if no edge exists between the
+pair of vertices.  we can also use Algorithm~\ref{labDirAlg}, by letting $L$ be
+the union of $\{0\}$ with the set of all labels that appear in the input
+graphs. Since there may be up to $n_G + n_H$ distinct labels, the loop in
+Algorithm~\ref{labDirAlg} may execute up to $n_G + n_H$ times, and this
+variation of the algorithm therefore requires $\BigO{n^2}$ time per search
+node.  It is possible to modify the algorithms to run in $\BigO{n \log n}$ per
+search node; we will give a brief outline of the method. First, run lines 17-19
+of Algorithm~\ref{jtAlg} to create a new label-class of vertices that are not
+adjacent to $v$ or $w$, and remove these vertices from $\langle P',T' \rangle$.
+Next, sort $P'$ and $T'$ in ascending order of the label on the edge from $v$
+or $w$ to each vertex. We can then create the label classes corresponding to
+each edge label by simulataneously traversing $P'$ and $T'$ from left to right.
 
-The labelled, directed case is similar to its undirected counterpart, except that
-each element $A[u][v]$ in an adjacency matrix is a pair $(l_1, l_2)$, where
-$l_1$ is the label on the edge $u \rightarrow v$ (or 0 if no edge exists) and $l_2$
-is the label on the reverse edge.
+\paragraph{Directed, with edge labels} This case is similar to its undirected
+counterpart, except that each element $A[u][v]$ in an adjacency matrix is a
+pair $(l_1, l_2)$, where $l_1$ is the label on the edge $u \rightarrow v$ (or 0
+if no edge exists) and $l_2$ is the label on the reverse edge.
 

### 2017-02-10 paper/paper.tex
Started addressing Patrick's comments

@@
 
-We assume that $G$ and $H$ are unlabelled, undirected graphs without loops
+We initially assume that $G$ and $H$ are unlabelled, undirected graphs without loops
 (Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
@@
 We use $n_G$ and $n_H$ to denote $|\V(G)|$ and $|\V(H)|$. We will occasionally use
-and $n$ in complexity results to denote $n_G + n_H$.
+$n$ in complexity results to denote $n_G + n_H$.
 
@@
 The most successful existing approaches to solving MCS are reformulation to the
-maximum clique problem on one hand, and constraint programming
-\cite{DBLP:conf/cp/McCreeshNPS16} on the other. In the clique encoding, an
-association graph is created with one node for each pair in the set $\{(u,v)
-\mid u \in \V(G), v \in \V(H) \}$ and an edge for each pair of mappings that
-may be chosen together. A maximum clique in this graph corresponds directly to
-a maximum common subgraph of $G$ and $H$.
+maximum clique problem on one hand \cite{DBLP:conf/cp/McCreeshNPS16}, and
+constraint programming on the other \cite{DBLP:conf/cp/NdiayeS11}. In the
+clique encoding, an association graph is created with one node for each pair in
+the set $\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and an edge for each pair of
+mappings that may be chosen together. A maximum clique in this graph
+corresponds directly to a maximum common subgraph of $G$ and $H$.
 
@@
 in which values are assigned to variables, and propagation rules are used to
-remove infeasible values from domains.
+remove infeasible values from domains.  A matching algorithm is used to propagate
+a soft $\mathit{allDifferent}$ constraint. ? Say this better
 

### 2017-02-10 paper/paper.tex
Wrote a bit more about extensions and experiments

@@
 \section{Introduction}
-
-\citet{UpcomingAAAIPaper}
-
-\citet{DBLP:conf/cp/McCreeshNPS16}
-
-\citet{DBLP:conf/cp/NdiayeS11}
 
@@
 
-For the variant in which the common subgraph must be connected, we consider
-only undirected graphs. We modify the algorithm by permitting branching only on
-a vertex $v$ that has at least one non-zero element in its bitstring label.
-(Cite the first paper that did this.) ? Should we say that smarter propagation
-might be possible?
+\paragraph{Maximum Common \emph{Connected} Subgraph} For the variant in which
+the common subgraph must be connected, we consider only undirected graphs. We
+modify the algorithm by permitting branching only on a vertex $v$ that has at
+least one non-zero element in its bitstring label.  (Cite the first paper that
+did this.) ? Should we say that smarter propagation might be possible?
+
+\paragraph{Almost Subgraph Isomorphism} \citet{UpcomingAAAIPaper} present a
+``top-down'' algorithm for Maximum Common Subgraph in which a solution of size
+$n_G$ is initially sought; if such a solution does not exist then this
+solutions of size $n_G-1, n_G-2, \dots$ are sought in sequence.  This strategy,
+when used with techniques based on a state-of-the-art subgraph isomorphism
+solver, outperformed CP and clique encodings on a large benchmark set of graph
+pairs.  We can modify the \McSplit\ algorithm to use the top-down strategy by
+calling the main \FuncSty{McSplit} method once per goal size ($n_G, n_G-1,
+n_G-2, \dots$).  We backtrack (\lineref{PruneSearch} of Algorithm~\ref{jtAlg})
+when the bound is strictly less than the goal size, and terminate when a
+solution of the goal size is found.
 
@@
 
-Dual Xeon E5-2640 v2 CPUs, 64GBytes RAM. Parallel experiments 32 threads, using
-all 16 physical cores with hyper-threading enabled.
+Our experiments were performed on a cluster of machines with Dual Xeon E5-2640
+v2 CPUs and 64GBytes RAM. Parallel experiments used 32 threads per machine,
+using all 16 physical cores with hyper-threading enabled.
+
+For all experiments except the top-down ones, we used a database of
+randomly-generated Maximum Common Subgraph instances, which contains graphs of
+order ranging from 10 to 100 vertices
+\cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}.  For
+unlabelled instances, we selected 10 per cent of the 41,100 instances with up
+to 40 vertices; labels were discarded. For labelled instances, we selected 10
+per cent of the full database of 81,400 instances, and used a labelling scheme
+in which then number of distict vertex labels and the number of distinct edge
+labels is approximately equal to 33 per cent of the number of vertices in each
+graph (? re-word, give original reference for labelling scheme).
+
+(? Ciaran, while I remember to ask: what should a program do if it is running
+undirected, labelled and the same edge appears twice in the input in opposite
+directions? My program just uses the one that appears later in the input file.)
 

### 2017-02-13 paper/paper.tex
Added Patrick's edits, excluding those to the introduction

@@
 
-\title{A Partitioning Algorithm for Maximum Common (Connected) Subgraphs\thanks{This work
+\title{A Partitioning Algorithm for Maximum Common Subgraphs\thanks{This work
 was supported by the Engineering and Physical Sciences Research Council [grant
@@
     Maximium Common Subgraph (MCS) is the problem of finding a
-    graph of maximium size that is an induced subgraph of each of the two input
+    graph of maximium order that is an induced subgraph of each of the two input
     graphs.  We introduce \McSplit, a branch and bound algorithm that
@@
     for each recursive call, greatly improving on all existing algorithms for
-    the problem.
-
-    Experimental results using two large graph databases show that \McSplit\ is
+    the problem. Experimental results using two large graph databases show that \McSplit\ is
     more than an order of magnitude faster than the leading existing algorithms
@@
 
-We initially assume that $G$ and $H$ are unlabelled, undirected graphs without loops
+We initially assume that graphs are unlabelled, undirected and without self-loops
 (Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
-
 The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.
@@
 \emph{neighbourhood} of $v$, and denoted $\N(v, G)$.
-
 We use $n_G$ and $n_H$ to denote $|\V(G)|$ and $|\V(H)|$. We will occasionally use
-$n$ in complexity results to denote $n_G + n_H$.
+$n$ in complexity results to denote $n_G + n_H$.  The graph $\overline{G}$ is the complement (aka the inverse) of $G$.
 
@@
 
-In the constraint programming approach, we have a variable corresponding to
+In the constraint programming approach  \cite{DBLP:conf/cp/NdiayeS11}, a variable corresponds to
 each vertex in $G$, and each variable initially has one value in its domain for
@@
 in which values are assigned to variables, and propagation rules are used to
-remove infeasible values from domains.  A matching algorithm is used to propagate
-a soft $\mathit{allDifferent}$ constraint. ? Say this better
+remove infeasible values from domains.  A matching algorithm propagates
+a soft global allDifferent constraint which maximizes the number variables that are assigned to values different from $\bot$, while ensuring they are all different when they are not assigned to $\bot$.
+Different constraint propagation techniques are compared in \cite{DBLP:conf/cp/NdiayeS11}.
+The combination ``MAC+Bound'' (resp.
+``FC+Bound'') obtains the best results on labelled (resp.
+non labelled) graphs and outperforms the approach of \cite{McGreg82} and the CP model of \cite{DBLP:conf/mco/VismaraV08}.
+The combination ``MAC+Bound'' maintains arc consistency~\cite{sabi94} of edge constraints, whereas the conbination ``FC+Bound'' simply performs forward-checking on these constraints.
+In both combinations, the ``Bound'' filtering checks whether it is possible to assign distinct values to enough variables to surpass the best cost found so far (it is a weaker version of GAC(\emph{softAllDiff}) \cite{peti01} which computes the maximum number of variables that can be assigned distinct values).
 
@@
 
-This paper's main contribution is a branch and bound algorithm, \McSplit,
+We now present \McSplit, a branch and bound algorithm 
 which uses a new vertex-labelling scheme to avoid explicitly maintaining the
-set of feasible mappings for each vertex in $\V(G)$. As we will explore in
+set of feasible mappings for each vertex in $\V(G)$. 
+\iffalse
+As we will explore in
 Section~\ref{sec:comparison}, the algorithm may be viewed as a version of the
@@
 %have the same label.
-
-The \McSplit\ algorithm finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
+\fi
+\McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
 (v_{|M^*|}, w_{|M^*|})\}$, where the $v_i$ are distinct members of $\V(G)$ and the
-$w_i$ are distinct members of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
+$w_i$ are distinct vertices of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
 in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. It is clear that,
@@
 algorithm.  These graphs have a maximum common subgraph with four vertices; one
-example is the mapping $\{1a, 2f, 3d, 5b\}$ (where vertex $1$ is assigned to
-vertex $a$, $2$ is assigned to $f$, and so on).
+example is the mapping $\{1a, 2f, 3d, 5b\}$ where vertex $1$ is assigned to
+vertex $a$, $2$ is assigned to $f$, $3$ to $d$ and $5$ to $b$.
 
@@
 our example we will arbitrarily choose vertex $1$. Each of the vertices in
-$\V(H)$ to which $1$ may be mapped will be tried in turn, and finally the
-possibility where $1$ remains unmatched will be tried.
+$\V(H)$ to which vertex $1$ may be mapped will be tried in turn, and finally the
+possibility where vertex $1$ remains unmatched will be tried.
 
-We will begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Label
-each unmatched vertex according to whether it is adjacent to $1$ (in $G$) or
-$a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent vertices have label
+We begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Now label
+each unmatched vertex according to whether it is adjacent to vertex $1$ (in $G$) or
+vertex $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent vertices have label
 1; non-adjacent vertices have label 0.  It is straightforward to see that we
 can extend $M$ with a vertex mapping $vw$, with $v \in \V(G)$ and $w \in
-\V(H)$, if and only if $v$ and $w$ have the same label.  This property that
-two vertices may be mapped together if and only if they share a label is the
+\V(H)$, if and only if $v$ and $w$ have the same label.  This property, that
+two vertices may be mapped together if and only if they share a label, is the
 algorithm's main invariant.
@@
 2d\}$ (Figure~\ref{fig:alg3}).  Each unmapped vertex $v \in \V(G)$ is labelled
-with a two-character bit string, indicating whether $v$ is adjacent to each of
+with a two-character bit string, indicating its adjacency to each of
 the two mapped vertices in $\V(G)$ (vertices $1$ and $2$).  For example, vertex
-$3$ is labelled $01$, indicating that it is not adjacent to $1$ but is adjacent
-to $2$.  Labels are given to unmapped vertices in $\V(H)$ in a similar fashion,
-showing whether each vertex is adjacent to $a$ and $d$.
+$3$ is labelled $01$, indicating that it is not adjacent to vertex $1$ but is adjacent
+to vertex $2$.  Labels are given to unmapped vertices in $\V(H)$ in a similar fashion,
+showing adjacency to matched vertices $a$ and $d$.
 
 Our invariant is maintained: we can extend $M$ by a vertex pairing if
-and only if the two vertices have the same label. Note also that each label is prefixed
+and only if the two vertices have the same label. Note also that each label (in Figure~\ref{fig:alg3}) is prefixed
 by the label at the previous level of the search tree (Figure~\ref{fig:alg2}).
@@
 
-The algorithm backtracks when the incumbent best mapping is at least as large
+The algorithm backtracks when the incumbent, i.e. the largest matching found so far, is at least as large
 as a calculated bound given $M$ and the current labelling. To demonstrate how
@@
 since labels may be up to $n_G$ bits in length. We can reduce the memory
-requirement to $\BigO{n}$ per level by storing a pair $\langle P,T \rangle$ for
+requirement to $\BigO{n}$ per level by storing a \emph{label class} as a pair $\langle P,T \rangle$ for
 each label $l$ that is used, where $P$ is the set of vertices in $\V(G)$
-labelled $l$, and $T$ is the set of vertices in $\V(H)$ labelled $l$. Since
-there are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$
-label-classes can exist at once, and the total length of the $P$ and $T$ sets
+labelled $l$, and $T$ is the set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology from subgraph isomorphism where one graph is a pattern graph $P$ and the other the target graph $T$.}.
+Since there are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$
+label classes can exist at once, and the total length of the $P$ and $T$ sets
 is at most $n_G + n_H$.  We can gain further efficiency by discarding during
-search any label-class that has an empty $P$ or $T$ set.
+search any label class that has an empty $P$ or $T$ set.
 
@@
 %\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
-\nl \lIf {$|M| > |\AlgVar{incumbent}|$}
-    {$save(matching)$} \label{StoreIncumbent}
+\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
 %\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
-\nl $bestFuture \gets 0$ \label{SetBestFutureToZero} \;
-\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}
-    {$\AlgVar{bestFuture} \gets \AlgVar{bestFuture} + \min(|P|,|T|)$}
-\nl \lIf {$|M|  + \AlgVar{bestFuture} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
+\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
+\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
+\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
 \bigskip
@@
 \nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
-\nl        \lIf {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
+\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
         {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
@@
 \nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
-\nl        \lIf {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
+\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
         {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
@@
 \nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
-\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\} \cup \{\langle P', T \rangle\}$\;
-\nl \lIf {$P' = \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P',T \rangle \}$}
+\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
+\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
 \nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
@@
 
-\paragraph{Full algorithm} The full algorithm is shown as
-Algorithm~\ref{jtAlg}.  The recursive procedure, $\FuncSty{expand}$, has
+\paragraph{Algorithm \ref{jtAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
 two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
@@
 to unseat the incumbent.  \Linerangeref{SetBestFutureToZero}{PruneSearch} prune
-the search when a calculated upper bound ($\AlgVar{bestFuture}$) is not larger
+the search when a calculated upper bound (best future $\AlgVar{bf}$ plus $|M|$) is not larger
 than the incumbent.
 
-The remainder of $\FuncSty{expand}$ performs the search.  A label-class
-$\langle P, T \rangle$ is selected from $\AlgVar{future}$
-(\lineref{SelectClass}) using some heuristic.  From this class, a pattern
-vertex $v$ is selected using a heuristic (\lineref{SelectVertex}). We now
+The remainder of $\FuncSty{expand}$ performs the search.  A label class
+$\langle P, T \rangle$ is selected from $\AlgVar{future}$, where $P \subseteq V(G)$ and $T \subseteq V(H)$,
+using some heuristic (\lineref{SelectClass}).  From this label class, a
+vertex $v$ is selected from $P$ using a heuristic (\lineref{SelectVertex}). We now
 iterate over all possible vertices in $T$ (\linerangeref{WLoop}{ShrinkM}). A
-vertex $w \in \V(H)$ is selected mapping $M$ is extended (\lineref{GrowM}). A
+vertex $w \in T$ is selected and mapping $M$ is extended (\lineref{GrowM}). A
 new set of label-classes, $\AlgVar{future'}$, is created (\lineref{NewFuture}).
@@
 these classes (\linerangeref{NewPWithEdge}{AddToFutureWithEdge}) contains
-pattern vertices adjacent to $v$ and target vertices adjacent to $w$.  This is
+vertices in $P$ adjacent to $v$ and vertices in $T$ adjacent to $w$.  This is
 added to $\AlgVar{future'}$ if both sets contain at least one vertex.  This is
@@
 which we remove the mapping $(v,w)$.  Having explored all possible mappings of
-$v$ its target set we now consider what happens if $v$ is not matched
+$v$ with vertices in $T$ we now consider what happens if $v$ is not matched
 (\linerangeref{RemoveV}{ExpandWithoutV}). Note that removing $v$ from its label
-class is equivalent in effect to assigning it the label $\bot$, which is never
-taken by any vertex in $\V(H)$.
+class is equivalent to matching vertex $v$ to the null vertex $\bot$.
 
@@
 
+\iffalse
 \paragraph{Heuristics} We ran a small experiment comparing lots of heuristics
@@
 as small a max(left size, right size) as possible seemed to work best. ? Why?
+\fi
 
 \section{Comparison with Existing Algorithms}\label{sec:comparison}
-
-The Lilybank algorithm explores exactly the same search tree as CP-FC, because \dots
-
-The new algorithm takes $\BigO{n}$ time at each search node. CP-FC takes $\BigO{?}$ time,
-and clique takes $\BigO{n^4}$ time.
+The  \McSplit\ algorithm may be considered as a version of the
+constraint programming, forward-checking (CP-FC) algorithm of \cite{DBLP:conf/cp/NdiayeS11}.
+The vertices in $P$ correspond to variables in CP-FC and the 
+vertices in $T$ to the domains of variables in CP-FC. A label class $\langle P,T \rangle$ then corresponds to a set 
+of variables ($P$) in CP-FC  with identical domains ($T$). Where CP-FC uses a soft allDifferent constraint to filter and compute 
+a bound, \McSplit\ computes a bound from each label class in the future, i.e. it performs Hall Set reasoning on the fly.
+\iffalse
+\McSplit\ takes $\BigO{n}$ time at each search node. Furthermore, \McSplit\ has a low start up cost. The clique encoding requires the
+construction of the association graph prior to search
 
@@
 
-? Mention partition backtracking
+%? Mention partition backtracking
+\fi
 
 \section{Extensions}\label{sec:extensions}
+\paragraph{Vertex labels and self-loops} If vertices are labelled, replace ${\langle
+V(G),V(H) \rangle}$ in \lineref{FirstExpandCall} with a set of label classes, one for each label that appears in both $\V(G)$ and
+$\V(H)$. If vertex labels are represented as positive integers, represent self-loops
+as negative integers.
 
-So far, we have considered only unlabelled, undirected graphs. In this section, we
-describe some straightforward changes to the algorithm that relax these restrictions.
-
-\paragraph{Vertex labels} If we have vertex labels, we replace ${\langle
-V(G),V(H) \rangle}$ in \lineref{FirstExpandCall} of the algorithm with a set
-containing a label class for each label that appears in both $\V(G)$ and
-$\V(H)$. Loops on vertices can be handled without difficulty; assuming without
-loss of generality that all vertex labels are positive integers, we can negate
-the labels of all vertices that have loops.
-
-The remainder of this section considers edges that are directed or labelled.
-
-\paragraph{Directed, without edge labels} Let $A_G$ and $A_H$ be the adjacency matrices
-of $G$ and $H$ respectively, as stored in memory. For each
+\paragraph{Directed graphs without edge labels} Let $A_G$ and $A_H$ be adjacency matrices
+of $G$ and $H$. For each
 vertex pair $(t,u)$ in graph $G$ or $H$, the adjacency matrix entry 
@@
 t$, and 3 if there are edges in both directions. Where lines 14-19 of the
-basic algorithm split the pair $\langle P',T' \rangle$ two ways, according to whether
-vertices are ajacent to $v$ (resp. $w$), we now perform a four-way split where
-each vertex is classified according to the label on its adjacency matrix from $v$
-(resp. $w$). This is shown in Algorithm~\ref{labDirAlg}, where $L=\{0,1,2,3\}$.
+basic algorithm split the label class $\langle P',T' \rangle$ according to whether
+vertices are adjacent to $v$ (resp. $w$), we now perform a four-way split where
+each vertex is classified according to the label on its adjacency matrix from $v$ (resp. $w$). 
+This is shown in Algorithm~\ref{labDirAlg}, where $L=\{0,1,2,3\}$.
 
@@
 
-\paragraph{Undirected, with edge labels} In this case, each adjacency-matrix
-entry contains the edge label, or a null entry 0 if no edge exists between the
-pair of vertices.  we can also use Algorithm~\ref{labDirAlg}, by letting $L$ be
+\paragraph{Undirected with edge labels} Each adjacency matrix
+entry contains an edge label, or a null entry (non-adjacent) $0$.
+We use Algorithm~\ref{labDirAlg}, by letting $L$ be
 the union of $\{0\}$ with the set of all labels that appear in the input
 graphs. Since there may be up to $n_G + n_H$ distinct labels, the loop in
-Algorithm~\ref{labDirAlg} may execute up to $n_G + n_H$ times, and this
-variation of the algorithm therefore requires $\BigO{n^2}$ time per search
+Algorithm~\ref{labDirAlg} may execute up to $n_G + n_H$ times and
+requires $\BigO{n^2}$ time per search
 node.  It is possible to modify the algorithms to run in $\BigO{n \log n}$ per
-search node; we will give a brief outline of the method. First, run lines 17-19
+search node.  First, run lines 17-19
 of Algorithm~\ref{jtAlg} to create a new label-class of vertices that are not
@@
 or $w$ to each vertex. We can then create the label classes corresponding to
-each edge label by simulataneously traversing $P'$ and $T'$ from left to right.
+each edge label by simultaneously traversing $P'$ and $T'$ from left to right.
 
-\paragraph{Directed, with edge labels} This case is similar to its undirected
+\paragraph{Directed with edge labels} This case is similar to its undirected
 counterpart, except that each element $A[u][v]$ in an adjacency matrix is a
@@
 
-\paragraph{Maximum Common \emph{Connected} Subgraph} For the variant in which
-the common subgraph must be connected, we consider only undirected graphs. We
-modify the algorithm by permitting branching only on a vertex $v$ that has at
-least one non-zero element in its bitstring label.  (Cite the first paper that
-did this.) ? Should we say that smarter propagation might be possible?
+\paragraph{Maximum Common \emph{Connected} Subgraph} 
+The common subgraph must be connected. We consider only undirected graphs. 
+Modify \McSplit\ by permitting branching only on a vertex $v$ that has at
+least one non-zero element in its bit-string label.
 
+\iffalse
 \paragraph{Almost Subgraph Isomorphism} \citet{UpcomingAAAIPaper} present a
@@
 solution of the goal size is found.
+\fi
 
-\section{Computational Experiments}
+\section{A Computational Study}
 
-Our experiments were performed on a cluster of machines with Dual Xeon E5-2640
-v2 CPUs and 64GBytes RAM. Parallel experiments used 32 threads per machine,
-using all 16 physical cores with hyper-threading enabled.
-
-For all experiments except the top-down ones, we used a database of
+Experiments were performed on a cluster of machines with Dual Xeon E5-2640
+v2 CPUs and 64GBytes RAM. 
+%Parallel experiments used 32 threads per machine,
+%using all 16 physical cores with hyper-threading enabled.
+All experiments used a database of
 randomly-generated Maximum Common Subgraph instances, which contains graphs of
@@
 per cent of the full database of 81,400 instances, and used a labelling scheme
-in which then number of distict vertex labels and the number of distinct edge
+in which the number of distinct vertex labels and the number of distinct edge
 labels is approximately equal to 33 per cent of the number of vertices in each
-graph (? re-word, give original reference for labelling scheme).
+graph.
+Small scale experiments (not presented here) were performed to identify a suitable heuristic for label class selection.
+The heuristic used selects the label class  with the smallest product $|P| \times |T|$ (a surrogate for expected branching factor) and from that selects the first vertex in $P$.
 
-(? Ciaran, while I remember to ask: what should a program do if it is running
-undirected, labelled and the same edge appears twice in the input in opposite
-directions? My program just uses the one that appears later in the input file.)
+\paragraph{No vertex labels, no edge labels, undirected}
+\cref{figure:plain-cumulative} shows a plot of cumulative run times (x-axis) against number of problem instances solved (y-axis). There are five contours: 
+the clique encoding \cite{DBLP:conf/cp/McCreeshNPS16}, CP-FC \cite{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.
+WE NEED CITATIONS FOR THESE AND ALSO NEED TO CLEAN UP THE LABELS ON THE GRAPHS.
+	What we see is that \McSplit\ is typically more than one order of magnitude better than its nearest competitor.
 
-\paragraph{No labels, undirected} are in \cref{figure:plain-cumulative}.
+\paragraph{Vertex labels, no edge labels, undirected} are in \cref{figure:33v-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
 
-\paragraph{Vertex labels but no edge labels, undirected} are in \cref{figure:33v-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
+\paragraph{Vertex labels,  edge labels, undirected} are in \cref{figure:33ve-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
 
-\paragraph{Vertex and edge labels, not directed} are in \cref{figure:33ve-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
+\paragraph{Vertex labels, edge labels, directed} are in \cref{figure:33ved-cumulative}.
 
-\paragraph{Vertex and edge labels, directed} are in \cref{figure:33ved-cumulative}.
+\paragraph{No Vertex labels, no edge labels, undirected, connected} are in \cref{figure:plain-connected-cumulative}.
 
-\paragraph{No labels, undirected, connected} are in \cref{figure:plain-connected-cumulative}.
-
-\paragraph{Vertex and edge labels, undirected, connected} are in \cref{figure:33ve-connected-cumulative}.
+\paragraph{Vertex labels, edge labels, undirected, connected} are in \cref{figure:33ve-connected-cumulative}.
 
@@
 
-We have introduced the \McSplit\ algorithm for Max Common Subgraph, and shown
-that it explores the same search tree as CP-FC with greatly reduced effort at
-each search node.  The new algorithm is more than an order of
-magnitude faster than the previous state of the art for unlabelled, undirected
-instances.
+We have introduced the \McSplit\ algorithm for Max Common Subgraph.
+The new algorithm is more than an order of
+magnitude faster than the previous state of the art for unlabelled and undirected
+instances. We have shown how the algorithm can be extended for graphs with labels on edges, labels on vertices, self-loops, directed edges and the requirement that the resultant graph be connected.
 
-\paragraph{Future work} We could create a hybrid of Lilybank and clique (for
-example, using Lilybank near the top of the search tree, but switching to a
+\paragraph{Future work} We could create a hybrid algorithm. This would use \McSplit\ near the top of the search tree, switching to a
 clique encoding when fewer than some threshold number of vertices remain to be
-selected). This might give us most of the benefits of the clique encoding for
+selected. This might deliver the benefits of the clique encoding for
 labelled graphs, while avoiding the high memory cost and colouring time of
@@
 
-Instead of branching on vertices in $\V(G)$, it would be equally valid to branch on
-a vertex in $\V(H)$ in lines 9-23 of the algorithm, since our data structure treats the
+We have yet to fully investigate variable and value ordering heuristics, i.e. the selection of a label class and the selection of a vertex from that label class.
+There is also the potential to branch on both sides, that is instead of branching on vertices in $\V(G)$, it would be equally valid to branch on
+a vertex in $\V(H)$ , since our data structure treats the
 two graphs symmetrically. More interestingly, we could choose which graph to branch on
 at each search node using some heuristic (perhaps choosing based on whether the $P$ or
-$T$ set is smaller).
+$T$ set is smaller), so long as we can ensure that the search remains complete.
 
-Parallel versions: bitset and multithreaded
+We have not yet explored parallel versions of the algorithm. The first paralellism is \emph{within-core}, using bitsets to encode sets. The second parallelism  is  multithreaded,
+exploited the ubiquitous multicore architecture.
 
-Application to induced subgraph isomorphism (with almost no changes). Applications
-of a similar technique to other problems???
+%Application to induced subgraph isomorphism (with almost no changes). Applications
+%of a similar technique to other problems???
 
+\newpage
 \bibliographystyle{named}

### 2017-02-13 paper/paper.tex
Added Ciaran's draft introduction (fixing citations)

@@
 
-Given graphs $G$ and $H$, the \emph{Maximum Common Subgraph (MCS)} problem is
-to find a graph with as many vertices as possible that is isomorphic to induced
-subgraphs of both $G$ and $H$.
-
-MCS has numerous applications, including finding the similarity between organic
-molecules, \dots
-
-This paper introduces a new branch and bound algorithm for MCS that improves
-state-of-the-art run times on unlabelled, undirected instances by well over an
-order of magnitude, and uses only quadratic memory in the order of the input
-graphs, allowing us to solve larger instances than previously possible.
+To determine the similarity or difference between two graphs, we must first
+find what they have in common
+\citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,KriegeThesis}.
+The \emph{maximum common subgraph} family of problems involves finding a large
+graph which is isomorphic to subgraphs of two given graphs simultaneously.
+Because graphs are widely used to model real-world phenomena, maximum common
+subgraph problems have arisen in ...  The problems are NP-hard, and remain
+challenging computationally. Recent practical progress has been made by using
+constraint programming
+\citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}, by reducing to the
+maximum clique problem \citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and
+by adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. This paper
+introduces a new branch and bound algorithm which exploits special properties
+of these problems to allow a much faster exploration of the search space,
+whilst retaining the filtering and bounding benefits of the constraint
+programming approach. We describe the algorithm for the basic maximum common
+subgraph problem, and discuss how it may be adapted to handle vertex labels,
+edge labels, and the requirement that the found subgraph be connected. We then
+present an empirical study of the algorithm, demonstrating that it improves the
+state of the art by more than an order of magnitude on the unlabelled variant
+of the problem, and showing that it can handle much larger instances than
+earlier constraint programming or clique approaches due to lower memory usage.
 

### 2017-02-14 code/james-cpp/mcsp.c
Removed spurious upper bound output on timeout when -b flag is used

@@ -466,7 +466,7 @@ vector<VtxPair> mcs(const Graph & g0, const Graph & g1) {
             auto domains_copy = domains;
             vector<VtxPair> current;
             solve(g0, g1, incumbent, current, domains_copy, left_copy, right_copy, goal);
-            if (incumbent.size() == goal) break;
+            if (incumbent.size() == goal || abort_due_to_timeout) break;
             if (!arguments.quiet) cout << "Upper bound: " << goal-1 << std::endl;
         }
 

### 2017-02-14 paper/paper.tex
Changed plot sizes, removed a couple of plots

@@
 graph.
-Small scale experiments (not presented here) were performed to identify a suitable heuristic for label class selection.
-The heuristic used selects the label class  with the smallest product $|P| \times |T|$ (a surrogate for expected branching factor) and from that selects the first vertex in $P$.
+
+Small scale experiments (not presented here) were performed to identify a
+suitable heuristic for label class selection. The heuristic used selects the
+label class  with the smallest $\min(|P|,|T|)$ (a surrogate for expected
+branching factor). From this class, a vertex is selected from $P$ with maximum
+degree if $H$ is sparse and with minimum degree if $H$ is dense.
 
@@
 
-\paragraph{Vertex labels, no edge labels, undirected} are in \cref{figure:33v-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
+\paragraph{Vertex labels, no edge labels, undirected} are in
+FIGURE REMOVED
+%\cref{figure:33v-cumulative},
+and are not in other papers so we should omit this unless we see something interesting.
 
-\paragraph{Vertex labels,  edge labels, undirected} are in \cref{figure:33ve-cumulative}, and are not in other papers so we should omit this unless we see something interesting.
+\paragraph{Vertex labels,  edge labels, undirected} are in
+FIGURE REMOVED
+%\cref{figure:33ve-cumulative},
+and are not in other papers so we should omit this unless we see something interesting.
 
@@
 
-\paragraph{Compared to CP FC} Nodes are in \cref{figure:plain-james-versus-cp-fc-nodes-scatter} and \cref{figure:33ved-james-versus-cp-fc-nodes-scatter}.
+\paragraph{Compared to CP FC} Nodes are in
+FIGURE REMOVED
+%\cref{figure:plain-james-versus-cp-fc-nodes-scatter}
+and \cref{figure:33ved-james-versus-cp-fc-nodes-scatter}.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics*{gen-graph-33v-cumulative.pdf}
-    \caption{MCS 33\% vertex labelled, no edge labels, undirected instances, cumulative runtimes}\label{figure:33v-cumulative}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics*{gen-graph-33ve-cumulative.pdf}
-    \caption{MCS 33\% vertex and edge labelled undirected instances, cumulative runtimes}\label{figure:33ve-cumulative}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    \includegraphics*{gen-graph-33v-cumulative.pdf}
+%    \caption{MCS 33\% vertex labelled, no edge labels, undirected instances, cumulative runtimes}\label{figure:33v-cumulative}
+%\end{figure}
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics*{gen-graph-33ve-cumulative.pdf}
+%    \caption{MCS 33\% vertex and edge labelled undirected instances, cumulative runtimes}\label{figure:33ve-cumulative}
+%\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{MCS Plain instances, James vs CP-FC, Nodes (TODO?? heatmapify this
-    and exclude timeouts)}\label{figure:plain-james-versus-cp-fc-nodes-scatter}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
+%    \caption{MCS Plain instances, James vs CP-FC, Nodes (TODO?? heatmapify this
+%    and exclude timeouts)}\label{figure:plain-james-versus-cp-fc-nodes-scatter}
+%\end{figure}
 

### 2017-02-14 paper/paper.tex
Added some subfigures

@@
 \usepackage{cleveref}
+
+\usepackage{subfigure}
 
@@
     \centering
+    \subfigure[][MCS unlabelled unconnected instances] {
+        \centering
         \includegraphics*{gen-graph-plain-cumulative.pdf}
-    \caption{MCS Plain instances, cumulative runtimes}\label{figure:plain-cumulative}
-\end{figure}
+        \label{figure:plain-cumulative}
+    }
 
-%\begin{figure}
-%    \centering
-%    \includegraphics*{gen-graph-33v-cumulative.pdf}
-%    \caption{MCS 33\% vertex labelled, no edge labels, undirected instances, cumulative runtimes}\label{figure:33v-cumulative}
-%\end{figure}
-%
-%\begin{figure}
-%    \centering
-%    \includegraphics*{gen-graph-33ve-cumulative.pdf}
-%    \caption{MCS 33\% vertex and edge labelled undirected instances, cumulative runtimes}\label{figure:33ve-cumulative}
-%\end{figure}
-
-\begin{figure}
+    \subfigure[][MCS 33\% vertex and edge labelled directed instances] {
         \centering
         \includegraphics*{gen-graph-33ved-cumulative.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, cumulative runtimes}\label{figure:33ved-cumulative}
+        \label{figure:33ved-cumulative}
+    }
+    \caption{Cumulative runtimes}\label{figure:mcs-cumulative}
 \end{figure}
@@
     \centering
+    \subfigure[][MCS unlabelled undirected instances] {
+        \centering
         \includegraphics*{gen-graph-plain-connected-cumulative.pdf}
-    \caption{Unlabelled undirected instances, connected, cumulative runtimes}\label{figure:plain-connected-cumulative}
-\end{figure}
+        \label{figure:plain-connected-cumulative}
+    }
 
-\begin{figure}
+    \subfigure[][MCS 33\% vertex and edge labelled undirected instances] {
         \centering
         \includegraphics*{gen-graph-33ve-connected-cumulative.pdf}
-    \caption{MCS 33\% vertex and edge labelled undirected instances, connected, cumulative runtimes}\label{figure:33ve-connected-cumulative}
+        \label{figure:33ve-connected-cumulative}
+    }
+    \caption{Cumulative runtimes, connected variant}
+    \label{figure:mcs-connected-cumulative}
 \end{figure}
@@
     \includegraphics*{gen-graph-sip-cumulative.pdf}
-    \caption{SIP instances, cumulative runtimes}\label{figure:sip-cumulative}
+    \caption{Cumulative runtimes, SIP instances}\label{figure:sip-cumulative}
 \end{figure}

### 2017-02-14 paper/paper.tex
Some writing about experiments

@@
 \paragraph{No vertex labels, no edge labels, undirected}
-\cref{figure:plain-cumulative} shows a plot of cumulative run times (x-axis) against number of problem instances solved (y-axis). There are five contours: 
-the clique encoding \cite{DBLP:conf/cp/McCreeshNPS16}, CP-FC \cite{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.
-WE NEED CITATIONS FOR THESE AND ALSO NEED TO CLEAN UP THE LABELS ON THE GRAPHS.
-	What we see is that \McSplit\ is typically more than one order of magnitude better than its nearest competitor.
+\cref{figure:plain-cumulative} shows a plot of cumulative run times (x-axis)
+against number of problem instances solved (y-axis).  There are five contours:
+the clique encoding \cite{DBLP:conf/cp/McCreeshNPS16}, CP-FC
+\cite{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.  WE NEED CITATIONS
+FOR THESE AND ALSO NEED TO CLEAN UP THE LABELS ON THE GRAPHS.  We may compare
+the speed of two algorithms using the horizontal distance between their curves.
+For example, we could solve 2000 of the 4110 unlabelled undirected instances
+using the \McSplit\ algorithm if a time limit of 0.5 seconds per instance were
+imposed.  Its nearest competitor, CP-FC, would require a time limit of over 24
+seconds per instance to solve the same number of instances.  For any given
+number of instances, \McSplit\ is comfortably more than an order of magnitude
+faster than its nearer competitor.  Moreover, there are few instances in this
+set for which \McSplit\ is not the fastest solver; of the 3506 instances that
+could be solved by at least one of \McSplit, CP-FC and clique in less than than
+1000 seconds, there were only 126 instances on which \McSplit\ was not strictly
+faster than the other two solvers.
+
+\paragraph{Vertex labels, edge labels, directed}
+Cumulative run times for this class of instances are in \cref{figre:33ved-cumulative}.
+Again, \McClique\ is over an order of magnitude faster than the best CP algorithm,
+which is CP-MAC in this case.  The clique encoding, however, outperforms the other
+algorithms on these labelled instances, except in the easy region at the left
+of the graph.
 
@@
 and are not in other papers so we should omit this unless we see something interesting.
-
-\paragraph{Vertex labels, edge labels, directed} are in \cref{figure:33ved-cumulative}.
 
@@
     \centering
-    \subfigure[][MCS unlabelled unconnected instances] {
+    \subfigure[][MCS unlabelled undirected instances] {
         \centering
@@
     }
-    \caption{Cumulative runtimes}\label{figure:mcs-cumulative}
+    \caption{Cumulative run times}\label{figure:mcs-cumulative}
 \end{figure}
@@
     }
-    \caption{Cumulative runtimes, connected variant}
+    \caption{Cumulative run times, connected variant}
     \label{figure:mcs-connected-cumulative}
@@
     \includegraphics*{gen-graph-sip-cumulative.pdf}
-    \caption{Cumulative runtimes, SIP instances}\label{figure:sip-cumulative}
+    \caption{Cumulative run times, SIP instances}\label{figure:sip-cumulative}
 \end{figure}

### 2017-02-14 more-analysis/on-how-many-instances-is-mcsplit-faster.sh

@@ -1,0 +1,4 @@
+echo "Number of instances where at least one of clique, CP-FC, McClique didn't time out"
+cat ../experiments/gpgnode-results/mcsplain/runtimes.data | awk 'NR>1 && ($2<1000000 || $3<1000 || $8<1000000)' | wc -l
+echo "Number of these where McClique wan't strictly the fastest"
+cat ../experiments/gpgnode-results/mcsplain/runtimes.data | awk 'NR>1 && ($2<1000000 || $3<1000 || $8<1000000) && ($2<=$8 || $3*1000<=$8)' | wc -l

### 2017-02-14 paper/paper.tex
Fixed a couple of errors

@@
 \paragraph{Vertex labels, edge labels, directed}
-Cumulative run times for this class of instances are in \cref{figre:33ved-cumulative}.
-Again, \McClique\ is over an order of magnitude faster than the best CP algorithm,
+Cumulative run times for this class of instances are in \cref{figure:33ved-cumulative}.
+Again, \McSplit\ is over an order of magnitude faster than the best CP algorithm,
 which is CP-MAC in this case.  The clique encoding, however, outperforms the other

### 2017-02-14 paper/paper.tex
Highlighted a section which may need to be re-worded

@@
 in which values are assigned to variables, and propagation rules are used to
-remove infeasible values from domains.  A matching algorithm propagates
+remove infeasible values from domains.
+\textit{A matching algorithm propagates
 a soft global allDifferent constraint which maximizes the number variables that are assigned to values different from $\bot$, while ensuring they are all different when they are not assigned to $\bot$.
@@
 The combination ``MAC+Bound'' maintains arc consistency~\cite{sabi94} of edge constraints, whereas the conbination ``FC+Bound'' simply performs forward-checking on these constraints.
-In both combinations, the ``Bound'' filtering checks whether it is possible to assign distinct values to enough variables to surpass the best cost found so far (it is a weaker version of GAC(\emph{softAllDiff}) \cite{peti01} which computes the maximum number of variables that can be assigned distinct values).
+In both combinations, the ``Bound'' filtering checks whether it is possible to assign distinct values to enough variables to surpass the best cost found so far (it is a weaker version of GAC(\emph{softAllDiff}) \cite{peti01} which computes the maximum number of variables that can be assigned distinct values).}
 

### 2017-02-14 paper/paper.tex
Added a bit more to the comparison with other algorithms

@@
 a bound, \McSplit\ computes a bound from each label class in the future, i.e. it performs Hall Set reasoning on the fly.
-\iffalse
+
 \McSplit\ takes $\BigO{n}$ time at each search node. Furthermore, \McSplit\ has a low start up cost. The clique encoding requires the
 construction of the association graph prior to search
-
-Our data structure lets us compute good heuristics cheaply, which is why we take
-slightly fewer search nodes than CP-FC.
-
-Clique gives us very strong propagation.
 
@@
 
-%? Mention partition backtracking
-\fi
+? Mention partition backtracking
+
+LARGELY FROM CIARAN'S EMAIL; NEEDS WORK:
+
+\begin{algorithm}
+\DontPrintSemicolon
+\nl    \For {$q \in \N(b,G)$}{
+\nl        remove $t$ from $D'_q$ \;
+\nl        remove the inverse neighbourhood of $t$ from $D'_q$ \;
+       }
+\nl    \For {$q \in \N(b,\overline{G})$}{
+\nl        remove $t$ from $D'_q$ \;
+\nl        remove the neighbourhood of $t$ from $D'_q$ \;
+       }
+\caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} in CP algorithm}
+\label{cpAlg}
+\end{algorithm}
+
+The first difference is in the underlying data structures: James uses a clever
+partitioning data structure, not a conventional domain store. With the
+algorithm described this way, I think the proof that domains are partitioned
+into disjoint sets is almost immediate.
+
+The second difference is in the bound function: the French code constructs a
+maximum cardinality matching, with some special weighting trickery for $\bot$,
+whilst James just sums the smaller of the two sides of each partition. If you
+think a bit about what the matching looks like, you can see that these two give
+the same answer, but at very different speeds.
+
+The third difference is in the details of the variable ordering heuristic.
 
@@
 
-\paragraph{Directed graphs without edge labels} Let $A_G$ and $A_H$ be adjacency matrices
-of $G$ and $H$. For each
-vertex pair $(t,u)$ in graph $G$ or $H$, the adjacency matrix entry 
-takes the value 0 if the vertices are
-not adjacent, 1 if the two vertices share a single edge in the direction $t
+\paragraph{Directed graphs without edge labels} Let $A_G$ and $A_H$ be
+adjacency matrices of $G$ and $H$. For each vertex pair $(t,u)$ in graph $G$ or
+$H$, the adjacency matrix entry takes the value 0 if the vertices are not
+adjacent, 1 if the two vertices share a single edge in the direction $t
 \rightarrow u$, 2 if they share a single edge in the direction $u \rightarrow
-t$, and 3 if there are edges in both directions. Where lines 14-19 of the
-basic algorithm split the label class $\langle P',T' \rangle$ according to whether
-vertices are adjacent to $v$ (resp. $w$), we now perform a four-way split where
-each vertex is classified according to the label on its adjacency matrix from $v$ (resp. $w$). 
+t$, and 3 if there are edges in both directions. Where
+\linerangeref{NewPWithEdge}{InnerLoopEnd} of the basic algorithm split the
+label class $\langle P',T' \rangle$ according to whether vertices are adjacent
+to $v$ (resp. $w$), we now perform a four-way split where each vertex is
+classified according to the label on its adjacency matrix from $v$ (resp. $w$).
 This is shown in Algorithm~\ref{labDirAlg}, where $L=\{0,1,2,3\}$.
@@
        }
-\caption{Replacement for lines 14--19 in directed in labelled cases}
+\caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} in directed and labelled cases}
 \label{labDirAlg}
@@
 
+? Portfolio algorithms---finding out at what level of labelledness clique becomes best etc.
+
 %Application to induced subgraph isomorphism (with almost no changes). Applications

### 2017-02-15 paper/paper.tex
More writing about the experiments

@@
 
-\paragraph{No vertex labels, no edge labels, undirected}
+\begin{figure}
+    \centering
+    \subfigure[][MCS unlabelled undirected instances] {
+        \centering
+        \includegraphics*{gen-graph-plain-cumulative.pdf}
+        \label{figure:plain-cumulative}
+    }
+
+    \subfigure[][MCS 33\% vertex and edge labelled directed instances] {
+        \centering
+        \includegraphics*{gen-graph-33ved-cumulative.pdf}
+        \label{figure:33ved-cumulative}
+    }
+    \caption{Cumulative run times}\label{figure:mcs-cumulative}
+\end{figure}
+
+\paragraph{No vertex or edge labels, undirected}
 \cref{figure:plain-cumulative} shows a plot of cumulative run times (x-axis)
@@
 
-\paragraph{Vertex labels, edge labels, directed}
-Cumulative run times for this class of instances are in \cref{figure:33ved-cumulative}.
-Again, \McSplit\ is over an order of magnitude faster than the best CP algorithm,
-which is CP-MAC in this case.  The clique encoding, however, outperforms the other
-algorithms on these labelled instances, except in the easy region at the left
-of the graph.
+\paragraph{Vertex and edge labels, directed} Cumulative run times for this
+class of instances are in \cref{figure:33ved-cumulative}.  Again, \McSplit\ is
+over an order of magnitude faster than the best existing CP algorithm, which is
+CP-MAC in this case.  The clique encoding outperforms the other
+algorithms---including \McSplit---on these labelled instances, except in the
+very easy region of instances that can be solved in well under 100 ms.
 
-\paragraph{Vertex labels, no edge labels, undirected} are in
-FIGURE REMOVED
-%\cref{figure:33v-cumulative},
-and are not in other papers so we should omit this unless we see something interesting.
+\paragraph{No Vertex or edge labels, undirected, connected} are in
+\cref{figure:plain-connected-cumulative}.  These results are very similar to
+the corresponding experiment in which the subgraph is not required to be
+connected (\cref{figure:plain-cumulative}).
 
-\paragraph{Vertex labels,  edge labels, undirected} are in
-FIGURE REMOVED
-%\cref{figure:33ve-cumulative},
-and are not in other papers so we should omit this unless we see something interesting.
-
-\paragraph{No Vertex labels, no edge labels, undirected, connected} are in \cref{figure:plain-connected-cumulative}.
-
-\paragraph{Vertex labels, edge labels, undirected, connected} are in \cref{figure:33ve-connected-cumulative}.
-
-\paragraph{Large subgraph isomorphism instances} are in \cref{figure:sip-cumulative}.
-
-\paragraph{Compared to CP FC} Nodes are in
-FIGURE REMOVED
-%\cref{figure:plain-james-versus-cp-fc-nodes-scatter}
-and \cref{figure:33ved-james-versus-cp-fc-nodes-scatter}.
-
-\paragraph{Nodes compared to kdown on SIP} are in \cref{figure:sip-james-versus-kdown-nodes-scatter}.
-
-\begin{figure}
-    \centering
-    \subfigure[][MCS unlabelled undirected instances] {
-        \centering
-        \includegraphics*{gen-graph-plain-cumulative.pdf}
-        \label{figure:plain-cumulative}
-    }
-
-    \subfigure[][MCS 33\% vertex and edge labelled directed instances] {
-        \centering
-        \includegraphics*{gen-graph-33ved-cumulative.pdf}
-        \label{figure:33ved-cumulative}
-    }
-    \caption{Cumulative run times}\label{figure:mcs-cumulative}
-\end{figure}
+\paragraph{Vertex and edge labels, undirected, connected} For the labelled,
+connected case, clique slightly outperforms \McSplit\ on harder instances
+(\cref{figure:33ve-connected-cumulative}).  The gap between the two algorithms
+is narrow; indeed, the cumulative curve for \McSplit\ briefly rises above
+the curve for clique at a runtime just below 100 seconds.
 
@@
 
-\begin{figure}
+\paragraph{Large subgraph isomorphism instances} We also ran the algorithms on
+a set of 5725 larger instances used in recent studies of subgraph
+isomorphism~\cite{thelionpaper} and maximum common
+subgraph~\cite{UpcomingAAAIPaper}.  This benchmark set includes real-world
+graphs, along with graphs generated using several random models.  Pattern
+graphs range from 4 vertices to 900 with a median of 80; target graphs range
+from 10 vertices to 6671 with a median of 561.  Cumulative runtimes including
+top-down algorithm from~\cite{UpcomingAAAIPaper} which was the previous state
+of the art for this benchmark, are in \cref{figure:sip-cumulative}.  This is a
+challenging set of instances, and more than half of the instances cannot be
+solved within a timeout of 1000 seconds by any of the solvers. \McSplit\ is the
+strongest solver overall for these instances.
+
+\begin{figure}[h]
     \centering
@@
 
-%\begin{figure}
-%    \centering
-%    \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
-%    \caption{MCS Plain instances, James vs CP-FC, Nodes (TODO?? heatmapify this
-%    and exclude timeouts)}\label{figure:plain-james-versus-cp-fc-nodes-scatter}
-%\end{figure}
+\paragraph{CP-FC and \McSplit\ search nodes on MCS instances}
+\cref{figure:33ved-james-versus-cp-fc-nodes-scatter} shows one point per instance,
+with search nodes visited by CP-FC on the horizontal axis and search nodes visited
+by \McSplit\ on the vertical axis.  There is a strong correlation, which reflects
+the similarity of the search space explored by the two algorithms.  \McSplit\
+tends to explore slightly fewer search nodes than CP-FC, because the label-classes
+allow us to (cheaply) compute a very good heuristics for variable selection.
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 
-\begin{figure}
+\paragraph{kdown and \McSplit\ search nodes on SIP instances}
+\cref{figure:sip-james-versus-kdown-nodes-scatter} shows the number of search
+nodes taken by kdown and \McSplit\ on each instance.  Although \McSplit\ is the
+faster algorithm overall, it explores more search nodes than kdown for most instances.
+The scatter plot has a concentration of points around the line $y=2x$. We conjecture
+that the two algorithms explore very similar search trees for instances close to this
+line.
+
+\begin{figure}[h]
     \centering
@@
 
-\newpage
 \bibliographystyle{named}

### 2017-02-15 paper/paper.tex
More introduction, and some tidying

@@
 Because graphs are widely used to model real-world phenomena, maximum common
-subgraph problems have arisen in ...  The problems are NP-hard, and remain
-challenging computationally. Recent practical progress has been made by using
-constraint programming
-\citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}, by reducing to the
-maximum clique problem \citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and
-by adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. This paper
-introduces a new branch and bound algorithm which exploits special properties
-of these problems to allow a much faster exploration of the search space,
-whilst retaining the filtering and bounding benefits of the constraint
-programming approach. We describe the algorithm for the basic maximum common
-subgraph problem, and discuss how it may be adapted to handle vertex labels,
-edge labels, and the requirement that the found subgraph be connected. We then
-present an empirical study of the algorithm, demonstrating that it improves the
-state of the art by more than an order of magnitude on the unlabelled variant
-of the problem, and showing that it can handle much larger instances than
-earlier constraint programming or clique approaches due to lower memory usage.
+subgraph problems have arisen in molecular science (where graphs represent
+molecules)
+\citep{DBLP:journals/jcamd/RaymondW02a,Ehrlich:2011,DAM2014,Grindley1993707},
+computer vision \cite{DBLP:journals/jair/CookH94}, discovery of repeated
+patterns in source code \cite{DBLP:journals/tkde/DjokoCH97} and circuits
+\cite{DBLP:journals/jair/CookH94}, and recognition of handwritten characters
+\cite{SIWEILU1991617}.
+
+Maximum common subgraph problems are NP-hard, and remain challenging
+computationally. Recent practical progress has been made by using constraint
+programming \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}, by
+reducing to the maximum clique problem
+\citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and by adapting subgraph isomorphism
+algorithms \citep{UpcomingAAAIPaper}. This paper introduces a new branch and
+bound algorithm which exploits special properties of these problems to allow a
+much faster exploration of the search space, whilst retaining the filtering and
+bounding benefits of the constraint programming approach. We describe the
+algorithm for the basic maximum common subgraph problem, and discuss how it may
+be adapted to handle vertex labels, edge labels, and the requirement that the
+found subgraph be connected. We then present an empirical study of the
+algorithm, demonstrating that it improves the state of the art by more than an
+order of magnitude on the unlabelled variant of the problem, and showing that
+it can handle much larger instances than earlier constraint programming or
+clique approaches due to lower memory usage.
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[ht]
 \DontPrintSemicolon
@@
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
@@
 subgraph~\cite{UpcomingAAAIPaper}.  This benchmark set includes real-world
-graphs, along with graphs generated using several random models.  Pattern
-graphs range from 4 vertices to 900 with a median of 80; target graphs range
-from 10 vertices to 6671 with a median of 561.  Cumulative runtimes including
-top-down algorithm from~\cite{UpcomingAAAIPaper} which was the previous state
-of the art for this benchmark, are in \cref{figure:sip-cumulative}.  This is a
-challenging set of instances, and more than half of the instances cannot be
-solved within a timeout of 1000 seconds by any of the solvers. \McSplit\ is the
-strongest solver overall for these instances.
+graphs and graphs generated using random models.  Pattern graphs range from 4
+vertices to 900 with a median of 80; target graphs range from 10 vertices to
+6671 with a median of 561.  Cumulative runtimes including the kdown algorithm
+from~\cite{UpcomingAAAIPaper} which was the previous state of the art for this
+benchmark, are in \cref{figure:sip-cumulative}.  This is a challenging set of
+instances, and more than half of the instances cannot be solved within a
+timeout of 1000 seconds by any of the solvers. Moreover the CP-FC algorithm and
+the clique encoding run out of memory on many of the instances.  \McSplit\
+is the strongest solver overall for this dataset.
 

### 2017-02-15 paper/paper.tex
More writing

@@
 
-We initially assume that graphs are unlabelled, undirected and without self-loops
+We initially assume that graphs are unlabelled, undirected and without loops
 (Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
-The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.
-The set of vertices adjacent to vertex $v$ in graph $G$ is called the
-\emph{neighbourhood} of $v$, and denoted $\N(v, G)$.
-We use $n_G$ and $n_H$ to denote $|\V(G)|$ and $|\V(H)|$. We will occasionally use
-$n$ in complexity results to denote $n_G + n_H$.  The graph $\overline{G}$ is the complement (aka the inverse) of $G$.
+The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.  The
+set of vertices adjacent to vertex $v$ in graph $G$ is called the
+\emph{neighbourhood} of $v$, and denoted $\N(v, G)$.  The graph $\overline{G}$
+is the complement of $G$, obtained copying the vertices of $G$ and adding an
+edge between each pair of distinct vertices if and only if $G$ does not
+have an edge between the corresponding pair.
+
+Let $G$ and $H$ be the two input graphs to our maximum common subgraph problem.
+The orders (vertex counts) of these graphs are denoted $n_G$ and $n_H$. We will
+use $n$ in complexity results to denote $\max(n_G + n_H)$.
 
@@
 $w_i$ are distinct vertices of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
-in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. It is clear that,
-given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{|M^*|}\}$
+in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. 
+Given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{|M^*|}\}$
 and the subgraph of $H$ induced by $\{w_i, \dots, w_{|M^*|}\}$ are a maximum
@@
 We begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Now label
-each unmatched vertex according to whether it is adjacent to vertex $1$ (in $G$) or
-vertex $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent vertices have label
-1; non-adjacent vertices have label 0.  It is straightforward to see that we
-can extend $M$ with a vertex mapping $vw$, with $v \in \V(G)$ and $w \in
-\V(H)$, if and only if $v$ and $w$ have the same label.  This property, that
-two vertices may be mapped together if and only if they share a label, is the
-algorithm's main invariant.
+each unmatched vertex according to whether it is adjacent to vertex $1$ (in
+$G$) or vertex $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent
+vertices have label 1; non-adjacent vertices have label 0.  We can extend $M$
+with a vertex mapping $vw$, with $v \in \V(G)$ and $w \in \V(H)$, if and only
+if $v$ and $w$ have the same label.  This property, that two vertices may be
+mapped together if and only if they share a label, is the algorithm's main
+invariant.
 
@@
 
-Our invariant is maintained: we can extend $M$ by a vertex pairing if
-and only if the two vertices have the same label. Note also that each label (in Figure~\ref{fig:alg3}) is prefixed
+Our invariant is maintained: we can extend $M$ by a vertex pairing if and only
+if the two vertices have the same label. Note also that the labels can be
+calculated efficiently, since each label (in Figure~\ref{fig:alg3}) is prefixed
 by the label at the previous level of the search tree (Figure~\ref{fig:alg2}).
@@
 since labels may be up to $n_G$ bits in length. We can reduce the memory
-requirement to $\BigO{n}$ per level by storing a \emph{label class} as a pair $\langle P,T \rangle$ for
-each label $l$ that is used, where $P$ is the set of vertices in $\V(G)$
-labelled $l$, and $T$ is the set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology from subgraph isomorphism where one graph is a pattern graph $P$ and the other the target graph $T$.}.
-Since there are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$
-label classes can exist at once, and the total length of the $P$ and $T$ sets
-is at most $n_G + n_H$.  We can gain further efficiency by discarding during
-search any label class that has an empty $P$ or $T$ set.
+requirement to $\BigO{n}$ per level by storing a \emph{label class} as a pair
+$\langle P,T \rangle$ for each label $l$ that is used, where $P$ is the set of
+vertices in $\V(G)$ labelled $l$, and $T$ is the set of vertices in $\V(H)$
+labelled $l$.\footnote{We adopt terminology from subgraph isomorphism where one
+graph is a pattern graph $P$ and the other the target graph $T$.} Since there
+are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$ label classes
+can exist at once, and there are at most $n_G + n_H$ vertices in the union of
+all of the $P$ and $T$ sets.  We can gain further efficiency by discarding
+during search any label class that has an empty $P$ or $T$ set.
 
@@
 \caption{\McSplit\ algorithm for Maximum Common Subgraph}
-\label{jtAlg}
+\label{McCliqueAlg}
 \end{algorithm}
 
-\paragraph{Algorithm \ref{jtAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
+\paragraph{Algorithm \ref{McCliqueAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
 two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
@@
 
-\section{Comparison with Existing Algorithms}\label{sec:comparison}
-The  \McSplit\ algorithm may be considered as a version of the
-constraint programming, forward-checking (CP-FC) algorithm of \cite{DBLP:conf/cp/NdiayeS11}.
-The vertices in $P$ correspond to variables in CP-FC and the 
-vertices in $T$ to the domains of variables in CP-FC. A label class $\langle P,T \rangle$ then corresponds to a set 
-of variables ($P$) in CP-FC  with identical domains ($T$). Where CP-FC uses a soft allDifferent constraint to filter and compute 
-a bound, \McSplit\ computes a bound from each label class in the future, i.e. it performs Hall Set reasoning on the fly.
+\section{Comparison with Existing Algorithms}\label{sec:comparison} The
+\McSplit\ algorithm may be considered as a version of the constraint
+programming, forward-checking (CP-FC) algorithm of
+\citep{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
+of domains.  (Recall that in the CP-FC algorithm, each vertex in $v \in \V(G)$
+is represented by a variable, whose domain corresponds to the set of vertices
+in $\V(H)$ to which $v$ may currently be mapped.)
 
-\McSplit\ takes $\BigO{n}$ time at each search node. Furthermore, \McSplit\ has a low start up cost. The clique encoding requires the
-construction of the association graph prior to search
+Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
+correspond to variables in CP-FC, all of which have domain equal to $T$.  The
+label-class representation of domains is possible because throughout the CP-FC
+algorithm for max common subgraph, the domains of any two variables are either
+identical or disjoint.  (To the best of our knowledge, this observation has
+not been made previously.)
 
-Our algorithm needs less space than CP-FC ($\BigO{n^2}$ rather than $\BigO{n^3}$?) and
-much less than clique ($\BigO{n^4}$?).
+Where CP-FC uses a soft allDifferent constraint to filter and compute a bound,
+\McSplit\ computes a bound from each label class in the future, i.e. it
+performs Hall Set reasoning on the fly.
+
+\McSplit\ requires $\BigO{n}$ time at each search node, and does not require
+any auxiliary data structures larger than the adjacency matrices of $G$ and $H$
+to be built prior to search.  CP-FC requires $\BigO(?????)$ \dots The clique
+encoding requires an association graph with $n_G n_H$ vertices to be
+constructed before search.  Modern maximum-clique algorithms typically perform
+a colouring step at each search node that is quadratic in the order of the
+graph, and thus $\BigO{n^4}$ in the order of the max common subgraph input
+graphs.
 
@@
 
-LARGELY FROM CIARAN'S EMAIL; NEEDS WORK:
+The correctness of \McSplit\ can be seen by noting its equivalence to CP-FC
+\citep{DBLP:conf/cp/NdiayeS11}, if we replace the label-class representation of
+$\mathit{future}$ with domain stores in Algorithm \ref{McCliqueAlg}.  To obtain
+the CP-FC algorithm, replace \linerangeref{InnerLoop}{InnerLoopEnd} of
+Algorithm \ref{McCliqueAlg} with Algorithm \ref{cpAlg}.  The pruning of domains
+carried out in these lines can be easily seen to be equivalent to the
+corresponding lines in Algorithm \ref{McCliqueAlg}.
 
@@
 \DontPrintSemicolon
-\nl    \For {$q \in \N(b,G)$}{
-\nl        remove $t$ from $D'_q$ \;
-\nl        remove the inverse neighbourhood of $t$ from $D'_q$ \;
+\nl    \For {$u \in \N(b,G)$}{
+\nl        remove $v$ from $D'_u$ \;
+\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
        }
-\nl    \For {$q \in \N(b,\overline{G})$}{
-\nl        remove $t$ from $D'_q$ \;
-\nl        remove the neighbourhood of $t$ from $D'_q$ \;
+\nl    \For {$u \in \N(b,\overline{G})$}{
+\nl        remove $v$ from $D'_u$ \;
+\nl        remove the neighbourhood of $t$ from $D'_u$ \;
        }
-\caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} in CP algorithm}
+\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
 \label{cpAlg}
@@
 
-The first difference is in the underlying data structures: James uses a clever
-partitioning data structure, not a conventional domain store. With the
-algorithm described this way, I think the proof that domains are partitioned
-into disjoint sets is almost immediate.
-
-The second difference is in the bound function: the French code constructs a
-maximum cardinality matching, with some special weighting trickery for $\bot$,
-whilst James just sums the smaller of the two sides of each partition. If you
-think a bit about what the matching looks like, you can see that these two give
-the same answer, but at very different speeds.
-
-The third difference is in the details of the variable ordering heuristic.
+TO WRITE PROPERLY: The second difference is in the bound function: the French
+code constructs a maximum cardinality matching, with some special weighting
+trickery for $\bot$, whilst James just sums the smaller of the two sides of
+each partition. If you think a bit about what the matching looks like, you can
+see that these two give the same answer, but at very different speeds.
 
@@
 search node.  First, run lines 17-19
-of Algorithm~\ref{jtAlg} to create a new label-class of vertices that are not
+of Algorithm~\ref{McCliqueAlg} to create a new label-class of vertices that are not
 adjacent to $v$ or $w$, and remove these vertices from $\langle P',T' \rangle$.
@@
 
-\iffalse
-\paragraph{Almost Subgraph Isomorphism} \citet{UpcomingAAAIPaper} present a
-``top-down'' algorithm for Maximum Common Subgraph in which a solution of size
-$n_G$ is initially sought; if such a solution does not exist then this
-solutions of size $n_G-1, n_G-2, \dots$ are sought in sequence.  This strategy,
-when used with techniques based on a state-of-the-art subgraph isomorphism
-solver, outperformed CP and clique encodings on a large benchmark set of graph
-pairs.  We can modify the \McSplit\ algorithm to use the top-down strategy by
-calling the main \FuncSty{McSplit} method once per goal size ($n_G, n_G-1,
-n_G-2, \dots$).  We backtrack (\lineref{PruneSearch} of Algorithm~\ref{jtAlg})
-when the bound is strictly less than the goal size, and terminate when a
-solution of the goal size is found.
-\fi
+\paragraph{Top-Down Strategy} I'VE RE-ADDED A SHORTER VERSION OF THIS
+PARAGRAPH, WHICH WAS COMMENTED OUT BY PATRICK. IT NEEDS WORK, BUT WE NEED TO
+EXPLAIN THE SIP EXPERIMENT. We can modify the \McSplit\ algorithm to use a
+top-down strategy based on \citet{UpcomingAAAIPaper} by calling the main
+\FuncSty{McSplit} method once per goal size ($n_G, n_G-1, n_G-2, \dots$).  We
+backtrack (\lineref{PruneSearch} of Algorithm~\ref{McCliqueAlg}) when the bound is
+strictly less than the goal size, and terminate when a solution of the goal
+size is found.
 
@@
 v2 CPUs and 64GBytes RAM.
+
+? MENTION PROGRAMMING LANGUAGES, COMPILER FLAGS, MCSa1 ALGORITHM
+
 %Parallel experiments used 32 threads per machine,
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
     \centering

### 2017-02-15 paper/paper.tex
s/self-loops/loops/g

@@
 \section{Extensions}\label{sec:extensions}
-\paragraph{Vertex labels and self-loops} If vertices are labelled, replace ${\langle
+\paragraph{Vertex labels and loops} If vertices are labelled, replace ${\langle
 V(G),V(H) \rangle}$ in \lineref{FirstExpandCall} with a set of label classes, one for each label that appears in both $\V(G)$ and
-$\V(H)$. If vertex labels are represented as positive integers, represent self-loops
+$\V(H)$. If vertex labels are represented as positive integers, represent loops
 as negative integers.
@@
 magnitude faster than the previous state of the art for unlabelled and undirected
-instances. We have shown how the algorithm can be extended for graphs with labels on edges, labels on vertices, self-loops, directed edges and the requirement that the resultant graph be connected.
+instances. We have shown how the algorithm can be extended for graphs with labels on edges, labels on vertices, loops, directed edges and the requirement that the resultant graph be connected.
 

### 2017-02-15 paper/paper.tex
Minor edits

@@
 \section{Extensions}\label{sec:extensions}
+
 \paragraph{Vertex labels and loops} If vertices are labelled, replace ${\langle
-V(G),V(H) \rangle}$ in \lineref{FirstExpandCall} with a set of label classes, one for each label that appears in both $\V(G)$ and
-$\V(H)$. If vertex labels are represented as positive integers, represent loops
-as negative integers.
+V(G),V(H) \rangle}$ in \lineref{FirstExpandCall} with a set of label classes,
+one for each label that appears on at least one vertex of both $G$ and $H$.
+Loops may be treated as a label modifier: assuming the original vertex labels
+are positive integers, we replace the label $l$ with $-l$ on each vertex that
+has a loop.
 
@@
 \linerangeref{NewPWithEdge}{InnerLoopEnd} of the basic algorithm split the
-label class $\langle P',T' \rangle$ according to whether vertices are adjacent
-to $v$ (resp. $w$), we now perform a four-way split where each vertex is
-classified according to the label on its adjacency matrix from $v$ (resp. $w$).
-This is shown in Algorithm~\ref{labDirAlg}, where $L=\{0,1,2,3\}$.
+label class $\langle P',T' \rangle$ in two, we now perform a four-way split
+where each vertex is classified according to the label on its adjacency matrix
+entry from $v$ or $w$.  This is shown in Algorithm~\ref{labDirAlg}, where
+$L=\{0,1,2,3\}$.
 
@@
 
-\paragraph{Maximum Common \emph{Connected} Subgraph} 
-The common subgraph must be connected. We consider only undirected graphs. 
-Modify \McSplit\ by permitting branching only on a vertex $v$ that has at
-least one non-zero element in its bit-string label.
+\paragraph{Maximum Common \emph{Connected} Subgraph} The common subgraph must
+be connected. We consider only undirected graphs.  Modify \McSplit\ by
+permitting branching only on a vertex $v$ that has at least one non-zero
+element in its bit-string label.  We can represent this information compactly,
+and without increasing time complexity at each search node, by storing an extra
+bit with each label class.  This bit takes the value $1$ if and only if the vertices in
+the class are adjacent to at least one vertex in $M$.
 
@@
 
-? MENTION PROGRAMMING LANGUAGES, COMPILER FLAGS, MCSa1 ALGORITHM
+? MENTION PROGRAMMING LANGUAGES, COMPILER FLAGS, MCSa ALGORITHM AND
+BITSET IMPLEMENTATION
 
@@
 %using all 16 physical cores with hyper-threading enabled.
-All experiments used a database of
+All ??? experiments used a database of
 randomly-generated Maximum Common Subgraph instances, which contains graphs of

### 2017-02-15 paper/paper.tex
New abstract

@@
 \begin{abstract}
-    Maximium Common Subgraph (MCS) is the problem of finding a
-    graph of maximium order that is an induced subgraph of each of the two input
-    graphs.  We introduce \McSplit, a branch and bound algorithm that
-    uses a vertex-labelling scheme to compactly represent the subsets of
-    vertices that may be matched together to create a larger common subgraph.
-    The new algorithm explores the same search space as the existing CP-FC
-    algorithm, but requires only \BigO{n^2} space in total and \BigO{n} time
-    for each recursive call, greatly improving on all existing algorithms for
-    the problem. Experimental results using two large graph databases show that \McSplit\ is
-    more than an order of magnitude faster than the leading existing algorithms
-    on undirected, unlabelled graphs, including the problem
-    variant in which the subgraph must be connected.
+    We introduce a new branch and bound algorithm for the maximum common
+    subgraph and maximum common connected subgraph problems which is based
+    around vertex labelling and partitioning. Our method in some ways resembles
+    a traditional constraint programming approach, but uses a novel compact
+    domain store and supporting inference algorithms which dramatically reduce
+    the memory and computation requirements during search, and allow better
+    dual viewpoint ordering heuristics to be calculated cheaply.  Experiments
+    show a speedup of more than an order of magnitude over the state of the
+    art, and demonstate that we can operate on much larger graphs without
+    running out of memory.
 \end{abstract}

### 2017-02-15 paper/paper.tex
Typo in abstract

@@
     show a speedup of more than an order of magnitude over the state of the
-    art, and demonstate that we can operate on much larger graphs without
+    art, and demonstrate that we can operate on much larger graphs without
     running out of memory.

### 2017-02-15 paper/paper.tex
Added Problems to paper title

@@
 
-\title{A Partitioning Algorithm for Maximum Common Subgraphs\thanks{This work
+\title{A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work
 was supported by the Engineering and Physical Sciences Research Council [grant

### 2017-02-15 paper/paper.tex
Anonymousify

--- 2017-02-15 paper/paper.tex
@@
 
-\title{A Partitioning Algorithm for Maximum Common Subgraphs\thanks{This work
-was supported by the Engineering and Physical Sciences Research Council [grant
-numbers EP/K503058/1 and EP/M508056/1]}}
-\author{Ciaran McCreesh \and Patrick Prosser \and James Trimble \\
-University of Glasgow, Glasgow, Scotland \\
-j.trimble.1@research.gla.ac.uk}
+\title{A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work
+was supported by a pile of money from a funding agency.}}
+\author{The Anonymous Authors of Paper 2431 \\
+Academy of Anonymous Academics, Some City, Some Country \\
+anonymous@anonymous.anonymous}
 

### 2017-02-15 paper/paper.tex
Some things that changed

@@
 \DeclareMathOperator{\N}{N}
+\DeclareMathOperator{\invN}{\overline{N}}
 \DeclareMathOperator{\vtxlabel}{label}
@@
 
-\section{Preliminaries}
+\section{The \McSplit\ algorithm}
 
@@
 set of vertices adjacent to vertex $v$ in graph $G$ is called the
-\emph{neighbourhood} of $v$, and denoted $\N(v, G)$.  The graph $\overline{G}$
-is the complement of $G$, obtained copying the vertices of $G$ and adding an
-edge between each pair of distinct vertices if and only if $G$ does not
-have an edge between the corresponding pair.
+\emph{neighbourhood} of $v$, denoted $\N(G, v)$. We denote by $\invN(G, v)$ the
+\emph{inverse neighbourhood} of $v$, being the set of the vertices not adjacent
+to $v$ (excluding $v$ itself).
 
@@
 
-\section{Prior Work}
-
-The most successful existing approaches to solving MCS are reformulation to the
-maximum clique problem on one hand \cite{DBLP:conf/cp/McCreeshNPS16}, and
-constraint programming on the other \cite{DBLP:conf/cp/NdiayeS11}. In the
-clique encoding, an association graph is created with one node for each pair in
-the set $\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and an edge for each pair of
-mappings that may be chosen together. A maximum clique in this graph
-corresponds directly to a maximum common subgraph of $G$ and $H$.
-
-In the constraint programming approach  \cite{DBLP:conf/cp/NdiayeS11}, a variable corresponds to
-each vertex in $G$, and each variable initially has one value in its domain for
-each vertex in $H$ to which it may be mapped, plus an additional value $\bot$
-signifying that the vertex remains unmapped. A backtracking search is performed
-in which values are assigned to variables, and propagation rules are used to
-remove infeasible values from domains.
-\textit{A matching algorithm propagates
-a soft global allDifferent constraint which maximizes the number variables that are assigned to values different from $\bot$, while ensuring they are all different when they are not assigned to $\bot$.
-Different constraint propagation techniques are compared in \cite{DBLP:conf/cp/NdiayeS11}.
-The combination ``MAC+Bound'' (resp.
-``FC+Bound'') obtains the best results on labelled (resp.
-non labelled) graphs and outperforms the approach of \cite{McGreg82} and the CP model of \cite{DBLP:conf/mco/VismaraV08}.
-The combination ``MAC+Bound'' maintains arc consistency~\cite{sabi94} of edge constraints, whereas the conbination ``FC+Bound'' simply performs forward-checking on these constraints.
-In both combinations, the ``Bound'' filtering checks whether it is possible to assign distinct values to enough variables to surpass the best cost found so far (it is a weaker version of GAC(\emph{softAllDiff}) \cite{peti01} which computes the maximum number of variables that can be assigned distinct values).}
-
-\section{The \McSplit\ algorithm}
-
-We now present \McSplit, a branch and bound algorithm 
-which uses a new vertex-labelling scheme to avoid explicitly maintaining the
-set of feasible mappings for each vertex in $\V(G)$. 
+With these definitions established, we now present \McSplit, a branch and bound
+algorithm which uses a new vertex-labelling scheme to avoid explicitly
+maintaining the set of feasible mappings for each vertex in $\V(G)$.
 \iffalse
@@
 \nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
-\nl        \bf{Let} $P'' \gets P' \cap N(v,G) \setminus \{v\}$ \label{NewPWithEdge} \;
-\nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
+\nl        \bf{Let} $P'' \gets P' \cap \N(G, G) \setminus \{v\}$ \label{NewPWithEdge} \;
+\nl        \bf{Let} $T'' \gets T' \cap \N(H, w) \setminus \{w\}$ \;
 \nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
         {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
-\nl        \bf{Let} $P'' \gets P' \cap N(v,\overline{G}) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
-\nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
+\nl        \bf{Let} $P'' \gets P' \cap \invN(G, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
+\nl        \bf{Let} $T'' \gets T' \cap \invN(H, w) \setminus \{w\}$ \;
 \nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
@@
 \citep{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
-of domains.  (Recall that in the CP-FC algorithm, each vertex in $v \in \V(G)$
+of domains.  In the CP-FC algorithm, each vertex in $v \in \V(G)$
 is represented by a variable, whose domain corresponds to the set of vertices
-in $\V(H)$ to which $v$ may currently be mapped.)
+in $\V(H)$ to which $v$ may currently be mapped.
 
@@
 \DontPrintSemicolon
-\nl    \For {$u \in \N(b,G)$}{
+\nl    \For {$u \in \N(G, b)$}{
 \nl        remove $v$ from $D'_u$ \;
@@
        }
-\nl    \For {$u \in \N(b,\overline{G})$}{
+\nl    \For {$u \in \invN(G, b)$}{
 \nl        remove $v$ from $D'_u$ \;

### 2017-02-15 paper/paper.tex
Notation fight

@@
 Let $G$ and $H$ be the two input graphs to our maximum common subgraph problem.
-The orders (vertex counts) of these graphs are denoted $n_G$ and $n_H$. We will
-use $n$ in complexity results to denote $\max(n_G + n_H)$.
+The orders (vertex counts) of these graphs are denoted $g$ and $h$. We will
+use $n$ in complexity results to denote $\max(g + h)$ ?? WTF James? Help?.
 
@@
 described so far requires $\BigO{n^2}$ space at each level of the search tree,
-since labels may be up to $n_G$ bits in length. We can reduce the memory
+since labels may be up to $g$ bits in length. We can reduce the memory
 requirement to $\BigO{n}$ per level by storing a \emph{label class} as a pair
@@
 graph is a pattern graph $P$ and the other the target graph $T$.} Since there
-are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$ label classes
-can exist at once, and there are at most $n_G + n_H$ vertices in the union of
+are $g + h$ vertices in the two graphs, at most $g + h$ label classes
+can exist at once, and there are at most $g + h$ vertices in the union of
 all of the $P$ and $T$ sets.  We can gain further efficiency by discarding
@@
 to be built prior to search.  CP-FC requires $\BigO(?????)$ \dots The clique
-encoding requires an association graph with $n_G n_H$ vertices to be
+encoding requires an association graph with $g h$ vertices to be
 constructed before search.  Modern maximum-clique algorithms typically perform
@@
 the union of $\{0\}$ with the set of all labels that appear in the input
-graphs. Since there may be up to $n_G + n_H$ distinct labels, the loop in
-Algorithm~\ref{labDirAlg} may execute up to $n_G + n_H$ times and
+graphs. Since there may be up to $g + h$ distinct labels, the loop in
+Algorithm~\ref{labDirAlg} may execute up to $g + h$ times and
 requires $\BigO{n^2}$ time per search
@@
 top-down strategy based on \citet{UpcomingAAAIPaper} by calling the main
-\FuncSty{McSplit} method once per goal size ($n_G, n_G-1, n_G-2, \dots$).  We
+\FuncSty{McSplit} method once per goal size ($g, g-1, g-2, \dots$).  We
 backtrack (\lineref{PruneSearch} of Algorithm~\ref{McCliqueAlg}) when the bound is

### 2017-02-16 paper/paper.tex
Define

@@
 \emph{inverse neighbourhood} of $v$, being the set of the vertices not adjacent
-to $v$ (excluding $v$ itself).
+to $v$ (excluding $v$ itself). A \emph{subgraph} of a graph $G$ is a graph
+consisting of some of the vertices of $G$, and all of the edges between these
+vertices (in this paper all subgraphs are induced). A \emph{common subgraph} of
+two graphs is a graph which is (isomorphic to) a subgraph of two graphs
+simultaneously, and a \emph{maximum} common subgraph is one with as many
+vertices as possible.
 
@@
 \nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
-\nl        \bf{Let} $P'' \gets P' \cap \N(G, G) \setminus \{v\}$ \label{NewPWithEdge} \;
+\nl        \bf{Let} $P'' \gets P' \cap \N(G, v) \setminus \{v\}$ \label{NewPWithEdge} \;
 \nl        \bf{Let} $T'' \gets T' \cap \N(H, w) \setminus \{w\}$ \;

### 2017-02-16 paper/paper.tex
Fixed a typo in the algorithm
Avoid superscripts on subscripts, this isn't an IP paper

@@
 \McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
-(v_{|M^*|}, w_{|M^*|})\}$, where the $v_i$ are distinct members of $\V(G)$ and the
+(v_{m}, w_{m})\}$ with $|M^*| = m$ vertex pairs, where the $v_i$ are distinct members of $\V(G)$ and the
 $w_i$ are distinct vertices of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
 in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. 
-Given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{|M^*|}\}$
-and the subgraph of $H$ induced by $\{w_i, \dots, w_{|M^*|}\}$ are a maximum
+Given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{m}\}$
+and the subgraph of $H$ induced by $\{w_i, \dots, w_{m}\}$ are a maximum
 common subgraph.

### 2017-02-16 paper/paper.tex
Avoid big spaces between words
cref

@@
 \newcommand{\nmax}{n_{\max}}
+
+% cref style
+\crefname{algorithm}{Algorithm}{Algorithms}
+\Crefname{algorithm}{Algorithm}{Algorithms}
+\crefname{figure}{Figure}{Figures}
+\Crefname{figure}{Figure}{Figures}
+\crefname{table}{Table}{Tables}
+\Crefname{table}{Table}{Tables}
+\crefname{section}{Section}{Sections}
+\Crefname{section}{Section}{Sections}
 
@@
 We initially assume that graphs are unlabelled, undirected and without loops
-(Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
+(\cref{sec:extensions} describes how these restrictions may be relaxed).
 The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.  The
@@
 As we will explore in
-Section~\ref{sec:comparison}, the algorithm may be viewed as a version of the
+\cref{sec:comparison}, the algorithm may be viewed as a version of the
 constraint programming forward-checking (CP-FC) algorithm using a compact data
@@
 
-We will use the graphs $G$ and $H$ in Figure~\ref{fig:alg1} to demonstrate the
+We will use the graphs $G$ and $H$ in \cref{fig:alg1} to demonstrate the
 algorithm.  These graphs have a maximum common subgraph with four vertices; one
@@
 each unmatched vertex according to whether it is adjacent to vertex $1$ (in
-$G$) or vertex $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent
+$G$) or vertex $a$ (in $H$), as shown in \cref{fig:alg2}.  Adjacent
 vertices have label 1; non-adjacent vertices have label 0.  We can extend $M$
@@
 \begin{minipage}[t]{0.3\linewidth}
-    Labelling of $G$
+    \centering Labelling of $G$
     \begin{tabular}[t]{cc}
@@
 \begin{minipage}[t]{0.3\linewidth}
-    Labelling of $H$
+    \centering Labelling of $H$
     \begin{tabular}[t]{cc}
@@
 same label; we will choose to map vertex $2$ to vertex $d$, giving $M=\{1a,
-2d\}$ (Figure~\ref{fig:alg3}).  Each unmapped vertex $v \in \V(G)$ is labelled
+2d\}$ (\cref{fig:alg3}).  Each unmapped vertex $v \in \V(G)$ is labelled
 with a two-character bit string, indicating its adjacency to each of
@@
 if the two vertices have the same label. Note also that the labels can be
-calculated efficiently, since each label (in Figure~\ref{fig:alg3}) is prefixed
-by the label at the previous level of the search tree (Figure~\ref{fig:alg2}).
+calculated efficiently, since each label (in \cref{fig:alg3}) is prefixed
+by the label at the previous level of the search tree (\cref{fig:alg2}).
 
@@
 this bound is calculated, we consider the situation one level deeper in the
-search tree shown in Figure~\ref{fig:alg4}.
+search tree shown in \cref{fig:alg4}.
 
@@
 
-\paragraph{Algorithm \ref{McCliqueAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
+\paragraph{\cref{McCliqueAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
 two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
@@
 \citep{DBLP:conf/cp/NdiayeS11}, if we replace the label-class representation of
-$\mathit{future}$ with domain stores in Algorithm \ref{McCliqueAlg}.  To obtain
+$\mathit{future}$ with domain stores in \cref{McCliqueAlg}.  To obtain
 the CP-FC algorithm, replace \linerangeref{InnerLoop}{InnerLoopEnd} of
-Algorithm \ref{McCliqueAlg} with Algorithm \ref{cpAlg}.  The pruning of domains
+\cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains
 carried out in these lines can be easily seen to be equivalent to the
-corresponding lines in Algorithm \ref{McCliqueAlg}.
+corresponding lines in \cref{McCliqueAlg}.
 
@@
 where each vertex is classified according to the label on its adjacency matrix
-entry from $v$ or $w$.  This is shown in Algorithm~\ref{labDirAlg}, where
+entry from $v$ or $w$.  This is shown in \cref{labDirAlg}, where
 $L=\{0,1,2,3\}$.
@@
 entry contains an edge label, or a null entry (non-adjacent) $0$.
-We use Algorithm~\ref{labDirAlg}, by letting $L$ be
+We use \cref{labDirAlg}, by letting $L$ be
 the union of $\{0\}$ with the set of all labels that appear in the input
 graphs. Since there may be up to $g + h$ distinct labels, the loop in
-Algorithm~\ref{labDirAlg} may execute up to $g + h$ times and
+\cref{labDirAlg} may execute up to $g + h$ times and
 requires $\BigO{n^2}$ time per search
@@
 search node.  First, run lines 17-19
-of Algorithm~\ref{McCliqueAlg} to create a new label-class of vertices that are not
+of \cref{McCliqueAlg} to create a new label-class of vertices that are not
 adjacent to $v$ or $w$, and remove these vertices from $\langle P',T' \rangle$.
@@
 \FuncSty{McSplit} method once per goal size ($g, g-1, g-2, \dots$).  We
-backtrack (\lineref{PruneSearch} of Algorithm~\ref{McCliqueAlg}) when the bound is
+backtrack (\lineref{PruneSearch} of \cref{McCliqueAlg}) when the bound is
 strictly less than the goal size, and terminate when a solution of the goal

### 2017-02-16 paper/paper.tex
Cut for space, we do better later anyway
ie

@@
 to vertex $2$.  Labels are given to unmapped vertices in $\V(H)$ in a similar fashion,
-showing adjacency to matched vertices $a$ and $d$.
-
-Our invariant is maintained: we can extend $M$ by a vertex pairing if and only
-if the two vertices have the same label. Note also that the labels can be
-calculated efficiently, since each label (in \cref{fig:alg3}) is prefixed
-by the label at the previous level of the search tree (\cref{fig:alg2}).
+showing adjacency to matched vertices $a$ and $d$.  Our invariant is
+maintained: we can extend $M$ by a vertex pairing if and only if the two
+vertices have the same label.
 
@@
 
-The algorithm backtracks when the incumbent, i.e. the largest matching found so far, is at least as large
+The algorithm backtracks when the incumbent (the largest matching found so far) is at least as large
 as a calculated bound given $M$ and the current labelling. To demonstrate how
@@
 Where CP-FC uses a soft allDifferent constraint to filter and compute a bound,
-\McSplit\ computes a bound from each label class in the future, i.e. it
+\McSplit\ computes a bound from each label class in the future, i.e.\ it
 performs Hall Set reasoning on the fly.
@@
 
-We have yet to fully investigate variable and value ordering heuristics, i.e. the selection of a label class and the selection of a vertex from that label class.
+We have yet to fully investigate variable and value ordering heuristics, i.e.\ the selection of a label class and the selection of a vertex from that label class.
 There is also the potential to branch on both sides, that is instead of branching on vertices in $\V(G)$, it would be equally valid to branch on

### 2017-02-16 paper/paper.tex
Fix silly line spacing
Punctuation

@@
 \begin{multline*}
-    \mathit{bound} = |M| + \sum_{l \in L} \min\big(|\{ v \in \V(G) : \vtxlabel(v)=l\}|, \\
+    \mathit{bound} = |M| + \sum_{l \in L} \min\big(|\{ v \in \V(G) : \vtxlabel(v)=l\}|, \\[-0.2cm]
         |\{ v \in \V(H) : \vtxlabel(v)=l \}|\big)
@@
 vertices in $\V(G)$ labelled $l$, and $T$ is the set of vertices in $\V(H)$
-labelled $l$.\footnote{We adopt terminology from subgraph isomorphism where one
-graph is a pattern graph $P$ and the other the target graph $T$.} Since there
+labelled $l$\footnote{We adopt terminology from subgraph isomorphism, where one
+graph is a pattern graph $P$ and the other the target graph $T$.}. Since there
 are $g + h$ vertices in the two graphs, at most $g + h$ label classes

### 2017-02-16 paper/paper.tex
I hate algorithm2e
microtype makes this look better
newcommand for algorithm example tables
Note to self. Or James. One of the two
Phrasing

@@
 \usepackage{cleveref}
+\usepackage{microtype}
 
@@
 \Crefname{algorithm}{Algorithm}{Algorithms}
+\crefname{algocf}{Algorithm}{Algorithms}
+\Crefname{algocf}{Algorithm}{Algorithms}
 \crefname{figure}{Figure}{Figures}
@@
 %}
+
+\newcommand{\LabelTables}[4] {
+  \centering
+  \begin{minipage}[t]{#1\linewidth}
+      Mapping
+
+      \bigskip
+
+      #2
+  \end{minipage}
+  \quad
+  \begin{minipage}[t]{0.3\linewidth}
+      \centering Labelling of $G$
+      \begin{tabular}[t]{cc}
+      \hline
+          Vertex & Label\\
+      \hline
+          #3
+      \hline
+      \end{tabular}
+  \end{minipage}
+  \quad
+  \begin{minipage}[t]{0.3\linewidth}
+      \centering Labelling of $H$
+      \begin{tabular}[t]{cc}
+      \hline
+          Vertex & Label\\
+      \hline
+          #4
+      \hline
+      \end{tabular}
+  \end{minipage}
+}
 
@@
 maintaining the set of feasible mappings for each vertex in $\V(G)$.
-\iffalse
-As we will explore in
-\cref{sec:comparison}, the algorithm may be viewed as a version of the
-constraint programming forward-checking (CP-FC) algorithm using a compact data
-structure that exploits the nature of the MCS problem.  It is sufficiently
-different, however, that it is helpful to describe the algorithm without
-reference to previous approaches.
 
-%At each node of the search tree, the algorithm
-%labels vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in
-%\V(G)$ and $w \in \V(H)$ can be added to the mapping if and only if $v$ and $w$
-%have the same label.
-\fi
 \McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
@@
 
-We will use the graphs $G$ and $H$ in \cref{fig:alg1} to demonstrate the
-algorithm.  These graphs have a maximum common subgraph with four vertices; one
-example is the mapping $\{1a, 2f, 3d, 5b\}$ where vertex $1$ is assigned to
-vertex $a$, $2$ is assigned to $f$, $3$ to $d$ and $5$ to $b$.
+Before discussing the algorithm in detail, we start with an illustration of the
+main concepts using the graphs $G$ and $H$ in \cref{fig:alg1}.  These graphs
+have a maximum common subgraph with four vertices; one example is the mapping
+$\{1a, 2f, 3d, 5b\}$ where vertex $1$ is assigned to vertex $a$, $2$ is
+assigned to $f$, $3$ to $d$ and $5$ to $b$.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
     \centering
-\begin{minipage}[t]{0.15\linewidth}
-    Mapping
-
-    \bigskip
-
-    $\{1a\}$
-\end{minipage}
-\quad
-\begin{minipage}[t]{0.3\linewidth}
-    \centering Labelling of $G$
-    \begin{tabular}[t]{cc}
-    \hline
-        Vertex & Label\\
-    \hline
-        $2$ & 0 \\
+    \subfigure[][After mapping $1$ to $a$] {
+      \LabelTables{.15}
+                  {$\{1a\}$}
+                  {$2$ & 0 \\
                    $3$ & 0 \\
                    $4$ & 1 \\
-        $5$ & 1 \\
-    \hline
-    \end{tabular}
-\end{minipage}
-\quad
-\begin{minipage}[t]{0.3\linewidth}
-    \centering Labelling of $H$
-    \begin{tabular}[t]{cc}
-    \hline
-        Vertex & Label\\
-    \hline
-        $b$ & 1 \\
+                   $5$ & 1 \\}
+                  {$b$ & 1 \\
                    $c$ & 1 \\
@@
                    $e$ & 1 \\
-        $f$ & 0 \\
-    \hline
-    \end{tabular}
-\end{minipage}
-    \caption{Labels on vertices after mapping $1$ to $a$}
+                   $f$ & 0 \\}
       \label{fig:alg2}
+    }
+
+    \subfigure[][After mapping $2$ to $d$] {
+      \LabelTables{.15}
+                  {$\{1a,2d\}$}
+                  {$3$ & 01 \\
+                   $4$ & 10 \\
+                   $5$ & 11 \\}
+                  {$b$ & 11 \\
+                   $c$ & 11 \\
+                   $e$ & 10 \\
+                   $f$ & 01 \\}
+      \label{fig:alg3}
+    }
+
+    \subfigure[][After mapping $3$ to $f$] {
+      \LabelTables{.2}
+                  {$\{1a,2d,3f\}$}
+                  {$4$ & 100 \\
+                   $5$ & 111 \\}
+                  {$b$ & 111 \\
+                   $c$ & 111 \\
+                   $e$ & 101 \\}
+      \label{fig:alg4}
+    }
+
+    \caption{Mapping $M$ and vertex labels during search on example graphs $G$ and $H$}
+    \label{figure:mcsplit-examples}
+\end{figure}
+
+\begin{figure}[ht]
 \end{figure}
@@
 
-\begin{figure}[ht]
-\centering
-\begin{minipage}[t]{0.15\linewidth}
-    Mapping
-
-    \bigskip
-
-    $\{1a, 2d\}$
-\end{minipage}
-\quad
-\begin{minipage}[t]{0.3\linewidth}
-    \centering Labelling of $G$
-    \begin{tabular}[t]{cc}
-    \hline
-        Vertex & Label\\
-    \hline
-        $3$ & 01 \\
-        $4$ & 10 \\
-        $5$ & 11 \\
-    \hline
-    \end{tabular}
-\end{minipage}
-\quad
-\begin{minipage}[t]{0.3\linewidth}
-    \centering Labelling of $H$
-    \begin{tabular}[t]{cc}
-    \hline
-        Vertex & Label\\
-    \hline
-        $b$ & 11 \\
-        $c$ & 11 \\
-        $e$ & 10 \\
-        $f$ & 01 \\
-    \hline
-    \end{tabular}
-\end{minipage}
-\caption{After mapping $2$ to $d$}
-\label{fig:alg3}
-\end{figure}
-
 The algorithm backtracks when the incumbent (the largest matching found so far) is at least as large
@@
 search tree shown in \cref{fig:alg4}.
-
-\begin{figure}[ht]
-\centering
-\begin{minipage}[t]{0.2\linewidth}
-    Mapping
-
-    \bigskip
-
-    $\{1a, 2d, 3f\}$
-\end{minipage}
-\quad
-\begin{minipage}[t]{0.3\linewidth}
-    \centering Labelling of $G$
-    \begin{tabular}[t]{cc}
-    \hline
-        Vertex & Label\\
-    \hline
-        $4$ & 100 \\
-        $5$ & 111 \\
-    \hline
-    \end{tabular}
-\end{minipage}
-\quad
-\begin{minipage}[t]{0.3\linewidth}
-    \centering Labelling of $H$
-    \begin{tabular}[t]{cc}
-    \hline
-        Vertex & Label\\
-    \hline
-        $b$ & 111 \\
-        $c$ & 111 \\
-        $e$ & 101 \\
-    \hline
-    \end{tabular}
-\end{minipage}
-\caption{After mapping $3$ to $f$}
-\label{fig:alg4}
-\end{figure}
 
@@
 during search any label class that has an empty $P$ or $T$ set.
+
+?? Talk about how we can use a pair of flat arrays and indices to the start of
+label classes, and throw in citation to partition backtracking here.
 

### 2017-02-16 paper/paper.tex
Subfigured algorithm example tables

@@
 \usepackage{cleveref}
-\usepackage{microtype}
 
@@
 \Crefname{algorithm}{Algorithm}{Algorithms}
-\crefname{algocf}{Algorithm}{Algorithms}
-\Crefname{algocf}{Algorithm}{Algorithms}
 \crefname{figure}{Figure}{Figures}
@@
 maintaining the set of feasible mappings for each vertex in $\V(G)$.
+\iffalse
+As we will explore in
+\cref{sec:comparison}, the algorithm may be viewed as a version of the
+constraint programming forward-checking (CP-FC) algorithm using a compact data
+structure that exploits the nature of the MCS problem.  It is sufficiently
+different, however, that it is helpful to describe the algorithm without
+reference to previous approaches.
 
+%At each node of the search tree, the algorithm
+%labels vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in
+%\V(G)$ and $w \in \V(H)$ can be added to the mapping if and only if $v$ and $w$
+%have the same label.
+\fi
 \McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
@@
 
-Before discussing the algorithm in detail, we start with an illustration of the
-main concepts using the graphs $G$ and $H$ in \cref{fig:alg1}.  These graphs
-have a maximum common subgraph with four vertices; one example is the mapping
-$\{1a, 2f, 3d, 5b\}$ where vertex $1$ is assigned to vertex $a$, $2$ is
-assigned to $f$, $3$ to $d$ and $5$ to $b$.
+We will use the graphs $G$ and $H$ in \cref{fig:alg1} to demonstrate the
+algorithm.  These graphs have a maximum common subgraph with four vertices; one
+example is the mapping $\{1a, 2f, 3d, 5b\}$ where vertex $1$ is assigned to
+vertex $a$, $2$ is assigned to $f$, $3$ to $d$ and $5$ to $b$.
 
@@
 during search any label class that has an empty $P$ or $T$ set.
-
-?? Talk about how we can use a pair of flat arrays and indices to the start of
-label classes, and throw in citation to partition backtracking here.
 

### 2017-02-16 paper/paper.tex
Muck around with spacing a bit
Tweak figure
Clean up a bit
Notation consistency
Lies
We've not spoken about null

--- 2017-02-16 paper/paper.tex
@@
       \end{tabular}
+      \medskip
   \end{minipage}
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
@@
 %    \examplexH{}{}{}{}{}{}
-\caption{Graphs $G$ and $H$}
+\caption{Example graphs $G$ and $H$.}
 \label{fig:alg1}
@@
 \nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
-\bigskip
+\medskip
 \nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
@@
 \nl \For {$w \in T$ \label{WLoop}} {
-\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
+    \nl    $M \gets M \cup \{(v,w)\}$ \label{GrowM} \;
 \nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
@@
 \nl        \bf{Let} $T'' \gets T' \cap \N(H, w) \setminus \{w\}$ \;
-\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
-        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
+\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
+\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
 \nl        \bf{Let} $P'' \gets P' \cap \invN(G, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
 \nl        \bf{Let} $T'' \gets T' \cap \invN(H, w) \setminus \{w\}$ \;
-\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
-        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
+\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
+\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
        }
@@
 The remainder of $\FuncSty{expand}$ performs the search.  A label class
-$\langle P, T \rangle$ is selected from $\AlgVar{future}$, where $P \subseteq V(G)$ and $T \subseteq V(H)$,
-using some heuristic (\lineref{SelectClass}).  From this label class, a
-vertex $v$ is selected from $P$ using a heuristic (\lineref{SelectVertex}). We now
-iterate over all possible vertices in $T$ (\linerangeref{WLoop}{ShrinkM}). A
-vertex $w \in T$ is selected and mapping $M$ is extended (\lineref{GrowM}). A
-new set of label-classes, $\AlgVar{future'}$, is created (\lineref{NewFuture}).
-Every label-class in $\AlgVar{future}$ can now be split
-(\linerangeref{InnerLoop}{InnerLoopEnd}) into two new classes. The first of
-these classes (\linerangeref{NewPWithEdge}{AddToFutureWithEdge}) contains
-vertices in $P$ adjacent to $v$ and vertices in $T$ adjacent to $w$.  This is
-added to $\AlgVar{future'}$ if both sets contain at least one vertex.  This is
-then repeated symmetrically for non-adjacency
-(\linerangeref{NewPWithoutEdge}{InnerLoopEnd}) using the compliments of the
-graphs.  A recursive call is then made (\lineref{ExpandWithV}), on return from
-which we remove the mapping $(v,w)$.  Having explored all possible mappings of
-$v$ with vertices in $T$ we now consider what happens if $v$ is not matched
-(\linerangeref{RemoveV}{ExpandWithoutV}). Note that removing $v$ from its label
-class is equivalent to matching vertex $v$ to the null vertex $\bot$.
+$\langle P, T \rangle$ is selected from $\AlgVar{future}$, where $P \subseteq
+V(G)$ and $T \subseteq V(H)$, using some heuristic (\lineref{SelectClass}).
+From this label class, a vertex $v$ is selected from $P$ using a heuristic
+(\lineref{SelectVertex}). We now iterate over all possible vertices in $T$
+(\linerangeref{WLoop}{ShrinkM}). A vertex $w \in T$ is selected and mapping $M$
+is extended (\lineref{GrowM}). A new set of label-classes, $\AlgVar{future'}$,
+is created (\lineref{NewFuture}).  Every label-class in $\AlgVar{future}$ can
+now be split (\linerangeref{InnerLoop}{InnerLoopEnd}) into two new classes. The
+first of these classes (\linerangeref{NewPWithEdge}{AddToFutureWithEdge})
+contains vertices in $P$ adjacent to $v$ and vertices in $T$ adjacent to $w$.
+This is added to $\AlgVar{future'}$ if both sets contain at least one vertex.
+This is then repeated symmetrically for non-adjacency
+(\linerangeref{NewPWithoutEdge}{InnerLoopEnd}). A recursive call is then made
+(\lineref{ExpandWithV}), on return from which we remove the mapping $(v,w)$.
+Having explored all possible mappings of $v$ with vertices in $T$ we now
+consider what happens if $v$ is not matched
+(\linerangeref{RemoveV}{ExpandWithoutV}).
 

### 2017-02-16 paper/paper.tex
A radical reordering

@@
 
-\iffalse
-\paragraph{Heuristics} We ran a small experiment comparing lots of heuristics
-such as \dots on random instances. The one where you choose a label class with
-as small a max(left size, right size) as possible seemed to work best. ? Why?
-\fi
+\subsection{Heuristics}
 
-\section{Comparison with Existing Algorithms}\label{sec:comparison} The
-\McSplit\ algorithm may be considered as a version of the constraint
-programming, forward-checking (CP-FC) algorithm of
-\citep{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
-of domains.  In the CP-FC algorithm, each vertex in $v \in \V(G)$
-is represented by a variable, whose domain corresponds to the set of vertices
-in $\V(H)$ to which $v$ may currently be mapped.
+?? Heuristics go here, defer justification until \cref{sec:comparison}
 
-Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
-correspond to variables in CP-FC, all of which have domain equal to $T$.  The
-label-class representation of domains is possible because throughout the CP-FC
-algorithm for max common subgraph, the domains of any two variables are either
-identical or disjoint.  (To the best of our knowledge, this observation has
-not been made previously.)
-
-Where CP-FC uses a soft allDifferent constraint to filter and compute a bound,
-\McSplit\ computes a bound from each label class in the future, i.e.\ it
-performs Hall Set reasoning on the fly.
-
-\McSplit\ requires $\BigO{n}$ time at each search node, and does not require
-any auxiliary data structures larger than the adjacency matrices of $G$ and $H$
-to be built prior to search.  CP-FC requires $\BigO(?????)$ \dots The clique
-encoding requires an association graph with $g h$ vertices to be
-constructed before search.  Modern maximum-clique algorithms typically perform
-a colouring step at each search node that is quadratic in the order of the
-graph, and thus $\BigO{n^4}$ in the order of the max common subgraph input
-graphs.
-
-? Mention partition backtracking
-
-The correctness of \McSplit\ can be seen by noting its equivalence to CP-FC
-\citep{DBLP:conf/cp/NdiayeS11}, if we replace the label-class representation of
-$\mathit{future}$ with domain stores in \cref{McCliqueAlg}.  To obtain
-the CP-FC algorithm, replace \linerangeref{InnerLoop}{InnerLoopEnd} of
-\cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains
-carried out in these lines can be easily seen to be equivalent to the
-corresponding lines in \cref{McCliqueAlg}.
-
-\begin{algorithm}
-\DontPrintSemicolon
-\nl    \For {$u \in \N(G, b)$}{
-\nl        remove $v$ from $D'_u$ \;
-\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
-       }
-\nl    \For {$u \in \invN(G, b)$}{
-\nl        remove $v$ from $D'_u$ \;
-\nl        remove the neighbourhood of $t$ from $D'_u$ \;
-       }
-\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
-\label{cpAlg}
-\end{algorithm}
-
-TO WRITE PROPERLY: The second difference is in the bound function: the French
-code constructs a maximum cardinality matching, with some special weighting
-trickery for $\bot$, whilst James just sums the smaller of the two sides of
-each partition. If you think a bit about what the matching looks like, you can
-see that these two give the same answer, but at very different speeds.
-
-\section{Extensions}\label{sec:extensions}
+\subsection{Extensions}\label{sec:extensions}
 
@@
 
+\section{Comparison with Existing Algorithms}\label{sec:comparison} The
+\McSplit\ algorithm may be considered as a version of the constraint
+programming, forward-checking (CP-FC) algorithm of
+\citet{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
+of domains.  In the CP-FC algorithm, each vertex in $v \in \V(G)$ is
+represented by a variable, whose domain corresponds to the set of vertices in
+$\V(H)$ to which $v$ may currently be mapped, with an additional special $\bot$
+value representing an unmapped vertex.
+
+Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
+correspond to variables in CP-FC, all of which have domain equal to $T \cup \{
+    \bot \}$.  The label-class representation of domains is possible because
+throughout the CP-FC algorithm for max common subgraph, the domains of any two
+variables are either identical or disjoint.  (To the best of our knowledge,
+this observation has not been made previously.)
+
+Where CP-FC uses a soft allDifferent constraint to filter and compute a bound,
+\McSplit\ computes a bound from each label class in the future, i.e.\ it
+performs Hall Set reasoning on the fly.
+
+\McSplit\ requires $\BigO{n}$ time at each search node, and does not require
+any auxiliary data structures larger than the adjacency matrices of $G$ and $H$
+to be built prior to search.  CP-FC requires $\BigO(?????)$ \dots The clique
+encoding requires an association graph with $g h$ vertices to be
+constructed before search.  Modern maximum-clique algorithms typically perform
+a colouring step at each search node that is quadratic in the order of the
+graph, and thus $\BigO{n^4}$ in the order of the max common subgraph input
+graphs.
+
+?? Mention partition backtracking
+
+?? Mention dual viewpoint heuristics
+
+The correctness of \McSplit\ can be seen by noting its equivalence to CP-FC
+\citep{DBLP:conf/cp/NdiayeS11}, if we replace the label-class representation of
+$\mathit{future}$ with domain stores in \cref{McCliqueAlg}.  To obtain
+the CP-FC algorithm, replace \linerangeref{InnerLoop}{InnerLoopEnd} of
+\cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains
+carried out in these lines can be easily seen to be equivalent to the
+corresponding lines in \cref{McCliqueAlg}.
+
+\begin{algorithm}
+\DontPrintSemicolon
+\nl    \For {$u \in \N(G, b)$}{
+\nl        remove $v$ from $D'_u$ \;
+\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
+       }
+\nl    \For {$u \in \invN(G, b)$}{
+\nl        remove $v$ from $D'_u$ \;
+\nl        remove the neighbourhood of $t$ from $D'_u$ \;
+       }
+\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
+\label{cpAlg}
+\end{algorithm}
+
+TO WRITE PROPERLY: The second difference is in the bound function: the French
+code constructs a maximum cardinality matching, with some special weighting
+trickery for $\bot$, whilst James just sums the smaller of the two sides of
+each partition. If you think a bit about what the matching looks like, you can
+see that these two give the same answer, but at very different speeds.
+
 \section{Conclusion}

### 2017-02-16 paper/paper.tex
Words

@@
 \paragraph{Efficient representation of label classes} The algorithm as
-described so far requires $\BigO{n^2}$ space at each level of the search tree,
-since labels may be up to $g$ bits in length. We can reduce the memory
-requirement to $\BigO{n}$ per level by storing a \emph{label class} as a pair
-$\langle P,T \rangle$ for each label $l$ that is used, where $P$ is the set of
-vertices in $\V(G)$ labelled $l$, and $T$ is the set of vertices in $\V(H)$
-labelled $l$\footnote{We adopt terminology from subgraph isomorphism, where one
-graph is a pattern graph $P$ and the other the target graph $T$.}. Since there
-are $g + h$ vertices in the two graphs, at most $g + h$ label classes
-can exist at once, and there are at most $g + h$ vertices in the union of
-all of the $P$ and $T$ sets.  We can gain further efficiency by discarding
-during search any label class that has an empty $P$ or $T$ set.
+described so far apparently requires $\BigO{n^2}$ space at each level of the
+search tree.  We can reduce the memory requirement to $\BigO{n}$ per level by
+storing a \emph{label class} as a pair $\langle P,T \rangle$ for each label $l$
+that is used, where $P$ is the set of vertices in $\V(G)$ labelled $l$, and $T$
+is the set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology
+from subgraph isomorphism, where one graph is a pattern graph $P$ and the other
+the target graph $T$.}. Since there are $g + h$ vertices in the two graphs, at
+most $g + h$ label classes can exist at once, and there are at most $g + h$
+vertices in the union of all of the $P$ and $T$ sets.  We can gain further
+efficiency by discarding during search any label class that has an empty $P$ or
+$T$ set.
 

### 2017-02-16 paper/paper.tex
https://xkcd.com/1015/
Notational consistency
I like it better when you're on top
Tie this together
Muck around and pretend it's progress

@@
 
-?? Heuristics go here, defer justification until \cref{sec:comparison}
+?? Reword for here, explain later Small scale experiments (not presented here)
+were performed to identify a suitable heuristic for label class selection. The
+heuristic used selects the label class  with the smallest $\min(|P|,|T|)$ (a
+surrogate for expected branching factor). From this class, a vertex is selected
+from $P$ with maximum degree if $H$ is sparse and with minimum degree if $H$ is
+dense.
 
 \subsection{Extensions}\label{sec:extensions}
+
+?? This needs a two sentence intro saying that people care about stuff and why.
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \DontPrintSemicolon
 \nl    \For {$l \in L$}{
-\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} \mid u \neq v \wedge A_G[v][u] = l \}$ \;
-\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} \mid u \neq w \wedge A_H[w][u] = l \}$ \;
-\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
+\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} : u \neq v \wedge A_G[v][u] = l \}$ \;
+\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} : u \neq w \wedge A_H[w][u] = l \}$ \;
+\nl        \If {$P^{''} \negmedspace\neq \emptyset$ \bf{and} $T^{''} \negmedspace \neq \emptyset$}{
+    \nl $\AlgVar{future^{'}} \gets \AlgVar{future^{'}} + \langle P^{''}\negmedspace, T^{''} \rangle$}
        }
@@
 
-\paragraph{Maximum Common \emph{Connected} Subgraph} The common subgraph must
-be connected. We consider only undirected graphs.  Modify \McSplit\ by
-permitting branching only on a vertex $v$ that has at least one non-zero
-element in its bit-string label.  We can represent this information compactly,
-and without increasing time complexity at each search node, by storing an extra
-bit with each label class.  This bit takes the value $1$ if and only if the vertices in
-the class are adjacent to at least one vertex in $M$.
+\paragraph{Maximum Common \emph{Connected} Subgraph} ?? In applications,
+something. The common subgraph must be connected. We consider only undirected
+graphs.  Modify \McSplit\ by permitting branching only on a vertex $v$ that has
+at least one non-zero element in its bit-string label.  We can represent this
+information compactly, and without increasing time complexity at each search
+node, by storing an extra bit with each label class.  This bit takes the value
+$1$ if and only if the vertices in the class are adjacent to at least one
+vertex in $M$. ?? Link to branching scheme papers.
 
@@
 
-Experiments were performed on a cluster of machines with Dual Xeon E5-2640
-v2 CPUs and 64GBytes RAM.
+Experiments were performed on machines with dual Intel Xeon E5-2640 v2 CPUs and
+64GBytes RAM. Our algorithm was implemented in C++ and compiled using g++
+5.3.0. For comparison, we used the implementations of ?? cite clique, kdown,
+CP.
 
-? MENTION PROGRAMMING LANGUAGES, COMPILER FLAGS, MCSa ALGORITHM AND
-BITSET IMPLEMENTATION
-
-%Parallel experiments used 32 threads per machine,
-%using all 16 physical cores with hyper-threading enabled.
-All ??? experiments used a database of
-randomly-generated Maximum Common Subgraph instances, which contains graphs of
-order ranging from 10 to 100 vertices
-\cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}.  For
-unlabelled instances, we selected 10 per cent of the 41,100 instances with up
-to 40 vertices; labels were discarded. For labelled instances, we selected 10
-per cent of the full database of 81,400 instances, and used a labelling scheme
-in which the number of distinct vertex labels and the number of distinct edge
-labels is approximately equal to 33 per cent of the number of vertices in each
-graph.
-
-Small scale experiments (not presented here) were performed to identify a
-suitable heuristic for label class selection. The heuristic used selects the
-label class  with the smallest $\min(|P|,|T|)$ (a surrogate for expected
-branching factor). From this class, a vertex is selected from $P$ with maximum
-degree if $H$ is sparse and with minimum degree if $H$ is dense.
+Our first set of experiments use a database of randomly-generated maximum
+common subgraph instances
+\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}.  For
+unlabelled instances, we selected the first ten instances from each family
+whose members have no more than 50 vertices, for a total of 4,100 instances.
+For labelled instances, we selected the first ten instances from every family,
+for a total of 8,140 instances with up to 100 vertices; like
+\citet{DBLP:conf/cp/McCreeshNPS16}, we use the labelling scheme in which the
+number of distinct vertex labels and the number of distinct edge labels is
+approximately equal to 33 per cent of the number of vertices in each graph.
 
@@
 
-\paragraph{No vertex or edge labels, undirected}
+\paragraph{Unlabelled, undirected}
 \cref{figure:plain-cumulative} shows a plot of cumulative run times (x-axis)

### 2017-02-16 paper/paper.tex
I'm plotting

@@
 
-\section{A Computational Study}
+\section{Experimental Evaluation}
 
@@
     \centering
-    \subfigure[][MCS unlabelled undirected instances] {
+    \subfigure[][Unlabelled, undirected instances] {
         \centering
@@
 
-    \subfigure[][MCS 33\% vertex and edge labelled directed instances] {
+    \subfigure[][Vertex and edge labelled directed instances] {
         \centering
@@
     }
-    \caption{Cumulative run times}\label{figure:mcs-cumulative}
+    \caption{Cumulative numbers of instances solved over time for the maximum
+    common subgraph problem.}\label{figure:mcs-cumulative}
 \end{figure}

### 2017-02-16 paper/paper.tex
Use g+h in big-O notation, avoiding need to define n

--- 2017-02-16 paper/paper.tex
@@
 Let $G$ and $H$ be the two input graphs to our maximum common subgraph problem.
-The orders (vertex counts) of these graphs are denoted $g$ and $h$. We will
-use $n$ in complexity results to denote $\max(g + h)$ ?? WTF James? Help?.
+The orders (vertex counts) of these graphs are denoted $g$ and $h$.
 
@@
 \paragraph{Efficient representation of label classes} The algorithm as
-described so far apparently requires $\BigO{n^2}$ space at each level of the
-search tree.  We can reduce the memory requirement to $\BigO{n}$ per level by
+described so far apparently requires $\BigO{g(g+h)}$ space at each level of the
+search tree.  We can reduce the memory requirement to $\BigO{g+h}$ per level by
 storing a \emph{label class} as a pair $\langle P,T \rangle$ for each label $l$
@@
 
-\paragraph{Undirected with edge labels} Each adjacency matrix
-entry contains an edge label, or a null entry (non-adjacent) $0$.
-We use \cref{labDirAlg}, by letting $L$ be
-the union of $\{0\}$ with the set of all labels that appear in the input
-graphs. Since there may be up to $g + h$ distinct labels, the loop in
-\cref{labDirAlg} may execute up to $g + h$ times and
-requires $\BigO{n^2}$ time per search
-node.  It is possible to modify the algorithms to run in $\BigO{n \log n}$ per
-search node.  First, run lines 17-19
-of \cref{McCliqueAlg} to create a new label-class of vertices that are not
-adjacent to $v$ or $w$, and remove these vertices from $\langle P',T' \rangle$.
-Next, sort $P'$ and $T'$ in ascending order of the label on the edge from $v$
-or $w$ to each vertex. We can then create the label classes corresponding to
-each edge label by simultaneously traversing $P'$ and $T'$ from left to right.
+\paragraph{Undirected with edge labels} Each adjacency matrix entry contains an
+edge label, or a null entry (non-adjacent) $0$.  We use \cref{labDirAlg}, by
+letting $L$ be the union of $\{0\}$ with the set of all labels that appear in
+the input graphs. Since there may be up to $g + h$ distinct labels, the loop in
+\cref{labDirAlg} may execute up to $g + h$ times, resulting in $\BigO{(g+h)^2}$
+time complexity per search node.  To achieve $\BigO{(g+h) \log (g+h)}$ time
+complexity per search node, we can modify the algorithms to use sorting rather
+than explicitly looping over all label classes, as follows.  First, run lines
+17-19 of \cref{McCliqueAlg} to create a new label-class of vertices that are
+not adjacent to $v$ or $w$, and remove these vertices from $\langle P',T'
+\rangle$.  Next, sort $P'$ and $T'$ in ascending order of the label on the edge
+from $v$ or $w$ to each vertex. We can then create the label classes
+corresponding to each edge label by simultaneously traversing $P'$ and $T'$
+from left to right.
 
@@
 
-\McSplit\ requires $\BigO{n}$ time at each search node, and does not require
+\McSplit\ requires $\BigO{g+h}$ time at each search node, and does not require
 any auxiliary data structures larger than the adjacency matrices of $G$ and $H$
@@
 a colouring step at each search node that is quadratic in the order of the
-graph, and thus $\BigO{n^4}$ in the order of the max common subgraph input
+graph, and thus $\BigO{g^2h^2}$ in the order of the max common subgraph input
 graphs.

### 2017-02-16 paper/paper.tex
Move things around, because I can

@@
 
-\paragraph{Top-Down Strategy} I'VE RE-ADDED A SHORTER VERSION OF THIS
-PARAGRAPH, WHICH WAS COMMENTED OUT BY PATRICK. IT NEEDS WORK, BUT WE NEED TO
-EXPLAIN THE SIP EXPERIMENT. We can modify the \McSplit\ algorithm to use a
-top-down strategy based on \citet{UpcomingAAAIPaper} by calling the main
-\FuncSty{McSplit} method once per goal size ($g, g-1, g-2, \dots$).  We
-backtrack (\lineref{PruneSearch} of \cref{McCliqueAlg}) when the bound is
-strictly less than the goal size, and terminate when a solution of the goal
-size is found.
-
 \section{Experimental Evaluation}
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-sip-cumulative.pdf}
-    \caption{Cumulative run times, SIP instances}\label{figure:sip-cumulative}
-\end{figure}
+?? Let's put this hereish: I'VE RE-ADDED A SHORTER VERSION OF THIS
+PARAGRAPH, WHICH WAS COMMENTED OUT BY PATRICK. IT NEEDS WORK, BUT WE NEED TO
+EXPLAIN THE SIP EXPERIMENT. We can modify the \McSplit\ algorithm to use a
+top-down strategy based on \citet{UpcomingAAAIPaper} by calling the main
+\FuncSty{McSplit} method once per goal size ($g, g-1, g-2, \dots$).  We
+backtrack (\lineref{PruneSearch} of \cref{McCliqueAlg}) when the bound is
+strictly less than the goal size, and terminate when a solution of the goal
+size is found.
 
-\paragraph{CP-FC and \McSplit\ search nodes on MCS instances}
-\cref{figure:33ved-james-versus-cp-fc-nodes-scatter} shows one point per instance,
-with search nodes visited by CP-FC on the horizontal axis and search nodes visited
-by \McSplit\ on the vertical axis.  There is a strong correlation, which reflects
-the similarity of the search space explored by the two algorithms.  \McSplit\
-tends to explore slightly fewer search nodes than CP-FC, because the label-classes
-allow us to (cheaply) compute a very good heuristics for variable selection.
+\section{Comparison with Existing Algorithms}\label{sec:comparison}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, James vs
-    CP-FC, Nodes (TODO?? heatmapify this and exclude
-    timeouts)}\label{figure:33ved-james-versus-cp-fc-nodes-scatter}
-\end{figure}
-
-\paragraph{kdown and \McSplit\ search nodes on SIP instances}
-\cref{figure:sip-james-versus-kdown-nodes-scatter} shows the number of search
-nodes taken by kdown and \McSplit\ on each instance.  Although \McSplit\ is the
-faster algorithm overall, it explores more search nodes than kdown for most instances.
-The scatter plot has a concentration of points around the line $y=2x$. We conjecture
-that the two algorithms explore very similar search trees for instances close to this
-line.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
-    \caption{SIP instances, James vs kdown, Nodes (TODO?? heatmapify this and exclude
-    timeouts)}\label{figure:sip-james-versus-kdown-nodes-scatter}
-\end{figure}
-
-\section{Comparison with Existing Algorithms}\label{sec:comparison} The
+The
 \McSplit\ algorithm may be considered as a version of the constraint
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-sip-cumulative.pdf}
+    \caption{Cumulative run times, SIP instances}\label{figure:sip-cumulative}
+\end{figure}
+
+\paragraph{CP-FC and \McSplit\ search nodes on MCS instances}
+\cref{figure:33ved-james-versus-cp-fc-nodes-scatter} shows one point per instance,
+with search nodes visited by CP-FC on the horizontal axis and search nodes visited
+by \McSplit\ on the vertical axis.  There is a strong correlation, which reflects
+the similarity of the search space explored by the two algorithms.  \McSplit\
+tends to explore slightly fewer search nodes than CP-FC, because the label-classes
+allow us to (cheaply) compute a very good heuristics for variable selection.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
+    \caption{MCS 33\% vertex and edge labelled directed instances, James vs
+    CP-FC, Nodes (TODO?? heatmapify this and exclude
+    timeouts)}\label{figure:33ved-james-versus-cp-fc-nodes-scatter}
+\end{figure}
+
+\paragraph{kdown and \McSplit\ search nodes on SIP instances}
+\cref{figure:sip-james-versus-kdown-nodes-scatter} shows the number of search
+nodes taken by kdown and \McSplit\ on each instance.  Although \McSplit\ is the
+faster algorithm overall, it explores more search nodes than kdown for most instances.
+The scatter plot has a concentration of points around the line $y=2x$. We conjecture
+that the two algorithms explore very similar search trees for instances close to this
+line.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
+    \caption{SIP instances, James vs kdown, Nodes (TODO?? heatmapify this and exclude
+    timeouts)}\label{figure:sip-james-versus-kdown-nodes-scatter}
+\end{figure}
+
 \section{Conclusion}

### 2017-02-16 paper/paper.tex
Progress

@@
 against number of problem instances solved (y-axis).  There are five contours:
-the clique encoding \cite{DBLP:conf/cp/McCreeshNPS16}, CP-FC
-\cite{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.  WE NEED CITATIONS
+the clique encoding \citep{DBLP:conf/cp/McCreeshNPS16}, CP-FC
+\citep{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.  WE NEED CITATIONS
 FOR THESE AND ALSO NEED TO CLEAN UP THE LABELS ON THE GRAPHS.  We may compare
@@
 over an order of magnitude faster than the best existing CP algorithm, which is
-CP-MAC in this case.  The clique encoding outperforms the other
-algorithms---including \McSplit---on these labelled instances, except in the
-very easy region of instances that can be solved in well under 100 ms.
+CP-MAC in this case. Matching the conclusions of
+\citet{DBLP:conf/cp/McCreeshNPS16}, we see that the clique encoding outperforms
+the other algorithms---including \McSplit---on these labelled instances, except
+in the very easy region of instances that can be solved in well under 100 ms.
 
-\paragraph{No Vertex or edge labels, undirected, connected} are in
-\cref{figure:plain-connected-cumulative}.  These results are very similar to
-the corresponding experiment in which the subgraph is not required to be
-connected (\cref{figure:plain-cumulative}).
+\paragraph{Unlabelled, undirected, connected} This class of instances are shown
+in \cref{figure:plain-connected-cumulative}.  These results are very similar to
+the corresponding experiment in \cref{figure:plain-cumulative} in which the
+subgraph is not required to be connected.
 
@@
 connected case, clique slightly outperforms \McSplit\ on harder instances
-(\cref{figure:33ve-connected-cumulative}).  The gap between the two algorithms
-is narrow; indeed, the cumulative curve for \McSplit\ briefly rises above
-the curve for clique at a runtime just below 100 seconds.
+(\cref{figure:33ve-connected-cumulative}). However, the gap between the two algorithms
+is very narrow, and is probably down to quality of implementation; indeed, the
+cumulative curve for \McSplit\ briefly rises above the curve for clique at a
+runtime just below 100 seconds. Additionally, \McSplit\ is the clear winner for
+easier instances, where the clique encoding is relatively expensive to
+construct but trivial to solve.
 
-\begin{figure}[ht]
+\begin{figure}[t]
     \centering

### 2017-02-16 paper/paper.tex
Move move move
Clean up experiments some more

@@
 
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 
-    \subfigure[][Vertex and edge labelled directed instances] {
+    \subfigure[][Vertex and edge labelled, directed instances] {
         \centering
@@
 
+?? kdown only does undirected unlabelled unconnected
+
+?? There are five contours:
+the clique encoding \citep{DBLP:conf/cp/McCreeshNPS16}, CP-FC
+\citep{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.  WE NEED CITATIONS
+FOR THESE AND ALSO NEED TO CLEAN UP THE LABELS ON THE GRAPHS.
+
 \paragraph{Unlabelled, undirected}
 \cref{figure:plain-cumulative} shows a plot of cumulative run times (x-axis)
-against number of problem instances solved (y-axis).  There are five contours:
-the clique encoding \citep{DBLP:conf/cp/McCreeshNPS16}, CP-FC
-\citep{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.  WE NEED CITATIONS
-FOR THESE AND ALSO NEED TO CLEAN UP THE LABELS ON THE GRAPHS.  We may compare
+against number of problem instances solved (y-axis).  We may compare
 the speed of two algorithms using the horizontal distance between their curves.
@@
 \paragraph{Vertex and edge labels, directed} Cumulative run times for this
-class of instances are in \cref{figure:33ved-cumulative}.  Again, \McSplit\ is
-over an order of magnitude faster than the best existing CP algorithm, which is
+class of instances are in \cref{figure:33ved-cumulative}. Again, \McSplit\ is over
+an order of magnitude faster than the best existing CP algorithm, which is
 CP-MAC in this case. Matching the conclusions of
@@
 \paragraph{Unlabelled, undirected, connected} This class of instances are shown
-in \cref{figure:plain-connected-cumulative}.  These results are very similar to
+in \cref{figure:plain-connected-cumulative} (again, ??kdown does not support the connected problem).  These results are very similar to
 the corresponding experiment in \cref{figure:plain-cumulative} in which the
@@
     \centering
-    \subfigure[][MCS unlabelled undirected instances] {
+    \subfigure[][Unlabelled, undirected, connected instances] {
         \centering
@@
 
-    \subfigure[][MCS 33\% vertex and edge labelled undirected instances] {
+    \subfigure[][Vertex and edge labelled, undirected, connected instances] {
         \centering
@@
     }
-    \caption{Cumulative run times, connected variant}
-    \label{figure:mcs-connected-cumulative}
+    \caption{Cumulative numbers of instances solved over time for the maximum
+    common connected subgraph problem.} \label{figure:mcs-connected-cumulative}
 \end{figure}
@@
 is the strongest solver overall for this dataset.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics*{gen-graph-sip-cumulative.pdf}
+    \caption{Cumulative run times, SIP instances}\label{figure:sip-cumulative}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-sip-cumulative.pdf}
-    \caption{Cumulative run times, SIP instances}\label{figure:sip-cumulative}
-\end{figure}
-
 \paragraph{CP-FC and \McSplit\ search nodes on MCS instances}

### 2017-02-16 paper/paper.tex
SIPping
title case
Cites

@@
 
-\section{The \McSplit\ algorithm}
+\section{The \McSplit\ Algorithm}
 
@@
 64GBytes RAM. Our algorithm was implemented in C++ and compiled using g++
-5.3.0. For comparison, we used the implementations of ?? cite clique, kdown,
-CP.
+5.3.0. For comparison, we compare against the best constraint programming
+implementations of \citet{DBLP:conf/cp/NdiayeS11} and
+\citet{DBLP:conf/cp/McCreeshNPS16} (CP-FC in the unlabelled cases, and CP-MAC
+in the labelled cases, using both branching and filtering for connected
+subgraphs), the clique encodings of \citet{DBLP:conf/cp/McCreeshNPS16}, and the
+$k\downarrow$ algorithm of \citet{UpcomingAAAIPaper} (which only supports
+unlabelled, undirected, unconnected instances).
 
@@
 \end{figure}
-
-?? kdown only does undirected unlabelled unconnected
-
-?? There are five contours:
-the clique encoding \citep{DBLP:conf/cp/McCreeshNPS16}, CP-FC
-\citep{DBLP:conf/cp/NdiayeS11}, \McSplit\, AND TWO OTHERS.  WE NEED CITATIONS
-FOR THESE AND ALSO NEED TO CLEAN UP THE LABELS ON THE GRAPHS.
 
@@
 a set of 5725 larger instances used in recent studies of subgraph
-isomorphism~\cite{thelionpaper} and maximum common
-subgraph~\cite{UpcomingAAAIPaper}.  This benchmark set includes real-world
+isomorphism~\citep{DBLP:conf/lion/KotthoffMS16} and maximum common
+subgraph~\citep{UpcomingAAAIPaper}.  This benchmark set includes real-world
 graphs and graphs generated using random models.  Pattern graphs range from 4
@@
 6671 with a median of 561.  Cumulative runtimes including the kdown algorithm
-from~\cite{UpcomingAAAIPaper} which was the previous state of the art for this
+from~\citep{UpcomingAAAIPaper} which was the previous state of the art for this
 benchmark, are in \cref{figure:sip-cumulative}.  This is a challenging set of
@@
     \includegraphics*{gen-graph-sip-cumulative.pdf}
-    \caption{Cumulative run times, SIP instances}\label{figure:sip-cumulative}
+    \caption{Cumulative numbers of instances solved over time for the maximum
+    common connected subgraph problem on the large subgraph isomorphism benchmark
+    suite.} \label{figure:sip-cumulative}
 \end{figure}

### 2017-02-16 paper/paper.tex
Describe the down and sip stuff

@@
 \paragraph{Unlabelled, undirected, connected} This class of instances are shown
-in \cref{figure:plain-connected-cumulative} (again, ??kdown does not support the connected problem).  These results are very similar to
+in \cref{figure:plain-connected-cumulative}.  These results are very similar to
 the corresponding experiment in \cref{figure:plain-cumulative} in which the
@@
 vertices to 900 with a median of 80; target graphs range from 10 vertices to
-6671 with a median of 561.  Cumulative runtimes including the kdown algorithm
-from~\citep{UpcomingAAAIPaper} which was the previous state of the art for this
-benchmark, are in \cref{figure:sip-cumulative}.  This is a challenging set of
-instances, and more than half of the instances cannot be solved within a
-timeout of 1000 seconds by any of the solvers. Moreover the CP-FC algorithm and
-the clique encoding run out of memory on many of the instances.  \McSplit\
-is the strongest solver overall for this dataset.
+6671 with a median of 561. Cumulative runtimes on these instances are shown in
+\cref{figure:sip-cumulative}.  This is a challenging set of instances, and more
+than half of the instances cannot be solved within a timeout of 1000 seconds by
+any solver. Furthermore, the CP-FC algorithm and the clique encoding run out of
+memory on many of the instances (these are treated as timeouts).
+
+The basic \McSplit\ is beaten by the $k\downarrow$ algorithm of
+\citet{UpcomingAAAIPaper} on this dataset. However, we can modify the \McSplit\
+algorithm to use a top-down strategy similar to that used by $k\downarrow$  by
+calling the main \FuncSty{McSplit} method once per goal size ($g, g-1, g-2,
+\dots$); we backtrack (\lineref{PruneSearch} of \cref{McCliqueAlg}) when the
+bound is strictly less than the goal size, and terminate when a solution of the
+goal size is found. We expect that this could do well because in many cases the
+matching covers nearly all of the smaller graph---indeed,
+\cref{figure:sip-cumulative} shows that this approach is the strongest on these
+instances.
 
@@
 \end{figure}
-
-?? Let's put this hereish: I'VE RE-ADDED A SHORTER VERSION OF THIS
-PARAGRAPH, WHICH WAS COMMENTED OUT BY PATRICK. IT NEEDS WORK, BUT WE NEED TO
-EXPLAIN THE SIP EXPERIMENT. We can modify the \McSplit\ algorithm to use a
-top-down strategy based on \citet{UpcomingAAAIPaper} by calling the main
-\FuncSty{McSplit} method once per goal size ($g, g-1, g-2, \dots$).  We
-backtrack (\lineref{PruneSearch} of \cref{McCliqueAlg}) when the bound is
-strictly less than the goal size, and terminate when a solution of the goal
-size is found.
 

### 2017-02-16 paper/paper.tex
Use booktabs
Thinging

@@
 \usepackage{microtype}
+\usepackage{booktabs}
 
@@
   \begin{minipage}[t]{0.3\linewidth}
-      \centering Labelling of $G$
+      \centering
+      Labelling of $G$
       \begin{tabular}[t]{cc}
-      \hline
+      \toprule
           Vertex & Label\\
-      \hline
+      \midrule
           #3
-      \hline
+      \bottomrule
       \end{tabular}
@@
   \begin{minipage}[t]{0.3\linewidth}
-      \centering Labelling of $H$
+      \centering
+      Labelling of $H$
       \begin{tabular}[t]{cc}
-      \hline
+      \toprule
           Vertex & Label\\
-      \hline
+      \midrule
           #4
-      \hline
+      \bottomrule
       \end{tabular}
@@
 
-The
-\McSplit\ algorithm may be considered as a version of the constraint
-programming, forward-checking (CP-FC) algorithm of
-\citet{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
-of domains.  In the CP-FC algorithm, each vertex in $v \in \V(G)$ is
-represented by a variable, whose domain corresponds to the set of vertices in
-$\V(H)$ to which $v$ may currently be mapped, with an additional special $\bot$
-value representing an unmapped vertex.
+Our results so far suggest that \McSplit\ has broadly similar performance
+trends to the constraint programming, forward-checking (CP-FC) algorithm of
+\citet{DBLP:conf/cp/NdiayeS11}, but with much lower constant factors. Indeed,
+in \cref{figure:33ved-james-versus-cp-fc-nodes-scatter} we plot the number of
+recursive calls made by our algorithm on the $y$-axis versus the number made by
+CP-FC on the $x$-axis, for the labelled, unconnected problem instances. We see
+a close correlation: we typically do slightly less work, and sometimes do more,
+but instances with more than one order of magnitude difference in search space
+size are rare.
 
-Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
-correspond to variables in CP-FC, all of which have domain equal to $T \cup \{
-    \bot \}$.  The label-class representation of domains is possible because
-throughout the CP-FC algorithm for max common subgraph, the domains of any two
-variables are either identical or disjoint.  (To the best of our knowledge,
-this observation has not been made previously.)
+Why is this? We do not see a similar correlation between our search space size
+and that of the clique spproach. The key observation is that \McSplit\ may be
+considered to be a different version of the constraint CP-FC algorithm, using
+an unconventional domain store and more efficient filtering algorithms. We now
+explore this relationship further.
 
-Where CP-FC uses a soft allDifferent constraint to filter and compute a bound,
-\McSplit\ computes a bound from each label class in the future, i.e.\ it
-performs Hall Set reasoning on the fly.
-
-\McSplit\ requires $\BigO{g+h}$ time at each search node, and does not require
-any auxiliary data structures larger than the adjacency matrices of $G$ and $H$
-to be built prior to search.  CP-FC requires $\BigO(?????)$ \dots The clique
-encoding requires an association graph with $g h$ vertices to be
-constructed before search.  Modern maximum-clique algorithms typically perform
-a colouring step at each search node that is quadratic in the order of the
-graph, and thus $\BigO{g^2h^2}$ in the order of the max common subgraph input
-graphs.
-
-?? Mention partition backtracking
-
-?? Mention dual viewpoint heuristics
-
-The correctness of \McSplit\ can be seen by noting its equivalence to CP-FC
-\citep{DBLP:conf/cp/NdiayeS11}, if we replace the label-class representation of
-$\mathit{future}$ with domain stores in \cref{McCliqueAlg}.  To obtain
-the CP-FC algorithm, replace \linerangeref{InnerLoop}{InnerLoopEnd} of
-\cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains
-carried out in these lines can be easily seen to be equivalent to the
-corresponding lines in \cref{McCliqueAlg}.
-
-\begin{algorithm}
-\DontPrintSemicolon
-\nl    \For {$u \in \N(G, b)$}{
-\nl        remove $v$ from $D'_u$ \;
-\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
-       }
-\nl    \For {$u \in \invN(G, b)$}{
-\nl        remove $v$ from $D'_u$ \;
-\nl        remove the neighbourhood of $t$ from $D'_u$ \;
-       }
-\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
-\label{cpAlg}
-\end{algorithm}
-
-TO WRITE PROPERLY: The second difference is in the bound function: the French
-code constructs a maximum cardinality matching, with some special weighting
-trickery for $\bot$, whilst James just sums the smaller of the two sides of
-each partition. If you think a bit about what the matching looks like, you can
-see that these two give the same answer, but at very different speeds.
-
-\paragraph{CP-FC and \McSplit\ search nodes on MCS instances}
-\cref{figure:33ved-james-versus-cp-fc-nodes-scatter} shows one point per instance,
-with search nodes visited by CP-FC on the horizontal axis and search nodes visited
-by \McSplit\ on the vertical axis.  There is a strong correlation, which reflects
-the similarity of the search space explored by the two algorithms.  \McSplit\
-tends to explore slightly fewer search nodes than CP-FC, because the label-classes
-allow us to (cheaply) compute a very good heuristics for variable selection.
-
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
     \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{MCS 33\% vertex and edge labelled directed instances, James vs
-    CP-FC, Nodes (TODO?? heatmapify this and exclude
-    timeouts)}\label{figure:33ved-james-versus-cp-fc-nodes-scatter}
+    \caption{Size of the search space for \McSplit versus CP-FC on the
+    labelled, directed, unconnected instances which were solved within the
+    timeout by both algorithms.}
+    \label{figure:33ved-james-versus-cp-fc-nodes-scatter}
 \end{figure}
 
-\paragraph{kdown and \McSplit\ search nodes on SIP instances}
-\cref{figure:sip-james-versus-kdown-nodes-scatter} shows the number of search
-nodes taken by kdown and \McSplit\ on each instance.  Although \McSplit\ is the
-faster algorithm overall, it explores more search nodes than kdown for most instances.
-The scatter plot has a concentration of points around the line $y=2x$. We conjecture
-that the two algorithms explore very similar search trees for instances close to this
-line.
+In the CP-FC algorithm, each vertex in $v \in \V(G)$ is represented by a
+variable, whose domain corresponds to the set of vertices in $\V(H)$ to which
+$v$ may currently be mapped, with an additional special $\bot$ value
+representing an unmapped vertex.
+Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
+correspond to variables in CP-FC. The label-class representation of domains is
+possible because throughout the CP-FC algorithm for max common subgraph, the
+domains of any two variables are either identical or disjoint (excluding
+$\bot$, which is slightly more complicated). To the best of our knowledge, this
+observation has not been made previously, and it is not exploited in other
+implementations.
+
+CP-FC uses a soft all-different constraint to compute a bound, which requires
+running a matching algorithm on a supporting compatibility graph.  Careful
+thought shows that \McSplit\ computes the same bound, but using a simple
+counting loop---this is only possible because of the disjoint nature of the
+domains.
+
+%%% This equivalence to CP-FC makes the correctness of \McSplit\ easy to establish:
+%%% if we replace the label-class representation of $\mathit{future}$ with
+%%% conventional domain stores, and replace \linerangeref{InnerLoop}{InnerLoopEnd}
+%%% of \cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains carried out in
+%%% these lines can be easily seen to be equivalent to the corresponding lines in
+%%% \cref{McCliqueAlg}.
+%%% 
+%%% 
+%%% \begin{algorithm}
+%%% \DontPrintSemicolon
+%%% \nl    \For {$u \in \N(G, b)$}{
+%%% \nl        remove $v$ from $D'_u$ \;
+%%% \nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
+%%%        }
+%%% \nl    \For {$u \in \invN(G, b)$}{
+%%% \nl        remove $v$ from $D'_u$ \;
+%%% \nl        remove the neighbourhood of $t$ from $D'_u$ \;
+%%%        }
+%%% \caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm ?? Not sure about this bit, maybe we don't need it}
+%%% \label{cpAlg}
+%%% \end{algorithm}
+
+A further advantage of our encoding is that it gives us efficient access to a
+better branching heuristic. The CP-FC algorithm uses smallest domain first,
+which in our algorithm corresponds to branching on a label class with smallest
+$|T|$. We instead branch on the label class with smallest $\min(|P|,|T|)$.
+This is empirically better, and accounts for much of the difference between
+the number of recursive calls made; branching on the smallest $|P| |T|$ gives
+very similar results. This can be viewed as exploiting both smallest domain first,
+and the dual viewpoint \citep{DBLP:conf/ecai/Geelen92} of smallest domain
+first, simultaneously, but we do not have the overheads of having to maintain
+and channel between the dual viewpoint that would be required when using a
+conventional domain store. ?? Check this!
+
+What about our relationship to the $k\downarrow$ of \citet{UpcomingAAAIPaper}?
+\Cref{figure:sip-james-versus-kdown-nodes-scatter} plots the number of
+recursive calls made by $k\downarrow$ and \McSplit\ on each of the subgraph
+isomorphism instances. Although \McSplit\ is the faster algorithm overall, it
+explores more search nodes than $k\downarrow$ for most instances.
+
+?? Talk about the classic tradeoff, and dynamically switching.
+
+?? The concentration of points, down to unit propagation and counting
 

### 2017-02-16 paper/paper.tex
OK I checked it
MIP

@@
 computationally. Recent practical progress has been made by using constraint
-programming \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}, by
-reducing to the maximum clique problem
-\citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and by adapting subgraph isomorphism
-algorithms \citep{UpcomingAAAIPaper}. This paper introduces a new branch and
-bound algorithm which exploits special properties of these problems to allow a
-much faster exploration of the search space, whilst retaining the filtering and
-bounding benefits of the constraint programming approach. We describe the
-algorithm for the basic maximum common subgraph problem, and discuss how it may
-be adapted to handle vertex labels, edge labels, and the requirement that the
-found subgraph be connected. We then present an empirical study of the
-algorithm, demonstrating that it improves the state of the art by more than an
-order of magnitude on the unlabelled variant of the problem, and showing that
-it can handle much larger instances than earlier constraint programming or
-clique approaches due to lower memory usage.
+programming \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16} and
+mathematical programming \citep{DBLP:journals/dam/BahienseMPS12}, by reducing
+to the maximum clique problem \citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and by
+adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. This paper
+introduces a new branch and bound algorithm which exploits special properties
+of these problems to allow a much faster exploration of the search space,
+whilst retaining the filtering and bounding benefits of the constraint
+programming approach. We describe the algorithm for the basic maximum common
+subgraph problem, and discuss how it may be adapted to handle vertex labels,
+edge labels, and the requirement that the found subgraph be connected. We then
+present an empirical study of the algorithm, demonstrating that it improves the
+state of the art by more than an order of magnitude on the unlabelled variant
+of the problem, and showing that it can handle much larger instances than
+earlier constraint programming or clique approaches due to lower memory usage.
 
@@
 and channel between the dual viewpoint that would be required when using a
-conventional domain store. ?? Check this!
+conventional domain store.
 

### 2017-02-16 paper/paper.tex
Induced

@@
 consisting of some of the vertices of $G$, and all of the edges between these
-vertices (in this paper all subgraphs are induced). A \emph{common subgraph} of
-two graphs is a graph which is (isomorphic to) a subgraph of two graphs
-simultaneously, and a \emph{maximum} common subgraph is one with as many
-vertices as possible.
+vertices. A \emph{common subgraph} of two graphs is a graph which is
+(isomorphic to) a subgraph of two graphs simultaneously, and a \emph{maximum}
+common subgraph is one with as many vertices as possible. (In this paper, all
+subgraphs are induced. The \emph{maximum common partial subgraph} problem
+instead asks for a common non-induced subgraph with as many \emph{edges} as
+possible.)
 

### 2017-02-16 paper/paper.tex
Enshrinken

@@
 
-%\newcommand{\examplexG}[5] {
-%    \begin{minipage}{.2\textwidth}
-%    \tikz {
-%        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1cm,
-%                clockwise=5] {
-%                    1[label=90:#1],2[label=0:#2],3[label=0:#3],4[label=180:#4],5[label=180:#5];
-%            1--4; 1--5; 2--3; 2--5; 3--5;
-%        };
-%    }
-%    \end{minipage}
-%}
-%\newcommand{\examplexH}[6] {
-%    \begin{minipage}{.2\textwidth}
-%    \tikz {
-%        \graph [nodes={draw, circle, minimum width=.6cm}, circular placement, radius=1.1cm,
-%                clockwise=6, phase=60] {
-%                    a[label=0:#1],b[label=0:#2],c[label=0:#3],d[label=180:#4],e[label=180:#5],f[label=180:#6];
-%            a--b; a--c; a--e; b--d; b--f; c--d; c--e; c--f; d--f; e--f;
-%        };
-%    }
-%    \end{minipage}
-%}
+\newcommand{\examplexG}[5] {
+    \begin{minipage}{.2\textwidth}
+    \tikz {
+        \graph [nodes={draw, circle, minimum width=.5cm, inner sep=1pt}, circular placement, radius=0.8cm,
+                clockwise=5] {
+                    1[label=90:#1],2[label=0:#2],3[label=0:#3],4[label=180:#4],5[label=180:#5];
+            1--4; 1--5; 2--3; 2--5; 3--5;
+        };
+    }
+    \end{minipage}
+}
+\newcommand{\examplexH}[6] {
+    \begin{minipage}{.2\textwidth}
+    \tikz {
+        \graph [nodes={draw, circle, minimum width=.5cm, inner sep=1pt}, circular placement, radius=0.8cm,
+                clockwise=6, phase=60] {
+                    a[label=0:#1],b[label=0:#2],c[label=0:#3],d[label=180:#4],e[label=180:#5],f[label=180:#6];
+            a--b; a--c; a--e; b--d; b--f; c--d; c--e; c--f; d--f; e--f;
+        };
+    }
+    \end{minipage}
+}
 
@@
 \centering
-    \includegraphics{graph_G}
-    \includegraphics{graph_H}
-%    \examplexG{}{}{}{}{}
-%    \examplexH{}{}{}{}{}{}
+    \examplexG{}{}{}{}{}
+    \examplexH{}{}{}{}{}{}
 \caption{Example graphs $G$ and $H$.}

### 2017-02-16 paper/paper.tex
Shrink for page layout, because I can
Shorter title
Link to the past

@@
 
-    \caption{Mapping $M$ and vertex labels during search on example graphs $G$ and $H$}
+    \caption{Mapping $M$ and vertex labels during search on example graphs $G$ and $H$ from \cref{fig:alg1}.}
     \label{figure:mcsplit-examples}
@@
 matching size is $|M| + 1 = 4$. The general formula for the upper bound is
-as follows, where $L$ is the set of labels used in both graphs.
-
 \begin{multline*}
-    \mathit{bound} = |M| + \sum_{l \in L} \min\big(|\{ v \in \V(G) : \vtxlabel(v)=l\}|, \\[-0.2cm]
-        |\{ v \in \V(H) : \vtxlabel(v)=l \}|\big)
-\end{multline*}
+    \mathit{bound} = |M| + \sum_{l \in L} \min\big(|\{ v \in \V(G) : \vtxlabel(v)=l\}|, \\[-0.3cm]
+        |\{ v \in \V(H) : \vtxlabel(v)=l \}|\big) \text{,}
+\end{multline*} where $L$ is the set of labels used in both graphs.
 
@@
 }
-\caption{\McSplit\ algorithm for Maximum Common Subgraph}
+\caption{Finding a maximum common subgraph.}
 \label{McCliqueAlg}

### 2017-02-16 paper/paper.tex
This commit might make things worse, but I think we should say something along these lines and make it more coherent tomorrow

@@
 
-Before discussing the algorithm in detail, we start with an illustration of the
-main concepts using the graphs $G$ and $H$ in \cref{fig:alg1}.  These graphs
-have a maximum common subgraph with four vertices; one example is the mapping
-$\{1a, 2f, 3d, 5b\}$ where vertex $1$ is assigned to vertex $a$, $2$ is
-assigned to $f$, $3$ to $d$ and $5$ to $b$.
+\paragraph{Walkthrough} Before discussing the algorithm in detail, we start
+with an illustration of the main concepts using the graphs $G$ and $H$ in
+\cref{fig:alg1}.  These graphs have a maximum common subgraph with four
+vertices; one example is the mapping $\{1a, 2f, 3d, 5b\}$ where vertex $1$ is
+assigned to vertex $a$, $2$ is assigned to $f$, $3$ to $d$ and $5$ to $b$.
 
@@
 
-\paragraph{Efficient representation of label classes} The algorithm as
-described so far apparently requires $\BigO{g(g+h)}$ space at each level of the
-search tree.  We can reduce the memory requirement to $\BigO{g+h}$ per level by
-storing a \emph{label class} as a pair $\langle P,T \rangle$ for each label $l$
-that is used, where $P$ is the set of vertices in $\V(G)$ labelled $l$, and $T$
-is the set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology
-from subgraph isomorphism, where one graph is a pattern graph $P$ and the other
-the target graph $T$.}. Since there are $g + h$ vertices in the two graphs, at
-most $g + h$ label classes can exist at once, and there are at most $g + h$
-vertices in the union of all of the $P$ and $T$ sets.  We can gain further
-efficiency by discarding during search any label class that has an empty $P$ or
-$T$ set.
+\paragraph{Label classes} We require only $\BigO{g+h}$ space per level of the
+search tree to store labelling information.  This is done by storing a
+\emph{label class} as a pair $\langle P,T \rangle$ for each label $l$ that is
+used, where $P$ is the set of vertices in $\V(G)$ labelled $l$, and $T$ is the
+set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology from
+subgraph isomorphism, where one graph is a pattern graph $P$ and the other the
+target graph $T$.}. Since there are $g + h$ vertices in the two graphs, at most
+$g + h$ label classes can exist at once, and there are at most $g + h$ vertices
+in the union of all of the $P$ and $T$ sets. Furthermore, we do not actually
+need to store the bits making up a label---we care only that like-labelled
+vertices are kept together, and the label itself is not used. Nor do we need to
+store any label class which is present only in one graph but not the other (or
+which is not present at all).  Together, these facts allow us to store all the
+necessary information in two pairs of flat arrays. One pair of arrays is used
+to store a permutation of the vertices in $P$ having like-labelled vertices
+stored consecutively, together with the indices of the start of each label in
+turn (in some arbitrary order). The second pair stores the permutation of
+vertices in $T$, together with label start indices \emph{in the same order} as
+they are used in the first pair of arrays. ?? This representation has
+similarities to data structures used in partition backtracking and bron
+kerbosch.
 
-?? Talk about how we can use a pair of flat arrays and indices to the start of
-label classes, and throw in citation to partition backtracking here.
-
-\begin{algorithm}[ht]
+\begin{algorithm}[t]
 \DontPrintSemicolon
@@
 
-\paragraph{\cref{McCliqueAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
+\paragraph{\cref{McCliqueAlg} in detail} The recursive procedure, $\FuncSty{expand}$, has
 two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,

### 2017-02-16 paper/paper.tex
Better caption
James?

@@
 $\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
-if and only if $v$ and $w$ belong to the same label class in $M$.
+if and only if $v$ and $w$ belong to the same label class in $M$ ?? shouldn't that be future?.
 
@@
        }
-\caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} in directed and labelled cases}
+    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McCliqueAlg} to handle directed and labelled cases.}
 \label{labDirAlg}

### 2017-02-16 paper/paper.tex
We can math inside algorithm

@@
 the search when a calculated upper bound (best future $\AlgVar{bf}$ plus $|M|$) is not larger
-than the incumbent.
+than the incumbent ?? rewrite this as one line using $\sum$?.
 

### 2017-02-16 paper/paper.tex
Caption

@@
     \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{Size of the search space for \McSplit versus CP-FC on the
+    \caption{Size of the search space for \McSplit\ versus CP-FC on the
     labelled, directed, unconnected instances which were solved within the
@@
     \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
-    \caption{SIP instances, James vs kdown, Nodes (TODO?? heatmapify this and exclude
-    timeouts)}\label{figure:sip-james-versus-kdown-nodes-scatter}
+    \caption{Size of the search space for \McSplit$\downarrow$\ versus $k\downarrow$ on the
+    large subgraph isomorphism instances which were solved within the
+    timeout by both algorithms.}\label{figure:sip-james-versus-kdown-nodes-scatter}
 \end{figure}
@@
 There is also the potential to branch on both sides, that is instead of branching on vertices in $\V(G)$, it would be equally valid to branch on
-a vertex in $\V(H)$ , since our data structure treats the
+a vertex in $\V(H)$, since our data structure treats the
 two graphs symmetrically. More interestingly, we could choose which graph to branch on

### 2017-02-16 paper/paper.tex
Conclude

@@
 
-We have introduced the \McSplit\ algorithm for Max Common Subgraph.
-The new algorithm is more than an order of
-magnitude faster than the previous state of the art for unlabelled and undirected
-instances. We have shown how the algorithm can be extended for graphs with labels on edges, labels on vertices, loops, directed edges and the requirement that the resultant graph be connected.
+We have introduced the \McSplit\ algorithm for Max Common Subgraph.  The new
+algorithm is more than an order of magnitude faster than the previous state of
+the art for unlabelled and undirected instances. We have shown how the
+algorithm can be extended for graphs with labels on edges, labels on vertices,
+loops, directed edges and the requirement that the resultant graph be
+connected.
 
-\paragraph{Future work} We could create a hybrid algorithm. This would use \McSplit\ near the top of the search tree, switching to a
-clique encoding when fewer than some threshold number of vertices remain to be
-selected. This might deliver the benefits of the clique encoding for
-labelled graphs, while avoiding the high memory cost and colouring time of
-encoding the full instance.
+To further improve the state of the art, we could try creating a hybrid
+algorithm.  This could use $k\downarrow$ initially, switching to \McSplit\ when
+the extra filtering is ineffective, and finally switching to a clique encoding
+when fewer than some threshold number of vertices remain to be selected. This
+might deliver the benefits of the clique encoding for labelled graphs, while
+avoiding the high memory cost and colouring time of encoding the full instance.
 
-We have yet to fully investigate variable and value ordering heuristics, i.e.\ the selection of a label class and the selection of a vertex from that label class.
-There is also the potential to branch on both sides, that is instead of branching on vertices in $\V(G)$, it would be equally valid to branch on
-a vertex in $\V(H)$, since our data structure treats the
-two graphs symmetrically. More interestingly, we could choose which graph to branch on
-at each search node using some heuristic (perhaps choosing based on whether the $P$ or
-$T$ set is smaller), so long as we can ensure that the search remains complete.
+We believe there is more to be discovered about branching heuristics. There is
+also the potential to branch on both sides, that is instead of branching on
+vertices in $\V(G)$, it would be equally valid to branch on a vertex in
+$\V(H)$, since our data structure treats the two graphs symmetrically. More
+interestingly, we could choose which graph to branch on at each search node
+using some heuristic (perhaps choosing based on whether the $P$ or $T$ set is
+smaller), so long as we can ensure that the search remains complete.
 
-We have not yet explored parallel versions of the algorithm. The first paralellism is \emph{within-core}, using bitsets to encode sets. The second parallelism  is  multithreaded,
-exploited the ubiquitous multicore architecture.
-
-? Portfolio algorithms---finding out at what level of labelledness clique becomes best etc.
-
-%Application to induced subgraph isomorphism (with almost no changes). Applications
-%of a similar technique to other problems???
+It would be interesting to see whether these techniques are more broadly
+applicable. Most obviously, we could solve the induced subgraph isomorphism
+problem in the same way (and with nearly no changes to the code), but we could
+not solve the non-induced version. However, we suspect that some other problems
+may have a similar branching structure which would also benefit from a
+different domain store representation.
 

### 2017-02-16 paper/paper.tex
s/McCliqueAlg/McSplitAlg/g

@@
 \caption{Finding a maximum common subgraph.}
-\label{McCliqueAlg}
+\label{McSplitAlg}
 \end{algorithm}
 
-\paragraph{\cref{McCliqueAlg} in detail} The recursive procedure, $\FuncSty{expand}$, has
+\paragraph{\cref{McSplitAlg} in detail} The recursive procedure, $\FuncSty{expand}$, has
 two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
@@
        }
-    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McCliqueAlg} to handle directed and labelled cases.}
+    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and labelled cases.}
 \label{labDirAlg}
@@
 than explicitly looping over all label classes, as follows.  First, run lines
-17-19 of \cref{McCliqueAlg} to create a new label-class of vertices that are
+17-19 of \cref{McSplitAlg} to create a new label-class of vertices that are
 not adjacent to $v$ or $w$, and remove these vertices from $\langle P',T'
@@
 calling the main \FuncSty{McSplit} method once per goal size ($g, g-1, g-2,
-\dots$); we backtrack (\lineref{PruneSearch} of \cref{McCliqueAlg}) when the
+\dots$); we backtrack (\lineref{PruneSearch} of \cref{McSplitAlg}) when the
 bound is strictly less than the goal size, and terminate when a solution of the
@@
 %%% conventional domain stores, and replace \linerangeref{InnerLoop}{InnerLoopEnd}
-%%% of \cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains carried out in
+%%% of \cref{McSplitAlg} with \cref{cpAlg}.  The pruning of domains carried out in
 %%% these lines can be easily seen to be equivalent to the corresponding lines in
-%%% \cref{McCliqueAlg}.
+%%% \cref{McSplitAlg}.
 %%% 

### 2017-02-16 paper/paper.tex
Well spotted Ciaran

@@
 $\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
-if and only if $v$ and $w$ belong to the same label class in $M$ ?? shouldn't that be future?.
+if and only if $v$ and $w$ belong to the same label class in $\AlgVar{future}$.
 

### 2017-02-16 paper/paper.tex
Heatmaps are cool

@@
 recursive calls made by our algorithm on the $y$-axis versus the number made by
-CP-FC on the $x$-axis, for the labelled, unconnected problem instances. We see
-a close correlation: we typically do slightly less work, and sometimes do more,
-but instances with more than one order of magnitude difference in search space
-size are rare.
+CP-FC on the $x$-axis, for the labelled, unconnected problem instances. (Rather
+than a simple scatter plot, we use darker colours to indicate a higher density
+of points around a location.) We see a close correlation: we typically do
+slightly less work, and sometimes do more, but instances with more than one
+order of magnitude difference in search space size are rare.
 

### 2017-02-16 paper/heatmapify.rb

@@ -1,0 +1,31 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+buckets = Hash.new(0)
+granularity = 100
+
+xmul = ARGV[3].to_i
+ymul = ARGV[4].to_i
+
+IO.readlines(ARGV[0]).each_with_index do | line, index |
+    if index == 0 then next end
+    words = line.split(' ')
+
+    next if words[ARGV[1].to_i] == "NaN" or words[ARGV[2].to_i] == "NaN"
+
+    x = words[ARGV[1].to_i].to_f * xmul
+    bx = if x < 1 then 0 else (granularity * (Math.log(x + 0.1) / Math.log(1e10))).round end
+
+    y = words[ARGV[2].to_i].to_f * ymul
+    by = if y < 1 then 0 else (granularity * (Math.log(y + 0.1) / Math.log(1e10))).round end
+
+    abort "bx is " + bx.to_s + " by is " + by.to_s + " on line " + (index + 1).to_s if bx < 0 || bx > granularity || by < 0 || by > granularity
+    buckets[[bx, by]] += 1
+end
+
+0.upto granularity do | x |
+    0.upto granularity do | y |
+        print(buckets[[x, y]].to_s + " ")
+    end
+    puts
+end

### 2017-02-16 paper/heatmapify.rb
Coarser granularity looks better

@@ -2,7 +2,7 @@
 # vim: set sw=4 sts=4 et tw=80 :
 
 buckets = Hash.new(0)
-granularity = 100
+granularity = 50
 
 xmul = ARGV[3].to_i
 ymul = ARGV[4].to_i

### 2017-02-16 paper/paper.tex
Variety is the spice of life

@@
 \citet{DBLP:conf/cp/NdiayeS11}, but with much lower constant factors. Indeed,
-in \cref{figure:33ved-james-versus-cp-fc-nodes-scatter} we plot the number of
+in \cref{figure:prettyheatmaps1,figure:prettyheatmaps2} we plot the number of
 recursive calls made by our algorithm on the $y$-axis versus the number made by
@@
 
-\begin{figure}[t]
-    \centering
+\begin{figure*}[t]
+    \centering\subfiguretopcaptrue
+    \subfigure[][Unlabelled, undirected instances] {\label{figure:prettyheatmaps1}
+    \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
+    }\subfigure[][Labelled, directed instances] {\label{figure:prettyheatmaps2}
     \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
-    \caption{Size of the search space for \McSplit\ versus CP-FC on the
-    labelled, directed, unconnected instances which were solved within the
-    timeout by both algorithms.}
-    \label{figure:33ved-james-versus-cp-fc-nodes-scatter}
-\end{figure}
+    }\subfigure[][Subgraph isomorphism instances]{\label{figure:prettyheatmaps3}
+    \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
+}
+    \caption{Relative search space sizes for instances which were solved by both algorithms within the timeout.}
+    \label{figure:prettyheatmaps}
+\end{figure*}
 
@@
 What about our relationship to the $k\downarrow$ of \citet{UpcomingAAAIPaper}?
-\Cref{figure:sip-james-versus-kdown-nodes-scatter} plots the number of
+\Cref{figure:prettyheatmaps3} plots the number of
 recursive calls made by $k\downarrow$ and \McSplit\ on each of the subgraph
@@
 ?? The concentration of points, down to unit propagation and counting
-
-\begin{figure}[h]
-    \centering
-    \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
-    \caption{Size of the search space for \McSplit$\downarrow$\ versus $k\downarrow$ on the
-    large subgraph isomorphism instances which were solved within the
-    timeout by both algorithms.}\label{figure:sip-james-versus-kdown-nodes-scatter}
-\end{figure}
 

### 2017-02-16 paper/paper.tex
Line range macro: choose word and or to

@@
 \newcommand{\lineref}[1]{line~\ref{#1}}
-\newcommand{\linerangeref}[2]{lines~\ref{#1} to~\ref{#2}}
+\newcommand{\linerangeref}[2]{\count255=\ref{#1}\advance\count255 by 1 \ifnum\count255=\ref{#2}lines~\ref{#1} and~\ref{#2}\else lines~\ref{#1} to~\ref{#2}\fi}
 \newcommand{\Lineref}[1]{Line~\ref{#1}}
-\newcommand{\Linerangeref}[2]{Lines~\ref{#1} to~\ref{#2}}
+\newcommand{\Linerangeref}[2]{\count255=\ref{#1}\advance\count255 by 1 \ifnum\count255=\ref{#2}Lines~\ref{#1} and~\ref{#2}\else Lines~\ref{#1} to~\ref{#2}\fi}
 
@@
 %\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
-\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
-\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
-\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
+\medskip
+\nl $\AlgVar{bound} \gets |M|  + \sum_{\langle P,T \rangle \in \AlgVar{future}} \min(|P|,|T|)$ \label{CalcBound} \;
+\nl \lIf {$\AlgVar{bound} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
 \medskip
@@
 
-\paragraph{\cref{McSplitAlg} in detail} The recursive procedure, $\FuncSty{expand}$, has
-two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
-each represented as a $\langle P, T \rangle$ pair as described above.  The
-parameter $M$ is the current mapping of vertices.  On each call to
-$\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
-if and only if $v$ and $w$ belong to the same label class in $\AlgVar{future}$.
+\paragraph{\cref{McSplitAlg} in detail} The recursive procedure,
+$\FuncSty{expand}$, has two parameters.  The parameter $\AlgVar{future}$ is a
+list of label classes, each represented as a $\langle P, T \rangle$ pair as
+described above.  The parameter $M$ is the current mapping of vertices.  On
+each call to $\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be
+added to $M$ if and only if $v$ and $w$ belong to the same label class in
+$\AlgVar{future}$.
 
-\Lineref{StoreIncumbent} stores the current mapping $M$ if it is large enough
-to unseat the incumbent.  \Linerangeref{SetBestFutureToZero}{PruneSearch} prune
-the search when a calculated upper bound (best future $\AlgVar{bf}$ plus $|M|$) is not larger
-than the incumbent ?? rewrite this as one line using $\sum$?.
+\Lineref{StoreIncumbent} stores the current mapping $M$ if it is larger than
+the incumbent.  \Linerangeref{CalcBound}{PruneSearch} prune the search when a
+calculated upper bound is not large enough to unseat the incumbent.
 

### 2017-02-16 paper/paper.tex
Minor re-ordering of words

@@
 
-\Lineref{StoreIncumbent} stores the current mapping $M$ if it is larger than
-the incumbent.  \Linerangeref{CalcBound}{PruneSearch} prune the search when a
-calculated upper bound is not large enough to unseat the incumbent.
+\Lineref{StoreIncumbent} stores the current mapping $M$ if it is large enough
+to unseat the incumbent.  \Linerangeref{CalcBound}{PruneSearch} prune the
+search when a calculated upper bound is not larger than the incumbent.
 

### 2017-02-16 paper/paper.tex
Bron Kerbosch citation
Muck around with citations

@@
 \citep{DBLP:journals/jcamd/RaymondW02a,Ehrlich:2011,DAM2014,Grindley1993707},
-computer vision \cite{DBLP:journals/jair/CookH94}, discovery of repeated
-patterns in source code \cite{DBLP:journals/tkde/DjokoCH97} and circuits
-\cite{DBLP:journals/jair/CookH94}, and recognition of handwritten characters
-\cite{SIWEILU1991617}.
+and also in other domains including malware detection
+\citep{DBLP:journals/compsec/ParkRS13}, source code analysis
+\cite{DBLP:journals/tkde/DjokoCH97}, and computer vision
+\cite{DBLP:journals/jair/CookH94}.
 
@@
 to the maximum clique problem \citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and by
-adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. This paper
-introduces a new branch and bound algorithm which exploits special properties
-of these problems to allow a much faster exploration of the search space,
-whilst retaining the filtering and bounding benefits of the constraint
+adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. Some
+special cases also have practial polynomial time algorithms
+\citep{DBLP:conf/mfcs/DroschinskyKM16,DBLP:conf/sofsem/DroschinskyKM17}.
+
+This paper introduces a new branch and bound algorithm which exploits special
+properties of these problems to allow a much faster exploration of the search
+space, whilst retaining the filtering and bounding benefits of the constraint
 programming approach. We describe the algorithm for the basic maximum common
@@
 they are used in the first pair of arrays. ?? This representation has
-similarities to data structures used in partition backtracking and bron
-kerbosch.
+similarities to data structures used in partition backtracking and the \citet{DBLP:journals/cacm/BronK73} clique enumeration algorithm.
 

### 2017-02-16 paper/paper.tex
Page layout mucking around
Remove the funding note for anon review, put it back in later and worry about page layout then
Last one

@@
 
-\title{A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work
-was supported by a pile of money from a funding agency.}}
+\title{A Partitioning Algorithm for Maximum Common Subgraph Problems}
 \author{The Anonymous Authors of Paper 2431 \\
@@
 anonymous@anonymous.anonymous}
+
+% \thanks{This work was supported by a pile of money from a funding agency.}
 
@@
 
+\begin{algorithm}[t]
+\DontPrintSemicolon
+\nl    \For {$l \in L$}{
+\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} : u \neq v \wedge A_G[v][u] = l \}$ \;
+\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} : u \neq w \wedge A_H[w][u] = l \}$ \;
+\nl        \If {$P^{''} \negmedspace\neq \emptyset$ \bf{and} $T^{''} \negmedspace \neq \emptyset$}{
+    \nl $\AlgVar{future^{'}} \gets \AlgVar{future^{'}} + \langle P^{''}\negmedspace, T^{''} \rangle$}
+       }
+    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and labelled cases.}
+\label{labDirAlg}
+\end{algorithm}
+
 ?? This needs a two sentence intro saying that people care about stuff and why.
@@
 $L=\{0,1,2,3\}$.
-
-\begin{algorithm}[t]
-\DontPrintSemicolon
-\nl    \For {$l \in L$}{
-\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} : u \neq v \wedge A_G[v][u] = l \}$ \;
-\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} : u \neq w \wedge A_H[w][u] = l \}$ \;
-\nl        \If {$P^{''} \negmedspace\neq \emptyset$ \bf{and} $T^{''} \negmedspace \neq \emptyset$}{
-    \nl $\AlgVar{future^{'}} \gets \AlgVar{future^{'}} + \langle P^{''}\negmedspace, T^{''} \rangle$}
-       }
-    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and labelled cases.}
-\label{labDirAlg}
-\end{algorithm}
 
@@
 
-Our results so far suggest that \McSplit\ has broadly similar performance
-trends to the constraint programming, forward-checking (CP-FC) algorithm of
-\citet{DBLP:conf/cp/NdiayeS11}, but with much lower constant factors. Indeed,
-in \cref{figure:prettyheatmaps1,figure:prettyheatmaps2} we plot the number of
-recursive calls made by our algorithm on the $y$-axis versus the number made by
-CP-FC on the $x$-axis, for the labelled, unconnected problem instances. (Rather
-than a simple scatter plot, we use darker colours to indicate a higher density
-of points around a location.) We see a close correlation: we typically do
-slightly less work, and sometimes do more, but instances with more than one
-order of magnitude difference in search space size are rare.
-
-Why is this? We do not see a similar correlation between our search space size
-and that of the clique spproach. The key observation is that \McSplit\ may be
-considered to be a different version of the constraint CP-FC algorithm, using
-an unconventional domain store and more efficient filtering algorithms. We now
-explore this relationship further.
-
 \begin{figure*}[t]
@@
 \end{figure*}
+
+Our results so far suggest that \McSplit\ has broadly similar performance
+trends to the constraint programming, forward-checking (CP-FC) algorithm of
+\citet{DBLP:conf/cp/NdiayeS11}, but with much lower constant factors. Indeed,
+in \cref{figure:prettyheatmaps1,figure:prettyheatmaps2} we plot the number of
+recursive calls made by our algorithm on the $y$-axis versus the number made by
+CP-FC on the $x$-axis, for the unlabelled and labelled, unconnected problem
+instances. (Rather than a simple scatter plot, we use darker colours to
+indicate a higher density of points around a location.) We see a close
+correlation: we typically do slightly less work, and sometimes do more, but
+instances with more than one order of magnitude difference in search space size
+are rare.
+
+Why is this? We do not see a similar correlation between our search space size
+and that of the clique spproach. The key observation is that \McSplit\ may be
+considered to be a different version of the constraint CP-FC algorithm, using
+an unconventional domain store and more efficient filtering algorithms. We now
+explore this relationship further.
 

### 2017-02-16 paper/paper.tex
Spelling
Promise URL for code

@@
 Experiments were performed on machines with dual Intel Xeon E5-2640 v2 CPUs and
-64GBytes RAM. Our algorithm was implemented in C++ and compiled using g++
-5.3.0. For comparison, we compare against the best constraint programming
-implementations of \citet{DBLP:conf/cp/NdiayeS11} and
-\citet{DBLP:conf/cp/McCreeshNPS16} (CP-FC in the unlabelled cases, and CP-MAC
-in the labelled cases, using both branching and filtering for connected
-subgraphs), the clique encodings of \citet{DBLP:conf/cp/McCreeshNPS16}, and the
-$k\downarrow$ algorithm of \citet{UpcomingAAAIPaper} (which only supports
-unlabelled, undirected, unconnected instances).
+64GBytes RAM. Our algorithm was implemented\footnote{URL removed for anonymous
+review} in C++ and compiled using g++ 5.3.0. For comparison, we compare against
+the best constraint programming implementations of
+\citet{DBLP:conf/cp/NdiayeS11} and \citet{DBLP:conf/cp/McCreeshNPS16} (CP-FC in
+the unlabelled cases, and CP-MAC in the labelled cases, using both branching
+and filtering for connected subgraphs), the clique encodings of
+\citet{DBLP:conf/cp/McCreeshNPS16}, and the $k\downarrow$ algorithm of
+\citet{UpcomingAAAIPaper} (which only supports unlabelled, undirected,
+unconnected instances).
 
@@
 Why is this? We do not see a similar correlation between our search space size
-and that of the clique spproach. The key observation is that \McSplit\ may be
+and that of the clique approach. The key observation is that \McSplit\ may be
 considered to be a different version of the constraint CP-FC algorithm, using

### 2017-02-16 paper/paper.tex
Shrunk example tables and moved subcaptions to top
Simplified example graph commands

@@
 
-\newcommand{\examplexG}[5] {
+\newcommand{\exampleG} {
     \begin{minipage}{.2\textwidth}
@@
                 clockwise=5] {
-                    1[label=90:#1],2[label=0:#2],3[label=0:#3],4[label=180:#4],5[label=180:#5];
+                    1,2,3,4,5;
             1--4; 1--5; 2--3; 2--5; 3--5;
@@
 }
-\newcommand{\examplexH}[6] {
+\newcommand{\exampleH} {
     \begin{minipage}{.2\textwidth}
@@
                 clockwise=6, phase=60] {
-                    a[label=0:#1],b[label=0:#2],c[label=0:#3],d[label=180:#4],e[label=180:#5],f[label=180:#6];
+                    a,b,c,d,e,f;
             a--b; a--c; a--e; b--d; b--f; c--d; c--e; c--f; d--f; e--f;
@@
 
-\newcommand{\LabelTables}[4] {
+\newcommand{\LabelTables}[3] {
+  {\footnotesize
     \centering
-  \begin{minipage}[t]{#1\linewidth}
+    \begin{minipage}[t]{.17\linewidth}
         Mapping
 
-      \bigskip
+        \bigskip\vspace{-4pt}
 
-      #2
+        #1
     \end{minipage}
     \quad
-  \begin{minipage}[t]{0.3\linewidth}
+    \begin{minipage}[t]{0.31\linewidth}
         \centering
-      Labelling of $G$
+        Labelling of $G$\vspace{-4pt}
+        \begin{tabular}[t]{cc}
+        \toprule
+            Vertex & Label\\
+        \midrule
+            #2
+        \bottomrule
+        \end{tabular}
+    \end{minipage}
+    \quad
+    \begin{minipage}[t]{0.31\linewidth}
+        \centering
+        Labelling of $H$\vspace{-4pt}
         \begin{tabular}[t]{cc}
@@
         \end{tabular}
-  \end{minipage}
-  \quad
-  \begin{minipage}[t]{0.3\linewidth}
-      \centering
-      Labelling of $H$
-      \begin{tabular}[t]{cc}
-      \toprule
-          Vertex & Label\\
-      \midrule
-          #4
-      \bottomrule
-      \end{tabular}
         \medskip
     \end{minipage}
+  }
 }
@@
 \centering
-    \examplexG{}{}{}{}{}
-    \examplexH{}{}{}{}{}{}
+    \exampleG
+    \exampleH
 \caption{Example graphs $G$ and $H$.}
@@
 \begin{figure}[h]
-    \centering
+    \centering\subfiguretopcaptrue
     \subfigure[][After mapping $1$ to $a$] {
-      \LabelTables{.15}
-                  {$\{1a\}$}
+      \LabelTables{$\{1a\}$}
                   {$2$ & 0 \\
@@
     \subfigure[][After mapping $2$ to $d$] {
-      \LabelTables{.15}
-                  {$\{1a,2d\}$}
+      \LabelTables{$\{1a,2d\}$}
                   {$3$ & 01 \\
@@
     \subfigure[][After mapping $3$ to $f$] {
-      \LabelTables{.2}
-                  {$\{1a,2d,3f\}$}
+      \LabelTables{$\{1a,2d,3f\}$}
                   {$4$ & 100 \\
@@
 \begin{figure}[t]
-    \centering
+    \centering\subfiguretopcaptrue
     \subfigure[][Unlabelled, undirected instances] {
@@
 \begin{figure}[t]
-    \centering
+    \centering\subfiguretopcaptrue
     \subfigure[][Unlabelled, undirected, connected instances] {

### 2017-02-17 paper/paper.tex
Un-superscripted primes on variables
Removed a word that shouldn't have been there

@@
 \nl    \For {$l \in L$}{
-\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} : u \neq v \wedge A_G[v][u] = l \}$ \;
-\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} : u \neq w \wedge A_H[w][u] = l \}$ \;
-\nl        \If {$P^{''} \negmedspace\neq \emptyset$ \bf{and} $T^{''} \negmedspace \neq \emptyset$}{
-    \nl $\AlgVar{future^{'}} \gets \AlgVar{future^{'}} + \langle P^{''}\negmedspace, T^{''} \rangle$}
+\nl        \bf{Let} $P'' \gets \{ u \in P' : u \neq v \wedge A_G[v][u] = l \}$ \;
+\nl        \bf{Let} $T'' \gets \{ u \in T' : u \neq w \wedge A_H[w][u] = l \}$ \;
+\nl        \If {$P'' \negmedspace\neq \emptyset$ \bf{and} $T'' \negmedspace \neq \emptyset$}{
+    \nl $\AlgVar{future'} \gets \AlgVar{future'} + \langle P''\negmedspace, T'' \rangle$}
        }
@@
 
-In the CP-FC algorithm, each vertex in $v \in \V(G)$ is represented by a
+In the CP-FC algorithm, each vertex $v \in \V(G)$ is represented by a
 variable, whose domain corresponds to the set of vertices in $\V(H)$ to which

### 2017-02-17 paper/paper.tex
Fixed centring of example graphs G and H

@@
 \newcommand{\exampleG} {
-    \begin{minipage}{.2\textwidth}
     \tikz {
@@
     }
-    \end{minipage}
 }
 \newcommand{\exampleH} {
-    \begin{minipage}{.2\textwidth}
     \tikz {
@@
     }
-    \end{minipage}
 }
@@
     \exampleG
+    \qquad
     \exampleH

### 2017-02-17 more-analysis/on-how-many-instances-is-mcsplit-faster.sh
I'd forgotten to include kdown in the test of how many instance McSplit was fastest on

@@ -1,4 +1,7 @@
-echo "Number of instances where at least one of clique, CP-FC, McClique didn't time out"
-cat ../experiments/gpgnode-results/mcsplain/runtimes.data | awk 'NR>1 && ($2<1000000 || $3<1000 || $8<1000000)' | wc -l
-echo "Number of these where McClique wan't strictly the fastest"
-cat ../experiments/gpgnode-results/mcsplain/runtimes.data | awk 'NR>1 && ($2<1000000 || $3<1000 || $8<1000000) && ($2<=$8 || $3*1000<=$8)' | wc -l
+echo "Number of instances where at least one of clique, CP-FC, McSplit didn't time out"
+cat ../experiments/gpgnode-results/mcsplain/runtimes.data | \
+  awk 'NR>1 && ($2<1000000 || $3<1000 || $5<1000000 || $8<1000000)' | wc -l
+
+echo "Number of these where McSplit wan't strictly the fastest"
+cat ../experiments/gpgnode-results/mcsplain/runtimes.data | \
+  awk 'NR>1 && ($2<1000000 || $3<1000 || $5<1000000 || $8<1000000) && ($2<=$8 || $3*1000<=$8 || $5<=$8)' | wc -l

### 2017-02-17 paper/paper.tex
Spelling
Added kdown to text about the instances where McSplit is fastest

@@
 adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. Some
-special cases also have practial polynomial time algorithms
+special cases also have practical polynomial time algorithms
 \citep{DBLP:conf/mfcs/DroschinskyKM16,DBLP:conf/sofsem/DroschinskyKM17}.
@@
 number of instances, \McSplit\ is comfortably more than an order of magnitude
-faster than its nearer competitor.  Moreover, there are few instances in this
+faster than its nearer competitor  Moreover, there are few instances in this
 set for which \McSplit\ is not the fastest solver; of the 3506 instances that
-could be solved by at least one of \McSplit, CP-FC and clique in less than than
-1000 seconds, there were only 126 instances on which \McSplit\ was not strictly
+could be solved by at least one of \McSplit, $k\downarrow$, CP-FC and clique in less than than
+1000 seconds, there were 426 instances on which \McSplit\ was not strictly
 faster than the other two solvers.

### 2017-02-17 paper/paper.tex
Added commas to 4-digit numbers

@@
 the speed of two algorithms using the horizontal distance between their curves.
-For example, we could solve 2000 of the 4110 unlabelled undirected instances
+For example, we could solve 2,000 of the 4,110 unlabelled undirected instances
 using the \McSplit\ algorithm if a time limit of 0.5 seconds per instance were
@@
 faster than its nearer competitor  Moreover, there are few instances in this
-set for which \McSplit\ is not the fastest solver; of the 3506 instances that
+set for which \McSplit\ is not the fastest solver; of the 3,506 instances that
 could be solved by at least one of \McSplit, $k\downarrow$, CP-FC and clique in less than than
-1000 seconds, there were 426 instances on which \McSplit\ was not strictly
+1,000 seconds, there were 426 instances on which \McSplit\ was not strictly
 faster than the other two solvers.
@@
 \paragraph{Large subgraph isomorphism instances} We also ran the algorithms on
-a set of 5725 larger instances used in recent studies of subgraph
+a set of 5,725 larger instances used in recent studies of subgraph
 isomorphism~\citep{DBLP:conf/lion/KotthoffMS16} and maximum common
@@
 vertices to 900 with a median of 80; target graphs range from 10 vertices to
-6671 with a median of 561. Cumulative runtimes on these instances are shown in
+6,671 with a median of 561. Cumulative runtimes on these instances are shown in
 \cref{figure:sip-cumulative}.  This is a challenging set of instances, and more
-than half of the instances cannot be solved within a timeout of 1000 seconds by
+than half of the instances cannot be solved within a timeout of 1,000 seconds by
 any solver. Furthermore, the CP-FC algorithm and the clique encoding run out of

### 2017-02-17 paper/paper.tex
Added page break before bibliography
Use mathcal graph names and G,H instead of P,T

@@
 \newcommand{\nmax}{n_{\max}}
+
+\newcommand{\graphG}{\mathcal{G}}
+\newcommand{\graphH}{\mathcal{H}}
+\newcommand{\setG}{G}
+\newcommand{\setH}{H}
 
@@
         \centering
-        Labelling of $G$\vspace{-4pt}
+        Labelling of $\graphG$\vspace{-4pt}
         \begin{tabular}[t]{cc}
@@
         \centering
-        Labelling of $H$\vspace{-4pt}
+        Labelling of $\graphH$\vspace{-4pt}
         \begin{tabular}[t]{cc}
@@
 (\cref{sec:extensions} describes how these restrictions may be relaxed).
-The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.  The
-set of vertices adjacent to vertex $v$ in graph $G$ is called the
-\emph{neighbourhood} of $v$, denoted $\N(G, v)$. We denote by $\invN(G, v)$ the
+The vertices and edge sets of a graph $\graphG$ are denoted $\V(\graphG)$ and $\E(\graphG)$.  The
+set of vertices adjacent to vertex $v$ in graph $\graphG$ is called the
+\emph{neighbourhood} of $v$, denoted $\N(\graphG, v)$. We denote by $\invN(\graphG, v)$ the
 \emph{inverse neighbourhood} of $v$, being the set of the vertices not adjacent
-to $v$ (excluding $v$ itself). A \emph{subgraph} of a graph $G$ is a graph
-consisting of some of the vertices of $G$, and all of the edges between these
+to $v$ (excluding $v$ itself). A \emph{subgraph} of a graph $\graphG$ is a graph
+consisting of some of the vertices of $\graphG$, and all of the edges between these
 vertices. A \emph{common subgraph} of two graphs is a graph which is
@@
 
-Let $G$ and $H$ be the two input graphs to our maximum common subgraph problem.
+Let $\graphG$ and $\graphH$ be the two input graphs to our maximum common subgraph problem.
 The orders (vertex counts) of these graphs are denoted $g$ and $h$.
@@
 algorithm which uses a new vertex-labelling scheme to avoid explicitly
-maintaining the set of feasible mappings for each vertex in $\V(G)$.
+maintaining the set of feasible mappings for each vertex in $\V(\graphG)$.
 
 \McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
-(v_{m}, w_{m})\}$ with $|M^*| = m$ vertex pairs, where the $v_i$ are distinct members of $\V(G)$ and the
-$w_i$ are distinct vertices of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
-in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. 
-Given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{m}\}$
-and the subgraph of $H$ induced by $\{w_i, \dots, w_{m}\}$ are a maximum
+(v_{m}, w_{m})\}$ with $|M^*| = m$ vertex pairs, where the $v_i$ are distinct members of $\V(\graphG)$ and the
+$w_i$ are distinct vertices of $\V(\graphH)$, such that $v_i$ and $v_j$ are adjacent
+in $\graphG$ if and only if $w_i$ and $w_j$ are adjacent in $\graphH$. 
+Given such a mapping, the subgraph of $\graphG$ induced by $\{v_i, \dots, v_{m}\}$
+and the subgraph of $\graphH$ induced by $\{w_i, \dots, w_{m}\}$ are a maximum
 common subgraph.
@@
 \paragraph{Walkthrough} Before discussing the algorithm in detail, we start
-with an illustration of the main concepts using the graphs $G$ and $H$ in
+with an illustration of the main concepts using the graphs $\graphG$ and $\graphH$ in
 \cref{fig:alg1}.  These graphs have a maximum common subgraph with four
@@
     \exampleH
-\caption{Example graphs $G$ and $H$.}
+\caption{Example graphs $\graphG$ and $\graphH$.}
 \label{fig:alg1}
@@
 The algorithm builds up a mapping $M$, starting with the empty mapping
-$\emptyset$. Select a vertex in $\V(G)$ as the first vertex to be mapped; in
+$\emptyset$. Select a vertex in $\V(\graphG)$ as the first vertex to be mapped; in
 our example we will arbitrarily choose vertex $1$. Each of the vertices in
-$\V(H)$ to which vertex $1$ may be mapped will be tried in turn, and finally the
+$\V(\graphH)$ to which vertex $1$ may be mapped will be tried in turn, and finally the
 possibility where vertex $1$ remains unmatched will be tried.
@@
 each unmatched vertex according to whether it is adjacent to vertex $1$ (in
-$G$) or vertex $a$ (in $H$), as shown in \cref{fig:alg2}.  Adjacent
+$\graphG$) or vertex $a$ (in $\graphH$), as shown in \cref{fig:alg2}.  Adjacent
 vertices have label 1; non-adjacent vertices have label 0.  We can extend $M$
-with a vertex mapping $vw$, with $v \in \V(G)$ and $w \in \V(H)$, if and only
+with a vertex mapping $vw$, with $v \in \V(\graphG)$ and $w \in \V(\graphH)$, if and only
 if $v$ and $w$ have the same label.  This property, that two vertices may be
@@
 
-    \caption{Mapping $M$ and vertex labels during search on example graphs $G$ and $H$ from \cref{fig:alg1}.}
+    \caption{Mapping $M$ and vertex labels during search on example graphs $\graphG$ and $\graphH$ from \cref{fig:alg1}.}
     \label{figure:mcsplit-examples}
@@
 
-Next, extend the mapping by pairing a vertex in $G$ with a vertex in $H$ of the
+Next, extend the mapping by pairing a vertex in $\graphG$ with a vertex in $\graphH$ of the
 same label; we will choose to map vertex $2$ to vertex $d$, giving $M=\{1a,
-2d\}$ (\cref{fig:alg3}).  Each unmapped vertex $v \in \V(G)$ is labelled
+2d\}$ (\cref{fig:alg3}).  Each unmapped vertex $v \in \V(\graphG)$ is labelled
 with a two-character bit string, indicating its adjacency to each of
-the two mapped vertices in $\V(G)$ (vertices $1$ and $2$).  For example, vertex
+the two mapped vertices in $\V(\graphG)$ (vertices $1$ and $2$).  For example, vertex
 $3$ is labelled $01$, indicating that it is not adjacent to vertex $1$ but is adjacent
-to vertex $2$.  Labels are given to unmapped vertices in $\V(H)$ in a similar fashion,
+to vertex $2$.  Labels are given to unmapped vertices in $\V(\graphH)$ in a similar fashion,
 showing adjacency to matched vertices $a$ and $d$.  Our invariant is
@@
 there is no way to add a pair of vertices with label 100 or 101 to the mapping.
-The final label, 111, appears once in $G$ and twice in $H$, and therefore at
+The final label, 111, appears once in $\graphG$ and twice in $\graphH$, and therefore at
 most one pair with this label can be added to $M$.  Thus, the upper bound on
@@
 \begin{multline*}
-    \mathit{bound} = |M| + \sum_{l \in L} \min\big(|\{ v \in \V(G) : \vtxlabel(v)=l\}|, \\[-0.3cm]
-        |\{ v \in \V(H) : \vtxlabel(v)=l \}|\big) \text{,}
+    \mathit{bound} = |M| + \sum_{l \in L} \min\big(|\{ v \in \V(\graphG) : \vtxlabel(v)=l\}|, \\[-0.3cm]
+        |\{ v \in \V(\graphH) : \vtxlabel(v)=l \}|\big) \text{,}
 \end{multline*} where $L$ is the set of labels used in both graphs.
@@
 %left unmatched, by giving $2$ the label $\bot$ and selecting another vertex in
-%$\V(G)$ to assign.
+%$\V(\graphG)$ to assign.
 
@@
 search tree to store labelling information.  This is done by storing a
-\emph{label class} as a pair $\langle P,T \rangle$ for each label $l$ that is
-used, where $P$ is the set of vertices in $\V(G)$ labelled $l$, and $T$ is the
-set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology from
-subgraph isomorphism, where one graph is a pattern graph $P$ and the other the
-target graph $T$.}. Since there are $g + h$ vertices in the two graphs, at most
+\emph{label class} as a pair $\langle \setG,\setH \rangle$ for each label $l$ that is
+used, where $\setG$ is the set of vertices in $\V(\graphG)$ labelled $l$, and $\setH$ is the
+set of vertices in $\V(\graphH)$ labelled $l$\footnote{We adopt terminology from
+subgraph isomorphism, where one graph is a pattern graph $\setG$ and the other the
+target graph $\setH$.}. Since there are $g + h$ vertices in the two graphs, at most
 $g + h$ label classes can exist at once, and there are at most $g + h$ vertices
-in the union of all of the $P$ and $T$ sets. Furthermore, we do not actually
+in the union of all of the $\setG$ and $\setH$ sets. Furthermore, we do not actually
 need to store the bits making up a label---we care only that like-labelled
@@
 necessary information in two pairs of flat arrays. One pair of arrays is used
-to store a permutation of the vertices in $P$ having like-labelled vertices
+to store a permutation of the vertices in $\setG$ having like-labelled vertices
 stored consecutively, together with the indices of the start of each label in
 turn (in some arbitrary order). The second pair stores the permutation of
-vertices in $T$, together with label start indices \emph{in the same order} as
+vertices in $\setH$, together with label start indices \emph{in the same order} as
 they are used in the first pair of arrays. ?? This representation has
@@
 \medskip
-\nl $\AlgVar{bound} \gets |M|  + \sum_{\langle P,T \rangle \in \AlgVar{future}} \min(|P|,|T|)$ \label{CalcBound} \;
+\nl $\AlgVar{bound} \gets |M|  + \sum_{\langle \setG,\setH \rangle \in \AlgVar{future}} \min(|\setG|,|\setH|)$ \label{CalcBound} \;
 \nl \lIf {$\AlgVar{bound} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
 \medskip
-\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
-\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
-\nl \For {$w \in T$ \label{WLoop}} {
+\nl $\langle \setG,\setH \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
+\nl $v \gets \FuncSty{SelectVertex}(\setG)$ \label{SelectVertex} \;
+\nl \For {$w \in \setH$ \label{WLoop}} {
     \nl    $M \gets M \cup \{(v,w)\}$ \label{GrowM} \;
 \nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
-\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
-\nl        \bf{Let} $P'' \gets P' \cap \N(G, v) \setminus \{v\}$ \label{NewPWithEdge} \;
-\nl        \bf{Let} $T'' \gets T' \cap \N(H, w) \setminus \{w\}$ \;
-\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
-\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
-\nl        \bf{Let} $P'' \gets P' \cap \invN(G, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
-\nl        \bf{Let} $T'' \gets T' \cap \invN(H, w) \setminus \{w\}$ \;
-\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
-\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
+\nl    \For {$\langle \setG',\setH'\rangle \in future$ \label{InnerLoop}}{
+\nl        \bf{Let} $\setG'' \gets \setG' \cap \N(\graphG, v) \setminus \{v\}$ \label{NewPWithEdge} \;
+\nl        \bf{Let} $\setH'' \gets \setH' \cap \N(\graphH, w) \setminus \{w\}$ \;
+\nl        \If {$\setG'' \neq \emptyset$ \bf{and} $\setH'' \neq \emptyset$}{
+\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle \setG'' , \setH'' \rangle$ \label{AddToFutureWithEdge}}
+\nl        \bf{Let} $\setG'' \gets \setG' \cap \invN(\graphG, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
+\nl        \bf{Let} $\setH'' \gets \setH' \cap \invN(\graphH, w) \setminus \{w\}$ \;
+\nl        \If {$\setG'' \neq \emptyset$ \bf{and} $\setH'' \neq \emptyset$}{
+\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle \setG'' , \setH'' \rangle$} \label{InnerLoopEnd}
        }
@@
   }
-\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
-\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
-\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
+\nl $\setG' \gets \setG \setminus \{v\}$ \label{RemoveV} \;
+\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle \setG,\setH \rangle\}$\;
+\nl \lIf {$\setG' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle \setG',\setH \rangle \}$}
 \nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
@@
 \;
-\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
+\nl $\FuncSty{McSplit}(\graphG,\graphH)$ \label{McSplitFun} \;
 \nl \Begin{
 \nl $\AlgVar{incumbent} \gets \emptyset$ \;
-\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
+\nl $\FuncSty{expand}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
 \nl return $|\AlgVar{incubent}|$ \;
@@
 $\FuncSty{expand}$, has two parameters.  The parameter $\AlgVar{future}$ is a
-list of label classes, each represented as a $\langle P, T \rangle$ pair as
+list of label classes, each represented as a $\langle \setG, \setH \rangle$ pair as
 described above.  The parameter $M$ is the current mapping of vertices.  On
@@
 The remainder of $\FuncSty{expand}$ performs the search.  A label class
-$\langle P, T \rangle$ is selected from $\AlgVar{future}$, where $P \subseteq
-V(G)$ and $T \subseteq V(H)$, using some heuristic (\lineref{SelectClass}).
-From this label class, a vertex $v$ is selected from $P$ using a heuristic
-(\lineref{SelectVertex}). We now iterate over all possible vertices in $T$
-(\linerangeref{WLoop}{ShrinkM}). A vertex $w \in T$ is selected and mapping $M$
+$\langle \setG, \setH \rangle$ is selected from $\AlgVar{future}$, where $\setG \subseteq
+V(\graphG)$ and $\setH \subseteq V(\graphH)$, using some heuristic (\lineref{SelectClass}).
+From this label class, a vertex $v$ is selected from $\setG$ using a heuristic
+(\lineref{SelectVertex}). We now iterate over all possible vertices in $\setH$
+(\linerangeref{WLoop}{ShrinkM}). A vertex $w \in \setH$ is selected and mapping $M$
 is extended (\lineref{GrowM}). A new set of label-classes, $\AlgVar{future'}$,
@@
 first of these classes (\linerangeref{NewPWithEdge}{AddToFutureWithEdge})
-contains vertices in $P$ adjacent to $v$ and vertices in $T$ adjacent to $w$.
+contains vertices in $\setG$ adjacent to $v$ and vertices in $\setH$ adjacent to $w$.
 This is added to $\AlgVar{future'}$ if both sets contain at least one vertex.
@@
 (\lineref{ExpandWithV}), on return from which we remove the mapping $(v,w)$.
-Having explored all possible mappings of $v$ with vertices in $T$ we now
+Having explored all possible mappings of $v$ with vertices in $\setH$ we now
 consider what happens if $v$ is not matched
@@
 We start our search at the function $\FuncSty{McSplit}$ (\lineref{McSplitFun}),
-with graphs $G$ and $H$ as inputs.  This function returns the size of the
+with graphs $\graphG$ and $\graphH$ as inputs.  This function returns the size of the
 largest mapping.  In \lineref{FirstExpandCall} the initial call is made to
@@
 were performed to identify a suitable heuristic for label class selection. The
-heuristic used selects the label class  with the smallest $\min(|P|,|T|)$ (a
+heuristic used selects the label class  with the smallest $\min(|\setG|,|\setH|)$ (a
 surrogate for expected branching factor). From this class, a vertex is selected
-from $P$ with maximum degree if $H$ is sparse and with minimum degree if $H$ is
+from $\setG$ with maximum degree if $\graphH$ is sparse and with minimum degree if $\graphH$ is
 dense.
@@
 \nl    \For {$l \in L$}{
-\nl        \bf{Let} $P'' \gets \{ u \in P' : u \neq v \wedge A_G[v][u] = l \}$ \;
-\nl        \bf{Let} $T'' \gets \{ u \in T' : u \neq w \wedge A_H[w][u] = l \}$ \;
-\nl        \If {$P'' \negmedspace\neq \emptyset$ \bf{and} $T'' \negmedspace \neq \emptyset$}{
-    \nl $\AlgVar{future'} \gets \AlgVar{future'} + \langle P''\negmedspace, T'' \rangle$}
+\nl        \bf{Let} $\setG'' \gets \{ u \in \setG' : u \neq v \wedge A_G[v][u] = l \}$ \;
+\nl        \bf{Let} $\setH'' \gets \{ u \in \setH' : u \neq w \wedge A_H[w][u] = l \}$ \;
+\nl        \If {$\setG'' \negmedspace\neq \emptyset$ \bf{and} $\setH'' \negmedspace \neq \emptyset$}{
+    \nl $\AlgVar{future'} \gets \AlgVar{future'} + \langle \setG''\negmedspace, \setH'' \rangle$}
        }
@@
 \paragraph{Vertex labels and loops} If vertices are labelled, replace ${\langle
-V(G),V(H) \rangle}$ in \lineref{FirstExpandCall} with a set of label classes,
-one for each label that appears on at least one vertex of both $G$ and $H$.
+V(\graphG),V(\graphH) \rangle}$ in \lineref{FirstExpandCall} with a set of label classes,
+one for each label that appears on at least one vertex of both $\graphG$ and $\graphH$.
 Loops may be treated as a label modifier: assuming the original vertex labels
@@
 \paragraph{Directed graphs without edge labels} Let $A_G$ and $A_H$ be
-adjacency matrices of $G$ and $H$. For each vertex pair $(t,u)$ in graph $G$ or
-$H$, the adjacency matrix entry takes the value 0 if the vertices are not
+adjacency matrices of $\graphG$ and $\graphH$. For each vertex pair $(t,u)$ in graph $\graphG$ or
+$\graphH$, the adjacency matrix entry takes the value 0 if the vertices are not
 adjacent, 1 if the two vertices share a single edge in the direction $t
@@
 \linerangeref{NewPWithEdge}{InnerLoopEnd} of the basic algorithm split the
-label class $\langle P',T' \rangle$ in two, we now perform a four-way split
+label class $\langle \setG',\setH' \rangle$ in two, we now perform a four-way split
 where each vertex is classified according to the label on its adjacency matrix
@@
 17-19 of \cref{McSplitAlg} to create a new label-class of vertices that are
-not adjacent to $v$ or $w$, and remove these vertices from $\langle P',T'
-\rangle$.  Next, sort $P'$ and $T'$ in ascending order of the label on the edge
+not adjacent to $v$ or $w$, and remove these vertices from $\langle \setG',\setH'
+\rangle$.  Next, sort $\setG'$ and $\setH'$ in ascending order of the label on the edge
 from $v$ or $w$ to each vertex. We can then create the label classes
-corresponding to each edge label by simultaneously traversing $P'$ and $T'$
+corresponding to each edge label by simultaneously traversing $\setG'$ and $\setH'$
 from left to right.
@@
 
-In the CP-FC algorithm, each vertex $v \in \V(G)$ is represented by a
-variable, whose domain corresponds to the set of vertices in $\V(H)$ to which
+In the CP-FC algorithm, each vertex $v \in \V(\graphG)$ is represented by a
+variable, whose domain corresponds to the set of vertices in $\V(\graphH)$ to which
 $v$ may currently be mapped, with an additional special $\bot$ value
 representing an unmapped vertex.
-Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
+Given a label class $\langle \setG,\setH \rangle$ in \McSplit\, the vertices in $\setG$
 correspond to variables in CP-FC. The label-class representation of domains is
@@
 %%% \DontPrintSemicolon
-%%% \nl    \For {$u \in \N(G, b)$}{
+%%% \nl    \For {$u \in \N(\graphG, b)$}{
 %%% \nl        remove $v$ from $D'_u$ \;
@@
 %%%        }
-%%% \nl    \For {$u \in \invN(G, b)$}{
+%%% \nl    \For {$u \in \invN(\graphG, b)$}{
 %%% \nl        remove $v$ from $D'_u$ \;
@@
 which in our algorithm corresponds to branching on a label class with smallest
-$|T|$. We instead branch on the label class with smallest $\min(|P|,|T|)$.
+$|\setH|$. We instead branch on the label class with smallest $\min(|\setG|,|\setH|)$.
 This is empirically better, and accounts for much of the difference between
-the number of recursive calls made; branching on the smallest $|P| |T|$ gives
+the number of recursive calls made; branching on the smallest $|\setG| |\setH|$ gives
 very similar results. This can be viewed as exploiting both smallest domain first,
@@
 also the potential to branch on both sides, that is instead of branching on
-vertices in $\V(G)$, it would be equally valid to branch on a vertex in
-$\V(H)$, since our data structure treats the two graphs symmetrically. More
+vertices in $\V(\graphG)$, it would be equally valid to branch on a vertex in
+$\V(\graphH)$, since our data structure treats the two graphs symmetrically. More
 interestingly, we could choose which graph to branch on at each search node
-using some heuristic (perhaps choosing based on whether the $P$ or $T$ set is
+using some heuristic (perhaps choosing based on whether the $\setG$ or $\setH$ set is
 smaller), so long as we can ensure that the search remains complete.
@@
 
+\newpage
+
 \bibliographystyle{named}

### 2017-02-17 paper/paper.tex
We no longer need to explain why P and T are called P and T

@@
 used, where $\setG$ is the set of vertices in $\V(\graphG)$ labelled $l$, and $\setH$ is the
-set of vertices in $\V(\graphH)$ labelled $l$\footnote{We adopt terminology from
-subgraph isomorphism, where one graph is a pattern graph $\setG$ and the other the
-target graph $\setH$.}. Since there are $g + h$ vertices in the two graphs, at most
+set of vertices in $\V(\graphH)$ labelled $l$. Since there are $g + h$ vertices in the two graphs, at most
 $g + h$ label classes can exist at once, and there are at most $g + h$ vertices

### 2017-02-17 paper/paper.tex
No need to mention x and y axes in text
Mouse

@@
 \author{The Anonymous Authors of Paper 2431 \\
-Academy of Anonymous Academics, Some City, Some Country \\
-anonymous@anonymous.anonymous}
+Academy of Anonymous Academics, Atlantis \\
+anonymous@anony.mouse}
 
@@
 \paragraph{Unlabelled, undirected}
-\cref{figure:plain-cumulative} shows a plot of cumulative run times (x-axis)
-against number of problem instances solved (y-axis).  We may compare
+\cref{figure:plain-cumulative} shows a plot of cumulative run times
+against number of problem instances solved.  We may compare
 the speed of two algorithms using the horizontal distance between their curves.

### 2017-02-17 paper/paper.tex
Cite

@@
 vertices in $\setH$, together with label start indices \emph{in the same order} as
-they are used in the first pair of arrays. ?? This representation has
-similarities to data structures used in partition backtracking and the \citet{DBLP:journals/cacm/BronK73} clique enumeration algorithm.
+they are used in the first pair of arrays. This representation has
+similarities to data structures used in partition backtracking
+\citep{DBLP:journals/jsc/McKayP14} and the \citet{DBLP:journals/cacm/BronK73}
+clique enumeration algorithm.
 

### 2017-02-17 paper/paper.tex
No space

@@
 and filtering for connected subgraphs), the clique encodings of
-\citet{DBLP:conf/cp/McCreeshNPS16}, and the $k\downarrow$ algorithm of
+\citet{DBLP:conf/cp/McCreeshNPS16}, and the $k{\downarrow}$ algorithm of
 \citet{UpcomingAAAIPaper} (which only supports unlabelled, undirected,
@@
 set for which \McSplit\ is not the fastest solver; of the 3,506 instances that
-could be solved by at least one of \McSplit, $k\downarrow$, CP-FC and clique in less than than
+could be solved by at least one of \McSplit, $k{\downarrow}$, CP-FC and clique in less than than
 1,000 seconds, there were 426 instances on which \McSplit\ was not strictly
@@
 
-The basic \McSplit\ is beaten by the $k\downarrow$ algorithm of
+The basic \McSplit\ is beaten by the $k{\downarrow}$ algorithm of
 \citet{UpcomingAAAIPaper} on this dataset. However, we can modify the \McSplit\
-algorithm to use a top-down strategy similar to that used by $k\downarrow$  by
+algorithm to use a top-down strategy similar to that used by $k{\downarrow}$  by
 calling the main \FuncSty{McSplit} method once per goal size ($g, g-1, g-2,
@@
 
-What about our relationship to the $k\downarrow$ of \citet{UpcomingAAAIPaper}?
+What about our relationship to the $k{\downarrow}$ of \citet{UpcomingAAAIPaper}?
 \Cref{figure:prettyheatmaps3} plots the number of
-recursive calls made by $k\downarrow$ and \McSplit\ on each of the subgraph
+recursive calls made by $k{\downarrow}$ and \McSplit\ on each of the subgraph
 isomorphism instances. Although \McSplit\ is the faster algorithm overall, it
-explores more search nodes than $k\downarrow$ for most instances.
+explores more search nodes than $k{\downarrow}$ for most instances.
 
@@
 To further improve the state of the art, we could try creating a hybrid
-algorithm.  This could use $k\downarrow$ initially, switching to \McSplit\ when
+algorithm.  This could use $k{\downarrow}$ initially, switching to \McSplit\ when
 the extra filtering is ineffective, and finally switching to a clique encoding

### 2017-02-17 paper/paper.tex
Axis of evil

@@
 in \cref{figure:prettyheatmaps1,figure:prettyheatmaps2} we plot the number of
-recursive calls made by our algorithm on the $y$-axis versus the number made by
-CP-FC on the $x$-axis, for the unlabelled and labelled, unconnected problem
+recursive calls made by our algorithm versus the number made by
+CP-FC, for the unlabelled and labelled, unconnected problem
 instances. (Rather than a simple scatter plot, we use darker colours to

### 2017-02-17 paper/paper.tex
Do or do not, there is no try
Open conclude

@@
 
-We have introduced the \McSplit\ algorithm for Max Common Subgraph.  The new
-algorithm is more than an order of magnitude faster than the previous state of
-the art for unlabelled and undirected instances. We have shown how the
-algorithm can be extended for graphs with labels on edges, labels on vertices,
-loops, directed edges and the requirement that the resultant graph be
-connected.
+We have introduced the \McSplit\ algorithm for maximum common subgraph
+problems.  This algorithm is more than an order of magnitude faster than the
+previous state of the art for unlabelled and undirected instances. We have
+shown how the algorithm can be extended for graphs with labels on edges, labels
+on vertices, loops, directed edges and the requirement that the resultant graph
+be connected.
 
-To further improve the state of the art, we could try creating a hybrid
+To further improve the state of the art, we could create a hybrid
 algorithm.  This could use $k{\downarrow}$ initially, switching to \McSplit\ when

### 2017-02-17 paper/paper.tex
New last paragraph

@@
 It would be interesting to see whether these techniques are more broadly
-applicable. Most obviously, we could solve the induced subgraph isomorphism
-problem in the same way (and with nearly no changes to the code), but we could
-not solve the non-induced version. However, we suspect that some other problems
-may have a similar branching structure which would also benefit from a
-different domain store representation.
+applicable---we suspect that some other problems may have a similar branching
+structure which would also benefit from a different domain store
+representation. Most obviously, we could solve the induced subgraph isomorphism
+problem in the same way (and with nearly no changes to the code), and our
+connected variant shows that certain side constraints can also be handled.
+However, we cannot solve the non-induced subgraph isomorphism this way, nor can
+we handle certain richer labelling schemes such as those used in temporal
+subgraph isomorphism \citep{DBLP:conf/asunam/RedmondC13}.
 

### 2017-02-17 paper/paper.tex
No forced new page
Removed an unnecessary 'the'

@@
 connected variant shows that certain side constraints can also be handled.
-However, we cannot solve the non-induced subgraph isomorphism this way, nor can
+However, we cannot solve non-induced subgraph isomorphism this way, nor can
 we handle certain richer labelling schemes such as those used in temporal
 subgraph isomorphism \citep{DBLP:conf/asunam/RedmondC13}.
-
-\newpage
 

### 2017-02-17 paper/paper.tex
Quickie heuristics

@@
 
-?? Reword for here, explain later Small scale experiments (not presented here)
-were performed to identify a suitable heuristic for label class selection. The
-heuristic used selects the label class  with the smallest $\min(|\setG|,|\setH|)$ (a
-surrogate for expected branching factor). From this class, a vertex is selected
-from $\setG$ with maximum degree if $\graphH$ is sparse and with minimum degree if $\graphH$ is
-dense.
+Small scale experiments (not presented here) were performed to identify
+suitable heuristics for selecting a label class abd a vertex within a label
+class selection. Our heuristic of choice selects the label class with the
+smallest $\min(|\setG|,|\setH|)$. From this class, a vertex is selected from
+$\setG$ with maximum degree if $\graphH$ is sparse and with minimum degree if
+$\graphH$ is dense. We discuss this further in \cref{sec:comparison}.
 

### 2017-02-17 paper/paper.tex
More cites and texts

@@
 computationally. Recent practical progress has been made by using constraint
-programming \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16} and
+programming \citep{DBLP:conf/mco/VismaraV08,DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16} and
 mathematical programming \citep{DBLP:journals/dam/BahienseMPS12}, by reducing
@@
 similarities to data structures used in partition backtracking
-\citep{DBLP:journals/jsc/McKayP14} and the \citet{DBLP:journals/cacm/BronK73}
+\citep{DBLP:conf/wea/Lopez-PresaA09,DBLP:journals/jsc/McKayP14} and the \citet{DBLP:journals/cacm/BronK73}
 clique enumeration algorithm.
@@
 
-?? This needs a two sentence intro saying that people care about stuff and why.
+Maximum common subgraph problems come in many variants. For example, often
+vertices have labels (for example, denoting what kind of atom they represent in
+a molecule), and vertices may only be mapped to like-labelled vertices. We now
+outline how to adapt \cref{McSplitAlg} to handle this and other cases.
 
@@
 
-\paragraph{Maximum Common \emph{Connected} Subgraph} ?? In applications,
-something. The common subgraph must be connected. We consider only undirected
-graphs.  Modify \McSplit\ by permitting branching only on a vertex $v$ that has
-at least one non-zero element in its bit-string label.  We can represent this
-information compactly, and without increasing time complexity at each search
-node, by storing an extra bit with each label class.  This bit takes the value
-$1$ if and only if the vertices in the class are adjacent to at least one
-vertex in $M$. ?? Link to branching scheme papers.
+\paragraph{Maximum Common \emph{Connected} Subgraph} In chemistry applications,
+it is sometimes desirable to require the common subgraph be connected
+\citep{Ehrlich:2011}. We consider only undirected graphs. We may modify \McSplit\ by
+permitting branching only on a vertex $v$ that has at least one non-zero
+element in its bit-string label, following the scheme described by
+\citet{DBLP:conf/mco/VismaraV08}.  We can represent this information compactly,
+and without increasing time complexity at each search node, by storing an extra
+bit with each label class.  This bit takes the value $1$ if and only if the
+vertices in the class are adjacent to at least one vertex in $M$.
 

### 2017-02-17 paper/paper.tex
Added conauto citation
More consistent naming of the problem we're solving

--- 2017-02-17 paper/paper.tex
@@
 
-\paragraph{Maximum Common \emph{Connected} Subgraph} ?? In applications,
+\paragraph{Maximum common \emph{connected} subgraph} ?? In applications,
 something. The common subgraph must be connected. We consider only undirected
@@
 correspond to variables in CP-FC. The label-class representation of domains is
-possible because throughout the CP-FC algorithm for max common subgraph, the
+possible because throughout the CP-FC algorithm for maximum common subgraph, the
 domains of any two variables are either identical or disjoint (excluding

### 2017-02-17 paper/paper.tex
Move around

--- 2017-02-17 paper/paper.tex
@@
 
-\paragraph{Maximum Common \emph{Connected} Subgraph} In chemistry applications,
+\paragraph{Maximum common \emph{connected} subgraph} In chemistry applications,
 it is sometimes desirable to require the common subgraph be connected
@@
 
-What about our relationship to the $k{\downarrow}$ of \citet{UpcomingAAAIPaper}?
-\Cref{figure:prettyheatmaps3} plots the number of
+What about our relationship to the $k{\downarrow}$ of
+\citet{UpcomingAAAIPaper}?  \Cref{figure:prettyheatmaps3} plots the number of
 recursive calls made by $k{\downarrow}$ and \McSplit\ on each of the subgraph
 isomorphism instances. Although \McSplit\ is the faster algorithm overall, it
-explores more search nodes than $k{\downarrow}$ for most instances.
-
-?? Talk about the classic tradeoff, and dynamically switching.
-
-?? The concentration of points, down to unit propagation and counting
+explores more search nodes than $k{\downarrow}$ for most instances (even taking
+into account that $k{\downarrow}$ uses a unit propagation loop, and so measures
+the search tree slightly differently). This is the classic tradeoff between
+speed and cleverness. A hybrid algorithm could be beneficial here: it could use
+$k{\downarrow}$ initially, switching to \McSplit\ when the extra filtering is
+ineffective, and finally switching to a clique encoding when fewer than some
+threshold number of vertices remain to be selected. This might deliver the
+benefits of the clique encoding for labelled graphs, while avoiding the high
+memory cost and colouring time of encoding the full instance.
 
@@
 be connected.
-
-To further improve the state of the art, we could create a hybrid
-algorithm.  This could use $k{\downarrow}$ initially, switching to \McSplit\ when
-the extra filtering is ineffective, and finally switching to a clique encoding
-when fewer than some threshold number of vertices remain to be selected. This
-might deliver the benefits of the clique encoding for labelled graphs, while
-avoiding the high memory cost and colouring time of encoding the full instance.
 

### 2017-02-17 paper/paper.tex
Deduplicate

--- 2017-02-17 paper/paper.tex
@@
 
-With these definitions established, we now present \McSplit, a branch and bound
-algorithm which uses a new vertex-labelling scheme to avoid explicitly
-maintaining the set of feasible mappings for each vertex in $\V(\graphG)$.
-
-\McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
-(v_{m}, w_{m})\}$ with $|M^*| = m$ vertex pairs, where the $v_i$ are distinct members of $\V(\graphG)$ and the
-$w_i$ are distinct vertices of $\V(\graphH)$, such that $v_i$ and $v_j$ are adjacent
-in $\graphG$ if and only if $w_i$ and $w_j$ are adjacent in $\graphH$. 
-Given such a mapping, the subgraph of $\graphG$ induced by $\{v_i, \dots, v_{m}\}$
-and the subgraph of $\graphH$ induced by $\{w_i, \dots, w_{m}\}$ are a maximum
-common subgraph.
+With these definitions established, we now present \McSplit. This algorithm
+finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots, (v_{m},
+w_{m})\}$ with $|M^*| = m$ vertex pairs, where the $v_i$ are distinct members
+of $\V(\graphG)$ and the $w_i$ are distinct vertices of $\V(\graphH)$, such
+that $v_i$ and $v_j$ are adjacent in $\graphG$ if and only if $w_i$ and $w_j$
+are adjacent in $\graphH$.  Given such a mapping, the subgraph of $\graphG$
+induced by $\{v_i, \dots, v_{m}\}$ and the subgraph of $\graphH$ induced by
+$\{w_i, \dots, w_{m}\}$ are a maximum common subgraph.
 

### 2017-02-17 paper/paper.tex
Cite

@@
 instead asks for a common non-induced subgraph with as many \emph{edges} as
-possible.)
+possible \citep{DBLP:conf/cp/NdiayeS11}.)
 

### 2017-02-17 paper/paper.tex
Expand to search

@@
 \DontPrintSemicolon
-\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
+\nl $\FuncSty{search}(\AlgVar{future},M)$ \;
 \nl \Begin{
@@
        }
-\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
+\nl   $\FuncSty{search}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
 \nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
@@
 \nl \lIf {$\setG' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle \setG',\setH \rangle \}$}
-\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
+\nl $\FuncSty{search}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
 }
@@
 \nl $\AlgVar{incumbent} \gets \emptyset$ \;
-\nl $\FuncSty{expand}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
+\nl $\FuncSty{search}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
 \nl return $|\AlgVar{incubent}|$ \;
@@
 \paragraph{\cref{McSplitAlg} in detail} The recursive procedure,
-$\FuncSty{expand}$, has two parameters.  The parameter $\AlgVar{future}$ is a
+$\FuncSty{search}$, has two parameters.  The parameter $\AlgVar{future}$ is a
 list of label classes, each represented as a $\langle \setG, \setH \rangle$ pair as
 described above.  The parameter $M$ is the current mapping of vertices.  On
-each call to $\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be
+each call to $\FuncSty{search}$, the invariant holds that a $(v,w)$ pair may be
 added to $M$ if and only if $v$ and $w$ belong to the same label class in
@@
 
-The remainder of $\FuncSty{expand}$ performs the search.  A label class
+The remainder of $\FuncSty{search}$ performs the search.  A label class
 $\langle \setG, \setH \rangle$ is selected from $\AlgVar{future}$, where $\setG \subseteq
@@
 largest mapping.  In \lineref{FirstExpandCall} the initial call is made to
-$\FuncSty{expand}$; at this point we have a single label-class containing all
+$\FuncSty{search}$; at this point we have a single label-class containing all
 vertices, and the mapping $M$ is empty.

### 2017-02-17 paper/paper.tex
Return the incumbent
Global
Return in bold
Incumbent
Compare comparison
So we win again

@@
 %\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
-\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
+\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$\AlgVar{incumbent} \gets M$} \label{StoreIncumbent}
 %\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
@@
 \nl $\AlgVar{bound} \gets |M|  + \sum_{\langle \setG,\setH \rangle \in \AlgVar{future}} \min(|\setG|,|\setH|)$ \label{CalcBound} \;
-\nl \lIf {$\AlgVar{bound} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
+\nl \lIf {$\AlgVar{bound} \leq |\AlgVar{incumbent}|$}{\KwSty{return}} \label{PruneSearch}
 \medskip
@@
 \nl \Begin{
-\nl $\AlgVar{incumbent} \gets \emptyset$ \;
+    \nl $\KwSty{global}~\AlgVar{incumbent} \gets \emptyset$ \;
 \nl $\FuncSty{search}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
-\nl return $|\AlgVar{incubent}|$ \;
+\nl $\KwSty{return}$~$\AlgVar{incumbent}$ \;
 }
@@
 64GBytes RAM. Our algorithm was implemented\footnote{URL removed for anonymous
-review} in C++ and compiled using g++ 5.3.0. For comparison, we compare against
+review} in C++ and compiled using g++ 5.3.0. We compare against
 the best constraint programming implementations of
@@
 the corresponding experiment in \cref{figure:plain-cumulative} in which the
-subgraph is not required to be connected.
+subgraph is not required to be connected: \McSplit\ is the clear winner by more
+than an order of magnitude.
 

### 2017-02-17 paper/paper.tex
We're not the first to do this
We are good

@@
 any solver. Furthermore, the CP-FC algorithm and the clique encoding run out of
-memory on many of the instances (these are treated as timeouts).
+memory on many of the instances (these are treated as timeouts, following
+\citet{UpcomingAAAIPaper}).
 
@@
 \cref{figure:sip-cumulative} shows that this approach is the strongest on these
-instances.
+instances, and \McSplit{$\downarrow$} is the best algorithm for every choice of timeout.
 

### 2017-02-17 paper/paper.tex
Add two words that James can't do

@@
 to store a permutation of the vertices in $\setG$ having like-labelled vertices
-stored consecutively, together with the indices of the start of each label in
+stored consecutively, together with the size and indices of the start of each label in
 turn (in some arbitrary order). The second pair stores the permutation of

### 2017-02-17 paper/paper.tex
Less RAM
Expand
Wee edits

--- 2017-02-17 paper/paper.tex
@@
 (\cref{sec:extensions} describes how these restrictions may be relaxed).
-The vertices and edge sets of a graph $\graphG$ are denoted $\V(\graphG)$ and $\E(\graphG)$.  The
+The vertex and edge sets of a graph $\graphG$ are denoted $\V(\graphG)$ and $\E(\graphG)$.  The
 set of vertices adjacent to vertex $v$ in graph $\graphG$ is called the
@@
 instead asks for a common non-induced subgraph with as many \emph{edges} as
-possible \citep{DBLP:conf/cp/NdiayeS11}.)
+possible.)
 
@@
 finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots, (v_{m},
-w_{m})\}$ with $|M^*| = m$ vertex pairs, where the $v_i$ are distinct members
-of $\V(\graphG)$ and the $w_i$ are distinct vertices of $\V(\graphH)$, such
+w_{m})\}$ with $|M^*| = m$ vertex pairs, where the $v_i$ are distinct vertices
+from $\V(\graphG)$ and the $w_i$ are distinct vertices from $\V(\graphH)$, such
 that $v_i$ and $v_j$ are adjacent in $\graphG$ if and only if $w_i$ and $w_j$
@@
 
-\paragraph{Walkthrough} Before discussing the algorithm in detail, we start
-with an illustration of the main concepts using the graphs $\graphG$ and $\graphH$ in
+\paragraph{Walkthrough} Before discussing the algorithm in detail, we illustrate
+the main concepts using the graphs $\graphG$ and $\graphH$ in
 \cref{fig:alg1}.  These graphs have a maximum common subgraph with four
@@
 
-We begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Now label
-each unmatched vertex according to whether it is adjacent to vertex $1$ (in
-$\graphG$) or vertex $a$ (in $\graphH$), as shown in \cref{fig:alg2}.  Adjacent
+We begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Now
+label each unmatched vertex in $\V(\graphG)$ according to whether it is adjacent to vertex $1$, and
+label each unmatched vertex in $\V(\graphH)$ according to whether it is adjacent to vertex $a$,
+as shown in \cref{fig:alg2}.  Adjacent
 vertices have label 1; non-adjacent vertices have label 0.  We can extend $M$
-with a vertex mapping $vw$, with $v \in \V(\graphG)$ and $w \in \V(\graphH)$, if and only
+with a mapping $vw$, with $v \in \V(\graphG)$ and $w \in \V(\graphH)$, if and only
 if $v$ and $w$ have the same label.  This property, that two vertices may be
@@
 \DontPrintSemicolon
-\nl $\FuncSty{search}(\AlgVar{future},M)$ \;
+\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
 \nl \Begin{
 %\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
-\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$\AlgVar{incumbent} \gets M$} \label{StoreIncumbent}
+\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$\AlgVar{incumbent} \gets \AlgVar{M}$} \label{StoreIncumbent}
 %\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
@@
 \nl $\AlgVar{bound} \gets |M|  + \sum_{\langle \setG,\setH \rangle \in \AlgVar{future}} \min(|\setG|,|\setH|)$ \label{CalcBound} \;
-\nl \lIf {$\AlgVar{bound} \leq |\AlgVar{incumbent}|$}{\KwSty{return}} \label{PruneSearch}
+\nl \lIf {$\AlgVar{bound} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
 \medskip
@@
 \nl        \If {$\setG'' \neq \emptyset$ \bf{and} $\setH'' \neq \emptyset$}{
-\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle \setG'' , \setH'' \rangle$ \label{AddToFutureWithEdge}}
+\nl            $\AlgVar{future'} \gets \AlgVar{future'} \cup \{\langle \setG'' , \setH'' \rangle\}$ \label{AddToFutureWithEdge}}
 \nl        \bf{Let} $\setG'' \gets \setG' \cap \invN(\graphG, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
@@
 \nl        \If {$\setG'' \neq \emptyset$ \bf{and} $\setH'' \neq \emptyset$}{
-\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle \setG'' , \setH'' \rangle$} \label{InnerLoopEnd}
+\nl            $\AlgVar{future'} \gets \AlgVar{future'} \cup \{\langle \setG'' , \setH'' \rangle\}$} \label{InnerLoopEnd}
        }
-\nl   $\FuncSty{search}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
+\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
 \nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
@@
 \nl \lIf {$\setG' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle \setG',\setH \rangle \}$}
-\nl $\FuncSty{search}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
+\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
 }
@@
 \nl \Begin{
-    \nl $\KwSty{global}~\AlgVar{incumbent} \gets \emptyset$ \;
-\nl $\FuncSty{search}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
-\nl $\KwSty{return}$~$\AlgVar{incumbent}$ \;
+\nl $\AlgVar{incumbent} \gets \emptyset$ \;
+\nl $\FuncSty{expand}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
+\nl return $|\AlgVar{incumbent}|$ \;
 }
@@
 \paragraph{\cref{McSplitAlg} in detail} The recursive procedure,
-$\FuncSty{search}$, has two parameters.  The parameter $\AlgVar{future}$ is a
+$\FuncSty{expand}$, has two parameters.  The parameter $\AlgVar{future}$ is a
 list of label classes, each represented as a $\langle \setG, \setH \rangle$ pair as
 described above.  The parameter $M$ is the current mapping of vertices.  On
-each call to $\FuncSty{search}$, the invariant holds that a $(v,w)$ pair may be
+each call to $\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be
 added to $M$ if and only if $v$ and $w$ belong to the same label class in
@@
 
-The remainder of $\FuncSty{search}$ performs the search.  A label class
+The remainder of $\FuncSty{expand}$ performs the search.  A label class
 $\langle \setG, \setH \rangle$ is selected from $\AlgVar{future}$, where $\setG \subseteq
@@
 largest mapping.  In \lineref{FirstExpandCall} the initial call is made to
-$\FuncSty{search}$; at this point we have a single label-class containing all
+$\FuncSty{expand}$; at this point we have a single label-class containing all
 vertices, and the mapping $M$ is empty.
@@
 64GBytes RAM. Our algorithm was implemented\footnote{URL removed for anonymous
-review} in C++ and compiled using g++ 5.3.0. We compare against
+review} in C++ and compiled using g++ 5.3.0. For comparison, we compare against
 the best constraint programming implementations of
@@
 the corresponding experiment in \cref{figure:plain-cumulative} in which the
-subgraph is not required to be connected: \McSplit\ is the clear winner by more
-than an order of magnitude.
+subgraph is not required to be connected.
 

### 2017-02-17 paper/paper.tex

--- 2017-02-17 paper/paper.tex
@@
 \nl        \If {$\setG'' \neq \emptyset$ \bf{and} $\setH'' \neq \emptyset$}{
-\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle \setG'' , \setH'' \rangle$} \label{InnerLoopEnd}
+\nl            $\AlgVar{future'} \gets \AlgVar{future'} \cup \{\langle \setG'' , \setH'' \rangle\}$} \label{InnerLoopEnd}
        }
@@
 trends to the constraint programming, forward-checking (CP-FC) algorithm of
-\citet{DBLP:conf/cp/NdiayeS11}, but with much lower constant factors. Indeed,
+\citet{DBLP:conf/cp/NdiayeS11}, but with much lower constant factors and memory usage. Indeed,
 in \cref{figure:prettyheatmaps1,figure:prettyheatmaps2} we plot the number of
@@
 In the CP-FC algorithm, each vertex $v \in \V(\graphG)$ is represented by a
-variable, whose domain corresponds to the set of vertices in $\V(\graphH)$ to which
-$v$ may currently be mapped, with an additional special $\bot$ value
-representing an unmapped vertex.
-Given a label class $\langle \setG,\setH \rangle$ in \McSplit\, the vertices in $\setG$
-correspond to variables in CP-FC. The label-class representation of domains is
-possible because throughout the CP-FC algorithm for maximum common subgraph, the
-domains of any two variables are either identical or disjoint (excluding
-$\bot$, which is slightly more complicated). To the best of our knowledge, this
-observation has not been made previously, and it is not exploited in other
-implementations.
+variable, whose domain corresponds to the set of vertices in $\V(\graphH)$ to
+which $v$ may currently be mapped, with an additional special $\bot$ value
+representing an unmapped vertex.  Given a label class $\langle \setG,\setH
+\rangle$ in \McSplit\, the vertices in $\setG$ correspond to variables in
+CP-FC, and the vertices in $\setH$ to domain values. The label-class
+representation of domains is possible because throughout the CP-FC algorithm
+for maximum common subgraph, the domains of any two variables are either
+identical or disjoint (excluding $\bot$, which is slightly more complicated).
+To the best of our knowledge, this observation has not been made previously,
+and it is not exploited in other implementations.
 

### 2017-02-17 paper/paper.tex
The algorithm returns the mapping, not its size
Nicer sentence about null entry in adj matrix

@@
 We start our search at the function $\FuncSty{McSplit}$ (\lineref{McSplitFun}),
-with graphs $\graphG$ and $\graphH$ as inputs.  This function returns the size of the
-largest mapping.  In \lineref{FirstExpandCall} the initial call is made to
+with graphs $\graphG$ and $\graphH$ as inputs.  This function returns a mapping of
+maximum cardinality.  In \lineref{FirstExpandCall} the initial call is made to
 $\FuncSty{search}$; at this point we have a single label-class containing all
@@
 \paragraph{Undirected with edge labels} Each adjacency matrix entry contains an
-edge label, or a null entry (non-adjacent) $0$.  We use \cref{labDirAlg}, by
+edge label, or a null entry $0$ indicating that no edge is present.  We use \cref{labDirAlg}, by
 letting $L$ be the union of $\{0\}$ with the set of all labels that appear in

### 2017-02-17 paper/paper.tex
liken to merge sort

@@
 corresponding to each edge label by simultaneously traversing $\setG'$ and $\setH'$
-from left to right.
+from left to right, in a manner that resembles the merging step of merge sort.
 

### 2017-02-17 paper/paper.tex
We've got room to make this a bit nicer

@@
 
-Let $\graphG$ and $\graphH$ be the two input graphs to our maximum common subgraph problem.
-The orders (vertex counts) of these graphs are denoted $g$ and $h$.
+Throughout, $\graphG$ and $\graphH$ will be the two input graphs to our maximum
+common subgraph problem.  The orders (number of vertices) of these graphs are
+denoted $g$ and $h$ respectively.
 

### 2017-02-17 paper/paper.tex
Change number to instances where we're fastest to percentage

--- 2017-02-17 paper/paper.tex
@@
 number of instances, \McSplit\ is comfortably more than an order of magnitude
-faster than its nearer competitor  Moreover, there are few instances in this
-set for which \McSplit\ is not the fastest solver; of the 3,506 instances that
-could be solved by at least one of \McSplit, $k{\downarrow}$, CP-FC and clique in less than than
-1,000 seconds, there were 426 instances on which \McSplit\ was not strictly
-faster than the other two solvers.
+faster than its nearer competitor.  Moreover, \McSplit\ was the fastest algorithm
+on 87\% of the 3,506 instances that
+could be solved by at least one of the four algorithm in less than than
+1,000 seconds.
 

### 2017-02-17 paper/paper.tex
Minor changes

@@
 
-\paragraph{Unlabelled, undirected, connected} This class of instances are shown
+\paragraph{Unlabelled, undirected, connected} This class of instances is shown
 in \cref{figure:plain-connected-cumulative}.  These results are very similar to
@@
 (\cref{figure:33ve-connected-cumulative}). However, the gap between the two algorithms
-is very narrow, and is probably down to quality of implementation; indeed, the
+is very narrow, and is probably down to minor implementation details; indeed, the
 cumulative curve for \McSplit\ briefly rises above the curve for clique at a

### 2017-02-17 paper/paper.tex
Spacing issues

@@
     \centering
-    \begin{minipage}[t]{.17\linewidth}
+    \begin{minipage}[t]{.20\linewidth}
         Mapping
 
-        \bigskip\vspace{-4pt}
+        \medskip
 
@@
         \centering
-        Labelling of $\graphG$\vspace{-4pt}
+        Labelling of $\graphG$
         \begin{tabular}[t]{cc}
@@
         \centering
-        Labelling of $\graphH$\vspace{-4pt}
+        Labelling of $\graphH$
         \begin{tabular}[t]{cc}
@@
 
-Our results so far suggest that \McSplit\ has broadly similar performance
+Our experimental results suggest that \McSplit\ has broadly similar performance
 trends to the constraint programming, forward-checking (CP-FC) algorithm of
@@
 indicate a higher density of points around a location.) We see a close
-correlation: we typically do slightly less work, and sometimes do more, but
-instances with more than one order of magnitude difference in search space size
+correlation: \McSplit\ typically does slightly less work, and sometimes does more, but
+instances with more than one order of magnitude difference in search tree size
 are rare.
 
-Why is this? We do not see a similar correlation between our search space size
+Why is this? We do not see a similar correlation between our search tree size
 and that of the clique approach. The key observation is that \McSplit\ may be
-considered to be a different version of the constraint CP-FC algorithm, using
+considered to be a different version of the CP-FC algorithm, using
 an unconventional domain store and more efficient filtering algorithms. We now

### 2017-02-17 paper/paper.tex
Wee changes to comparison section
Typo. I'm sure it was James who did it.

@@
 Small scale experiments (not presented here) were performed to identify
-suitable heuristics for selecting a label class abd a vertex within a label
+suitable heuristics for selecting a label class and a vertex within a label
 class selection. Our heuristic of choice selects the label class with the
@@
 for maximum common subgraph, the domains of any two variables are either
-identical or disjoint (excluding $\bot$, which is slightly more complicated).
+identical or disjoint (excluding $\bot$, which is either present in all domains or in none).
 To the best of our knowledge, this observation has not been made previously,
-and it is not exploited in other implementations.
+and it is not exploited in other algorithms for maximum common subgraph.
 

### 2017-02-17 paper/paper.tex
More wee changes to comparison section

@@
 Small scale experiments (not presented here) were performed to identify
-suitable heuristics for selecting a label class and a vertex within a label
+suitable heuristics for selecting a label class abd a vertex within a label
 class selection. Our heuristic of choice selects the label class with the

### 2017-02-17 paper/paper.tex
percent is one word

--- 2017-02-17 paper/paper.tex
@@
 number of distinct vertex labels and the number of distinct edge labels is
-approximately equal to 33 per cent of the number of vertices in each graph.
+approximately equal to 33 percent of the number of vertices in each graph.
 
@@
 using some heuristic (perhaps choosing based on whether the $\setG$ or $\setH$ set is
-smaller), so long as we can ensure that the search remains complete.
+smaller).
 
@@
 applicable---we suspect that some other problems may have a similar branching
-structure which would also benefit from a different domain store
+structure which would also benefit from a partitioning domain store
 representation. Most obviously, we could solve the induced subgraph isomorphism

### 2017-02-17 paper/paper.tex
Wee changes to conclusion

@@
 number of distinct vertex labels and the number of distinct edge labels is
-approximately equal to 33 percent of the number of vertices in each graph.
+approximately equal to 33 per cent of the number of vertices in each graph.
 

### 2017-02-17 paper/paper.tex
A tense moment

--- 2017-02-17 paper/paper.tex
@@
 class selection. Our heuristic of choice selects the label class with the
-smallest $\min(|\setG|,|\setH|)$. From this class, a vertex is selected from
+smallest $\max(|\setG|,|\setH|)$. From this class, a vertex is selected from
 $\setG$ with maximum degree if $\graphH$ is sparse and with minimum degree if
@@
 number of instances, \McSplit\ is comfortably more than an order of magnitude
-faster than its nearer competitor.  Moreover, \McSplit\ was the fastest algorithm
+faster than its nearer competitor.  Moreover, \McSplit\ is the fastest algorithm
 on 87\% of the 3,506 instances that

### 2017-02-17 paper/paper.tex
Runtimes is one word

@@
 \paragraph{Unlabelled, undirected}
-\cref{figure:plain-cumulative} shows a plot of cumulative run times
+\cref{figure:plain-cumulative} shows a plot of cumulative runtimes
 against number of problem instances solved.  We may compare
@@
 
-\paragraph{Vertex and edge labels, directed} Cumulative run times for this
+\paragraph{Vertex and edge labels, directed} Cumulative runtimes for this
 class of instances are in \cref{figure:33ved-cumulative}. Again, \McSplit\ is over

### 2017-02-17 paper/paper.tex
Heuristic smallest max not min oops

--- 2017-02-17 paper/paper.tex
@@
 number of instances, \McSplit\ is comfortably more than an order of magnitude
-faster than its nearer competitor.  Moreover, \McSplit\ is the fastest algorithm
+faster than its nearer competitor.  Moreover, \McSplit\ was the fastest algorithm
 on 87\% of the 3,506 instances that

### 2017-02-18 paper/paper.tex
No letting
Our code always branches on a vertex of maximum degree
Description of how we store partitions is now more like the program

--- 2017-02-17 paper/paper.tex
@@
 which is not present at all).  Together, these facts allow us to store all the
-necessary information in two pairs of flat arrays. One pair of arrays is used
-to store a permutation of the vertices in $\setG$ having like-labelled vertices
-stored consecutively, together with the size and indices of the start of each label in
-turn (in some arbitrary order). The second pair stores the permutation of
-vertices in $\setH$, together with label start indices \emph{in the same order} as
-they are used in the first pair of arrays. This representation has
-similarities to data structures used in partition backtracking
+necessary information in three arrays.  The first 
+stores a permutation of $\V(\graphG)$ in which like-labelled vertices appear
+consecutively.  The second array, similarly, stores a permutation
+of $\V(\graphH)$ with like-labelled vertices together.  The third array
+contains a record for each label class $\langle \setG,\setH \rangle$.  Each
+record contains start and end pointers to the portion of the first
+array that contains $\setG$, and start and end pointers to the portion
+of the second array that contains $\setH$.  This representation has
+similarities to data structures used in partition backtracking for graph isomorphism
 \citep{DBLP:conf/wea/Lopez-PresaA09,DBLP:journals/jsc/McKayP14} and the \citet{DBLP:journals/cacm/BronK73}
@@
     \nl    $M \gets M \cup \{(v,w)\}$ \label{GrowM} \;
-\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
+\nl    $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
 \nl    \For {$\langle \setG',\setH'\rangle \in future$ \label{InnerLoop}}{
-\nl        \bf{Let} $\setG'' \gets \setG' \cap \N(\graphG, v) \setminus \{v\}$ \label{NewPWithEdge} \;
-\nl        \bf{Let} $\setH'' \gets \setH' \cap \N(\graphH, w) \setminus \{w\}$ \;
+\nl        $\setG'' \gets \setG' \cap \N(\graphG, v) \setminus \{v\}$ \label{NewPWithEdge} \;
+\nl        $\setH'' \gets \setH' \cap \N(\graphH, w) \setminus \{w\}$ \;
 \nl        \If {$\setG'' \neq \emptyset$ \bf{and} $\setH'' \neq \emptyset$}{
 \nl            $\AlgVar{future'} \gets \AlgVar{future'} \cup \{\langle \setG'' , \setH'' \rangle\}$ \label{AddToFutureWithEdge}}
-\nl        \bf{Let} $\setG'' \gets \setG' \cap \invN(\graphG, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
-\nl        \bf{Let} $\setH'' \gets \setH' \cap \invN(\graphH, w) \setminus \{w\}$ \;
+\nl        $\setG'' \gets \setG' \cap \invN(\graphG, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
+\nl        $\setH'' \gets \setH' \cap \invN(\graphH, w) \setminus \{w\}$ \;
 \nl        \If {$\setG'' \neq \emptyset$ \bf{and} $\setH'' \neq \emptyset$}{
@@
 class selection. Our heuristic of choice selects the label class with the
-smallest $\max(|\setG|,|\setH|)$. From this class, a vertex is selected from
-$\setG$ with maximum degree if $\graphH$ is sparse and with minimum degree if
-$\graphH$ is dense. We discuss this further in \cref{sec:comparison}.
+smallest $\max(|\setG|,|\setH|)$. From this class, a vertex with maximum degree
+is selected from $\setG$. We discuss this further in \cref{sec:comparison}.
 
@@
 \nl    \For {$l \in L$}{
-\nl        \bf{Let} $\setG'' \gets \{ u \in \setG' : u \neq v \wedge A_G[v][u] = l \}$ \;
-\nl        \bf{Let} $\setH'' \gets \{ u \in \setH' : u \neq w \wedge A_H[w][u] = l \}$ \;
+\nl        $\setG'' \gets \{ u \in \setG' : u \neq v \wedge A_G[v][u] = l \}$ \;
+\nl        $\setH'' \gets \{ u \in \setH' : u \neq w \wedge A_H[w][u] = l \}$ \;
 \nl        \If {$\setG'' \negmedspace\neq \emptyset$ \bf{and} $\setH'' \negmedspace \neq \emptyset$}{

### 2017-02-18 paper/paper.tex
Typo

@@
 on 87\% of the 3,506 instances that
-could be solved by at least one of the four algorithm in less than than
+could be solved by at least one of the four algorithms in less than than
 1,000 seconds.

### 2017-02-19 paper/paper.tex
Typo

@@
 are adjacent in $\graphH$.  Given such a mapping, the subgraph of $\graphG$
-induced by $\{v_i, \dots, v_{m}\}$ and the subgraph of $\graphH$ induced by
-$\{w_i, \dots, w_{m}\}$ are a maximum common subgraph.
+induced by $\{v_1, \dots, v_{m}\}$ and the subgraph of $\graphH$ induced by
+$\{w_1, \dots, w_{m}\}$ are a maximum common subgraph.
 
@@
 
-The remainder of $\FuncSty{search}$ performs the search.  A label class
-$\langle \setG, \setH \rangle$ is selected from $\AlgVar{future}$, where $\setG \subseteq
-V(\graphG)$ and $\setH \subseteq V(\graphH)$, using some heuristic (\lineref{SelectClass}).
-From this label class, a vertex $v$ is selected from $\setG$ using a heuristic
-(\lineref{SelectVertex}). We now iterate over all possible vertices in $\setH$
-(\linerangeref{WLoop}{ShrinkM}). A vertex $w \in \setH$ is selected and mapping $M$
-is extended (\lineref{GrowM}). A new set of label-classes, $\AlgVar{future'}$,
-is created (\lineref{NewFuture}).  Every label-class in $\AlgVar{future}$ can
+The remainder of the procedure performs the search.  A label class
+$\langle \setG, \setH \rangle$ is selected from $\AlgVar{future}$, where $\setG
+\subseteq V(\graphG)$ and $\setH \subseteq V(\graphH)$, using some heuristic
+(\lineref{SelectClass}), and from this label class, a vertex $v$ is selected
+from $\setG$ (\lineref{SelectVertex}). We now iterate over all 
+vertices $w$ in $\setH$ , exploring the consequences of adding $(v,w)$ to $M$
+(\linerangeref{WLoop}{ShrinkM}).  A new set of label-classes,
+$\AlgVar{future'}$,
+is created (\lineref{NewFuture}); this will be the labelling that results
+from adding $(v,w)$ to our mapping.  Every label-class in $\AlgVar{future}$ can
 now be split (\linerangeref{InnerLoop}{InnerLoopEnd}) into two new classes. The
@@
 suitable heuristics for selecting a label class and a vertex within a label
-class selection. Our heuristic of choice selects the label class with the
+class. Our heuristic of choice selects the label class with the
 smallest $\max(|\setG|,|\setH|)$. From this class, a vertex with maximum degree
@@
 \nl        \If {$\setG'' \negmedspace\neq \emptyset$ \bf{and} $\setH'' \negmedspace \neq \emptyset$}{
-    \nl $\AlgVar{future'} \gets \AlgVar{future'} + \langle \setG''\negmedspace, \setH'' \rangle$}
+    \nl $\AlgVar{future'} \gets \AlgVar{future'} \cup \{\langle \setG''\negmedspace, \setH'' \rangle\}$}
        }
-    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and labelled cases.}
+    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and edge-labelled cases.}
 \label{labDirAlg}
@@
 
-Our first set of experiments use a database of randomly-generated maximum
+Our first set of experiments uses a database of randomly-generated maximum
 common subgraph instances
@@
 \paragraph{Unlabelled, undirected}
-\cref{figure:plain-cumulative} shows a plot of cumulative runtimes
-against number of problem instances solved.  We may compare
+\cref{figure:plain-cumulative} shows a plot of cumulative 
+number of instances solved against runtime.  We may compare
 the speed of two algorithms using the horizontal distance between their curves.
@@
 which in our algorithm corresponds to branching on a label class with smallest
-$|\setH|$. We instead branch on the label class with smallest $\min(|\setG|,|\setH|)$.
+$|\setH|$. We instead branch on the label class with smallest $\max(|\setG|,|\setH|)$.
 This is empirically better, and accounts for much of the difference between

### 2017-02-19 paper/paper.tex
Fix plenking

@@
 from $\setG$ (\lineref{SelectVertex}). We now iterate over all
-vertices $w$ in $\setH$ , exploring the consequences of adding $(v,w)$ to $M$
+vertices $w$ in $\setH$, exploring the consequences of adding $(v,w)$ to $M$
 (\linerangeref{WLoop}{ShrinkM}).  A new set of label-classes,

### 2017-02-19 paper/paper.tex
Capitalised Search procedure

@@
 \DontPrintSemicolon
-\nl $\FuncSty{search}(\AlgVar{future},M)$ \;
+\nl $\FuncSty{Search}(\AlgVar{future},M)$ \;
 \nl \Begin{
@@
        }
-\nl   $\FuncSty{search}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
+\nl   $\FuncSty{Search}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
 \nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
@@
 \nl \lIf {$\setG' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle \setG',\setH \rangle \}$}
-\nl $\FuncSty{search}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
+\nl $\FuncSty{Search}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
 }
@@
     \nl $\KwSty{global}~\AlgVar{incumbent} \gets \emptyset$ \;
-\nl $\FuncSty{search}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
+\nl $\FuncSty{Search}(\{\langle V(\graphG),V(\graphH) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
 \nl $\KwSty{return}$~$\AlgVar{incumbent}$ \;
@@
 \paragraph{\cref{McSplitAlg} in detail} The recursive procedure,
-$\FuncSty{search}$, has two parameters.  The parameter $\AlgVar{future}$ is a
+$\FuncSty{Search}$, has two parameters.  The parameter $\AlgVar{future}$ is a
 list of label classes, each represented as a $\langle \setG, \setH \rangle$ pair as
 described above.  The parameter $M$ is the current mapping of vertices.  On
-each call to $\FuncSty{search}$, the invariant holds that a $(v,w)$ pair may be
+each call to $\FuncSty{Search}$, the invariant holds that a $(v,w)$ pair may be
 added to $M$ if and only if $v$ and $w$ belong to the same label class in
@@
 maximum cardinality.  In \lineref{FirstExpandCall} the initial call is made to
-$\FuncSty{search}$; at this point we have a single label-class containing all
+$\FuncSty{Search}$; at this point we have a single label-class containing all
 vertices, and the mapping $M$ is empty.

### 2017-02-19 paper/paper.tex
Stated line number of counting loop in comparison with CP section

@@
 thought shows that \McSplit\ computes the same bound, but using a simple
-counting loop---this is only possible because of the disjoint nature of the
-domains.
+counting loop (\cref{McSplitAlg}, \lineref{CalcBound})---this is only possible
+because of the disjoint nature of the domains.
 

### 2017-02-19 paper/paper.tex
Removed a small amount of unnecessary stuff from the detailed description of McSplit

@@
 The remainder of the procedure performs the search.  A label class
-$\langle \setG, \setH \rangle$ is selected from $\AlgVar{future}$, where $\setG
-\subseteq V(\graphG)$ and $\setH \subseteq V(\graphH)$, using some heuristic
-(\lineref{SelectClass}), and from this label class, a vertex $v$ is selected
+$\langle \setG, \setH \rangle$ is selected from $\AlgVar{future}$
+using some heuristic
+(\lineref{SelectClass}); from this label class, a vertex $v$ is selected
 from $\setG$ (\lineref{SelectVertex}). We now iterate over all
@@
 This is then repeated symmetrically for non-adjacency
-(\linerangeref{NewPWithoutEdge}{InnerLoopEnd}). A recursive call is then made
+(\linerangeref{NewPWithoutEdge}{InnerLoopEnd}). A recursive call is made
 (\lineref{ExpandWithV}), on return from which we remove the mapping $(v,w)$.

### 2017-02-19 code/james-cpp/mcsp.c
Added a code comment about the way vertices are re-ordered in our code

@@ -534,6 +534,12 @@ int main(int argc, char** argv) {
     vector<int> g0_deg = calculate_degrees(g0);
     vector<int> g1_deg = calculate_degrees(g1);
 
+    // As implemented here, g1_dense and g0_dense are false for all instances
+    // in the Experimental Evaluation section of the paper.  Thus,
+    // we always sort the vertices in descending order of degree (or total degree,
+    // in the case of directed graphs.  Improvements could be made here: it would
+    // be nice if the program explored exactly the same search tree if both
+    // input graphs were complemented.
     vector<int> vv0(g0.n);
     std::iota(std::begin(vv0), std::end(vv0), 0);
     bool g1_dense = sum(g1_deg) > g1.n*(g1.n-1);

### 2017-02-22 code/james-cpp-parallel/mcsp.cc
Maybe this is a bit better

@@ -223,19 +223,26 @@ namespace
 
     using PerThreadIncumbents = std::map<std::thread::id, vector<VtxPair> >;
 
-    const constexpr int split_levels = 3;
+    const constexpr int split_levels = 10;
 
     struct Position
     {
         std::array<unsigned, split_levels + 1> values;
+        unsigned depth;
 
         Position()
         {
             std::fill(values.begin(), values.end(), 0);
+            depth = 0;
         }
 
         bool operator< (const Position & other) const
         {
+            if (depth < other.depth)
+                return true;
+            else if (depth > other.depth)
+                return false;
+
             for (unsigned p = 0 ; p < split_levels + 1 ; ++p)
                 if (values.at(p) < other.values.at(p))
                     return true;
@@ -247,6 +254,7 @@ namespace
 
         void add(unsigned d, unsigned v)
         {
+            depth = d;
             if (d <= split_levels)
                 values[d] = v;
         }
@@ -256,18 +264,19 @@ namespace
     {
         struct Task
         {
-            const std::function<void ()> * func;
+            const std::function<void (unsigned long long &)> * func;
             int pending;
         };
 
-        std::mutex mutex;
+        std::mutex general_mutex;
         std::condition_variable cv;
         std::map<Position, Task> tasks;
         std::atomic<bool> finish;
 
-        std::vector<std::thread> threads;
+        vector<std::thread> threads;
 
         std::list<std::chrono::milliseconds> times;
+        std::list<unsigned long long> nodes;
 
         HelpMe(int n_threads) :
             finish(false)
@@ -275,8 +284,9 @@ namespace
             for (int t = 0 ; t < n_threads ; ++t)
                 threads.emplace_back([this, n_threads, t] {
                         std::chrono::milliseconds total_work_time = std::chrono::milliseconds::zero();
+                        unsigned long long this_thread_nodes = 0;
                         while (! finish.load()) {
-                            std::unique_lock<std::mutex> guard(mutex);
+                        std::unique_lock<std::mutex> guard(general_mutex);
                             bool did_something = false;
                             for (auto task = tasks.begin() ; task != tasks.end() ; ++task) {
                                 if (task->second.func) {
@@ -286,7 +296,7 @@ namespace
 
                                     auto start_work_time = std::chrono::steady_clock::now(); // local start time
 
-                                    (*f)();
[…]

### 2017-02-22 code/james-cpp-parallel/mcsp.cc
Possibly more parallelism?

@@ -655,8 +655,24 @@ void solve(const Graph & g0, const Graph & g1,
     if (! funcs.empty()) {
         std::atomic<int> shared_b{ 0 };
         auto this_thread_function = [&] (unsigned long long & active_nodes) {
-            for (int b = shared_b++, b_end = funcs.size() ; b < b_end /* not != */ ; b = shared_b++) {
+            int b = shared_b++, b_end = funcs.size();
+            for ( ; b < b_end /* not != */ ; b = shared_b++) {
                 (*std::next(funcs.begin(), b))(active_nodes);
+            }
+
+            if (b == b_end) {
+                auto new_domains = domains;
+                new_domains[bd_idx].right_len++;
+                if (bd.left_len == 0)
+                    remove_bidomain(new_domains, bd_idx);
+                auto new_position = position;
+                new_position.add(depth + 1, new_domains[bd_idx].right_len + 1);
+
+                if (depth > split_levels)
+                    solve_nopar(g0, g1, global_incumbent, per_thread_incumbents.find(std::this_thread::get_id())->second,
+                            current, new_domains, left, right, matching_size_goal, this_thread_nodes);
+                else
+                    solve(g0, g1, global_incumbent, per_thread_incumbents, current, new_domains, left, right, matching_size_goal, this_thread_nodes, depth + 1, new_position, help_me);
             }
         };
 
@@ -665,18 +681,6 @@ void solve(const Graph & g0, const Graph & g1,
         else
             this_thread_function(this_thread_nodes);
     }
-
-    bd.right_len++;
-    if (bd.left_len == 0)
-        remove_bidomain(domains, bd_idx);
-    auto new_position = position;
-    new_position.add(depth + 1, bd.right_len + 1);
-
-    if (depth > split_levels)
-        solve_nopar(g0, g1, global_incumbent, per_thread_incumbents.find(std::this_thread::get_id())->second,
-                current, domains, left, right, matching_size_goal, this_thread_nodes);
-    else
-        solve(g0, g1, global_incumbent, per_thread_incumbents, current, domains, left, right, matching_size_goal, this_thread_nodes, depth + 1, new_position, help_me);
 }
 
 vector<VtxPair> mcs(const Graph & g0, const Graph & g1) {

### 2017-02-22 code/james-cpp-parallel/mcsp.cc
Uh I think this is right

@@ -666,7 +666,7 @@ void solve(const Graph & g0, const Graph & g1,
                 if (bd.left_len == 0)
                     remove_bidomain(new_domains, bd_idx);
                 auto new_position = position;
-                new_position.add(depth + 1, new_domains[bd_idx].right_len + 1);
+                new_position.add(depth + 1, bd.right_len + 2);
 
                 if (depth > split_levels)
                     solve_nopar(g0, g1, global_incumbent, per_thread_incumbents.find(std::this_thread::get_id())->second,

### 2017-05-18 paper/paper.tex
Clarify that we're not using a CP toolkit
Corrected a typo identified by reviewer 1

@@
 \citet{UpcomingAAAIPaper} (which only supports unlabelled, undirected,
-unconnected instances).
+unconnected instances).  Each of these comparator programs is an optimised,
+dedicated implementation and does not use a general-purpose constraint programming
+toolkit.  We used the original authors' code in each case.
 
@@
 number of instances, \McSplit\ is comfortably more than an order of magnitude
-faster than its nearer competitor.  Moreover, \McSplit\ is the fastest algorithm
+faster than its nearest competitor.  Moreover, \McSplit\ is the fastest algorithm
 on 87\% of the 3,506 instances that

### 2017-05-18 paper/paper.tex
Expanded caption to give an example of what labels mean

@@
 
-    \caption{Mapping $M$ and vertex labels during search on example graphs $\graphG$ and $\graphH$ from \cref{fig:alg1}.}
+    \caption{Mapping $M$ and vertex labels during search on example graphs $\graphG$ and
+    $\graphH$ from \cref{fig:alg1}.  Labels represent adjacencies; for example, the label
+    101 on vertex $e$ in the final table signifies that $e$ is adjacent to the first
+    and third mapped vertices of $\graphH$ ($a$ and $f$) but not adjacent to the second
+    mapped vertex ($d$).}
     \label{figure:mcsplit-examples}

### 2017-05-18 paper/paper.tex
Revert "Remove the funding note for anon review, put it back in later and worry about page layout then"
This reverts commit a9c4b9a91c3e3cce66d213f68d9d871ec69b010f.
Revert "Anonymousify"
This reverts commit 3f2c6bf50f8b99e5abe9ca588507ef697ad93aae.

@@
 
-\title{A Partitioning Algorithm for Maximum Common Subgraph Problems}
-\author{The Anonymous Authors of Paper 2431 \\
-Academy of Anonymous Academics, Atlantis \\
-anonymous@anony.mouse}
-
-% \thanks{This work was supported by a pile of money from a funding agency.}
+\title{A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work
+was supported by the Engineering and Physical Sciences Research Council [grant
+numbers EP/K503058/1 and EP/M508056/1]}}
+\author{Ciaran McCreesh \and Patrick Prosser \and James Trimble \\
+University of Glasgow, Glasgow, Scotland \\
+j.trimble.1@research.gla.ac.uk}
 

### 2017-05-19 paper/paper.tex
Tweaked preliminaries, and moved the statement that we're considering INDUCED subgraphs to intro

@@
 
-This paper introduces a new branch and bound algorithm which exploits special
-properties of these problems to allow a much faster exploration of the search
-space, whilst retaining the filtering and bounding benefits of the constraint
-programming approach. We describe the algorithm for the basic maximum common
-subgraph problem, and discuss how it may be adapted to handle vertex labels,
-edge labels, and the requirement that the found subgraph be connected. We then
+This paper considers the \emph{maximum common induced subgraph} problem, in
+which the objective is to find a graph with as many vertices as possible which
+is an induced subgraph of each of two input graphs.  (The \emph{maximum common
+partial subgraph} problem instead asks for a common non-induced subgraph with
+as many \emph{edges} as possible \citep{DBLP:conf/cp/NdiayeS11}.) We introduce
+a new branch and bound algorithm which exploits special properties of the
+problem to allow a much faster exploration of the search space, whilst
+retaining the filtering and bounding benefits of the constraint programming
+approach. We describe the algorithm for the basic maximum common subgraph
+problem, and discuss how it may be adapted to handle vertex labels, edge
+labels, and the requirement that the found subgraph be connected. We then
 present an empirical study of the algorithm, demonstrating that it improves the
@@
 We initially assume that graphs are unlabelled, undirected and without loops
-(\cref{sec:extensions} describes how these restrictions may be relaxed).
-The vertex and edge sets of a graph $\graphG$ are denoted $\V(\graphG)$ and $\E(\graphG)$.  The
-set of vertices adjacent to vertex $v$ in graph $\graphG$ is called the
-\emph{neighbourhood} of $v$, denoted $\N(\graphG, v)$. We denote by $\invN(\graphG, v)$ the
-\emph{inverse neighbourhood} of $v$, being the set of the vertices not adjacent
-to $v$ (excluding $v$ itself). A \emph{subgraph} of a graph $\graphG$ is a graph
-consisting of some of the vertices of $\graphG$, and all of the edges between these
-vertices. A \emph{common subgraph} of two graphs is a graph which is
-(isomorphic to) a subgraph of two graphs simultaneously, and a \emph{maximum}
-common subgraph is one with as many vertices as possible. (In this paper, all
-subgraphs are induced. The \emph{maximum common partial subgraph} problem
-instead asks for a common non-induced subgraph with as many \emph{edges} as
-possible \citep{DBLP:conf/cp/NdiayeS11}.)
+(\cref{sec:extensions} describes how these restrictions may be relaxed).  The
+vertex and edge sets of a graph $\graphG$ are denoted $\V(\graphG)$ and
+$\E(\graphG)$.  The set of vertices adjacent to vertex $v$ in graph $\graphG$
+is called the \emph{neighbourhood} of $v$, denoted $\N(\graphG, v)$. We denote
+by $\invN(\graphG, v)$ the \emph{inverse neighbourhood} of $v$, being the set
+of the vertices not adjacent to $v$ (excluding $v$ itself). A \emph{subgraph}
+of a graph $\graphG$ is a graph consisting of some of the vertices of
+$\graphG$, and all of the edges between these vertices. (All subgraphs in this
+paper are induced subgraphs.) A \emph{common subgraph} of two graphs is a graph
+which is (isomorphic to) a subgraph of two graphs simultaneously, and a
+\emph{maximum} common subgraph is one with as many vertices as possible.
 

### 2017-05-19 paper/paper.tex
Mention in walkthrough that we add a pair to the mapping at each level of the search tree
matching -> mapping

@@
 The algorithm builds up a mapping $M$, starting with the empty mapping
-$\emptyset$. Select a vertex in $\V(\graphG)$ as the first vertex to be mapped; in
-our example we will arbitrarily choose vertex $1$. Each of the vertices in
-$\V(\graphH)$ to which vertex $1$ may be mapped will be tried in turn, and finally the
-possibility where vertex $1$ remains unmatched will be tried.
+$\emptyset$ and adding a $(v_i, w_i)$ pair at each level of the search tree.
+Select a vertex in $\V(\graphG)$ as the first vertex to be mapped; in our
+example we will arbitrarily choose vertex $1$. Each of the vertices in
+$\V(\graphH)$ to which vertex $1$ may be mapped will be tried in turn, and
+finally the possibility where vertex $1$ remains unmatched will be tried.
 
@@
 
-The algorithm backtracks when the incumbent (the largest matching found so far) is at least as large
+The algorithm backtracks when the incumbent (the largest mapping found so far) is at least as large
 as a calculated bound given $M$ and the current labelling. To demonstrate how
@@
 most one pair with this label can be added to $M$.  Thus, the upper bound on
-matching size is $|M| + 1 = 4$. The general formula for the upper bound is
+mapping size is $|M| + 1 = 4$. The general formula for the upper bound is
 \begin{multline*}
@@
 goal size is found. We expect that this could do well because in many cases the
-matching covers nearly all of the smaller graph---indeed,
+maximum common subgraph covers nearly all of the smaller graph---indeed,
 \cref{figure:sip-cumulative} shows that this approach is the strongest on these

### 2017-05-19 paper/paper.tex
Distinguish between labelled input graphs and our bitstring labels

@@
 
-\section{The \McSplit\ Algorithm}
+\section{The \McSplit\ Algorithm \label{sec:mcsplit}}
 
@@
 
-\paragraph{Vertex labels and loops} If vertices are labelled, replace ${\langle
-V(\graphG),V(\graphH) \rangle}$ in \lineref{FirstExpandCall} with a set of label classes,
-one for each label that appears on at least one vertex of both $\graphG$ and $\graphH$.
-Loops may be treated as a label modifier: assuming the original vertex labels
-are positive integers, we replace the label $l$ with $-l$ on each vertex that
-has a loop.
+\paragraph{Vertex labels and loops} If vertices in the input graphs have labels
+(as distinct from the bit-string labels described in \cref{sec:mcsplit}),
+replace ${\langle V(\graphG),V(\graphH) \rangle}$ in \lineref{FirstExpandCall}
+with a set of label classes, one for each label that appears on at least one
+vertex of both $\graphG$ and $\graphH$.  If some vertices have loops, we create
+two label classes for each input label: one class containing vertices with
+loops, and the other containing vertices without loops.
 

### 2017-05-19 paper/paper.tex
Clarified description of adjacency matrix storage in memory

@@
 \nl    \For {$l \in L$}{
-\nl        $\setG'' \gets \{ u \in \setG' : u \neq v \wedge A_G[v][u] = l \}$ \;
-\nl        $\setH'' \gets \{ u \in \setH' : u \neq w \wedge A_H[w][u] = l \}$ \;
+\nl        $\setG'' \gets \{ u \in \setG' : u \neq v \wedge A_\graphG[v][u] = l \}$ \;
+\nl        $\setH'' \gets \{ u \in \setH' : u \neq w \wedge A_\graphH[w][u] = l \}$ \;
 \nl        \If {$\setG'' \negmedspace\neq \emptyset$ \bf{and} $\setH'' \negmedspace \neq \emptyset$}{
@@
 
-\paragraph{Directed graphs without edge labels} Let $A_G$ and $A_H$ be
-adjacency matrices of $\graphG$ and $\graphH$. For each vertex pair $(t,u)$ in graph $\graphG$ or
-$\graphH$, the adjacency matrix entry takes the value 0 if the vertices are not
+\paragraph{Directed graphs without edge labels} Before running the algorithm,
+we create two-dimensional arrays $A_\graphG$ and $A_\graphH$ representing adjacencies in
+$\graphG$ and $\graphH$ respectively.  These store the same information as the graphs'
+adjacency matrices, but allow us to determine in a single memory access which of
+the two possible edges exist between a pair of vertices.
+We now describe the entries of $A_\graphG$.
+For each vertex pair $(t,u)$ in $\graphG$,
+$A_\graphG[t][u]$ takes the value 0 if $t$ and $u$ are not
 adjacent, 1 if the two vertices share a single edge in the direction $t
@@
 label class $\langle \setG',\setH' \rangle$ in two, we now perform a four-way split
-where each vertex is classified according to the label on its adjacency matrix
-entry from $v$ or $w$.  This is shown in \cref{labDirAlg}, where
+where each vertex is classified according to the label on its array entry indexed by
+$v$ and $w$.  This is shown in \cref{labDirAlg}, where
 $L=\{0,1,2,3\}$.
 
-\paragraph{Undirected with edge labels} Each adjacency matrix entry contains an
+\paragraph{Undirected with edge labels} Each entry of $A_\graphG$ or $A_\graphH$ contains an
 edge label, or a null entry $0$ indicating that no edge is present.  We use \cref{labDirAlg}, by
@@
 \paragraph{Directed with edge labels} This case is similar to its undirected
-counterpart, except that each element $A[u][v]$ in an adjacency matrix is a
+counterpart, except that each element $A_\graphG[u][v]$ or $A_\graphH[u][v]$ is a
 pair $(l_1, l_2)$, where $l_1$ is the label on the edge $u \rightarrow v$ (or 0

### 2017-05-19 paper/paper.tex
Gave molecular bond example for labelled edges

@@
 Maximum common subgraph problems come in many variants. For example, often
-vertices have labels (for example, denoting what kind of atom they represent in
+vertices or edges have labels (for example, denoting what kind of atom or bond they represent in
 a molecule), and vertices may only be mapped to like-labelled vertices. We now

### 2017-05-19 paper/paper.tex
Wrote a bit more about our heuristics

@@
 Small scale experiments (not presented here) were performed to identify
-suitable heuristics for selecting a label class and a vertex within a label
-class. Our heuristic of choice selects the label class with the
-smallest $\max(|\setG|,|\setH|)$. From this class, a vertex with maximum degree
-is selected from $\setG$. We discuss this further in \cref{sec:comparison}.
+suitable heuristics for the $\FuncSty{SelectLabelClass}$ and
+$\FuncSty{SelectVertex}$ functions.  Our $\FuncSty{SelectLabelClass}$
+function chooses a label class with the smallest $\max(|\setG|,|\setH|)$,
+breaking ties by selecting a class containing a vertex in $G$ with the largest degree. From
+the selected class, $\FuncSty{SelectVertex}$ chooses a vertex in $\setG$ with
+maximum degree. We further discuss the effectiveness of our
+$\FuncSty{SelectLabelClass}$ function in \cref{sec:comparison}.
 

### 2017-05-19 paper/paper.tex
Added a sentence about McSplit-down not being good for the small instances

@@
 instances, and \McSplit{$\downarrow$} is the best algorithm for every choice of timeout.
+(By contrast, the optimal solutions for the instances in
+\cref{figure:mcs-cumulative} typically cover a smaller proportion of the
+input graphs, and \McSplit{$\downarrow$} is often more than a magnitude slower than
+the plain \McSplit\ algorithm for these instances; the \McSplit{$\downarrow$} results
+are not shown for these instances.)
 

### 2017-05-19 paper/paper.tex
Added summary to experiments section

@@
 
+To summarise these experimental results, we find that \McSplit\ improves on the
+previous state of the art by more than an order of magnitude for small,
+unlabelled graphs.  On a benchmark suite of larger instances, the
+\McSplit{$\downarrow$} variant of the algorithm comfortably outperforms the
+state of the art.  On labelled instances, the clique encoding remains the
+strongest solver, except on the most trivial instances---particularly where
+there is no requirement for the common subgraph to be connected.
+
 \section{Comparison with Existing Algorithms}\label{sec:comparison}

### 2017-05-19 paper/paper.tex
A couple of Christine's suggestions

@@
 induced by $\{v_1, \dots, v_{m}\}$ and the subgraph of $\graphH$ induced by
-$\{w_1, \dots, w_{m}\}$ are a maximum common subgraph.
+$\{w_1, \dots, w_{m}\}$ are isomorphic and correspond to a maximum common subgraph.
 
@@
 \nl \For {$w \in \setH$ \label{WLoop}} {
-    \nl    $M \gets M \cup \{(v,w)\}$ \label{GrowM} \;
 \nl    $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
@@
        }
-\nl   $\FuncSty{Search}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
-\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
+\nl   $\FuncSty{Search}(\AlgVar{future'},M\cup \{(v,w)\})$ \label{ExpandWithV} \;
   }
@@
 vertices $w$ in $\setH$, exploring the consequences of adding $(v,w)$ to $M$
-(\linerangeref{WLoop}{ShrinkM}).  A new set of label-classes,
+(\linerangeref{WLoop}{ExpandWithV}).  A new set of label-classes,
 $\AlgVar{future'}$,

### 2017-05-19 paper/paper.tex
First draft of equivalence of McSplit and CP-FC bounds

@@
 \newcommand{\nmax}{n_{\max}}
+
+\newcommand{\graphB}{\mathcal{B}}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[htb]
     \centering
@@
 CP-FC uses a soft all-different constraint to compute a bound, which requires
-running a matching algorithm on a supporting compatibility graph.  Careful
-thought shows that \McSplit\ computes the same bound, but using a simple
-counting loop (\cref{McSplitAlg}, \lineref{CalcBound})---this is only possible
+running a matching algorithm on a supporting compatibility graph.  We now
+show that \McSplit\ computes the same bound, but using a simple
+counting loop (\cref{McSplitAlg}, \lineref{CalcBound})---this is possible
 because of the disjoint nature of the domains.
+
+To illustrate the method used by CP-FC to calculate a bound, we consider the
+input graphs $\graphG$ and $\graphH$ in \cref{fig:alg1}. Suppose that the
+variable corresponding to vertex 1 has been matched to the value corresponding
+to vertex $a$, and that no other assignments have been made.  In the
+terminology of \McSplit, we now have two label classes: $\langle \{2,3\},
+\{d,f\} \rangle$ and $\langle \{4,5\}, \{b,c,e\} \rangle$.  CP-FC represents
+this state by storing the domain of each remaining variable: 2 and 3 each
+have the domain $\{d,f\}$; 4 and 5 each have the domain $\{b,c,e\}$.
+
+As described previously, the \McSplit\ algorithm finds the size of the smaller
+set in each label class, and adds these sizes to the size of the current
+mapping, giving a bound of 5.  CP-FC uses the compatibility graph $\graphB$ in
+\cref{fig:cpfc-bipartite} to calculate an upper bound; variables are
+represented by vertices on the left, values are represented by vertices on the
+right, and a variable may take a value if and only if the corresponding
+vertices are adjacent.  The upper bound is calculated by CP-FC by computing a
+maximum matching on this bipartite graph, and adding the size of this matching
+to the size of the current mapping.
+
+The vertices in the bipartite graph are coloured to emphasise how they
+correspond to the two label classes.  Each label class corresponse to a
+complete bipartite graph, and there are no edges between the components
+corresponding to different label classes.  Clearly, a maximum matching in
+$\graphB$ must be the union of the maximum matching in each complete bipartite
+subgraph, and each of these matchings in turn have the same size as the smaller
+of the left and right sets of vertices.  It follows that the CP-FC bound is
+identical to the \McSplit\ bound.
+
+\begin{figure}[htb]
+\centering
+% Bipartite graph TikZ based on code from
+% https://tex.stackexchange.com/a/15094/61367
+\definecolor{myblue}{RGB}{240,240,240}
+\definecolor{mygreen}{RGB}{120,120,120}
+\usetikzlibrary{positioning,chains,fit,shapes,calc}
+
+\begin{tikzpicture}[thick,
+  every node/.style={draw,circle},
+  class1node/.style={fill=myblue},
+  class2node/.style={fill=mygreen},
+  every fit/.style={ellipse,draw,inner sep=-2pt,text width=2cm}
+]
+
+\begin{scope}[start chain=going below,node distance=3mm]
+\foreach \i in {2,3}
+  \node[class1node,on chain] (f\i) [label=left: \i] {};
+\end{scope}
+
+\begin{scope}[yshift=-1.8cm,start chain=going below,node distance=3mm]
+\foreach \i in {4,5}
+  \node[class2node,on chain] (f\i) [label=left: \i] {};
+\end{scope}
+
+\begin{scope}[xshift=2.2cm,start chain=going below,node distance=3mm]
+\foreach \i in {d,f}
+  \node[class1node,on chain] (s\i) [label=right: \i] {};
+\end{scope}
+
+\begin{scope}[xshift=2.2cm,yshift=-1.5cm,start chain=going below,node distance=3mm]
+\foreach \i in {b,c,e}
+  \node[class2node,on chain] (s\i) [label=right: \i] {};
+\end{scope}
+
+% the edges
+\draw (f2) -- (sd);
+\draw (f3) -- (sd);
+\draw (f2) -- (sf);
+\draw (f3) -- (sf);
+
+\draw (f4) -- (sb);
+\draw (f4) -- (sc);
+\draw (f4) -- (se);
+\draw (f5) -- (sb);
+\draw (f5) -- (sc);
+\draw (f5) -- (se);
+\end{tikzpicture}
+
+\caption{A bipartite graph of the type used to calculate a bound in CP-FC}
+\label{fig:cpfc-bipartite}
+\end{figure}
 

### 2017-05-19 paper/paper.tex
Figure 6(c) compares k-down with McSplit-down, not with McSplit

@@
 \citet{UpcomingAAAIPaper}?  \Cref{figure:prettyheatmaps3} plots the number of
-recursive calls made by $k{\downarrow}$ and \McSplit\ on each of the subgraph
-isomorphism instances. Although \McSplit\ is the faster algorithm overall, it
+recursive calls made by $k{\downarrow}$ and \McSplit{$\downarrow$} on each of the subgraph
+isomorphism instances. Although \McSplit{$\downarrow$} is the faster of the two algorithms overall, it
 explores more search nodes than $k{\downarrow}$ for most instances (even taking

### 2017-05-19 paper/paper.tex
Added another downarrow

@@
 speed and cleverness. A hybrid algorithm could be beneficial here: it could use
-$k{\downarrow}$ initially, switching to \McSplit\ when the extra filtering is
+$k{\downarrow}$ initially, switching to \McSplit{$\downarrow$} when the extra filtering is
 ineffective, and finally switching to a clique encoding when fewer than some

### 2017-05-19 paper/paper.tex
Added an often

@@
 Because graphs are widely used to model real-world phenomena, maximum common
-subgraph problems have arisen in molecular science (where graphs represent
+subgraph problems have arisen in molecular science (where graphs often represent
 molecules)

### 2017-05-19 paper/paper.tex
A eww arr ell

@@
 
-Maximum common subgraph problems come in many variants. For example, often
-vertices or edges have labels (for example, denoting what kind of atom or bond they represent in
-a molecule), and vertices may only be mapped to like-labelled vertices. We now
-outline how to adapt \cref{McSplitAlg} to handle this and other cases.
+Maximum common subgraph problems come in many variants. Often
+vertices or edges have labels (for example, denoting the kind of atom or bond they represent in
+a molecule \cite{Ehrlich:2011}), and the induced subgraphs of the two input graphs are
+required to have identical labels.  Directed edges are used in an application to systems
+of biochemical reactions \cite{DAM2014}.
+We now outline how to adapt \cref{McSplitAlg} to handle these cases.
 
@@
 Experiments were performed on machines with dual Intel Xeon E5-2640 v2 CPUs and
-64GBytes RAM. Our algorithm was implemented\footnote{URL removed for anonymous
-review} in C++ and compiled using g++ 5.3.0. We compare against
-the best constraint programming implementations of
-\citet{DBLP:conf/cp/NdiayeS11} and \citet{DBLP:conf/cp/McCreeshNPS16} (CP-FC in
-the unlabelled cases, and CP-MAC in the labelled cases, using both branching
-and filtering for connected subgraphs), the clique encodings of
-\citet{DBLP:conf/cp/McCreeshNPS16}, and the $k{\downarrow}$ algorithm of
-\citet{UpcomingAAAIPaper} (which only supports unlabelled, undirected,
-unconnected instances).  Each of these comparator programs is an optimised,
-dedicated implementation and does not use a general-purpose constraint programming
-toolkit.  We used the original authors' code in each case.
+64GBytes RAM. Our algorithm was
+implemented\footnote{Source code, instances, experimental scripts and raw
+results are available at
+https://github.com/jamestrimble/ijcai2017-partitioning-common-subgraph}
+in C++ and compiled using g++ 5.3.0. We compare against the best constraint
+programming implementations of \citet{DBLP:conf/cp/NdiayeS11} and
+\citet{DBLP:conf/cp/McCreeshNPS16} (CP-FC in the unlabelled cases, and CP-MAC
+in the labelled cases, using both branching and filtering for connected
+subgraphs), the clique encodings of \citet{DBLP:conf/cp/McCreeshNPS16}, and the
+$k{\downarrow}$ algorithm of \citet{UpcomingAAAIPaper} (which only supports
+unlabelled, undirected, unconnected instances).  Each of these comparator
+programs is an optimised, dedicated implementation and does not use a
+general-purpose constraint programming toolkit.  We used the original authors'
+code in each case.
 

### 2017-05-19 paper/paper.tex
Don't microtype
bigskip before summary of experiments

--- 2017-05-19 paper/paper.tex
@@
 \usepackage{cleveref}
-\usepackage{microtype}
+%\usepackage{microtype}
 \usepackage{booktabs}
@@
 
-To summarise these experimental results, we find that \McSplit\ improves on the
+\bigskip
+
+Overall, we find that \McSplit\ improves on the
 previous state of the art by more than an order of magnitude for small,

### 2017-05-19 paper/paper.tex
Oh so purdy

@@
 
-The algorithm builds up a mapping $M$, starting with the empty mapping
+The algorithm builds up a mapping $M$ using a depth-first search, starting with the empty mapping
 $\emptyset$ and adding a $(v_i, w_i)$ pair at each level of the search tree.
@@
 % https://tex.stackexchange.com/a/15094/61367
-\definecolor{myblue}{RGB}{240,240,240}
-\definecolor{mygreen}{RGB}{120,120,120}
+\definecolor{myblue}{RGB}{79.0, 18.0, 123.0}
+\definecolor{mygreen}{RGB}{251.0, 135.0, 97.0}
 \usetikzlibrary{positioning,chains,fit,shapes,calc}
@@
 
-\caption{A bipartite graph of the type used to calculate a bound in CP-FC}
+\caption{A bipartite compatibility graph of the type used to calculate a bound in the CP-FC algorithm.}
 \label{fig:cpfc-bipartite}

### 2017-05-19 paper/paper.tex
Compare to clique bound

--- 2017-05-19 paper/paper.tex
@@
 
+This bound is incomparable to the clique bound, which is given by a greedy
+colouring of a microstructure-like graph \citep{DBLP:conf/cp/McCreeshNPS16}.
+The clique bound is potentially stronger, in that it can reason about
+compatibilities between individual variable-vertex assignments. The clique
+bound is also able to make use of edge label information, which CP-FC and our
+equivalent bound cannot---this comes at the expense of much higher memory
+requirements, and also longer calculation times for unlabelled graphs. The
+clique bound is also potentially weaker: it is not even guaranteed to be less
+than or equal to the number of remaining variables, in the CP-FC sense.
+
 A further advantage of our encoding is that it gives us efficient access to a

### 2017-05-19 paper/paper.tex
Clarify in intro that we discuss only the induced variant

@@
 partial subgraph} problem instead asks for a common non-induced subgraph with
-as many \emph{edges} as possible \citep{DBLP:conf/cp/NdiayeS11}.) We introduce
+as many \emph{edges} as possible \citep{DBLP:conf/cp/NdiayeS11}; we discuss
+only the induced variant in this paper.) We introduce
 a new branch and bound algorithm which exploits special properties of the

### 2017-05-19 paper/paper.tex
Page layout

@@
 \documentclass[letterpaper]{article}
-%\usepackage[pass,showframe]{geometry}
+\usepackage[showframe,pass]{geometry}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[tb]
     \centering\subfiguretopcaptrue
@@
     \label{figure:mcsplit-examples}
-\end{figure}
-
-\begin{figure}[ht]
 \end{figure}
@@
 
-\begin{algorithm}[t]
-\DontPrintSemicolon
-\nl    \For {$l \in L$}{
-\nl        $\setG'' \gets \{ u \in \setG' : u \neq v \wedge A_\graphG[v][u] = l \}$ \;
-\nl        $\setH'' \gets \{ u \in \setH' : u \neq w \wedge A_\graphH[w][u] = l \}$ \;
-\nl        \If {$\setG'' \negmedspace\neq \emptyset$ \bf{and} $\setH'' \negmedspace \neq \emptyset$}{
-    \nl $\AlgVar{future'} \gets \AlgVar{future'} \cup \{\langle \setG''\negmedspace, \setH'' \rangle\}$}
-       }
-    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and edge-labelled cases.}
-\label{labDirAlg}
-\end{algorithm}
-
 Maximum common subgraph problems come in many variants. Often
@@
 loops, and the other containing vertices without loops.
+
+\begin{algorithm}[t]
+\DontPrintSemicolon
+\nl    \For {$l \in L$}{
+\nl        $\setG'' \gets \{ u \in \setG' : u \neq v \wedge A_\graphG[v][u] = l \}$ \;
+\nl        $\setH'' \gets \{ u \in \setH' : u \neq w \wedge A_\graphH[w][u] = l \}$ \;
+\nl        \If {$\setG'' \negmedspace\neq \emptyset$ \bf{and} $\setH'' \negmedspace \neq \emptyset$}{
+    \nl $\AlgVar{future'} \gets \AlgVar{future'} \cup \{\langle \setG''\negmedspace, \setH'' \rangle\}$}
+       }
+    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and edge-labelled cases.}
+\label{labDirAlg}
+\end{algorithm}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[p]
     \centering\subfiguretopcaptrue
-    \subfigure[][Unlabelled, undirected instances] {
+    \subfigure[][Unlabelled, undirected, not connected] {
         \centering
@@
 
-    \subfigure[][Vertex and edge labelled, directed instances] {
+    \subfigure[][Vertex and edge labelled, directed, not connected] {
         \centering
@@
     }
-    \caption{Cumulative numbers of instances solved over time for the maximum
-    common subgraph problem.}\label{figure:mcs-cumulative}
+
+    \subfigure[][Unlabelled, undirected, connected] {
+        \centering
+        \includegraphics*{gen-graph-plain-connected-cumulative.pdf}
+        \label{figure:plain-connected-cumulative}
+    }
+
+    \subfigure[][Vertex and edge labelled, undirected, connected] {
+        \centering
+        \includegraphics*{gen-graph-33ve-connected-cumulative.pdf}
+        \label{figure:33ve-connected-cumulative}
+    }
+    \caption{Cumulative numbers of instances solved over time for maximum
+    common subgraph problems.}\label{figure:mcs-cumulative}
 \end{figure}
@@
 
-\begin{figure}[t]
-    \centering\subfiguretopcaptrue
-    \subfigure[][Unlabelled, undirected, connected instances] {
+\begin{figure}[b]
     \centering
-        \includegraphics*{gen-graph-plain-connected-cumulative.pdf}
-        \label{figure:plain-connected-cumulative}
-    }
-
-    \subfigure[][Vertex and edge labelled, undirected, connected instances] {
-        \centering
-        \includegraphics*{gen-graph-33ve-connected-cumulative.pdf}
-        \label{figure:33ve-connected-cumulative}
-    }
+    \includegraphics*{gen-graph-sip-cumulative.pdf}
     \caption{Cumulative numbers of instances solved over time for the maximum
-    common connected subgraph problem.} \label{figure:mcs-connected-cumulative}
+    common connected subgraph problem on the large subgraph isomorphism benchmark
+    suite.} \label{figure:sip-cumulative}
 \end{figure}
@@
 \citet{UpcomingAAAIPaper}).
+
+\begin{figure*}[t]
+    \centering\subfiguretopcaptrue
+    \subfigure[][Unlabelled, undirected instances] {\label{figure:prettyheatmaps1}
+    \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
+    }\subfigure[][Labelled, directed instances] {\label{figure:prettyheatmaps2}
+    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
+    }\subfigure[][Subgraph isomorphism instances]{\label{figure:prettyheatmaps3}
+    \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
+}
+    \caption{Relative search space sizes for instances which were solved by both algorithms within the timeout.}
+    \label{figure:prettyheatmaps}
+\end{figure*}
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics*{gen-graph-sip-cumulative.pdf}
-    \caption{Cumulative numbers of instances solved over time for the maximum
-    common connected subgraph problem on the large subgraph isomorphism benchmark
-    suite.} \label{figure:sip-cumulative}
-\end{figure}
-
 \bigskip
@@
 \section{Comparison with Existing Algorithms}\label{sec:comparison}
-
-\begin{figure*}[t]
-    \centering\subfiguretopcaptrue
-    \subfigure[][Unlabelled, undirected instances] {\label{figure:prettyheatmaps1}
-    \includegraphics*{gen-graph-plain-james-versus-cp-fc-nodes-scatter.pdf}
-    }\subfigure[][Labelled, directed instances] {\label{figure:prettyheatmaps2}
-    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
-    }\subfigure[][Subgraph isomorphism instances]{\label{figure:prettyheatmaps3}
-    \includegraphics*{gen-graph-sip-james-versus-kdown-nodes-scatter.pdf}
-}
-    \caption{Relative search space sizes for instances which were solved by both algorithms within the timeout.}
-    \label{figure:prettyheatmaps}
-\end{figure*}
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[tb]
 \centering

### 2017-05-19 paper/paper.tex
More tweaks

@@
 \documentclass[letterpaper]{article}
-\usepackage[showframe,pass]{geometry}
+\usepackage[pass]{geometry}
 
@@
       \LabelTables{$\{1a\}$}
-                  {$2$ & 0 \\
-                   $3$ & 0 \\
-                   $4$ & 1 \\
-                   $5$ & 1 \\}
-                  {$b$ & 1 \\
-                   $c$ & 1 \\
-                   $d$ & 0 \\
-                   $e$ & 1 \\
-                   $f$ & 0 \\}
+                  {$2$ & $0$ \\
+                   $3$ & $0$ \\
+                   $4$ & $1$ \\
+                   $5$ & $1$ \\}
+                  {$b$ & $1$ \\
+                   $c$ & $1$ \\
+                   $d$ & $0$ \\
+                   $e$ & $1$ \\
+                   $f$ & $0$ \\}
       \label{fig:alg2}
@@
       \LabelTables{$\{1a,2d\}$}
-                  {$3$ & 01 \\
-                   $4$ & 10 \\
-                   $5$ & 11 \\}
-                  {$b$ & 11 \\
-                   $c$ & 11 \\
-                   $e$ & 10 \\
-                   $f$ & 01 \\}
+                  {$3$ & $01$ \\
+                   $4$ & $10$ \\
+                   $5$ & $11$ \\}
+                  {$b$ & $11$ \\
+                   $c$ & $11$ \\
+                   $e$ & $10$ \\
+                   $f$ & $01$ \\}
       \label{fig:alg3}
@@
       \LabelTables{$\{1a,2d,3f\}$}
-                  {$4$ & 100 \\
-                   $5$ & 111 \\}
-                  {$b$ & 111 \\
-                   $c$ & 111 \\
-                   $e$ & 101 \\}
+                  {$4$ & $100$ \\
+                   $5$ & $111$ \\}
+                  {$b$ & $111$ \\
+                   $c$ & $111$ \\
+                   $e$ & $101$ \\}
       \label{fig:alg4}
@@
 \foreach \i in {2,3}
-  \node[class1node,on chain] (f\i) [label=left: \i] {};
+  \node[class1node,on chain] (f\i) [label=left: $\i$] {};
 \end{scope}
@@
 \foreach \i in {4,5}
-  \node[class2node,on chain] (f\i) [label=left: \i] {};
+  \node[class2node,on chain] (f\i) [label=left: $\i$] {};
 \end{scope}
@@
 \foreach \i in {d,f}
-  \node[class1node,on chain] (s\i) [label=right: \i] {};
+  \node[class1node,on chain] (s\i) [label=right: $\i$] {};
 \end{scope}
@@
 \foreach \i in {b,c,e}
-  \node[class2node,on chain] (s\i) [label=right: \i] {};
+  \node[class2node,on chain] (s\i) [label=right: $\i$] {};
 \end{scope}

### 2017-05-21 camera-ready-paper/paper.tex
Added camera-ready source files

--- 2017-05-19 paper/paper.tex
@@
 The algorithm builds up a mapping $M$ using a depth-first search, starting with the empty mapping
-$\emptyset$ and adding a $(v_i, w_i)$ pair at each level of the search tree.
+$\emptyset$ and adding a $(v_i, w_i)$ pair or choosing to leave a vertex in $\V(\graphG)$ unmatched at each level of the search tree.
 Select a vertex in $\V(\graphG)$ as the first vertex to be mapped; in our
@@
 The vertices in the bipartite graph are coloured to emphasise how they
-correspond to the two label classes.  Each label class corresponse to a
+correspond to the two label classes.  Each label class corresponds to a
 complete bipartite graph, and there are no edges between the components

### 2017-08-09 poster/poster.tex
Started poster

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :

\documentclass[a0paper]{tikzposter}

\usepackage{algorithm2e,algpseudocode}
\usepackage{complexity}
\usepackage{wrapfig}
\usepackage{microtype}
\usepackage{gnuplot-lua-tikz}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{sfmath}

\usepackage{lmodern}
\renewcommand*\familydefault{\sfdefault}
\usepackage[T1]{fontenc}

\newcommand{\McSplit}{\textproc{McSplit}}

\title{A Partitioning Algorithm for Maximum Common Subgraph Problems}
\author{Ciaran McCreesh, Patrick Prosser and James Trimble}
\institute{University of Glasgow, Glasgow, Scotland}
\titlegraphic{\includegraphics[keepaspectratio=true,scale=3.5]{UoG_keyline.pdf}}

\settitle{
    \begin{tikzpicture}
        \node (T) [inner sep=0pt] {\begin{minipage}{\linewidth}
                \color{titlefgcolor}
                {\bfseries \Huge \hspace*{10mm}A Partitioning Algorithm for Maximum \\
                \hspace*{10mm}Common Subgraph Problems \par}
                \vspace*{1em}
                {\Large {\bfseries \hspace{10mm}\@author}, \@institute}
        \end{minipage}};

        \node at (T.east) [anchor=center, inner sep=0pt, xshift=-12cm] {\@titlegraphic};
    \end{tikzpicture}
}

% University of Glasgow standard colours
\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}

\definecolorstyle{UofG}{
}{
    % Background Colors
    \colorlet{backgroundcolor}{uofgsandstone!80!white}
    \colorlet{framecolor}{black}
    % Title Colors
    \colorlet{titlefgcolor}{white}
    \colorlet{titlebgcolor}{uofguniversityblue}
    % Block Colors
    \colorlet{blocktitlebgcolor}{white}
    \colorlet{blocktitlefgcolor}{uofguniversityblue}
    \colorlet{blockbodybgcolor}{white}
    \colorlet{blockbodyfgcolor}{black}
    % Innerblock Colors
    \colorlet{innerblocktitlebgcolor}{uofguniversityblue}
    \colorlet{innerblocktitlefgcolor}{black}
    \colorlet{innerblockbodybgcolor}{uofgsandstone}
    \colorlet{innerblockbodyfgcolor}{black}
    % Note colors
    \colorlet{notefgcolor}{black}
    \colorlet{notebgcolor}{uofgrust}
    \colorlet{noteframecolor}{red}
}

\usetheme{Autumn}
\usecolorstyle{UofG}

\tikzposterlatexaffectionproofoff

\useblockstyle[bodyverticalshift=-1cm, roundedcorners=1]{Default}

\renewcommand{\Huge}{\fontsize{77.2}{96}\selectfont}

% Styles for drawings

\tikzset{edge/.style={line width=3pt, color=uofgsandstone}}
\tikzset{ledge/.style={line width=3pt, color=uofgsandstone!40!white}}
\tikzset{hedge/.style={line width=3pt, color=uofgsandstone, dashed}}

\setlength\intextsep{0pt}

\begin{document}
\maketitle

{
    \colorlet{blockbodybgcolor}{uofgcobalt}
    \colorlet{blocktitlebgcolor}{uofgcobalt}
    \block[bodyverticalshift=0cm, bodyinnersep=3mm]{}{
        \centering\begin{minipage}{0.94\textwidth}
            We introduce a new branch and bound algorithm for the \textbf{maximum common
            subgraph} and \textbf{maximum common connected subgraph} problems which is based
            around vertex labelling and partitioning. Our method resembles
            a constraint programming approach, but uses a novel compact
            domain store to reduce
            the memory and computation requirements during search.  Experiments
            show a speedup of more than an order of magnitude over the state of the
            art, and demonstrate that we can operate on much larger graphs without
            running out of memory.
%            We show how to generate \textbf{really hard} random instances for \textbf{subgraph
%            isomorphism} problems. For the non-induced variant, we predict and observe a phase
%            transition between satisfiable and unsatisfiable instances, with a corresponding
%            complexity peak seen in \textbf{three different solvers}. For the induced variant, much
%            richer behaviour is observed, and \textbf{constrainedness} gives a better measure of
%            difficulty than does proximity to a phase transition. We also discuss \textbf{variable
%            and value ordering} heuristics, and their relationship to the \textbf{expected number of
%            solutions}.
        \end{minipage}
    }
}

\begin{columns}
\column{0.5}

\block{The Maximum Common Subgraph Problem}{
\begin{wrapfigure}[7]{r}{0.46\linewidth}
    \begin{center}\begin{tikzpicture}[scale=1.4]%{{{
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nc) at (1, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nd) at (3, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (Ne) at (3, -2)
        {\vphantom{0}};

        \draw [edge] (Na) -- (Nb);
        \draw [edge] (Nb) -- (Nc);
        \draw [edge] (Nc) -- (Nd);
        \draw [edge] (Nb) -- (Nd);
        \draw [ledge] (Na) -- (Ne);
        \draw [ledge] (Nb) -- (Ne);
        \draw [ledge] (Nd) -- (Ne);

        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};

        \draw [ledge] (N1) -- (N2);
        \draw [edge] (N1) -- (N3);
        \draw [ledge] (N1) -- (N4);
        \draw [ledge] (N2) -- (N4);
        \draw [edge] (N3) -- (N5);
        \draw [edge] (N3) -- (N6);
        \draw [ledge] (N4) -- (N6);
        \draw [edge] (N5) -- (N6);
        \draw [ledge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{center}

\end{wrapfigure}

The \textbf{maximum common subgraph} family of problems involves finding a large
graph which is isomorphic to subgraphs of two given graphs simultaneously.
Because graphs are widely used to model real-world phenomena, maximum common
subgraph problems have arisen in molecular science (where graphs often represent
molecules)
and also in other domains including malware detection, source code analysis, and computer vision.

\bigskip

Maximum common subgraph problems are NP-hard, and remain challenging
computationally. Recent practical progress has been made by using constraint
programming and
mathematical programming, by reducing
to the maximum clique problem, and by
adapting subgraph isomorphism algorithms. Some
special cases also have practical polynomial time algorithms.

\bigskip

We consider the \textbf{maximum common induced subgraph} problem, in
which the objective is to find a graph with as many vertices as possible which
is an induced subgraph of each of two input graphs.  We introduce
a new branch and bound algorithm which exploits special properties of the
problem to allow a much faster exploration of the search space, whilst
retaining the filtering and bounding benefits of the constraint programming
approach. 
}

\block{What goes here?}{

%If we randomly generate a pattern graph with $p$ vertices and density $d_p$, and a target graph
%with $t$ vertices and density $d_t$, the \textbf{expected number of solutions} to the non-induced
%isomorphism problem is \[
%    \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot {d_t}^{d_p \cdot
%    \binom{p}{2}} } \textnormal{.} \]

Walkthrough?

\bigskip

Parallel version??

%By considering when $\mathsf{\langle Sol \rangle = 1}$, we can \textbf{estimate the location} of the
%satisfiable / unsatisfiable phase transition. (This is not entirely accurate due to variance,
%particularly when the pattern is very sparse or very dense).

\bigskip

Extensions to labelled and connected cases???

%We can also use this formula to recover \textbf{variable} and \textbf{value ordering heuristics}: we
%should make branching decisions which maximise the expected number of solutions during search. This
%tells us to prefer small domains and pattern vertices of high degree, but target vertices of low
%degree, which matches the empirically-determined heuristics used in state of the art solvers.

}

\column{0.5}

\block{Horse Race}{
We implemented \McSplit in C++. We compare against the best constraint
programming implementations 
(CP-FC in the unlabelled cases, and CP-MAC
in the labelled cases, using both branching and filtering for connected
subgraphs), the encodings, and the
$k{\downarrow}$ algorithm (which only supports
unlabelled, undirected, unconnected instances).  Each of these comparator
programs is an optimised, dedicated implementation and does not use a
general-purpose constraint programming toolkit.

Our first set of experiments uses a database of randomly-generated maximum
common subgraph instances.  For
unlabelled instances, we selected the first ten instances from each family
whose members have no more than 50 vertices, for a total of 4,100 instances.
For labelled instances, we selected the first ten instances from every family,
for a total of 8,140 instances with up to 100 vertices;
we use the labelling scheme in which the
number of distinct vertex labels and the number of distinct edge labels is
approximately equal to 33 percent of the number of vertices in each graph.

    \vspace*{1em}

    \begin{center}
        \small
        \input{gen-graph-plain-cumulative}
        %\input{gen-graph-phase-transition}
    \end{center}

%    We can also look at what happens if we vary both the pattern and the target density, for
%    different orders of pattern. The top row shows satisfiability---the black lines show where
%    $\mathsf{\langle Sol \rangle = 1}$. The bottom row shows the search cost for one of the three
%    algorithms we consider in the paper. As expected, the ``really hard'' instances are near the
%    phase transition.

    \vspace*{0.9em}

    \begin{center}
        \small
        %\input{gen-graph-non-induced}
    \end{center}
}

\end{columns}

\block{Is this Just a Tweaked Version of CP-FC?}{
\begin{wrapfigure}[24]{r}{0.64\linewidth}\begin{center}
    %\small\input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
\end{center}\end{wrapfigure}

    Yes.

    \bigskip

    We show that here with a nice heatmap and a discussion of the soft all-diff propagator,
    maybe with a bipartite graph figure.

%    For induced isomorphisms, the behaviour is much richer. We can derive an expected number of
%    solutions of \[
%        \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot
%                    {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} } \textnormal{.} \]
%
%    \bigskip
%
%    Again this predicts a sharp phase transition. When the pattern is small, we see hard instances
%    near the phase transition, and easy instances elsewhere. However, when the pattern is larger,
%    the central unsatisfiable region \textbf{remains very hard}, even though it is no longer near a
%    phase transition.
%
%    \bigskip
%
%    This is not just a weakness of current subgraph isomorphism algorithms: the region is also hard
%    for pseudo-boolean and boolean satisfiability solvers, and under reduction to the clique
%    problem.
%
%    \bigskip
%
%    The central hard region \emph{is} predicted correctly by \textbf{constrainedness}, as we
%    show in the third row of plots. Although far from a phase transition, this region is only
%    slightly overconstrained.
%
%    \bigskip
%
%    Looking to maximise $\mathsf{\langle Sol \rangle}$ derives existing variable and value ordering
%    heuristics, in the case that both the pattern and target graphs are sparse. The formula suggests
%    that algorithms should swap heuristics when the pattern and target are dense---empirically, this
%    does indeed give an improvement. However, maximising the expected number of solutions is
%    \textbf{not enough to select the best heuristic} in two eighths of the search space (and
%    constrainedness does not help).
}

\begin{columns}
\column{0.5}

\block{See the Paper For\ldots}{
    \begin{itemize}
        \item ~ How the algorithm can be (straightforwardly extended) to handle vertex and edge labels
        \item ~ Experimental results for the labelled and connected cases
        \item ~ ...???
    \end{itemize}
}

\column{0.5}

\block{Future Work}{
    \begin{itemize}
        \item ~ Better branching heuristics
        \item ~ Branching on vertices in both graphs, not just the first one
        \item ~ Applying the partitioning data structure to other problems
    \end{itemize}
}

\end{columns}

{
    \colorlet{blockbodybgcolor}{uofgsandstone!80!white}
    \colorlet{blocktitlebgcolor}{uofgsandstone!80!white}
    \block[bodyverticalshift=-0.5cm]{}{
        This work was supported by the Engineering and Physical Sciences Research Council [grant
        number EP/K503058/1]. \hfill \texttt{\textcolor{white}{c.mccreesh.1@research.gla.ac.uk}}
    }
}

\end{document}

### 2017-08-09 poster/poster.tex
First draft of poster

@@
 
+\begin{columns}
+\column{1}
 {
@@
     \colorlet{blocktitlebgcolor}{uofgcobalt}
-    \block[bodyverticalshift=0cm, bodyinnersep=3mm]{}{
+    \block[bodyverticalshift=0.5cm, bodyinnersep=3mm]{}{
         \centering\begin{minipage}{0.94\textwidth}
-            We introduce a new branch and bound algorithm for the \textbf{maximum common
+            \Large We introduce a new branch and bound algorithm for the \textbf{maximum common
             subgraph} and \textbf{maximum common connected subgraph} problems which is based
@@
             domain store to reduce
-            the memory and computation requirements during search.  Experiments
+            the memory and computation requirements.  Experiments
             show a speedup of more than an order of magnitude over the state of the
@@
         \end{minipage}
+        \vspace*{0.4cm}
     }
 }
+\end{columns}
 
@@
 \block{The Maximum Common Subgraph Problem}{
-\begin{wrapfigure}[7]{r}{0.46\linewidth}
-    \begin{center}\begin{tikzpicture}[scale=1.4]%{{{
+\begin{wrapfigure}[7]{r}{0.48\linewidth}
+    \begin{flushright}\begin{tikzpicture}[scale=1.6]%{{{
         \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
@@
         \draw [ledge] (N8) -- (N9);
-    \end{tikzpicture}\end{center}
+    \end{tikzpicture}\end{flushright}
 
@@
 The \textbf{maximum common subgraph} family of problems involves finding a large
-graph which is isomorphic to subgraphs of two given graphs simultaneously.
-Because graphs are widely used to model real-world phenomena, maximum common
-subgraph problems have arisen in molecular science (where graphs often represent
-molecules)
-and also in other domains including malware detection, source code analysis, and computer vision.
+graph which is isomorphic to subgraphs of two input graphs.
+Such problems have arisen in molecular science---where graphs often represent
+molecules---malware detection, source code analysis, and computer vision.
 
@@
 
-Maximum common subgraph problems are NP-hard, and remain challenging
+    Maximum common subgraph problems are \mbox{NP-hard}, and remain challenging
 computationally. Recent practical progress has been made by using constraint
@@
 which the objective is to find a graph with as many vertices as possible which
-is an induced subgraph of each of two input graphs.  We introduce
-a new branch and bound algorithm which exploits special properties of the
-problem to allow a much faster exploration of the search space, whilst
-retaining the filtering and bounding benefits of the constraint programming
-approach. 
+is an induced subgraph of each of two input graphs.  The diagram shows a simple instance
+with an optimal solution highlighted.
 }
 
-\block{What goes here?}{
+\block{The \McSplit\ Algorithm}{
 
-%If we randomly generate a pattern graph with $p$ vertices and density $d_p$, and a target graph
-%with $t$ vertices and density $d_t$, the \textbf{expected number of solutions} to the non-induced
-%isomorphism problem is \[
-%    \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot {d_t}^{d_p \cdot
-%    \binom{p}{2}} } \textnormal{.} \]
-
-Walkthrough?
+We introduce a branch and bound algorithm which exploits special properties of the
+problem to allow a much faster exploration of the search space, whilst
+retaining the filtering and bounding benefits of a constraint programming
+approach.
 
@@
 
-Parallel version??
-
-%By considering when $\mathsf{\langle Sol \rangle = 1}$, we can \textbf{estimate the location} of the
-%satisfiable / unsatisfiable phase transition. (This is not entirely accurate due to variance,
-%particularly when the pattern is very sparse or very dense).
+Using depth-first search, the \McSplit\ algorithm builds up an injective mapping from a subset
+of vertices in the pattern graph to a subset of vertices in the target graph.  
+Remaining vertices are partitioned into \textbf{label classes} according to the set of mapped
+vertices to which they are adjacent; each pattern-graph vertex may only be mapped to target-graph
+vertices in the same label class.  We thus avoid
+the need for a conventional domain store in which the possible mappings for each
+vertex are stored explicitly.
 
@@
 
-Extensions to labelled and connected cases???
+Straightforward extensions can be made to the algorithm to handle edge- and vertex-labelled graphs,
+and directed graphs.  We can also solve the problem where the subgraph must be connected,
+by branching only on vertices adjacent to at least one currently-mapped vertex.
+}
 
-%We can also use this formula to recover \textbf{variable} and \textbf{value ordering heuristics}: we
-%should make branching decisions which maximise the expected number of solutions during search. This
-%tells us to prefer small domains and pattern vertices of high degree, but target vertices of low
-%degree, which matches the empirically-determined heuristics used in state of the art solvers.
+\block{Relationship to Prior Constraint Programming Encodings}{
 
+    \bigskip
+
+    Our experimental results suggest that \McSplit\ has broadly
+    similar performance trends to the constraint programming,
+    forward-checking (CP-FC) algorithm of Ndiaye and Solnon
+    (2011), but with much lower constant factors and memory
+    usage.  In the figure below, we plot the number
+    of recursive calls made by our algorithm versus the number
+    made by CP-FC, for unlabelled, unconnected
+    problem instances. (Rather than a simple scatter plot,
+    darker colours are used to indicate a higher density of points around
+    a location.) We see a close correlation: \McSplit\ typically
+    does slightly less work, and sometimes does more, but instances with
+    more than one order of magnitude difference in
+    search tree size are rare.
+
+    Why is this?  The key
+    observation is that \McSplit\ may be considered to be a different version of the CP-FC algorithm,
+    using an unconventional domain store and more efficient filtering algorithms.
+    We explore this relationship in the paper.
+
+\begin{center}
+    \small\input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
+\end{center}
 }
@@
 
-\block{Horse Race}{
-We implemented \McSplit in C++. We compare against the best constraint
-programming implementations 
-(CP-FC in the unlabelled cases, and CP-MAC
-in the labelled cases, using both branching and filtering for connected
-subgraphs), the encodings, and the
-$k{\downarrow}$ algorithm (which only supports
-unlabelled, undirected, unconnected instances).  Each of these comparator
+\block{Experiments}{
+We compare our implementation of \McSplit\ against the best existing constraint
+programming implementations, clique encodings, and the
+$k{\downarrow}$ algorithm of McCreesh et al. (2017).  Each of these comparator
 programs is an optimised, dedicated implementation and does not use a
@@
 
-Our first set of experiments uses a database of randomly-generated maximum
-common subgraph instances.  For
-unlabelled instances, we selected the first ten instances from each family
-whose members have no more than 50 vertices, for a total of 4,100 instances.
-For labelled instances, we selected the first ten instances from every family,
-for a total of 8,140 instances with up to 100 vertices;
-we use the labelling scheme in which the
-number of distinct vertex labels and the number of distinct edge labels is
-approximately equal to 33 percent of the number of vertices in each graph.
+\bigskip
 
-    \vspace*{1em}
+For the unlabelled problem, we used 4,100 instances from a
+database of randomly-generated maximum common subgraph instances.  For
+the labelled variant, we used 8,140 instances from the same database.  \McSplit\ is
+more than an order of magnitude faster than its nearest competitor on unlabelled instances.
+On labelled instances, the clique encoding remains the single best solver.
+
+    \vspace*{2em}
 
@@
         \input{gen-graph-plain-cumulative}
-        %\input{gen-graph-phase-transition}
     \end{center}
 
-%    We can also look at what happens if we vary both the pattern and the target density, for
-%    different orders of pattern. The top row shows satisfiability---the black lines show where
-%    $\mathsf{\langle Sol \rangle = 1}$. The bottom row shows the search cost for one of the three
-%    algorithms we consider in the paper. As expected, the ``really hard'' instances are near the
-%    phase transition.
-
-    \vspace*{0.9em}
+    \vspace*{2em}
 
@@
         \small
-        %\input{gen-graph-non-induced}
+        \input{gen-graph-33ved-cumulative}
     \end{center}
+
+    \vspace*{3em}
+
+We also ran the algorithms on a set of 5,725 larger instances used in recent
+studies of subgraph isomorphism and maximum common subgraph; a modified version
+of \McSplit\ is the single best solver on this set of instances.
+
+    \vspace*{3em}
+
 }
@@
 \end{columns}
-
-\block{Is this Just a Tweaked Version of CP-FC?}{
-\begin{wrapfigure}[24]{r}{0.64\linewidth}\begin{center}
-    %\small\input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
-\end{center}\end{wrapfigure}
-
-    Yes.
-
-    \bigskip
-
-    We show that here with a nice heatmap and a discussion of the soft all-diff propagator,
-    maybe with a bipartite graph figure.
-
-%    For induced isomorphisms, the behaviour is much richer. We can derive an expected number of
-%    solutions of \[
-%        \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot
-%                    {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} } \textnormal{.} \]
-%
-%    \bigskip
-%
-%    Again this predicts a sharp phase transition. When the pattern is small, we see hard instances
-%    near the phase transition, and easy instances elsewhere. However, when the pattern is larger,
-%    the central unsatisfiable region \textbf{remains very hard}, even though it is no longer near a
-%    phase transition.
-%
-%    \bigskip
-%
-%    This is not just a weakness of current subgraph isomorphism algorithms: the region is also hard
-%    for pseudo-boolean and boolean satisfiability solvers, and under reduction to the clique
-%    problem.
-%
-%    \bigskip
-%
-%    The central hard region \emph{is} predicted correctly by \textbf{constrainedness}, as we
-%    show in the third row of plots. Although far from a phase transition, this region is only
-%    slightly overconstrained.
-%
-%    \bigskip
-%
-%    Looking to maximise $\mathsf{\langle Sol \rangle}$ derives existing variable and value ordering
-%    heuristics, in the case that both the pattern and target graphs are sparse. The formula suggests
-%    that algorithms should swap heuristics when the pattern and target are dense---empirically, this
-%    does indeed give an improvement. However, maximising the expected number of solutions is
-%    \textbf{not enough to select the best heuristic} in two eighths of the search space (and
-%    constrainedness does not help).
-}
 
@@
     \begin{itemize}
-        \item ~ How the algorithm can be (straightforwardly extended) to handle vertex and edge labels
-        \item ~ Experimental results for the labelled and connected cases
-        \item ~ ...???
+        \item ~ A walk-through of the algorithm
+        \item ~ How to extend the algorithm to handle vertex and edge labels
+        \item ~ Additional experimental results
     \end{itemize}
@@
         \item ~ Better branching heuristics
-        \item ~ Branching on vertices in both graphs, not just the first one
+        \item ~ Branching on vertices in both graphs
         \item ~ Applying the partitioning data structure to other problems
     \end{itemize}
+    \vspace*{0.25em}
 }

### 2017-08-09 poster/poster.tex
Poster tweaks

@@
 
-\useblockstyle[bodyverticalshift=-1cm, roundedcorners=1]{Default}
+\useblockstyle[bodyverticalshift=-1cm, roundedcorners=0]{Default}
 
@@
         \centering\begin{minipage}{0.94\textwidth}
-            \Large We introduce a new branch and bound algorithm for the \textbf{maximum common
+            \Large We introduce a new \textbf{branch and bound} algorithm for the \textbf{maximum common
             subgraph} and \textbf{maximum common connected subgraph} problems which is based
@@
             the memory and computation requirements.  Experiments
-            show a speedup of more than an order of magnitude over the state of the
-            art, and demonstrate that we can operate on much larger graphs without
+            show a \textbf{speedup of more than an order of magnitude} over the state of the
+            art, and demonstrate that we can operate on \textbf{much larger graphs} without
             running out of memory.
@@
 \block{The Maximum Common Subgraph Problem}{
-\begin{wrapfigure}[7]{r}{0.48\linewidth}
+\begin{wrapfigure}[7]{r}{0.47\linewidth}
     \begin{flushright}\begin{tikzpicture}[scale=1.6]%{{{
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Na) at (1,  0)
         {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Nb) at (1, -2)
         {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nc) at (1, -4)
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Nc) at (1, -4)
         {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nd) at (3, -4)
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Nd) at (3, -4)
         {\vphantom{0}};
-        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (Ne) at (3, -2)
+        \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (Ne) at (3, -2)
         {\vphantom{0}};
@@
 
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
-        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
-        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
-        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
-        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
-        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
+        \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
+        \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
+        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
+        \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
+        \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
+        \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};
 
@@
 The \textbf{maximum common subgraph} family of problems involves finding a large
-graph which is isomorphic to subgraphs of two input graphs.
+graph that appears in each of two input graphs.
 Such problems have arisen in molecular science---where graphs often represent
@@
 vertex are stored explicitly.  
-
-\bigskip
-
-Straightforward extensions can be made to the algorithm to handle edge- and vertex-labelled graphs,
-and directed graphs.  We can also solve the problem where the subgraph must be connected,
-by branching only on vertices adjacent to at least one currently-mapped vertex.
 }
 
-\block{Relationship to Prior Constraint Programming Encodings}{
+\block{The Relationship to Constraint Programming}{
 
@@
     using an unconventional domain store and more efficient filtering algorithms.
-    We explore this relationship in the paper.
+    This is possible because in the CP model, \textbf{any two domains are either identical
+    or disjoint}, excluding $\bot$.  We explore this relationship further in the paper.
+
+    \vspace*{1em}
 
@@
 \end{center}
+}
+
+\block{See the Paper For\ldots}{
+    \begin{itemize}
+        \item ~ How to extend the algorithm to handle vertex and edge labels
+        \item ~ Additional experimental results
+    \end{itemize}
 }
@@
 
-    \vspace*{3em}
+    \vspace*{2em}
 
@@
 
-    \vspace*{3em}
+    \vspace*{2em}
 
+    \begin{center}
+        \small
+        \input{gen-graph-sip-cumulative}
+    \end{center}
 }
@@
 
-\begin{columns}
-\column{0.5}
-
-\block{See the Paper For\ldots}{
-    \begin{itemize}
-        \item ~ A walk-through of the algorithm
-        \item ~ How to extend the algorithm to handle vertex and edge labels
-        \item ~ Additional experimental results
-    \end{itemize}
-}
-
-\column{0.5}
-
-\block{Future Work}{
-    \begin{itemize}
-        \item ~ Better branching heuristics
-        \item ~ Branching on vertices in both graphs
-        \item ~ Applying the partitioning data structure to other problems
-    \end{itemize}
-    \vspace*{0.25em}
-}
-
-\end{columns}
+%\begin{columns}
+%\column{0.5}
+%
+%\column{0.5}
+%
+%%\block{Future Work}{
+%%    \begin{itemize}
+%%        \item ~ Branching on vertices in both graphs
+%%        \item ~ Applying the partitioning data structure to other problems
+%%    \end{itemize}
+%%    \vspace*{0.25em}
+%%}
+%
+%\end{columns}
 
@@
     \block[bodyverticalshift=-0.5cm]{}{
-        This work was supported by the Engineering and Physical Sciences Research Council [grant
-        number EP/K503058/1]. \hfill \texttt{\textcolor{white}{c.mccreesh.1@research.gla.ac.uk}}
+        This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1 and
+        EP/M508056/1]. \hfill \texttt{\textcolor{white}{j.trimble.1@research.gla.ac.uk}}
     }

### 2017-08-11 poster/poster.tex
Tweaked poster

@@
 
+\definecolor{magmamagenta}{rgb}{0.709804,0.211765,0.478431}
+
 \definecolorstyle{UofG}{
@@
             running out of memory.
-%            We show how to generate \textbf{really hard} random instances for \textbf{subgraph
-%            isomorphism} problems. For the non-induced variant, we predict and observe a phase
-%            transition between satisfiable and unsatisfiable instances, with a corresponding
-%            complexity peak seen in \textbf{three different solvers}. For the induced variant, much
-%            richer behaviour is observed, and \textbf{constrainedness} gives a better measure of
-%            difficulty than does proximity to a phase transition. We also discuss \textbf{variable
-%            and value ordering} heuristics, and their relationship to the \textbf{expected number of
-%            solutions}.
         \end{minipage}
@@
     \begin{flushright}\begin{tikzpicture}[scale=1.6]%{{{
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Na) at (1,  0)
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (Na) at (1,  0)
         {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Nb) at (1, -2)
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (Nb) at (1, -2)
         {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Nc) at (1, -4)
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (Nc) at (1, -4)
         {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (Nd) at (3, -4)
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (Nd) at (3, -4)
         {\vphantom{0}};
@@
 
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
         \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
         \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
-        \node[draw, circle, fill=uofgrose, inner sep=8pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
+        \node[draw, circle, fill=magmamagenta, inner sep=8pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
         \node[draw, circle, fill=white, inner sep=8pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
@@
     \begin{itemize}
-        \item ~ How to extend the algorithm to handle vertex and edge labels
+        \item ~ The algorithm in all its magnificent glory
         \item ~ Additional experimental results

### 2017-08-11 talk/talk.tex
Added talk

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 foldmethod=marker foldmarker={{{,}}} :

\documentclass{beamer}

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}

\newcommand{\McSplit}{\textproc{McSplit}}

\usetikzlibrary{shapes, arrows, shadows, calc, positioning, fit}
\usetikzlibrary{decorations.pathreplacing, decorations.pathmorphing, shapes.misc}
\usetikzlibrary{tikzmark}

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black, ultra thick}}
\tikzset{ledge/.style={color=uofgsandstone!50!white, ultra thick}}
\tikzset{bedge/.style={ultra thick}}

% {{{ theme things
\useoutertheme[footline=authortitle]{miniframes}
\useinnertheme{rectangles}

\setbeamerfont{block title}{size={}}
\setbeamerfont{title}{size=\large,series=\bfseries}
\setbeamerfont{section title}{size=\large,series=\mdseries}
\setbeamerfont{author}{size=\normalsize,series=\mdseries}
\setbeamercolor*{structure}{fg=uofguniversityblue}
\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=uofguniversityblue}
\setbeamercolor*{palette quaternary}{fg=white,bg=black}

\setbeamercolor*{titlelike}{parent=palette primary}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{title page}
{
    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };

        \node [anchor=west, xshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

\setbeamertemplate{section page}
{
    \begin{centering}
        \begin{beamercolorbox}[sep=12pt,center]{part title}
            \usebeamerfont{section title}\insertsection\par
        \end{beamercolorbox}
    \end{centering}
}

\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\makeatletter
\setbeamertemplate{footline}
{%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
        \hfill%
        {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        {\usebeamerfont{title in head/foot}\insertshorttitle}%
        \hfill%
        {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
}

\makeatletter
\newenvironment{nearlyplainframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}\vspace*{3pt}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{#2}
    }{
    \end{frame}
}
\makeatother

% }}}

\title[A Partitioning Algorithm for Maximum Common Subgraph Problems]{A Partitioning Algorithm for \\
Maximum Common Subgraph Problems}
\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{Ciaran McCreesh, Patrick
Prosser and \\ James Trimble}

\begin{document}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics*[keepaspectratio=true, width=\paperwidth]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\begin{frame}{Maximum Common Subgraph}

    \begin{center}\begin{tikzpicture}
        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (Nc) at (1, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (Nd) at (3, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (Ne) at (3, -2)
        {\vphantom{0}};

        \draw [edge] (Na) -- (Nb);
        \draw [edge] (Nb) -- (Nc);
        \draw [edge] (Nc) -- (Nd);
        \draw [edge] (Nb) -- (Nd);
        \draw [ledge] (Na) -- (Ne);
        \draw [ledge] (Nb) -- (Ne);
        \draw [ledge] (Nd) -- (Ne);

        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=uofgpumpkin, inner sep=5pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};

        \draw [ledge] (N1) -- (N2);
        \draw [edge] (N1) -- (N3);
        \draw [ledge] (N1) -- (N4);
        \draw [ledge] (N2) -- (N4);
        \draw [edge] (N3) -- (N5);
        \draw [edge] (N3) -- (N6);
        \draw [ledge] (N4) -- (N6);
        \draw [edge] (N5) -- (N6);
        \draw [ledge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{center}

\end{frame}

\begin{frame}{Previous Algorithms}

    \begin{itemize}
        \item \textbf{Constraint programming} --- good for unlabelled instances
        \item \textbf{Clique encoding} --- good for labelled instances
        \item $\mathbf{k{\downarrow}}$ --- good for big instances
    \end{itemize}

\end{frame}

\begin{frame}{CP Model}

    \begin{itemize}
        \item A variable for each vertex in the pattern graph
        \item A value for each vertex in the target graph, and $\bot$
        \item Soft all-different constraint
        \item Maximise the number of non-$\bot$ values
    \end{itemize}

\end{frame}

\begin{frame}{The Observation Behind our Algorithm}

    \begin{itemize}
        \item During search any two domains are identical or
            disjoint (with the exception of $\bot$)
    \end{itemize}

\end{frame}

\begin{frame}{Experiments}

    \begin{itemize}
        \item Unlabelled variant: 4,100 instances from a
          database of randomly-generated maximum common subgraph instances (up~to 50 vertices)
        \item Labelled variant: 8,140 instances from the same database (up~to 100 vertices)
        \item We also ran the algorithms on a set of 5,725 larger instances (up~to 6,671 vertices)
    \end{itemize}

\end{frame}

\begin{nearlyplainframe}[t]{Unlabelled Instances}
    \centering
        \input{gen-graph-plain-cumulative}
\end{nearlyplainframe}

\begin{nearlyplainframe}[t]{Labelled, Directed Instances}
    \centering
        \input{gen-graph-33ved-cumulative}
\end{nearlyplainframe}

\begin{nearlyplainframe}[t]{Large Instances}
    \centering
        \input{gen-graph-sip-cumulative}
\end{nearlyplainframe}

\begin{frame}{In the Paper}

    \begin{itemize}
        \item A detailed explanation of the algorithm
        \item Interesting things about dual viewpoint heuristics
        \item Extensions to labelled and directed graphs
        \item Maximum common \textbf{connected} subgraph
    \end{itemize}

\end{frame}

\begin{frame}[plain,noframenumbering]
    \begin{center}
        \vspace*{4em}
        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
        \vspace{1em}
        \href{mailto:ciaran.mccreesh@gla.ac.uk}{\nolinkurl{ciaran.mccreesh@gla.ac.uk}} \\
    \end{center}

    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };
    \end{tikzpicture}
\end{frame}

\begin{frame}{Branching Heuristics}

    \begin{itemize}
        \item Select a label class with minimum $\max(|G|,|H|)$
        \item From this class, choose a vertex with minimum degree
        \item Value ordering: choose a vertex of minimum degree first
    \end{itemize}

\end{frame}

\begin{frame}{Soft All-Different Constraint}

\usetikzlibrary{positioning,chains,fit,shapes,calc}

\begin{center}
\begin{tikzpicture}[thick,
  every node/.style={draw,circle},
  class1node/.style={fill=uofgcobalt},
  class2node/.style={fill=uofgthistle},
  every fit/.style={ellipse,draw,inner sep=-2pt,text width=2cm}
]

\begin{scope}[start chain=going below,node distance=3mm]
\foreach \i in {2,3}
  \node[class1node,on chain] (f\i) [label=left: $\i$] {};
\end{scope}

\begin{scope}[yshift=-1.8cm,start chain=going below,node distance=3mm]
\foreach \i in {4,5}
  \node[class2node,on chain] (f\i) [label=left: $\i$] {};
\end{scope}

\begin{scope}[xshift=2.2cm,start chain=going below,node distance=3mm]
\foreach \i in {d,f}
  \node[class1node,on chain] (s\i) [label=right: $\i$] {};
\end{scope}

\begin{scope}[xshift=2.2cm,yshift=-1.5cm,start chain=going below,node distance=3mm]
\foreach \i in {b,c,e}
  \node[class2node,on chain] (s\i) [label=right: $\i$] {};
\end{scope}

% the edges
\draw (f2) -- (sd);
\draw (f3) -- (sd);
\draw (f2) -- (sf);
\draw (f3) -- (sf);

\draw (f4) -- (sb);
\draw (f4) -- (sc);
\draw (f4) -- (se);
\draw (f5) -- (sb);
\draw (f5) -- (sc);
\draw (f5) -- (se);
\end{tikzpicture}
\end{center}

\end{frame}

\begin{nearlyplainframe}[t]{}
    \centering
        \input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
\end{nearlyplainframe}

\end{document}

### 2017-08-15 talk/talk.tex
Muck around

@@
 
-        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
-            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
+        \node (logo) [anchor=north east, shift={(-0.3cm,-0.6cm)}] at (current page.north east) {
+            \includegraphics*[keepaspectratio=true,scale=0.65]{UoG_keyline.pdf}
         };
@@
 
+\begin{frame}{The McSplit Algorithm}
+    \begin{itemize}
+        \item We can use a compact backtrackable data structure which splits vertices into
+            partitions
+        \item Propagation is just refining partitions
+        \item Soft all-different can be calculated by counting rather than matching
+    \end{itemize}
+\end{frame}
+
 \begin{frame}{Experiments}
@@
         \vspace{1em}
-        \href{mailto:ciaran.mccreesh@gla.ac.uk}{\nolinkurl{ciaran.mccreesh@gla.ac.uk}} \\
+        \href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}} \\
     \end{center}

### 2017-08-23 talk/talk.tex
Think I want this slide
I'm plotting a better talk

@@
         \item Labelled variant: 8,140 instances from the same database (up~to 100 vertices)
-        \item We also ran the algorithms on a set of 5,725 larger instances (up~to 6,671 vertices)
+        \item We also ran the algorithms on a set of 5,725 larger instances (up~to 6,671 vertices,
+            one graph is often smaller than the other)
     \end{itemize}
@@
         \input{gen-graph-plain-cumulative}
+\end{nearlyplainframe}
+
+\begin{nearlyplainframe}[t]{}
+    \centering
+        \input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
+\end{nearlyplainframe}
+
+\begin{nearlyplainframe}[t]{}
+    \centering
+        \input{gen-graph-plain-james-versus-cp-fc-runtimes-scatter}
 \end{nearlyplainframe}
@@
         \input{gen-graph-sip-cumulative}
+\end{nearlyplainframe}
+
+\begin{nearlyplainframe}[t]{}
+    \centering
+        \input{gen-graph-sip-james-versus-kdown-nodes-scatter}
 \end{nearlyplainframe}
@@
 
-\begin{nearlyplainframe}[t]{}
-    \centering
-        \input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
-\end{nearlyplainframe}
-
 \end{document}

### 2017-08-23 paper/heatmapify.rb

@@ -14,10 +14,10 @@ IO.readlines(ARGV[0]).each_with_index do | line, index |
     next if words[ARGV[1].to_i] == "NaN" or words[ARGV[2].to_i] == "NaN"
 
     x = words[ARGV[1].to_i].to_f * xmul
-    bx = if x < 1 then 0 else (granularity * (Math.log(x + 0.1) / Math.log(1e10))).round end
+    bx = if x < 1 then 0 else (granularity * (Math.log(x + 0.1) / Math.log(ARGV[5].to_f))).round end
 
     y = words[ARGV[2].to_i].to_f * ymul
-    by = if y < 1 then 0 else (granularity * (Math.log(y + 0.1) / Math.log(1e10))).round end
+    by = if y < 1 then 0 else (granularity * (Math.log(y + 0.1) / Math.log(ARGV[5].to_f))).round end
 
     abort "bx is " + bx.to_s + " by is " + by.to_s + " on line " + (index + 1).to_s if bx < 0 || bx > granularity || by < 0 || by > granularity
     buckets[[bx, by]] += 1

### 2017-08-23 talk/talk.tex
A bold change
Geez a job

@@
 Maximum Common Subgraph Problems}
-\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{Ciaran McCreesh, Patrick
-Prosser and \\ James Trimble}
+\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick
+Prosser, and \\ James Trimble}
 
@@
 
-\begin{frame}{The Observation Behind our Algorithm}
+\begin{frame}{The Observation Behind Our Algorithm}
 
@@
         \url{http://www.dcs.gla.ac.uk/~ciaran} \\
-        \vspace{1em}
+        \vspace{0.2em}
         \href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}} \\
+        \vspace{2em}
+        \textcolor{uofgpillarbox}{
+            Are you finding Melbourne to be a bit too warm and dry?  We are hiring a postdoc to work on
+            high level constraint modelling for graph problems as part of a ``subgraphs modulo
+            theories'' project:} \\
+        \vspace{0.2em}
+        \href{https://www.vacancies.st-andrews.ac.uk/}{\nolinkurl{https://www.vacancies.st-andrews.ac.uk/}}
+        \\ reference AR1972SB
     \end{center}
